---
title: ""
output:
  pdf_document:
    latex_engine: xelatex
    toc: false
    toc_depth: 3
geometry: "left=1cm,right=1cm,top=1.1cm,bottom=1.1cm"
fontsize: 12pt
params:
  country: "Default"
header-includes:
  - \usepackage{enumitem}
  - \geometry{a4paper, left=15mm, right=15mm, top=15mm, bottom=20mm}
  - \setlist[itemize,1]{label=\textbullet}
  - \setlist[itemize,2]{label=\(\circ\)}
  - \setlist[itemize,3]{label=\(\infty\)}
  - \setmainfont{Arial}
  - \usepackage{hyperref}
  - \hypersetup{colorlinks=true, linkcolor=blue, urlcolor=blue, breaklinks=true}
  - \usepackage{tcolorbox}
  - \usepackage{fancyhdr}
  - \setlength{\headheight}{15pt}
  - \addtolength{\topmargin}{-3pt}
  - \definecolor{unicefblue}{RGB}{28, 171, 226}
  - \pagestyle{fancy}
  - \fancyhf{}
  - \fancyhead[L]{}
  - \fancyhead[C]{}
  - \fancyhead[R]{\hyperlink{top}{\small \color{black} $\uparrow$ Back to top}}
  - \fancyfoot[C]{\thepage}
  - \usepackage{tocloft}
  - \renewcommand{\cftsecfont}{\normalfont}
  - \renewcommand{\cftsecpagefont}{\normalfont}
  - \renewcommand{\cftsecleader}{\cftdotfill{\cftdotsep}}
---

\hypertarget{top}{}

```{r setup, include=FALSE}
options(scipen = 999)

# filter to selected country and its region peers
df <- wuenic_dta %>% filter(country == params$country) %>% 
  filter(!vaccine == "pol3") # we are no longer producing estimates for POL3
hpv_df <- hpv_dta %>% filter(country == params$country)
has_hpv_data <- nrow(hpv_df) > 0
has_hpvc <- has_hpv_data && any(hpv_df$vaccine_code == "PRHPVC_F")

add_region_names <- function(df) {
  df <- df %>% 
    mutate(Region_Name = case_match(Region,
      "ROSA"          ~ "South Asia",
      "ESAR"          ~ "Eastern and Southern Africa",
      "ECAR"          ~ "Europe and Central Asia",
      "MENA"          ~ "Middle East and North Africa",
      "LACR"          ~ "Latin America and the Caribbean",
      "WCAR"          ~ "West and Central Africa",
      "EAPR"          ~ "East Asia and the Pacific",
      "Non-programme" ~ "Non-programme countries",
      .default        = Region
    ))
}
df <- add_region_names(df)
hpv_df <- add_region_names(hpv_df)

# pull region and iso3c for this country
ctry_iso   <- unique(df$iso3c)
region <- unique(df$Region)
ctry_regn <- df %>%
  distinct(Region_Name, Region) %>%
  mutate(formatted = str_glue("{Region_Name} ({Region})")) %>%
  pull(formatted)

# regional peers (all countries in same region, for comparisons)
reg_peers <- wuenic_dta %>% filter(Region == region, lvl_1 == "country")
hpv_peers <- hpv_dta %>% filter(Region == region, lvl_1 == "country")

# get available vaccines for this country
available_vaccines <- unique(df$vaccine)

# subnational data (admin1 within country) -- update lvl_2 value to match your data
sub_df <- wuenic_dta %>% filter(iso3c == ctry_iso)

wuenic_src <- str_glue("Source: WHO/UNICEF Estimates of National Immunization Coverage, {rev_yr} revision")

# countries that should have "the" before them
the_countries <- c(
  "United Arab Emirates", "Bahamas", "Central African Republic", "DRC", "Congo",
  "Dominican Republic", "Gambia", "Maldives", "Marshall Islands", "Netherlands (Kingdom of the)",
  "Philippines", "Republic of Korea", "Republic of Moldova", "Russian Federation",
  "Solomon Islands", "State of Palestine", "United Kingdom", "United States"
)

df <- df %>%
  mutate(the_country = ifelse(country %in% the_countries, paste("the", country), country))

ctry_name  <- unique(df$country)
the_ctry_name <- unique(df$the_country)

# whether this country has introduced hpv
hpv_intro <- wiise_hpv_intro_yrs %>%
  filter(iso3c == ctry_iso) %>%
  mutate(introduced = case_when(nationwide == "yes" | partially == "yes" ~ 1, TRUE ~ 0))
  
hpv_intro_flag <- ifelse(nrow(hpv_intro) > 0 && any(hpv_intro$introduced == 1, na.rm = TRUE), 1, 0)

# get flag for header
# flag download
if (!require("countrycode", quietly = TRUE)) install.packages("countrycode")
library(countrycode)
iso2 <- tolower(countrycode(ctry_name, origin = "country.name", destination = "iso2c"))

# set micronesia manually
if (ctry_name == "Micronesia") {
  iso2 <- "fm"
}

flag_path <- paste0("flags/", iso2, ".png")

# download if not already present
if (!file.exists(flag_path)) {
  dir.create("flags", showWarnings = FALSE)
  download.file(paste0("https://flagcdn.com/160x120/", iso2, ".png"), destfile = flag_path, mode = "wb")
}
```


```{r banner, echo=FALSE, results='asis'}
cat(paste0('
\\begin{tcolorbox}[
  colback=unicefblue,
  colframe=unicefblue,
  arc=6pt,
  boxrule=0pt,
  left=4mm,
  right=4mm,
  top=3mm,
  bottom=3mm,
  halign=center,
  valign=center
]
\\begin{center}
  {\\LARGE \\color{white} \\textbf{WUENIC Talking Points: ', ctry_name, '}} \\hspace{0.2cm}\\includegraphics[height=0.6cm]{', flag_path, '}\\\\[0.4em]
  {\\large \\color{white} Revision Year: ', rev_yr, '}
\\end{center}
\\end{tcolorbox}
\\vspace{0.4in}
'))
```

```{r toc, echo=FALSE, results='asis'}
cat("\\tableofcontents")
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}

region_for_country <- base_map_df %>% 
  filter(iso3c == ctry_iso) %>%
  pull(map_region)

map_df <- base_map_df %>%
  filter(map_region == region_for_country)

# shift coordinates for oceania (use 0–360 system)
if (!is.na(region_for_country) && region_for_country == "Oceania") {
  map_df <- st_shift_longitude(map_df)
}

# flag country
map_df <- map_df %>%
  mutate(highlight = ifelse(iso3c == ctry_iso, "Country", "Region"))

# coord limits for maps
if (!is.na(region_for_country)) {
  limits <- switch(region_for_country,
                   "Africa"   = list(x = c(-24, 56),    y = c(-40, 46)),
                   "Americas" = list(x = c(-170, -23),  y = c(-60, 85)),
                   "Asia"     = list(x = c(25, 150),    y = c(-15, 60)),
                   "Europe" = list(x = c(-25, 100), y = c(34, 75)),
                   "Oceania"  = list(x = c(110, 240),   y = c(-50, 5)),  # 0-360 space!
                   list(x = c(-180, 180), y = c(-90, 90))
  )
  
  xlim_val <- limits$x
  ylim_val <- limits$y
}

if (!is.na(region_for_country) && region_for_country == "Oceania") {
  pacific_crs <- st_crs("+proj=longlat +lon_0=180")
  map_df <- st_transform(map_df, pacific_crs)
  xlim_val <- c(-70, 15)
  ylim_val <- c(-50, 3)
}

if (params$country == "Russian Federation") {
  x_lim_val <- c(-25, 200)
  ylim_val <- c(34, 85)
}

# compute centroid to calculate country size and circle small ones on map
target_country_geo <- map_df %>% filter(highlight == "Country")

is_small_country <- as.numeric(st_area(target_country_geo)) < 5e9 

if (is_small_country) {
  country_centroid <- st_centroid(target_country_geo)
}

# plot
country_map <- ggplot(map_df) +
  geom_sf(aes(fill = highlight), color = "white", size = 0.2) +
  { if (is_small_country) 
    geom_sf(data = country_centroid, inherit.aes = FALSE, size = 6, shape = 21, color = "#1CABE2", stroke = 1.5)
  } + 
  scale_fill_manual(values = c("Country" = "#1CABE2", "Region" = "#d7d7d7"), name = NULL) +
  coord_sf(xlim = xlim_val, ylim = ylim_val, expand = FALSE) +
  theme_void() +
  theme(legend.position = "none", 
        plot.title = element_markdown(hjust = 0.5, size = 14, face = "bold"),
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.5)) +
  labs(title = paste0("<span style='color:#1CABE2;'>", ctry_name, "</span> in '", region_for_country, "' region")) #+
  # annotation_north_arrow(location = "tr", which_north = "true", style = north_arrow_fancy_orienteering(),
  #                        height = unit(1, "cm"), width = unit(1, "cm")) +
  # annotation_scale(location = "bl", width_hint = 0.3)

```

\vspace{2em}
```{r ggplot-country-map, echo=FALSE, warning=FALSE, fig.align='center'}
suppressMessages(print(country_map))
```

```{r setup-vax, include=FALSE}
# key vaccines and years
vax <- df %>%
  filter(vaccine %in% c("dtp1", "dtp3", "mcv1", "mcv2"), year %in% c(comp_yr, rev_yr - 1, rev_yr))

# unvaccinated labels
unvac_lbl <- vax %>%
  select(Region, iso3c, country, year, vaccine, unvaccinated_lbl) %>%
  pivot_wider(names_from = year, values_from = unvaccinated_lbl) %>%
  rename_with(~paste0(., "_lbl"), .cols = (ncol(.) - 2):ncol(.))

# unvaccinated raw values
unvac <- vax %>%
  select(Region, iso3c, country, year, vaccine, unvaccinated) %>%
  pivot_wider(names_from = year, values_from = unvaccinated) %>%
  left_join(., unvac_lbl) %>%
  distinct() %>%
  mutate(pchng = round((.[[7]] / .[[5]] - 1) * 100, 0),
         pchng_txt = case_when(
           .[[8]] == .[[10]] ~ str_glue("remained relatively stable between {comp_yr} and {rev_yr} at {.[[10]]}"),
           pchng < 0 ~ str_glue("declined {abs(pchng)}% from {.[[8]]} in {comp_yr} to {.[[10]]} in {rev_yr}"),
           pchng > 0 ~ str_glue("increased {pchng}% from {.[[8]]} in {comp_yr} to {.[[10]]} in {rev_yr}"),
           TRUE ~ NA_character_))

# vaccinated
vacc <- vax %>%
  select(Region, iso3c, country, year, vaccine, vaccinated) %>%
  pivot_wider(names_from = year, values_from = vaccinated) %>%
  mutate(pchng = round((.[[7]] / .[[5]] - 1) * 100, 1),
         pchng_txt = case_when(pchng > 0 ~ "increased", pchng < 0 ~ "decreased", TRUE ~ "remained the same")) %>%
  label_vals_millions(.[[5]], "vacc_comp_yr_lbl") %>%
  label_vals_millions(.[[7]], "vacc_rev_yr_lbl")

# coverage
cvg <- vax %>%
  select(Region, iso3c, country, year, vaccine, coverage) %>%
  pivot_wider(names_from = year, values_from = coverage)

# target population (surviving infants), dtp1 only
tgt <- vax %>%
  filter(vaccine == "dtp1") %>%
  select(Region, iso3c, country, year, vaccine, target) %>%
  pivot_wider(names_from = year, values_from = target) %>%
  mutate(diff = .[[7]] - .[[5]],
         moreless = case_when(diff < 0 ~ "fewer", diff > 0 ~ "more", TRUE ~ "same"),
         to_vax = case_when(
           diff < 0 ~ "for vaccine coverage to increase, the number of children vaccinated needs to either increase or decline at a slower rate than the decline in surviving infant target population",
           diff > 0 & diff < 1000000 ~ "vaccinating approximately the same number of children as currently reached will be sufficient due to the anticipated decrease or relatively small increases in the birth cohort",
           diff > 0 & diff >= 1000000 ~ "maintaining current coverage requires vaccinating an increasing number of children, which will require substantial increases in immunization programme and health system capacity")) %>%
  label_vals(diff, "diff_lbl") %>%
  mutate(diff_abs = abs(diff)) %>%
  label_vals_millions(diff_abs, "diff_abs_lbl")
```

```{r setup-overview, include=FALSE}
# vaccines achieving >= 90% in this country in rev_yr
gt90 <- df %>%
  select(-target_grp) %>%
  distinct() %>%
  filter(year == rev_yr) %>%
  mutate(gt90 = ifelse(coverage >= 90, 1, 0))

n_vax_total <- n_distinct(gt90$vaccine)
n_vax_gt90  <- sum(gt90$gt90, na.rm = TRUE)

# coverage change prev_yr to rev_yr for the key vaccines
cvg_chng <- vax %>%
  filter(year %in% c(rev_yr, rev_yr - 1)) %>%
  select(iso3c, country, vaccine, year, coverage) %>%
  pivot_wider(names_from = year, values_from = coverage) %>%
  mutate(diff = .[[5]] - .[[4]],
         diff_txt = case_when(
           diff > 0 ~ str_glue("increased from {.[[4]]}% in {rev_yr-1} to {.[[5]]}% in {rev_yr}"),
           diff < 0 ~ str_glue("declined from {.[[4]]}% in {rev_yr-1} to {.[[5]]}% in {rev_yr}"),
           TRUE ~ str_glue("was the same in {rev_yr} as it was in {rev_yr-1} ({.[[5]]}%)")))
```

```{r setup-progress-ia2030-zd-target, include=FALSE}
# zero-dose (unvaccinated dtp1) for this country (dtp1 is used as a proxy for the IA2030 goal)
zd <- df %>%
  filter(vaccine == "dtp1", year >= 2000) %>%
  select(iso3c, country, year, unvaccinated)

# ia2030 annual target (halve 2019 zero-dose by 2030)
zd_2019 <- zd %>% filter(year == 2019) %>% pull(unvaccinated)
ia2030_annual_target <- zd_2019 / 2

target <- zd %>%
  filter(year == 2019) %>% # baseline data only
  mutate(target_2030 = unvaccinated / 2, # defines the goal: 50% reduction in unvaccinated people by 2030
         factor = (unvaccinated - target_2030) / (2030 - 2019)) %>% # annual reduction rate needed: total drop divided by the 11-year gap
  crossing(year_new = 2020:2030) %>% # "join" that duplicates each country's row for every year from 2020 to 2030
  mutate(unvaccinated_new = unvaccinated - factor * (year_new - 2019)) %>% # applies linear formula to calculate the target number of zero-dose children
  select(iso3c, country, year = year_new, unvaccinated_new)

# text for ia2030 comparison in rev_yr
zd_rev_yr <- zd %>% filter(year == rev_yr) %>% pull(unvaccinated)
ia2030_rev_yr <- target %>% filter(year == rev_yr) %>% pull(unvaccinated_new)
zd_pcnt_diff  <- round((zd_rev_yr / ia2030_rev_yr - 1) * 100, 0)
zd_diff_txt   <- case_when(zd_pcnt_diff > 0 ~ "higher than", zd_pcnt_diff < 0 ~ "lower than", TRUE ~ "the same as")

zd_rev_yr_lbl     <- label_vals_millions_num(zd_rev_yr)
ia2030_rev_yr_lbl <- label_vals_millions_num(ia2030_rev_yr)
```

<!-- ===================== PAGE 1: OVERVIEW ===================== -->

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{Overview}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:overview}{}")
cat("\\addcontentsline{toc}{section}{Overview}")
```

- `r if (n_vax_gt90 == 0) "None" else n_vax_gt90` of the `r n_vax_total` childhood vaccines tracked for `r the_ctry_name` achieved coverage of 90% or higher in `r rev_yr`.

- Coverage of DTP1 `r cvg_chng %>% filter(vaccine == "dtp1") %>% pull(diff_txt)`, DTP3 `r cvg_chng %>% filter(vaccine == "dtp3") %>% pull(diff_txt)`, and MCV1 `r cvg_chng %>% filter(vaccine == "mcv1") %>% pull(diff_txt)`.

- In `r rev_yr`, the number of zero-dose children for DTP in `r the_ctry_name` (`r zd_rev_yr_lbl`) was approximately `r abs(zd_pcnt_diff)`% `r zd_diff_txt` the annual goal to achieve the IA2030 target of halving zero-dose children for DTP by 2030 (`r ia2030_rev_yr_lbl`).


```{r setup-heatmap-overview, include=FALSE}

# filter hpv data to bind with other vaccines
if (has_hpvc) {
  hpv_merge <- hpv_df %>% 
    filter(vaccine_code == "PRHPVC_F") %>%
    select(-vaccine) %>% 
    rename(vaccine = vaccine_code)
  
  # only bind columns that exist in both
  shared_cols <- intersect(names(df), names(hpv_merge))
  df <- bind_rows(
    df %>% select(all_of(shared_cols)),
    hpv_merge %>% select(all_of(shared_cols))
  )
}
# heatmap for all vaccines, this country only
heatmap_ovw_df <- df %>%
  mutate(year = as.character(year),
         vaccine = case_when(
           vaccine == "hepbb" ~ "HepBB", vaccine == "hepb3" ~ "HepB3",
           vaccine == "hib3"  ~ "Hib3",  vaccine == "rotac" ~ "RotaC",
           vaccine == "menga" ~ "MengA", vaccine == "PRHPVC_F" ~ "HPVc",
           TRUE ~ toupper(vaccine))) %>%
  mutate(cvg_cat = case_when(
    coverage < 60 ~ "<60", coverage %in% 60:69 ~ "60-69",
    coverage %in% 70:79 ~ "70-79", coverage %in% 80:89 ~ "80-89",
    coverage %in% 90:94 ~ "90-94", coverage >= 95 ~ ">=95", TRUE ~ NA_character_))

heatmap_ovw_df$vaccine <- factor(heatmap_ovw_df$vaccine,
  levels = c("BCG", "HepBB", "DTP1", "DTP3", "Hib3", "HepB3", "PCVC", "RotaC", "IPV1", "IPVC", "MCV1", "RCV1", "MCV2", "YFV", "MengA", "HPVc"))

heatmap_overview <-
  heatmap_ovw_df %>%
  ggplot(aes(x = year, y = vaccine, fill = coverage)) +
  geom_tile(color = "white") +
  geom_text(data = heatmap_ovw_df, aes(label = coverage), colour = "white", vjust = 0.5, size = 3) +
  labs(x = "", y = "",
       title = str_glue("Vaccine coverage (%), {ctry_name}, 2000–{rev_yr}"),
       caption = wuenic_src) +
  scale_x_discrete(position = "top") +
  scale_y_discrete(limits = rev) +
  theme_minimal() +
  theme(panel.grid = element_blank(), 
        legend.position = "bottom", 
        legend.key.width = unit(1.2, "cm"),
        plot.title = element_text(size = 14, hjust = 0.5, face = "bold")) +
  scale_fill_gradientn(
    colours = c("#B50800", "#E2231A", "#F26A21", "#FFC20E", "#80BD41", "#00833D"),
    values  = scales::rescale(c(0, 59, 69, 79, 89, 94, 100)),
    guide = "colorbar", limits = c(0, 100),
    name = "Vaccine coverage (%)  ",
    breaks = c(0, 59, 69, 79, 89, 100), labels = c("0", "60", "70", "80", "90", "100"))
```

```{r overview-line-plot, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center'}
# prep plot data from heatmap_ovw_df (already has cleaned vaccine names and categories)
all_vax_plot <- heatmap_ovw_df %>%
  mutate(year = as.numeric(year)) %>%
  filter(!is.na(coverage))

# color palette - one color per vaccine
vaccine_colors <- c(
  "BCG"    = "#1CABE2", "HepBB"  = "#0058AB", "DTP1"   = "#00833D",
  "DTP3"   = "#80BD41", "Hib3"   = "#6A1E74", "HepB3"  = "#961A49",
  "PCVC"   = "#E2231A", "RotaC"  = "#F26A21", "IPV1"   = "#B50800", 
  "IPVC"   = "#FF8C00", "MCV1"   = "#2E7EBB", "RCV1"   = "#9E9E9E", 
  "MCV2"   = "#FFB6C1", "YFV"    = "#5B4A9C", "MengA"  = "#C8A900", 
  "HPVc"   = "#D45F9E"
)

# first data point per vaccine for labels
first_points <- all_vax_plot %>%
  group_by(vaccine) %>%
  filter(year == min(year)) %>%
  ungroup()

num_vaccines <- n_distinct(all_vax_plot$vaccine)

min_coverage <- min(all_vax_plot$coverage, na.rm = TRUE)

all_vax_plot <- ggplot(all_vax_plot, aes(x = year, y = coverage, color = vaccine, group = vaccine)) +
  geom_line(size = 0.8, alpha = 0.7) +
  geom_point(size = 1.5, alpha = 0.7) +
  geom_hline(yintercept = 90, linetype = "dashed") +
  geom_text(data = first_points, aes(label = vaccine, color = vaccine, face = "bold"),
            hjust = 1.2, size = 2.5, show.legend = FALSE) +
  scale_color_manual(values = vaccine_colors) +
  scale_x_continuous(breaks = seq(2000, rev_yr, 1), expand = expansion(add = c(2, 0.5))) +
  scale_y_continuous(limits = c(min_coverage - 5, 100), breaks = seq(0, 100, by = 10), name = "Coverage (%)") +
  labs(
    title = str_glue("Vaccine coverage over time, {ctry_name}, 2000–{rev_yr}"),
    caption = paste0("Dotted line shows the IA2030 target of 90% coverage\n", wuenic_src),
    x = "Year", color = NULL
  ) +
  theme_unicef() +
  theme(
    panel.grid.minor = element_blank(), panel.grid.major.x = element_line(),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
    plot.title = element_text(size = 14, hjust = 0.5, face = "bold"),
    legend.position = "right",
    legend.text = element_text(size = 8),
    legend.key.height = unit(0.6, "cm")
  ) +
  guides(color = guide_legend(nrow = num_vaccines))
```

```{r overview-plot, echo=FALSE, fig.width=10, fig.height=5.5}
print(heatmap_overview)
```

```{r all_vax_plot, echo=FALSE, fig.width=10, fig.height=5.5}
print(all_vax_plot)
```

\newpage

<!-- ===================== PAGE 2: DTP1 ===================== -->

```{r setup-dtp1, include=FALSE}
v <- "dtp1"
unvac_v <- unvac %>% filter(vaccine == v)
cvg_v   <- cvg   %>% filter(vaccine == v)
vacc_v  <- vacc  %>% filter(vaccine == v)

# dtp1 coverage for this country in rev_yr and comp_yr
ctry_cvg_rev  <- cvg_v %>% pull(7)   # rev_yr column
ctry_cvg_comp <- cvg_v %>% pull(5)   # comp_yr column

# regional rank (among peers) in rev_yr
reg_rank <- reg_peers %>%
  filter(vaccine == v, year == rev_yr) %>%
  arrange(desc(coverage)) %>%
  mutate(rank = row_number()) %>%
  filter(iso3c == ctry_iso)

n_peers <- reg_peers %>% filter(vaccine == v, year == rev_yr) %>% n_distinct("iso3c")

# # subnational variation
# sub_dtp1 <- sub_df %>%
#   filter(vaccine == v, year == rev_yr) %>%
#   summarise(min_cvg = min(coverage, na.rm = TRUE), max_cvg = max(coverage, na.rm = TRUE),
#             range_cvg = max(coverage, na.rm = TRUE) - min(coverage, na.rm = TRUE),
#             n_admin1 = n_distinct(admin1_master))
# 
# sub_variation_txt <- case_when(
#   sub_dtp1$range_cvg >= 20 ~ str_glue("Subnational variation is high, with coverage ranging from {sub_dtp1$min_cvg}% to {sub_dtp1$max_cvg}% across {sub_dtp1$n_admin1} admin1 areas"),
#   sub_dtp1$range_cvg >= 10 ~ str_glue("Subnational variation is moderate, with coverage ranging from {sub_dtp1$min_cvg}% to {sub_dtp1$max_cvg}% across {sub_dtp1$n_admin1} admin1 areas"),
#   TRUE ~ str_glue("Subnational variation is low, with coverage ranging from {sub_dtp1$min_cvg}% to {sub_dtp1$max_cvg}% across {sub_dtp1$n_admin1} admin1 areas"))

# dtp1->dtp3 dropout
dropout_dtp <- df %>%
  filter(year %in% c(comp_yr, rev_yr), vaccine %in% c("dtp1", "dtp3")) %>%
  select(iso3c, country, year, vaccine, vaccinated) %>%
  pivot_wider(names_from = vaccine, values_from = vaccinated) %>%
  mutate(dropout = round(((dtp1 - dtp3) / dtp1) * 100, 0),
         dropout_txt = case_when(
           dropout < 5  ~ "Low DTP dropout rates imply good ability to provide a complete vaccine series early in life",
           dropout <= 10 ~ "Medium DTP dropout rates imply moderate ability to provide a complete vaccine series early in life",
           dropout > 10  ~ "High DTP dropout rates imply poor ability to provide a complete vaccine series early in life"))

dropout_comp_dtp <- dropout_dtp %>%
  select(-c(dtp1, dtp3, dropout_txt)) %>%
  pivot_wider(names_from = year, values_from = dropout) %>%
  mutate(diff = .[[4]] - .[[3]],
         diff_txt = case_when(
           diff == 1 ~ str_glue("{diff} percentage point higher than in {comp_yr} ({.[[3]]}%)"),
           diff > 0 ~ str_glue("{diff} percentage points higher than in {comp_yr} ({.[[3]]}%)"),
           diff == -1 ~ str_glue("{abs(diff)} percentage point lower than in {comp_yr} ({.[[3]]}%)"),
           diff < -1 ~ str_glue("{abs(diff)} percentage points lower than in {comp_yr} ({.[[3]]}%)"),
           TRUE ~ str_glue("the same as in {comp_yr}")))
```


```{r dtp1-line-plot, echo=FALSE, warning=FALSE, message=FALSE}
# prep dtp1 coverage over time
plot_dtp1 <- df %>% filter(vaccine == "dtp1") %>% select(year, coverage)

# get number unvaccinated for secondary y-axis
plot_zd_dtp <- df %>%
  filter(vaccine %in% c("dtp1")) %>% 
   select(year, unvaccinated)

# scaling factor for secondary axis
scale_factor <- max(plot_dtp1$coverage, na.rm = TRUE) / max(plot_zd_dtp$unvaccinated, na.rm = TRUE)

# plot
dtp_plot <- ggplot() +
  geom_col(data = plot_zd_dtp, aes(x = year, y = unvaccinated * scale_factor, fill = "Zero-dose children"),
           alpha = 0.4, width = 0.6) +
  geom_line(data = plot_dtp1, aes(x = year, y = coverage, color = "DTP1 coverage"), size = 0.8) +
  geom_point(data = plot_dtp1, aes(x = year, y = coverage, color = "DTP1 coverage"), size = 1.5) +
  geom_hline(yintercept = 90, linetype = "dashed") +
  scale_color_manual(name = NULL, values = c("DTP1 coverage" = "#1CABE2")) +
  scale_fill_manual(name = NULL, values = c("Zero-dose children" = "#E2231A")) +
  scale_x_continuous(breaks = seq(min(plot_dtp1$year), max(plot_dtp1$year), 1)) +
  scale_y_continuous(limits = c(0, 100), name = "Coverage (%)",
                     sec.axis = sec_axis(~ . / scale_factor,
                                         name = "Zero-dose children",
                                         labels = scales::label_comma())) +
  labs(
    title = str_glue("DTP1 coverage and zero-dose children over time, {ctry_name}, 2000–{rev_yr}"),
    caption = paste0("Dotted line shows the IA2030 target of 90% coverage\n", wuenic_src),
    x = "Year"
  ) +
  guides(color = guide_legend(order = 1), fill = guide_legend(order = 2)) +
  theme_unicef() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_line(),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
    plot.title = element_text(size = 14, hjust = 0.5, face = "bold"),
    legend.position = "bottom"
  )
```

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{DTP1: Reaching zero-dose children}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:overview}{}")
cat("\\addcontentsline{toc}{section}{DTP1: Reaching zero-dose children}")
```


- In `r rev_yr`, DTP1 coverage in `r the_ctry_name` was `r ctry_cvg_rev`%, compared to `r ctry_cvg_comp`% in `r comp_yr`.

- In `r rev_yr`, the number of zero-dose children in `r the_ctry_name` (`r zd_rev_yr_lbl`) was approximately `r abs(zd_pcnt_diff)`% `r zd_diff_txt` the annual IA2030 target (`r ia2030_rev_yr_lbl`).

- With `r tgt$diff_abs_lbl` `r tgt$moreless` surviving infants in `r rev_yr` compared to `r comp_yr`, `r tgt$to_vax`.

- Among `r ctry_regn` region peers, `r the_ctry_name` ranked `r reg_rank$rank` out of `r n_peers` countries in DTP1 coverage in `r rev_yr` (`r ctry_cvg_rev`% vs. regional peers).

- DTP1-to-DTP3 dropout in `r rev_yr` was `r dropout_dtp %>% filter(year == rev_yr) %>% pull(dropout)`%, `r dropout_comp_dtp %>% pull(diff_txt)`. `r dropout_dtp %>% filter(year == rev_yr) %>% pull(dropout_txt)`.

\vspace{1em}

```{r dtp1, echo=FALSE, fig.width=10, fig.height=6.2}
print(dtp_plot)
```


```{r setup-heatmap-dtp1-sub, include=FALSE}
# # heatmap of dtp1 by admin1 over time
# heatmap_sub_dtp1 <-
#   sub_df %>%
#   filter(vaccine == "dtp1", year >= 2000) %>%
#   mutate(year = as.factor(year)) %>%
#   ggplot(aes(x = year, y = reorder(admin1_master, coverage), fill = coverage)) +
#   geom_tile(color = "white") +
#   geom_text(aes(label = coverage), colour = "black", size = 2.8) +
#   labs(x = "", y = "",
#        title = str_glue("DTP1 coverage (%), by admin1, {ctry_name}, 2000-{rev_yr}"),
#        caption = wuenic_src) +
#   scale_x_discrete(position = "top") +
#   theme_minimal() +
#   theme(panel.grid = element_blank(), legend.position = "bottom", legend.key.width = unit(1.2, "cm"),
#         axis.text.x = element_text(angle = 45, hjust = 0, size = 9)) +
#   scale_fill_gradientn(
#     colours = c("#B50800", "#E2231A", "#F26A21", "#FFC20E", "#80BD41", "#00833D"),
#     values  = scales::rescale(c(0, 59, 69, 79, 89, 94, 100)),
#     guide = "colorbar", limits = c(0, 100),
#     name = "Vaccine coverage (%)  ",
#     breaks = c(0, 59, 69, 79, 89, 100), labels = c("0", "60", "70", "80", "90", "100"))
```

```{r ggplot-dtp1-sub, echo=FALSE, fig.width=10, fig.height=6.5}
#print(heatmap_sub_dtp1)
```

<!-- ===================== PAGE 3: DTP3 ===================== -->

```{r setup-dtp3, include=FALSE}
v <- "dtp3"
unvac_v <- unvac %>% filter(vaccine == v)
cvg_v   <- cvg   %>% filter(vaccine == v)
vacc_v  <- vacc  %>% filter(vaccine == v)

ctry_cvg_rev_dtp3  <- cvg_v %>% pull(7)
ctry_cvg_comp_dtp3 <- cvg_v %>% pull(5)

# regional rank
reg_rank_dtp3 <- reg_peers %>%
  filter(vaccine == v, year == rev_yr) %>%
  arrange(desc(coverage)) %>%
  mutate(rank = row_number()) %>%
  filter(iso3c == ctry_iso)

# un- and under-vaccinated (unvaccinated for dtp3)
unvac_dtp3_rev  <- unvac_v %>% pull(7)
unvac_dtp3_lbl  <- label_vals_millions_num(unvac_dtp3_rev)

# population growth context
tgt_chng <- tgt %>%
  mutate(pop_grwth = round((.[[7]] / .[[5]] - 1) * 100, 0)) %>%
  label_vals_millions(.[[7]], "tgt_rev_yr") %>%
  label_vals_millions(.[[5]], "tgt_comp_yr")

pop_chng_txt <- if (ctry_cvg_rev_dtp3 < ctry_cvg_comp_dtp3 & tgt[[7]] > tgt[[5]]) {
  str_glue("Due to population growth of approximately {tgt_chng$pop_grwth}% between {comp_yr} and {rev_yr}, DTP3 coverage in {rev_yr} ({ctry_cvg_rev_dtp3}%) was lower than in {comp_yr} ({ctry_cvg_comp_dtp3}%)")
} else { "" }

# # subnational variation for dtp3
# sub_dtp3 <- sub_df %>%
#   filter(vaccine == v, year == rev_yr) %>%
#   summarise(min_cvg = min(coverage, na.rm = TRUE), max_cvg = max(coverage, na.rm = TRUE),
#             range_cvg = max(coverage, na.rm = TRUE) - min(coverage, na.rm = TRUE),
#             n_admin1 = n_distinct(admin1_master))
# 
# sub_var_dtp3_txt <- case_when(
#   sub_dtp3$range_cvg >= 20 ~ str_glue("Subnational variation is high, with coverage ranging from {sub_dtp3$min_cvg}% to {sub_dtp3$max_cvg}% across {sub_dtp3$n_admin1} admin1 areas"),
#   sub_dtp3$range_cvg >= 10 ~ str_glue("Subnational variation is moderate, with coverage ranging from {sub_dtp3$min_cvg}% to {sub_dtp3$max_cvg}% across {sub_dtp3$n_admin1} admin1 areas"),
#   TRUE ~ str_glue("Subnational variation is low, with coverage ranging from {sub_dtp3$min_cvg}% to {sub_dtp3$max_cvg}% across {sub_dtp3$n_admin1} admin1 areas"))
```

```{r dtp3-line-plot, echo=FALSE, warning=FALSE, message=FALSE}
# prep dtp3 coverage over time
plot_dtp3 <- df %>% filter(vaccine == "dtp3") %>% select(year, coverage)

# get number unvaccinated for secondary y-axis
plot_zd_dtp3 <- df %>%
  filter(vaccine %in% c("dtp3")) %>% 
   select(year, unvaccinated)

# scaling factor for secondary axis
scale_factor <- max(plot_dtp3$coverage, na.rm = TRUE) / max(plot_zd_dtp3$unvaccinated, na.rm = TRUE)

# plot
dtp3_plot <- ggplot() +
  geom_col(data = plot_zd_dtp, aes(x = year, y = unvaccinated * scale_factor, fill = "Un- or Under-Vaccinated Children"),
           alpha = 0.4, width = 0.6) +
  geom_line(data = plot_dtp3, aes(x = year, y = coverage, color = "DTP3 coverage"), size = 0.8) +
  geom_point(data = plot_dtp3, aes(x = year, y = coverage, color = "DTP3 coverage"), size = 1.5) +
  geom_hline(yintercept = 90, linetype = "dashed") +
  scale_color_manual(name = NULL, values = c("DTP3 coverage" = "#1CABE2")) +
  scale_fill_manual(name = NULL, values = c("Un- or Under-Vaccinated Children" = "#E2231A")) +
  geom_hline(yintercept = 90, linetype = "dashed") +
  scale_color_manual(values = c("DTP3 coverage" = "#1CABE2", "Un- or Under-Vaccinated Children" = "#E2231A")) +
  scale_x_continuous(breaks = seq(min(plot_dtp3$year), max(plot_dtp3$year), 1)) +
  scale_y_continuous(limits = c(0, 100), name = "Coverage (%)", sec.axis = sec_axis(~ . / scale_factor, 
                                                                                    name = "Un- or Under-Vaccinated Children", 
                                                                                    labels = scales::label_comma())) +
  labs(
    title = str_glue("DTP3 coverage and un- or under-vaccinated children over time, {ctry_name}, 2000–{rev_yr}"),
    caption = paste0("Dotted line shows the IA2030 target of 90% coverage\n", wuenic_src),
    x = "Year",
    color = NULL
  ) +
  theme_unicef() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_line(),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
    plot.title = element_text(size = 14, hjust = 0.5, face = "bold"),
    legend.position = "bottom"
  )
```

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{DTP3: A marker of routine immunization service delivery to children}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:overview}{}")
cat("\\addcontentsline{toc}{section}{DTP3: A marker of routine immunization service delivery to children}")
```

- The number of children vaccinated with DTP3 `r vacc_v %>% pull(pchng_txt)` from `r vacc_v %>% pull(vacc_comp_yr_lbl)` in `r comp_yr` to `r vacc_v %>% pull(vacc_rev_yr_lbl)` in `r rev_yr`. `r pop_chng_txt`

- In `r rev_yr`, `r dropout_dtp %>% filter(year == rev_yr) %>% pull(dropout)`% of children who received DTP1 did not receive DTP3. This `r dropout_dtp %>% filter(year == rev_yr) %>% pull(dropout)`% dropout was `r dropout_comp_dtp %>% pull(diff_txt)`. `r dropout_dtp %>% filter(year == rev_yr) %>% pull(dropout_txt)`.

- In `r rev_yr`, DTP3 coverage in `r the_ctry_name` was `r ctry_cvg_rev_dtp3`%, leaving `r unvac_dtp3_lbl` children un- or under-vaccinated.

- Among `r ctry_regn` region peers, `r the_ctry_name` ranked `r reg_rank_dtp3$rank` out of `r n_peers` countries in DTP3 coverage in `r rev_yr`.

\vspace{1em}

```{r dtp3, echo=FALSE, fig.width=10, fig.height=6.2}
print(dtp3_plot)
```

```{r setup-heatmap-dtp3-sub, include=FALSE}
# heatmap_sub_dtp3 <-
#   sub_df %>%
#   filter(vaccine == "dtp3", year >= 2000) %>%
#   mutate(year = as.factor(year)) %>%
#   ggplot(aes(x = year, y = reorder(admin1_master, coverage), fill = coverage)) +
#   geom_tile(color = "white") +
#   geom_text(aes(label = coverage), colour = "black", size = 2.8) +
#   labs(x = "", y = "",
#        title = str_glue("DTP3 coverage (%), by admin1, {ctry_name}, 2000-{rev_yr}"),
#        caption = wuenic_src) +
#   scale_x_discrete(position = "top") +
#   theme_minimal() +
#   theme(panel.grid = element_blank(), legend.position = "bottom", legend.key.width = unit(1.2, "cm"),
#         axis.text.x = element_text(angle = 45, hjust = 0, size = 9)) +
#   scale_fill_gradientn(
#     colours = c("#B50800", "#E2231A", "#F26A21", "#FFC20E", "#80BD41", "#00833D"),
#     values  = scales::rescale(c(0, 59, 69, 79, 89, 94, 100)),
#     guide = "colorbar", limits = c(0, 100),
#     name = "Vaccine coverage (%)  ",
#     breaks = c(0, 59, 69, 79, 89, 100), labels = c("0", "60", "70", "80", "90", "100"))
```

```{r ggplot-dtp3-sub, echo=FALSE, fig.width=10, fig.height=6.5}
#print(heatmap_sub_dtp3)
```

<!-- ===================== PAGE 4: MCV ===================== -->

```{r setup-mcv, include=FALSE}

if ("mcv2" %in% available_vaccines) {
  v_mcv <- c("mcv1", "mcv2")
} else {
  v_mcv <- c("mcv1")
}
unvac_v <- unvac %>% filter(vaccine %in% v_mcv)
cvg_v   <- cvg   %>% filter(vaccine %in% v_mcv)

# mcv1 coverage change
cvg_mcv1 <- cvg_v %>%
  filter(vaccine == "mcv1") %>%
  mutate(diff = .[[7]] - .[[5]],
         cvg_chng_txt = case_when(
           diff %in% c(-1:1) ~ str_glue("remained relatively stable, within 1 percentage point of the {comp_yr} level"),
           diff < -1 ~ str_glue("was lower than in {comp_yr} ({.[[5]]}%)"),
           diff > 1  ~ str_glue("was higher than in {comp_yr} ({.[[5]]}%)") ))

# mcv2 coverage change
if ("mcv2" %in% available_vaccines) {
  cvg_mcv2 <- cvg_v %>%
    filter(vaccine == "mcv2") %>%
    mutate(diff = .[[7]] - .[[5]],
           cvg_chng_txt = case_when(
             diff %in% c(-1:1) ~ str_glue("remained relatively stable, within 1 percentage point of the {comp_yr} level"),
             diff < -1 ~ str_glue("decreased from {.[[5]]}% in {comp_yr} to {.[[7]]}% in {rev_yr}"),
             diff > 1  ~ str_glue("increased from {.[[5]]}% in {comp_yr} to {.[[7]]}% in {rev_yr}") ))
  
  # unvaccinated mcv1 vs mcv2
  unvac_mcv <- unvac_v %>%
    select(iso3c, country, vaccine, 7) %>%
    pivot_wider(names_from = vaccine, values_from = last_col()) %>%
    mutate(diff = mcv2 - mcv1) %>%
    label_vals_millions(mcv1, "mcv1_lbl") %>%
    label_vals_millions(mcv2, "mcv2_lbl") %>%
    label_vals_millions(diff, "diff_lbl")

} else {
  cvg_mcv2 <- cvg_v %>% 
    mutate(cvg_chng_txt = paste0("data is not available in ", ctry_name, "'s WUENIC data"))
    
  unvac_mcv <- NULL
}  

# regional rank for mcv1
reg_rank_mcv1 <- reg_peers %>%
  filter(vaccine == "mcv1", year == rev_yr) %>%
  arrange(desc(coverage)) %>%
  mutate(rank = row_number()) %>%
  filter(iso3c == ctry_iso)

# dtp1->mcv1 dropout
dropout_mcv <- df %>%
  filter(year %in% c(comp_yr, rev_yr), vaccine %in% c("dtp1", "mcv1")) %>%
  select(iso3c, country, year, vaccine, vaccinated) %>%
  pivot_wider(names_from = vaccine, values_from = vaccinated) %>%
  mutate(dropout = round(((dtp1 - mcv1) / dtp1) * 100, 0),
         dropout_txt = case_when(
           dropout < 5  ~ "Low DTP-MCV dropout rates imply good retention and ability to provide a full course of vaccines in infancy",
           dropout <= 10 ~ "Medium DTP-MCV dropout rates imply moderate retention in immunization programmes",
           dropout > 10  ~ "High DTP-MCV dropout rates imply poor retention in immunization programmes"))

dropout_comp_mcv <- dropout_mcv %>%
  select(-c(dtp1, mcv1, dropout_txt)) %>%
  pivot_wider(names_from = year, values_from = dropout) %>%
  mutate(diff = .[[4]] - .[[3]],
         diff_txt = case_when(
           diff == 1 ~ str_glue("{diff} percentage point higher than in {comp_yr} ({.[[3]]}%)"),
           diff > 0 ~ str_glue("{diff} percentage points higher than in {comp_yr} ({.[[3]]}%)"),
           diff == -1 ~ str_glue("{abs(diff)} percentage point lower than in {comp_yr} ({.[[3]]}%)"),
           diff < -1 ~ str_glue("{abs(diff)} percentage points lower than in {comp_yr} ({.[[3]]}%)"),
           TRUE ~ str_glue("the same as in {comp_yr}")))

# persistently low mcv1 (last 3 years < 80%)
nyrs <- 3; thrshold <- 80
cvg_trend_mcv1 <- df %>%
  filter(vaccine == "mcv1", year %in% seq(rev_yr - nyrs + 1, rev_yr)) %>%
  mutate(ltx = ifelse(coverage < thrshold, 1, 0)) %>%
  summarise(sum = sum(ltx), mean_cvg = mean(coverage, na.rm = TRUE))

capital_the_name <- if_else(str_starts(the_ctry_name, "the "), str_replace(the_ctry_name, "^the ", "The "), the_ctry_name)
persistent_low_txt <- if (cvg_trend_mcv1$sum == nyrs) {
  str_glue("{capital_the_name} has had MCV1 coverage below {thrshold}% for all of the last {nyrs} years")
} else { str_glue("{capital_the_name} has not had MCV1 coverage below {thrshold}% for any of the last {nyrs} years") }

# # subnational variation for mcv1
# sub_mcv1 <- sub_df %>%
#   filter(vaccine == "mcv1", year == rev_yr) %>%
#   summarise(min_cvg = min(coverage, na.rm = TRUE), max_cvg = max(coverage, na.rm = TRUE),
#             range_cvg = max(coverage, na.rm = TRUE) - min(coverage, na.rm = TRUE),
#             n_admin1 = n_distinct(admin1_master))
# 
# sub_var_mcv1_txt <- case_when(
#   sub_mcv1$range_cvg >= 20 ~ str_glue("Subnational variation is high, with MCV1 coverage ranging from {sub_mcv1$min_cvg}% to {sub_mcv1$max_cvg}% across {sub_mcv1$n_admin1} admin1 areas"),
#   sub_mcv1$range_cvg >= 10 ~ str_glue("Subnational variation is moderate, with MCV1 coverage ranging from {sub_mcv1$min_cvg}% to {sub_mcv1$max_cvg}% across {sub_mcv1$n_admin1} admin1 areas"),
#   TRUE ~ str_glue("Subnational variation is low, with MCV1 coverage ranging from {sub_mcv1$min_cvg}% to {sub_mcv1$max_cvg}% across {sub_mcv1$n_admin1} admin1 areas"))
```

```{r mcv-line-plot, echo=FALSE, warning=FALSE, message=FALSE}

# prep mcv1 coverage
plot_mcv1 <- df %>% filter(vaccine == "mcv1") %>% select(year, coverage)

# prep mcv2 coverage only if available
if ("mcv2" %in% available_vaccines) {
  plot_mcv2 <- df %>% filter(vaccine == "mcv2") %>% select(year, coverage)
}

# prep zero-dose (mcv1-based)
plot_zd_mcv <- df %>% filter(vaccine == "mcv1") %>% select(year, unvaccinated)

# scaling factor for secondary axis
scale_factor <- max(plot_mcv1$coverage, na.rm = TRUE) / max(plot_zd_mcv$unvaccinated, na.rm = TRUE)
  
# plot
mcv_plot <- ggplot() +
  geom_col(data = plot_zd_mcv, aes(x = year, y = unvaccinated * scale_factor, fill = "MCV zero-dose children"),
           alpha = 0.4, width = 0.6) +
  # mcv1
  geom_line(data = plot_mcv1, aes(x = year, y = coverage, color = "MCV1 coverage"), size = 0.8) +
  geom_point(data = plot_mcv1, aes(x = year, y = coverage, color = "MCV1 coverage"), size = 1.5) +
  # mcv2 (only if exists)
  {if ("mcv2" %in% available_vaccines)
    list(
      geom_line(data = plot_mcv2, aes(x = year, y = coverage, color = "MCV2 coverage"), size = 0.8),
      geom_point(data = plot_mcv2, aes(x = year, y = coverage, color = "MCV2 coverage"), size = 1.5)
    )} +
  geom_hline(yintercept = 90, linetype = "dashed") +
  scale_color_manual(name = NULL, values = c("MCV1 coverage" = "#1CABE2", "MCV2 coverage" = "#6A1E74")) +
  scale_fill_manual(name = NULL, values = c("MCV zero-dose children" = "#E2231A")) +
  scale_x_continuous(breaks = seq(min(plot_mcv1$year), max(plot_mcv1$year), 1)) +
  scale_y_continuous(limits = c(0, 100), name = "Coverage (%)",
    sec.axis = sec_axis(~ . / scale_factor, name = "MCV zero-dose children", labels = scales::label_comma())) +
  guides(color = guide_legend(order = 1), fill = guide_legend(order = 2)) +
  labs(title = str_glue("MCV coverage and MCV zero-dose children over time, {ctry_name}, 2000–{rev_yr}"),
    caption = paste0("Dotted line shows the IA2030 target of 90% coverage\n", wuenic_src),
    x = "Year") +
  theme_unicef() +
  theme(panel.grid.minor = element_blank(), panel.grid.major.x = element_line(),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
    plot.title = element_text(size = 14, hjust = 0.5, face = "bold"),
    legend.position = "bottom")
```

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{MCV: Canary in the coal mine}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:overview}{}")
cat("\\addcontentsline{toc}{section}{MCV: Canary in the coal mine}")
```

\vspace{1em}

- MCV1 coverage in `r rev_yr` (`r cvg_mcv1 %>% pull(7)`%) `r cvg_mcv1 %>% pull(cvg_chng_txt)`.

- MCV2 coverage `r cvg_mcv2 %>% pull(cvg_chng_txt)`. This leaves `r unvac_mcv$mcv1_lbl` children without any protection against measles and another `r unvac_mcv$diff_lbl` with only partial protection.

- In `r rev_yr`, `r dropout_mcv %>% filter(year == rev_yr) %>% pull(dropout)`% of children who received DTP1 did not receive MCV1. DTP-MCV dropout was `r dropout_comp_mcv %>% pull(diff_txt)`. `r dropout_mcv %>% filter(year == rev_yr) %>% pull(dropout_txt)`.

- `r persistent_low_txt`.

- Among `r ctry_regn` region peers, `r the_ctry_name` ranked `r reg_rank_mcv1$rank` out of `r n_peers` countries in MCV1 coverage in `r rev_yr`.

\vspace{1em}

```{r mcv, echo=FALSE, fig.width=10, fig.height=6.2}
print(mcv_plot)
```

```{r setup-heatmap-mcv1-sub, include=FALSE}
# heatmap_sub_mcv1 <-
#   sub_df %>%
#   filter(vaccine == "mcv1", year >= 2000) %>%
#   mutate(year = as.factor(year)) %>%
#   ggplot(aes(x = year, y = reorder(admin1_master, coverage), fill = coverage)) +
#   geom_tile(color = "white") +
#   geom_text(aes(label = coverage), colour = "black", size = 2.8) +
#   labs(x = "", y = "",
#        title = str_glue("MCV1 coverage (%), by admin1, {ctry_name}, 2000-{rev_yr}"),
#        caption = wuenic_src) +
#   scale_x_discrete(position = "top") +
#   theme_minimal() +
#   theme(panel.grid = element_blank(), legend.position = "bottom", legend.key.width = unit(1.2, "cm"),
#         axis.text.x = element_text(angle = 45, hjust = 0, size = 9)) +
#   scale_fill_gradientn(
#     colours = c("#B50800", "#E2231A", "#F26A21", "#FFC20E", "#80BD41", "#00833D"),
#     values  = scales::rescale(c(0, 59, 69, 79, 89, 94, 100)),
#     guide = "colorbar", limits = c(0, 100),
#     name = "Vaccine coverage (%)  ",
#     breaks = c(0, 59, 69, 79, 89, 100), labels = c("0", "60", "70", "80", "90", "100"))
```

```{r ggplot-mcv1-sub, echo=FALSE, fig.width=10, fig.height=6.5}
#print(heatmap_sub_mcv1)
```

<!-- ===================== PAGE 5: HPV ===================== -->

```{r setup-hpv, include=FALSE}
if(has_hpv_data) {
  
  # check which hpv vaccines are available
  available_hpv <- unique(hpv_df$vaccine_code)
  has_hpvc      <- "PRHPVC_F" %in% available_hpv
  has_hpv1      <- "PRHPV1_F" %in% available_hpv
  
  # check if enough years exist for comparison (need at least hpv_rev_yr and hpv_rev_yr - 1)
  hpvc_yrs <- hpv_df %>% filter(vaccine_code == "PRHPVC_F") %>% pull(year)
  hpv1_yrs <- hpv_df %>% filter(vaccine_code == "PRHPV1_F") %>% pull(year)
  
  has_hpvc_comparison <- has_hpvc && all(c(hpv_rev_yr, hpv_rev_yr - 1) %in% hpvc_yrs)
  has_hpv1_comparison <- has_hpv1 && all(c(hpv_rev_yr, hpv_rev_yr - 1) %in% hpv1_yrs)
  
  # use hpvc if available and has enough years, otherwise fall back to hpv1
  primary_hpv_code  <- if (has_hpvc_comparison) "PRHPVC_F" else "PRHPV1_F"
  primary_hpv_label <- if (has_hpvc_comparison) "HPVc" else "HPV1"
  
  # add var for if hpvc in rev_yr is available
  has_hpvc_for_rev_yr <- has_hpvc && hpv_rev_yr %in% hpvc_yrs
  
  # hpv coverage change for this country
  hpv <- hpv_df %>%
    filter(vaccine_code == primary_hpv_code) %>%
    select(iso3c, country, vaccine_code, year, coverage) %>%
    filter(year %in% c(hpv_rev_yr, hpv_rev_yr - 1)) %>%
    pivot_wider(names_from = year, values_from = coverage) %>%
    mutate(diff = .[[5]] - .[[4]],
           chng = case_when(diff > 0 ~ "increased", diff < 0 ~ "decreased", TRUE ~ "remained constant"),
           diff_txt = case_when(
             chng %in% c("increased", "decreased") ~ str_glue("{chng} from {.[[4]]}% in {hpv_rev_yr-1} to {.[[5]]}% in {hpv_rev_yr}"),
             TRUE ~ str_glue("{chng} at {.[[5]]}% between {hpv_rev_yr-1} and {hpv_rev_yr}")),
           gt90 = case_when(
             .[[5]] < 80  & chng == "increased" ~ str_glue("Despite the increase, {primary_hpv_label} coverage remains below the 90% target"),
             .[[5]] < 80  & chng != "increased" ~ str_glue("{primary_hpv_label} coverage remains below the 90% target"),
             .[[5]] %in% 80:89 & chng == "increased" ~ str_glue("Despite the increase, {primary_hpv_label} coverage has not yet reached the 90% target"),
             .[[5]] %in% 80:89 & chng != "increased" ~ str_glue("{primary_hpv_label} coverage has not yet reached the 90% target"),
             .[[5]] == 90 ~ str_glue("Coverage of {primary_hpv_label} achieved the 90% target"),
             .[[5]] > 90  ~ str_glue("Coverage of {primary_hpv_label} has exceeded the 90% target"),
             TRUE ~ NA_character_))
  
  # note for when hpvc is not available
  hpvc_unavailable_note <- if (!has_hpvc) str_glue("Note: HPVc (last dose) data is not available for {ctry_name} in WUENIC. HPV1 data is presented instead.") else NULL
  
  # unvaccinated females (hpv1 and/or hpvc)
  hpv_unvac <- hpv_df %>%
    filter(year == hpv_rev_yr) %>%
    select(iso3c, country, vaccine_code, unvaccinated) %>%
    pivot_wider(names_from = vaccine_code, values_from = unvaccinated)
  
  # add missing columns if not present
  if (!"PRHPVC_F" %in% names(hpv_unvac)) hpv_unvac$PRHPVC_F <- NA
  if (!"PRHPV1_F" %in% names(hpv_unvac)) hpv_unvac$PRHPV1_F <- NA
  
  hpv_unvac <- hpv_unvac %>%
    mutate(diff = PRHPVC_F - PRHPV1_F) %>%
    label_vals_millions(PRHPVC_F, "PRHPVC_F_lbl") %>%
    label_vals_millions(PRHPV1_F, "PRHPV1_F_lbl") %>%
    label_vals_millions(diff, "diff_lbl")
  
  has_hpv_num_unvac <- ifelse(!is.na(hpv_unvac$PRHPV1_F) | !is.na(hpv_unvac$PRHPVC_F), TRUE, FALSE)
  
  # get first year of hpv data
  first_hpv_year <- hpv_intro %>% filter(introduced == 1) %>% pull(year) %>% min(na.rm = TRUE)
  
  # regional rank for hpv (use primary vaccine code)
  reg_rank_hpv <- hpv_peers %>%
    filter(vaccine_code == primary_hpv_code, year == hpv_rev_yr) %>%
    arrange(desc(coverage)) %>%
    mutate(rank = row_number()) %>%
    filter(iso3c == ctry_iso)
  
  n_peers_hpv <- hpv_peers %>% filter(vaccine_code == primary_hpv_code, year == hpv_rev_yr) %>% n_distinct("iso3c")
}
```


```{r hpv-line-plot, echo=FALSE, warning=FALSE, message=FALSE}
if (has_hpv_data) {
  
  # prep plot data - hpv1 always shown if available
  plot_hpv1 <- if (has_hpv1) {
    hpv_df %>% filter(vaccine_code == "PRHPV1_F", year >= first_hpv_year) %>% select(year, coverage)
  } else NULL
  
  # hpvc only if available
  plot_hpvc <- if (has_hpvc) {
    hpv_df %>% filter(vaccine_code == "PRHPVC_F", year >= first_hpv_year) %>% select(year, coverage)
  } else NULL
  
  # reference df for x-axis range (use whichever is available)
  plot_hpv_ref <- if (has_hpvc) plot_hpvc else plot_hpv1
  
  # scale factor for unvaccinated axis (use primary vaccine code)
  if (has_hpv_num_unvac) {
    plot_zd_hpv <- hpv_df %>%
      filter(vaccine_code == primary_hpv_code, year >= first_hpv_year) %>%
      select(year, unvaccinated)
    
    hpv_scale_factor <- 100 / max(plot_zd_hpv$unvaccinated, na.rm = TRUE)
  }
  
  hpv_plot <- ggplot() +
    # hpv1 (if available)
    {if (has_hpv1 && nrow(plot_hpv1) > 1) list(
        geom_line(data = plot_hpv1, aes(x = year, y = coverage, color = "HPV1 coverage"), size = 0.8),
        geom_point(data = plot_hpv1, aes(x = year, y = coverage, color = "HPV1 coverage"), size = 1.5)
    ) else if (has_hpv1 && nrow(plot_hpv1) == 1) list(
        geom_point(data = plot_hpv1, aes(x = year, y = coverage, color = "HPV1 coverage"), size = 1.5)
    )} +
    # hpvc (if available)
    {if (has_hpvc && nrow(plot_hpvc) > 1) list(
        geom_line(data = plot_hpvc, aes(x = year, y = coverage, color = "HPVc coverage"), size = 0.8),
        geom_point(data = plot_hpvc, aes(x = year, y = coverage, color = "HPVc coverage"), size = 1.5)
    ) else if (has_hpvc && nrow(plot_hpvc) == 1) list(
        geom_point(data = plot_hpvc, aes(x = year, y = coverage, color = "HPVc coverage"), size = 1.5)
    )} +
    # unvaccinated females (if available)
    {if (has_hpv_num_unvac && nrow(plot_zd_hpv) > 1) list(
       geom_col(data = plot_zd_hpv, aes(x = year, y = unvaccinated * hpv_scale_factor, fill = "Unvaccinated females"),
           alpha = 0.4, width = 0.6)
    ) else if (has_hpv_num_unvac && nrow(plot_zd_hpv) == 1) list(
        geom_point(data = plot_zd_hpv, aes(x = year, y = unvaccinated * hpv_scale_factor, color = "Unvaccinated females"), size = 1.5)
    )} +
    geom_hline(yintercept = 90, linetype = "dashed") +
    scale_color_manual(values = c("HPV1 coverage" = "#1CABE2", "HPVc coverage" = "#6A1E72")) +
    scale_fill_manual(name = NULL, values = c("Unvaccinated females" = "#E2231A")) +
    scale_x_continuous(breaks = seq(min(plot_hpv_ref$year), max(plot_hpv_ref$year), 1)) +
    {if (has_hpv_num_unvac)
  scale_y_continuous(limits = c(0, 100), name = "Coverage (%)",
    sec.axis = sec_axis(~ . / hpv_scale_factor, name = "Unvaccinated females",
      labels = scales::label_comma()))
    else
      scale_y_continuous(limits = c(0, 100), name = "Coverage (%)")} +
    {if (has_hpv_num_unvac)
      guides(color = guide_legend(order = 1), fill = guide_legend(order = 2))} +
    labs(
      title = str_glue(
        "{if (has_hpvc) 'HPV1 and HPVc coverage' else 'HPV1 coverage'}{if (has_hpv_num_unvac) ' and unvaccinated females' else ''} over time, {ctry_name}, {first_hpv_year}–{hpv_rev_yr}"),
      caption = paste0(
        "Dotted line shows the IA2030 target of 90% coverage\n",
        if (!has_hpvc) "HPVc (last dose) data is not available for this country in WUENIC\n" else "",
        if (!has_hpv_num_unvac) "Number of unvaccinated females not available for this country\n" else "",
        wuenic_src),
      x = "Year", color = NULL) +
    theme_unicef() +
    theme(panel.grid.minor = element_blank(), panel.grid.major.x = element_line(),
      axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
      plot.title = element_text(size = 14, hjust = 0.5, face = "bold"),
      legend.position = "bottom")
}
```
\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{HPV}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:overview}{}")
cat("\\addcontentsline{toc}{section}{HPV}")
```

\vspace{1em}

```{r hpv-intro-flag, results='asis', echo=FALSE}
if (hpv_intro_flag == 0) {
  cat(paste0("- ", ctry_name, " had not introduced HPV vaccination as of ", rev_yr, ".\n"))
  cat("\\vspace{0.6em}\n")
} else if (hpv_intro_flag == 1 & has_hpv_data == FALSE) {
  cat(paste0("- ", ctry_name, " had introduced HPV vaccination as of ", rev_yr, " but has no ", rev_yr, " WUENIC HPV data.\n"))
  cat("\\vspace{0.6em}\n")
} else {
  if (has_hpvc) {
    cat(glue::glue("- Programme coverage of HPV1 among females {hpv %>% filter(vaccine_code == 'PRHPV1_F') %>% pull(diff_txt)}, and coverage of the last dose (HPVc) {hpv %>% filter(vaccine_code == 'PRHPVC_F') %>% pull(diff_txt)}.\n"))
    cat("\\vspace{0.6em}\n")
    cat(glue::glue("- {hpv %>% filter(vaccine_code == 'PRHPVC_F') %>% pull(gt90)}.\n"))
    cat("\\vspace{0.6em}\n")
  } else {
    cat(glue::glue("- Programme coverage of HPV1 among females {hpv %>% filter(vaccine_code == 'PRHPV1_F') %>% pull(diff_txt)}.\n"))
    cat("\\vspace{0.6em}\n")
    cat(glue::glue("- {hpv %>% filter(vaccine_code == 'PRHPV1_F') %>% pull(gt90)}.\n"))
    cat("\\vspace{0.6em}\n")
    if (exists("has_hpvc_comparison") && !has_hpvc_comparison) {
      cat(glue::glue("- HPVc (last dose) data is available for {ctry_name} in WUENIC but only for one year, so a year-on-year comparison is not possible and HPVc is not included in this report.\n"))
    } else {
      cat(glue::glue("- HPVc (last dose) coverage data is not available for {ctry_name} in WUENIC data.\n"))
    }
    cat("\\vspace{0.6em}\n")
  }
  if (has_hpv_num_unvac) {
    cat(glue::glue("- In {hpv_rev_yr}, {hpv_unvac$PRHPV1_F_lbl} females in {the_ctry_name} did not receive any HPV vaccination.\n"))
    cat("\\vspace{0.6em}\n")
  } else {
    cat(glue::glue("- The number of unvaccinated females for HPV in {the_ctry_name} in {hpv_rev_yr} is not available in WUENIC data.\n"))
    cat("\\vspace{0.6em}\n")
  }
}
if (exists("reg_rank_hpv") && nrow(reg_rank_hpv) > 0) {
  cat(glue::glue("- Among {ctry_regn} region peers, {the_ctry_name} ranked {reg_rank_hpv$rank} out of {n_peers_hpv} countries in {primary_hpv_label} coverage in {hpv_rev_yr}.\n"))
  cat("\\vspace{0.6em}\n")
}
invisible(NULL)
```


```{r hpv, echo=FALSE, warning=FALSE, fig.width=10, fig.height=6.2}
if (exists("hpv_plot")) {
  print(hpv_plot)
}
```

<!-- ===================== PAGE 6: REGIONAL COMPARISON ===================== -->

```{r setup-regional-comparison, include=FALSE}
reg_comp_vax <- reg_peers %>%
  filter(vaccine %in% c("dtp1", "dtp3", "mcv1"), year == rev_yr) %>%
  select(iso3c, country, vaccine, coverage)

reg_comp_hpv <- hpv_peers %>%
  filter(vaccine_code == "PRHPVC_F", year == hpv_rev_yr) %>%
  select(iso3c, country, coverage) %>%
  mutate(vaccine = "hpvc")

if (has_hpv_data && has_hpvc_for_rev_yr) {
  reg_comp <- bind_rows(reg_comp_vax, reg_comp_hpv) %>%
    mutate(
      highlight = ifelse(iso3c == ctry_iso, ctry_name, "Peers"),
      vaccine = toupper(vaccine),
      vaccine = factor(vaccine, levels = c("DTP1", "DTP3", "MCV1", "HPVC"))
    )
} else {
  reg_comp <- reg_comp_vax %>%
    mutate(
      highlight = ifelse(iso3c == ctry_iso, ctry_name, "Peers"),
      vaccine = toupper(vaccine),
      vaccine = factor(vaccine, levels = c("DTP1", "DTP3", "MCV1"))
    )
}

# regional average per vaccine
reg_avg <- reg_comp %>%
  group_by(vaccine) %>%
  summarise(avg = round(mean(coverage, na.rm = TRUE), 0))

plt_reg_comp <- reg_comp %>%
  mutate(country = reorder_within(country, coverage, vaccine)) %>%
  ggplot(aes(x = coverage, y = country, colour = highlight, size = highlight)) +
  geom_point() +
  geom_vline(data = reg_avg, aes(xintercept = avg), linetype = "dashed", colour = "grey50", linewidth = 0.5) +
  facet_wrap(~vaccine, ncol = 2, scales = "free_y") +
  scale_y_reordered() +
  scale_colour_manual(values = setNames(c("#E2231A", "#A8D5E2"), c(ctry_name, "Peers")), name = "") +
  scale_size_manual(values = setNames(c(4, 2.5), c(ctry_name, "Peers")), name = "") +
  labs(x = "Coverage (%)", y = "",
       title = str_glue("Coverage comparison: {ctry_name} vs. {region} peers, {rev_yr}"),
       subtitle = "Dashed line = regional average",
       caption = wuenic_src) +
  theme_unicef() +
  theme(legend.position = "bottom",
        strip.text = element_text(face = "bold"),
        panel.grid.minor = element_blank(),
        axis.text.y = element_text(size = 9))
```

```{r calc-height, echo=FALSE}
# get number unique countries to set plot height dynamically
n_countries <- length(unique(reg_comp$country))

# set a base height (for titles/legend) plus a per-country height
# 2.5 is the base, 0.2 is the height per country
dynamic_height <- 2.5 + (n_countries * 0.28)

# minimum and maximum height to prevent tiny or massive plots
dynamic_height <- max(4, min(dynamic_height, 12))
```

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{Regional Comparison}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:overview}{}")
cat("\\addcontentsline{toc}{section}{Regional Comparison}")
```

\vspace{1em}

The chart below shows how `r the_ctry_name` compares to other countries in the `r ctry_regn` region on DTP1, DTP3, MCV1`r if (has_hpv_data && has_hpvc_for_rev_yr) ", and HPVc" else ""` coverage in `r rev_yr`. The dashed line represents the regional average for each vaccine.

\vspace{1em}

```{r, echo=FALSE, results='asis'}
if (has_hpv_data) {
  if (has_hpvc && !has_hpvc_for_rev_yr) {
    cat(glue::glue("HPVc (last dose) data is available for {ctry_name} in WUENIC but not for {rev_yr}, so HPVc coverage is not included in the regional 
    comparison.\n"))
  } else if (!has_hpvc) {
    cat(glue::glue("HPVc (last dose) data is not available for {ctry_name} in WUENIC, so HPVc coverage is not included in the regional comparison.\n"))
  }
  cat("\\vspace{0.6em}\n")
}
```

```{r ggplot-reg-comp, echo=FALSE, fig.width=10, fig.height=dynamic_height}
print(plt_reg_comp)
```

\newpage

<!-- ===================== PAGE 7: ADDITIONAL VACCINES ===================== -->

```{r setup-additional-vax, include=FALSE}
# additional vaccines to check: bcg, hib3 or hepb3, ipv1, pcv3 or rotac
# -- only include if the vaccine has non-missing coverage in this country
add_vax <- c("bcg", "hib3", "hepb3", "ipv1", "pcvc", "rotac")

add_vax_df <- df %>%
  filter(vaccine %in% add_vax, year %in% c(comp_yr, rev_yr)) %>%
  select(iso3c, country, vaccine, year, coverage) %>%
  pivot_wider(names_from = year, values_from = coverage) %>%
  filter(!is.na(.[[4]]) | !is.na(.[[5]])) %>%    # keep only vaccines with data
  mutate(diff = .[[5]] - .[[4]],
         diff_txt = case_when(
           diff > 0  ~ str_glue("increased from {.[[4]]}% in {comp_yr} to {.[[5]]}% in {rev_yr}"),
           diff < 0  ~ str_glue("decreased from {.[[4]]}% in {comp_yr} to {.[[5]]}% in {rev_yr}"),
           diff == 0 ~ str_glue("remained the same ({.[[4]]}%) between {comp_yr} and {rev_yr}"),
           TRUE ~ NA_character_),
         vaccine_label = case_when(
           vaccine == "bcg"   ~ "BCG",   vaccine == "hib3"  ~ "Hib3",
           vaccine == "hepb3" ~ "HepB3", vaccine == "ipv1"  ~ "IPV1",
           vaccine == "pcvc"  ~ "PCV3",  vaccine == "rotac" ~ "RotaC",
           TRUE ~ toupper(vaccine)),
         gt90 = ifelse(.[[5]] >= 90, "was above", "was not above"))

# summary heatmap for additional vaccines over time
add_vax_heatmap_df <- df %>%
  filter(vaccine %in% add_vax, year >= 2000) %>%
  mutate(vaccine_label = case_when(
    vaccine == "bcg" ~ "BCG", vaccine == "hib3" ~ "Hib3", vaccine == "hepb3" ~ "HepB3",
    vaccine == "ipv1" ~ "IPV1", vaccine == "pcvc" ~ "PCV3", vaccine == "rotac" ~ "RotaC",
    TRUE ~ toupper(vaccine)))

plt_add_vax <-
  add_vax_heatmap_df %>%
  ggplot(aes(x = as.factor(year), y = vaccine_label, fill = coverage)) +
  geom_tile(color = "white") +
  geom_text(data = add_vax_heatmap_df, aes(label = coverage), colour = "white", size = 3) +
  labs(x = "", y = "",
       title = str_glue("Additional vaccine coverage (%), {ctry_name}, 2000-{rev_yr}"),
       caption = wuenic_src) +
  scale_x_discrete(position = "top") +
  scale_y_discrete(limits = rev) +
  theme_minimal() +
  theme(panel.grid = element_blank(), 
        legend.position = "bottom", 
        legend.key.width = unit(1.2, "cm"),
        axis.text.x = element_text(size = 9),
        plot.title = element_text(size = 14, hjust = 0.5, face = "bold")) +
  scale_fill_gradientn(
    colours = c("#B50800", "#E2231A", "#F26A21", "#FFC20E", "#80BD41", "#00833D"),
    values  = scales::rescale(c(0, 59, 69, 79, 89, 94, 100)),
    guide = "colorbar", limits = c(0, 100),
    name = "Vaccine coverage (%)  ",
    breaks = c(0, 59, 69, 79, 89, 100), labels = c("0", "60", "70", "80", "90", "100"))
```

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{Additional Vaccines}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:overview}{}")
cat("\\addcontentsline{toc}{section}{Additional Vaccines}")
```

\vspace{1em}

```{r additional-vax-bullets, results='asis', echo=FALSE}
# dynamically generate one bullet per vaccine with data
for (i in seq_len(nrow(add_vax_df))) {
  row <- add_vax_df[i, ]
  cat(str_glue("-   {row$vaccine_label} coverage {row$diff_txt} and {row$gt90} the 90% target in {rev_yr}."), "\n\n")
}
```

```{r ggplot-add-vax, echo=FALSE, fig.width=10, fig.height=4}
print(plt_add_vax)
```

<!-- ===================== APPENDIX ===================== -->

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{Appendix}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:overview}{}")
cat("\\addcontentsline{toc}{section}{Appendix}")
```

\vspace{1em}

\textcolor{cyan}{\large WHO/UNICEF Estimates of National Immunization Coverage (WUENIC)}

\vspace{1em}

Each year, WHO and UNICEF review national immunization data from Member States, including administrative and official coverage, survey estimates and contextual information, and use rules-based data triangulation approach to assess coverage. Estimates are produced independently, without borrowing data from other countries, and are not based on ad hoc adjustments. In some cases, they rely on a single source, typically nationally reported coverage. When data for a specific country-vaccine-year are unavailable, values from adjacent years are interpolated or extrapolated, and conflicting sources are assessed to determine the most reliable estimate, considering potential biases and expert opinion.

\vspace{0.5em}
Methods:

\begin{itemize}
  \item \begin{sloppypar}
        \textcolor{blue}{\href{https://pubmed.ncbi.nlm.nih.gov/19649368/}{Burton et al. 2009. WHO and UNICEF estimates of national infant immunization coverage: methods and processes}}
        \end{sloppypar}
  \item \begin{sloppypar}
        \textcolor{blue}{\href{https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0047806}{Burton et al. 2012. A formal representation of the WHO and UNICEF estimates of national immunization coverage: a computational logic approach}}
        \end{sloppypar}
  \item \begin{sloppypar}
        \textcolor{blue}{\href{https://openpublichealthjournal.com/VOLUME/6/PAGE/73/ABSTRACT/}{Brown et al. 2013. An introduction to the grade of confidence used to characterize uncertainty around the WHO and UNICEF estimates of national immunization coverage}}
        \end{sloppypar}
\end{itemize}

\vspace{0.5em}

These talking points present the latest WUENIC estimates (published 15 July `r rev_yr+1`).

\vspace{1.5em}

\textcolor{cyan}{\fontsize{14}{16}\selectfont Definitions of immunization terms:}

\vspace{0.5em}

-   Bacillus Calmette-Guerin (BCG): vaccine against tuberculosis
\vspace{0.5em}
-   Hepatitis B birth dose, given within 24 hours after birth (HepBB)
\vspace{0.5em}
-   Diphtheria, tetanus, and pertussis vaccine, first dose (DTP1) and third dose (DTP3)
\vspace{0.5em}
-   Hepatitis B vaccine, third dose (HepB3)
\vspace{0.5em}
-   Haemophilus influenzae type B vaccine, third dose (Hib3)
\vspace{0.5em}
-   Poliomyelitis vaccine, third dose (Polio3)
\vspace{0.5em}
-   Inactivated polio vaccine, first dose (IPV1) and second dose (IPV2): second dose is only shown for oral polio vaccine (OPV) using countries
\vspace{0.5em}
-   Measles containing vaccine, first dose (MCV1) and second dose (MCV2)
\vspace{0.5em}
-   Rotavirus vaccine, last dose (RotaC)
\vspace{0.5em}
-   Pneumococcal vaccine, third dose (PCV3)
\vspace{0.5em}
-   Yellow Fever vaccine (YFV)
\vspace{0.5em}
-   Meningococcal A vaccine (MengA)
\vspace{0.5em}
-   Human Papillomavirus vaccine, first dose (HPV1) and last dose (HPVc)

\vspace{1.5em}

\textcolor{cyan}{\fontsize{14}{16}\selectfont Additional resources:} \textcolor{blue}{\href{https://data.unicef.org/resources/immunization/}{\underline{https://data.unicef.org/resources/immunization/}}}
