---
title: ""
output:
  pdf_document:
    keep_tex: false
    latex_engine: xelatex
    toc: false
    toc_depth: 3
geometry: "left=1cm,right=1cm,top=1.1cm,bottom=1.1cm"
fontsize: 12pt
params:
  country: "Default"
  language: "en"
  plot_language: "en" # so we can make arabic plot labels english
header-includes:
  - \usepackage{enumitem}
  - \geometry{a4paper, left=15mm, right=15mm, top=15mm, bottom=20mm}
  - \setlist[itemize,1]{label=\textbullet}
  - \setlist[itemize,2]{label=\(\circ\)}
  - \setlist[itemize,3]{label=\(\infty\)}
  - \setmainfont{Arial}
  - \usepackage{hyperref}
  - \hypersetup{colorlinks=true, linkcolor=blue, urlcolor=blue, breaklinks=true}
  - \usepackage{tcolorbox}
  - \usepackage{fancyhdr}
  - \setlength{\headheight}{15pt}
  - \addtolength{\topmargin}{-3pt}
  - \definecolor{unicefblue}{RGB}{28, 171, 226}
  - \pagestyle{fancy}
  - \fancyhf{}
  - \fancyhead[L]{}
  - \fancyhead[C]{}
  - \fancyhead[R]{\hyperlink{top}{\small \color{black} $\uparrow$ Back to top}}
  - \fancyfoot[C]{\thepage}
  - \usepackage{tocloft}
  - \renewcommand{\cftsecfont}{\normalfont}
  - \renewcommand{\cftsecpagefont}{\normalfont}
  - \renewcommand{\cftsecleader}{\cftdotfill{\cftdotsep}}
---

\hypertarget{top}{}

```{r setup, include=FALSE}
options(scipen = 999)

# ── TRANSLATION SETUP ─────────────────────────────────────────────────────────
source(file.path(directory, type, "talking-points/country-specific-talking-points", "txt_vars_talking_points_countries.R"))
# loads text_vars_en
source(file.path(utils, "R/narrative-translation_deepl_grace.R"))    # loads text_vars (translated)

language <- params$language

comparison_col <- as.character(comp_yr)
recent_col  <- as.character(rev_yr)
minusone_col <- as.character(rev_yr - 1)

# ── COUNTRY DATA ──────────────────────────────────────────────────────────────
df <- wuenic_dta %>% filter(country == current_country) %>%
  filter(lvl_1 == "country") %>% 
  filter(!vaccine == "pol3") %>% 
  # new labels
  label_vals_millions(target, "target_lbl") %>%
  label_vals_millions(vaccinated, "vaccinated_lbl") %>%
  label_vals_millions(unvaccinated, "unvaccinated_lbl")
hpv_df <- hpv_dta %>% filter(country == current_country) %>%
  label_vals_millions(target, "target_lbl") %>%
  label_vals_millions(vaccinated, "vaccinated_lbl") %>%
  label_vals_millions(unvaccinated, "unvaccinated_lbl")
has_hpv_data <- nrow(hpv_df) > 0
has_hpvc     <- has_hpv_data && any(hpv_df$vaccine_code == "PRHPVC_F")

add_region_names <- function(df) {
  df %>%
    mutate(Region_Name = case_match(Region,
      "ROSA"          ~ "South Asia",
      "ESAR"          ~ "Eastern and Southern Africa",
      "ECAR"          ~ "Europe and Central Asia",
      "MENA"          ~ "Middle East and North Africa",
      "LACR"          ~ "Latin America and the Caribbean",
      "WCAR"          ~ "West and Central Africa",
      "EAPR"          ~ "East Asia and the Pacific",
      "Non-programme" ~ "Non-programme countries",
      .default        = Region))
}
df     <- add_region_names(df)
hpv_df <- add_region_names(hpv_df)

ctry_iso  <- unique(df$iso3c)
region    <- unique(df$Region)
region_translated <- t_lookup(tolower(region), language)
ctry_regn <- df %>%
  distinct(Region_Name, Region) %>%
  mutate(formatted = str_glue("{Region_Name} ({Region})")) %>%
  pull(formatted)

reg_peers <- wuenic_dta %>% filter(Region == region, lvl_1 == "country")
hpv_peers <- hpv_dta    %>% filter(Region == region, lvl_1 == "country")

available_vaccines <- unique(df$vaccine)

sub_df <- wuenic_dta %>% filter(iso3c == ctry_iso)

#wuenic_src <- str_glue("Source: WHO/UNICEF Estimates of National Immunization Coverage, {rev_yr} revision")

the_countries <- c(
  "United Arab Emirates", "Bahamas", "Central African Republic", "DRC", "Congo",
  "Dominican Republic", "Gambia", "Maldives", "Marshall Islands",
  "Netherlands (Kingdom of the)", "Philippines", "Republic of Korea",
  "Republic of Moldova", "Russian Federation", "Solomon Islands",
  "State of Palestine", "United Kingdom", "United States")

df <- df %>%
  mutate(the_country = ifelse(country %in% the_countries, paste("the", country), country))

ctry_name     <- unique(df$country)
the_ctry_name <- unique(df$the_country)

hpv_intro <- wiise_hpv_intro_yrs %>%
  filter(iso3c == ctry_iso) %>%
  mutate(introduced = case_when(nationwide == "yes" | partially == "yes" ~ 1, TRUE ~ 0))

hpv_intro_flag <- ifelse(nrow(hpv_intro) > 0 && any(hpv_intro$introduced == 1, na.rm = TRUE), 1, 0)

if (!require("countrycode", quietly = TRUE)) install.packages("countrycode")
library(countrycode)
iso2 <- tolower(countrycode(ctry_name, origin = "country.name", destination = "iso2c"))
if (ctry_name == "Micronesia") iso2 <- "fm"

flag_path <- paste0("flags/", iso2, ".png")
if (!file.exists(flag_path)) {
  dir.create("flags", showWarnings = FALSE)
  download.file(paste0("https://flagcdn.com/160x120/", iso2, ".png"),
                destfile = flag_path, mode = "wb")
}

country_lookup <- x
if(country == "Andorra") {
  country_lookup <- "and_code"
}

dtp1_translated <- ifelse(language == "fr", "DTC1", "DTP1")
dtp_translated <- ifelse(language == "fr", "DTC", "DTP")
dtp3_translated <- ifelse(language == "fr", "DTC3", "DTP3")
mcv1_translated <- ifelse(language == "fr", "VAR1", "MCV1")
mcv2_translated <- ifelse(language == "fr", "VAR2", "MCV2")
mcv_translated <- ifelse(language == "fr", "VAR", "MCV")
hpv_translated <- ifelse(language == "es", "VPH", "HPV")
hpv1_translated <- ifelse(language == "es", "VPH1", "HPV1")
hpvc_translated <- ifelse(language == "es", "VPHc", "HPVc")
```


```{r banner, echo=FALSE, results='asis'}
cat(paste0('
\\begin{tcolorbox}[
  colback=unicefblue,
  colframe=unicefblue,
  arc=6pt,
  boxrule=0pt,
  left=4mm,
  right=4mm,
  top=3mm,
  bottom=3mm,
  halign=center,
  valign=center
]
\\begin{center}
  {\\LARGE \\color{white} \\textbf{', t_lookup("txt_sec_banner_title", language), ': ', t_lookup(country_lookup, language), '}} \\hspace{0.2cm}\\includegraphics[height=0.6cm]{', flag_path, '}\\\\[0.4em]
  {\\large \\color{white} ', t_lookup("txt_sec_banner_revision", language), ' ', rev_yr, '}
\\end{center}
\\end{tcolorbox}
\\vspace{0.4in}
'))
```

```{r toc, echo=FALSE, results='asis'}
cat(paste0("\\renewcommand{\\contentsname}{", get_text2("txt_contents", text_vars), "}"))
cat("\\tableofcontents")
```


```{r country-map, echo=FALSE, warning=FALSE, message=FALSE}
region_for_country <- base_map_df %>%
  filter(iso3c == ctry_iso) %>%
  pull(map_region)

map_df <- base_map_df %>% filter(map_region == region_for_country)

if (!is.na(region_for_country) && region_for_country == "Oceania") {
  map_df <- st_shift_longitude(map_df)
}

map_df <- map_df %>%
  mutate(highlight = ifelse(iso3c == ctry_iso, get_text2("txt_map_country", text_vars), get_text2("txt_map_region",  text_vars)))

if (!is.na(region_for_country)) {
  limits <- switch(region_for_country,
    "Africa"   = list(x = c(-24, 56),   y = c(-40, 46)),
    "Americas" = list(x = c(-170, -23), y = c(-60, 85)),
    "Asia"     = list(x = c(25, 150),   y = c(-15, 60)),
    "Europe"   = list(x = c(-25, 100),  y = c(34, 75)),
    "Oceania"  = list(x = c(110, 240),  y = c(-50, 5)),
    list(x = c(-180, 180), y = c(-90, 90)))
  xlim_val <- limits$x
  ylim_val <- limits$y
}

if (!is.na(region_for_country) && region_for_country == "Oceania") {
  pacific_crs <- st_crs("+proj=longlat +lon_0=180")
  map_df      <- st_transform(map_df, pacific_crs)
  xlim_val    <- c(-70, 15)
  ylim_val    <- c(-50, 3)
}

if (current_country == "Russian Federation") {
  xlim_val <- c(-25, 200)
  ylim_val <- c(34, 85)
}

target_country_geo <- map_df %>% filter(highlight == get_text2("txt_map_country", text_vars))
is_small_country   <- as.numeric(st_area(target_country_geo)) < 3e10
if (is_small_country) country_centroid <- st_centroid(target_country_geo)

country_map <- ggplot(map_df) +
  geom_sf(aes(fill = highlight), color = "white", size = 0.2) +
  { if (is_small_country)
      geom_sf(data = country_centroid, inherit.aes = FALSE,
              size = 6, shape = 21, color = "#1CABE2", stroke = 1.5) } +
  scale_fill_manual(
    values = setNames(c("#1CABE2", "#d7d7d7"),
                      c(get_text2("txt_map_country", text_vars),
                        get_text2("txt_map_region",  text_vars))),
    name = NULL) +
  coord_sf(xlim = xlim_val, ylim = ylim_val, expand = FALSE) +
  theme_void() +
  theme(legend.position = "none",
        plot.title  = element_markdown(hjust = 0.5, size = 14, face = "bold"),
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.5)) +
  labs(title = str_glue(
    get_text2("txt_map_title_full", plot_vars),
    ctry_name          = ctry_name,
    region_for_country = region_for_country
  ))
```

\vspace{2em}
```{r ggplot-country-map, echo=FALSE, warning=FALSE, fig.align='center'}
suppressMessages(print(country_map))
```

```{r setup-vax, include=FALSE}
vax <- df %>%
  filter(vaccine %in% c("dtp1", "dtp3", "mcv1", "mcv2"),
         year %in% c(comp_yr, rev_yr - 1, rev_yr))

unvac_lbl <- vax %>%
  select(Region, iso3c, country, year, vaccine, unvaccinated_lbl) %>%
  pivot_wider(names_from = year, values_from = unvaccinated_lbl) %>%
  rename_with(~paste0(., "_lbl"), .cols = (ncol(.) - 2):ncol(.))

unvac <- vax %>%
  select(Region, iso3c, country, year, vaccine, unvaccinated) %>%
  pivot_wider(names_from = year, values_from = unvaccinated) %>%
  left_join(., unvac_lbl) %>%
  distinct()

# get variable names
comparison_lbl_val <- unvac[[paste0(comparison_col, "_lbl")]]
recent_lbl_val     <- unvac[[paste0(recent_col, "_lbl")]]

unvac <- unvac %>%
  mutate(
    pchng = round((.data[[recent_col]] / .data[[comparison_col]] - 1) * 100, 0)
  ) %>%
  rowwise() %>%
  mutate(
    comparison_lbl = .data[[paste0(comparison_col, "_lbl")]],
    recent_lbl     = .data[[paste0(recent_col, "_lbl")]],
    abs_pchng      = abs(pchng),
    pchng_txt = case_when(
      comparison_lbl == recent_lbl ~ str_glue(get_text2("txt_unvac_stable", text_vars)),
      pchng < 0                    ~ str_glue(get_text2("txt_unvac_declined", text_vars)),
      pchng > 0                    ~ str_glue(get_text2("txt_unvac_increased", text_vars)),
      TRUE                         ~ NA_character_
    )
  ) %>%
  ungroup()

vacc <- vax %>%
  select(Region, iso3c, country, year, vaccine, vaccinated) %>%
  pivot_wider(names_from = year, values_from = vaccinated) %>%
  mutate(
    pchng     = round((.data[[recent_col]] / .data[[comparison_col]] - 1) * 100, 1),
    pchng_txt = case_when(
      pchng > 0 ~ t_lookup("txt_vacc_increased", language),
      pchng < 0 ~ t_lookup("txt_vacc_decreased", language),
      TRUE      ~ t_lookup("txt_vacc_same", language))) %>%
  label_vals_millions(.[[comparison_col]], "vacc_comp_yr_lbl") %>%
  label_vals_millions(.[[recent_col]], "vacc_rev_yr_lbl")

cvg <- vax %>%
  select(Region, iso3c, country, year, vaccine, coverage) %>%
  pivot_wider(names_from = year, values_from = coverage)

tgt <- vax %>%
  filter(vaccine == "dtp1") %>%
  select(Region, iso3c, country, year, vaccine, target) %>%
  pivot_wider(names_from = year, values_from = target) %>%
  mutate(
    diff     = .data[[recent_col]] - .data[[comparison_col]],
    moreless = case_when(
      diff < 0 ~ t_lookup("txt_moreless_fewer", language),
      diff > 0 ~ t_lookup("txt_moreless_more",  language),
      TRUE ~ t_lookup("txt_moreless_same", language)),
    to_vax = case_when(
      diff < 0 ~ t_lookup("txt_surv_inf_vax_3", language),
      diff > 0 & diff < 1000000 ~ t_lookup("txt_surv_inf_vax_1", language),
      diff > 0 & diff >= 1000000 ~ t_lookup("txt_surv_inf_vax_2", language))) |> 
  label_vals(diff, "diff_lbl") %>%
  mutate(diff_abs = abs(diff)) %>%
  label_vals_millions(diff_abs, "diff_abs_lbl")
```

```{r setup-overview, include=FALSE}
gt90 <- df %>%
  select(-target_grp) %>%
  distinct() %>%
  filter(year == rev_yr) %>%
  mutate(gt90 = ifelse(coverage >= 90, 1, 0))

n_vax_total <- n_distinct(gt90$vaccine)
n_vax_gt90  <- sum(gt90$gt90, na.rm = TRUE)

cvg_chng <- vax %>%
  filter(year %in% c(rev_yr, rev_yr - 1)) %>%
  select(iso3c, country, vaccine, year, coverage) %>%
  pivot_wider(names_from = year, values_from = coverage) %>%
  rowwise() %>%
  mutate(
    diff = .data[[recent_col]] - .data[[minusone_col]],
    comp_yr      = rev_yr - 1,
    minusone_lbl = .data[[minusone_col]],
    recent_lbl   = .data[[recent_col]],
    diff_txt = case_when(
      diff > 0 ~ str_glue(get_text2("txt_cvg_increased_short", text_vars)),
      diff < 0 ~ str_glue(get_text2("txt_cvg_decreased_short", text_vars)),
      TRUE     ~ str_glue(get_text2("txt_cvg_same_short",      text_vars))
    )
  ) %>%
  ungroup()

# pre-build overview bullets
bullet_ovw_1 <- if (n_vax_gt90 == 0) {
  str_glue(get_text2("txt_ovw_vax_none_gt90", text_vars))
} else {
  str_glue(get_text2("txt_ovw_vax_gt90", text_vars))
}

bullet_ovw_2 <- str_glue(
  get_text2("txt_ovw_dtp1_cvg", text_vars),
  cvg_chng_dtp1 = cvg_chng %>% filter(vaccine == "dtp1") %>% pull(diff_txt),
  cvg_chng_dtp3 = cvg_chng %>% filter(vaccine == "dtp3") %>% pull(diff_txt),
  cvg_chng_mcv1 = cvg_chng %>% filter(vaccine == "mcv1") %>% pull(diff_txt))
```

```{r setup-progress-ia2030-zd-target, include=FALSE}
zd <- df %>%
  filter(vaccine == "dtp1", year >= 2000) %>%
  select(iso3c, country, year, unvaccinated)

zd_2019            <- zd %>% filter(year == comp_yr) %>% pull(unvaccinated)
ia2030_annual_target <- zd_2019 / 2

target <- zd %>%
  filter(year == comp_yr) %>%
  mutate(target_2030 = unvaccinated / 2,
         factor      = (unvaccinated - target_2030) / (2030 - comp_yr)) %>%
  crossing(year_new = 2020:2030) %>%
  mutate(unvaccinated_new = unvaccinated - factor * (year_new - comp_yr)) %>%
  select(iso3c, country, year = year_new, unvaccinated_new)

zd_rev_yr         <- zd     %>% filter(year == rev_yr) %>% pull(unvaccinated)
ia2030_rev_yr     <- target %>% filter(year == rev_yr) %>% pull(unvaccinated_new)
zd_pcnt_diff      <- round((zd_rev_yr / ia2030_rev_yr - 1) * 100, 0)
zd_rev_yr_lbl     <- label_vals_millions_num(zd_rev_yr)
ia2030_rev_yr_lbl <- label_vals_millions_num(ia2030_rev_yr)

# bullet 3 translation
target_key <- case_when(
  zd_pcnt_diff > 0 ~ "txt_ovw_zd_ia2030_higher",
  zd_pcnt_diff < 0 ~ "txt_ovw_zd_ia2030_lower",
  TRUE             ~ "txt_ovw_zd_ia2030_same"
)
bullet3_translated <- get_text2(target_key, text_vars)
bullet_ovw_3 <- str_glue(
  bullet3_translated,
  abs_zd_pcnt_diff = abs(zd_pcnt_diff),
  rev_yr           = rev_yr,
  the_ctry_name    = the_ctry_name,
  zd_rev_yr_lbl    = zd_rev_yr_lbl,
  ia2030_rev_yr_lbl = ia2030_rev_yr_lbl
)
```

<!-- ===================== PAGE 1: OVERVIEW ===================== -->

\newpage

\newpage
\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{`r t_lookup("txt_sec_overview", language)`}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:overview}{}")
cat(paste0("\\addcontentsline{toc}{section}{", t_lookup("txt_sec_overview", language), "}"))
```

- `r bullet_ovw_1`

- `r bullet_ovw_2`

- `r bullet_ovw_3`


```{r setup-heatmap-overview, include=FALSE}
if (has_hpvc) {
  hpv_merge <- hpv_df %>%
    filter(vaccine_code == "PRHPVC_F") %>%
    select(-vaccine) %>%
    rename(vaccine = vaccine_code)
  shared_cols <- intersect(names(df), names(hpv_merge))
  df <- bind_rows(df %>% select(all_of(shared_cols)),
                  hpv_merge %>% select(all_of(shared_cols)))
}

heatmap_ovw_df <- df %>%
  mutate(
    year    = as.character(year),
    vaccine = case_when(
      vaccine == "hepbb"    ~ "HepBB", vaccine == "hepb3"    ~ "HepB3",
      vaccine == "hib3"     ~ "Hib3",  vaccine == "rotac"    ~ "RotaC",
      vaccine == "menga"    ~ "MengA", vaccine == "PRHPVC_F" ~ "HPVc",
      TRUE ~ toupper(vaccine)),
    cvg_cat = case_when(
      coverage < 60              ~ "<60",
      coverage %in% 60:69        ~ "60-69",
      coverage %in% 70:79        ~ "70-79",
      coverage %in% 80:89        ~ "80-89",
      coverage %in% 90:94        ~ "90-94",
      coverage >= 95             ~ ">=95",
      TRUE                       ~ NA_character_))

heatmap_ovw_df$vaccine <- factor(heatmap_ovw_df$vaccine,
  levels = c("BCG","HepBB","DTP1","DTP3","Hib3","HepB3",
             "PCVC","RotaC","IPV1","IPVC","MCV1","RCV1",
             "MCV2","YFV","MengA","HPVc"))

heatmap_ovw_df <- translate_vaccine_names(heatmap_ovw_df)

heatmap_overview <- heatmap_ovw_df %>%
  ggplot(aes(x = year, y = vaccine, fill = coverage)) +
  geom_tile(color = "white") +
  geom_text(aes(label = coverage), colour = "white", vjust = 0.5, size = 3) +
  labs(x = "", y = "",
       title   = str_glue(get_text2("txt_plt_ovw_heatmap_title", plot_vars)),
       caption = str_glue(get_text2("txt_wuenic_source", plot_vars))) +
  scale_x_discrete(position = "top") +
  scale_y_discrete(limits = rev) +
  theme_minimal() +
  theme(panel.grid       = element_blank(),
        legend.position  = "bottom",
        legend.key.width = unit(1.2, "cm"),
        plot.title       = element_text(size = 14, hjust = 0.5, face = "bold")) +
  scale_fill_gradientn(
    colours = c('#B50800','#E2231A','#F26A21','#FFC20E','#80BD41','#00833D'),
    values = scales::rescale(c(0, 59, 69, 79, 88, 90, 100)),
    guide = "colorbar",
    limits = c(0, 100),
    name    = t_lookup("txt_plt_cvg_pct_legend", language),
    breaks = c(0, 60, 70, 80, 90, 100),
    labels = c("0", "60", "70", "80", "90", "100"))
```

```{r overview-line-plot, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center'}

# chart showing the progress towards the IA2030 goal of reducing ZD children by half (bar chart with line showing linear trajectory annual targets)

###### plt_ia2030_zd_proj_bar_line :: progress towards zd target :: bar-chart, linechart ######
zd <- df %>%
  filter(vaccine == 'dtp1', year >= 2000) %>%
  mutate(measure = 'unvaccinated') %>%
  select(lvl_1:vaccine, measure, unvaccinated)

zd_target <- zd %>%
  filter(year == 2019) %>%
  mutate(unvaccinated = unvaccinated/2,
         measure = t_lookup("plt_ia2030_zd_proj_bar_line_legend1", language),
         year = 2030)

# data for bars
zd2 <- zd %>%
  mutate(measure = t_lookup("plt_ia2030_zd_proj_bar_line_legend2", language)) %>%
  rbind(zd_target) %>%
  mutate(chart = 'bar',
         unvaccinated = round(unvaccinated, 0))

# target 
target <- zd %>%
  filter(vaccine == 'dtp1',
         year %in% c(2019)) %>%
  select(lvl_1:country, vaccine, year, unvaccinated) %>%
  pivot_wider(., names_from = year, values_from = unvaccinated) %>%
  mutate(target = `2019`/2) %>%
  mutate(`2020` = `2019`) %>%
  mutate(factor = (`2020`-target)/(2030-2020)) %>%
  mutate(`2021` = `2020` - factor) %>%
  mutate(`2022` = `2021` - factor) %>%
  mutate(`2023` = `2022` - factor) %>%
  mutate(`2024` = `2023` - factor) %>%
  mutate(`2025` = `2024` - factor) %>%
  mutate(`2026` = `2025` - factor) %>%
  mutate(`2027` = `2026` - factor) %>%
  mutate(`2028` = `2027` - factor) %>%
  mutate(`2029` = `2028` - factor) %>%
  mutate(`2030` = `2029` - factor) %>%
  pivot_longer(-c(lvl_1:vaccine), names_to = 'year', values_to = 'zerodose_mt') %>%
  filter(year != 'factor', year != 'target') %>%
  mutate(year = as.numeric(year)) %>%
  select(lvl_1:year, unvaccinated_new = zerodose_mt) %>%
  left_join(., zd) %>%
  filter(year %in% c(2020:2030)) %>%
  mutate(diff = unvaccinated - unvaccinated_new,
         measure = "unvaccinated")  %>%
  select(lvl_1:year, measure, unvaccinated_new, unvaccinated, diff)

# data for lines
target_lines <- target %>%
  select(lvl_1:unvaccinated_new) %>%
  rename(unvaccinated = unvaccinated_new) %>%
  mutate(chart = 'line',
         measure = 'IA2030')

ia2030_2019 <- zd2 %>%
  filter(year == 2019) %>%
  mutate(chart = 'line',
         measure = 'IA2030')

# append
zd2 <- rbind(zd2, target_lines, ia2030_2019)  %>%
  filter(!(year == 2019 & chart == 'line')) %>%
  distinct()

plt_ia2030_zd_proj_bar_line <- gg <-
  ggplot() + 
  geom_col(data = zd2 %>% filter(chart=='bar'), aes(y=unvaccinated, x=year, fill=measure)) +
  geom_line(data = zd2 %>% filter(chart=='line'), aes(y=unvaccinated, x=year)) +
  geom_point(data = zd2 %>% filter(chart=='line'), aes(y=unvaccinated, x=year)) +
  # geom_text(data = zd2 %>% filter(chart=='bar'), aes(y=unvaccinated, x=year, label = format(unvaccinated, big.mark=",")), vjust = -0.5, size = 3) +
  labs(y = get_text2("txt_ia2030_zd_proj_bar_line_y", plot_vars),
       x = "",
       title = str_glue(get_text2("txt_ia2030_zd_proj_bar_line_title", plot_vars)),
       caption = paste0(str_glue(get_text2("txt_wuenic_source", plot_vars)), "\n", str_glue(get_text2("txt_ia2030_zd_proj_bar_line_cpt", plot_vars)))) +
  theme_minimal()+
  theme_unicef()+
  theme(legend.position="bottom",
        legend.title = element_blank(),
        #axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=8.5),
        axis.text.y = element_text(size=9),
        legend.text = element_text(size=11),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor   = element_blank(),
        panel.grid.major.x = element_line(),
        axis.text.x        = element_text(angle = 45, hjust = 1, size = 9),
        plot.title         = element_text(size = 14, hjust = 0.5, face = "bold")) +
  scale_fill_manual(values = c("#0058AB","#69DBFF"),name = "") +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = c(seq(2000,rev_yr,1), 2025, 2030))
```

```{r overview-plot, echo=FALSE, fig.width=10, fig.height=5.3}
print(heatmap_overview)
```

```{r all_vax_plot, echo=FALSE, fig.width=10, fig.height=4.9}
print(plt_ia2030_zd_proj_bar_line)
```

\newpage

<!-- ===================== PAGE 2: DTP1 ===================== -->

```{r setup-dtp1, include=FALSE}
v       <- "dtp1"
unvac_v <- unvac %>% filter(vaccine == v)
cvg_v   <- cvg   %>% filter(vaccine == v)
vacc_v  <- vacc  %>% filter(vaccine == v)
ctry_cvg_rev  <- cvg_v %>% pull(recent_col)
ctry_cvg_comp <- cvg_v %>% pull(comparison_col)
ctry_cvg_minusone <- cvg_v %>% pull(minusone_col)

dropout_dtp <- df %>%
  filter(year %in% c(comp_yr, minusone_col, rev_yr), vaccine %in% c("dtp1", "dtp3")) %>%
  select(iso3c, country, year, vaccine, vaccinated) %>%
  pivot_wider(names_from = vaccine, values_from = vaccinated) %>%
  mutate(
    dropout     = round(((dtp1 - dtp3) / dtp1) * 100, 0),
    dropout_txt = case_when(
      dropout < 5  ~ str_glue(get_text2("txt_dropout_low", text_vars), vax_pair = dtp_translated),
      dropout <= 10 ~ str_glue(get_text2("txt_dropout_medium", text_vars), vax_pair = dtp_translated),
      dropout > 10 ~ str_glue(get_text2("txt_dropout_high", text_vars), vax_pair = dtp_translated)
    )
  )
dropout_comp_dtp <- dropout_dtp %>%
  select(-c(dtp1, dtp3, dropout_txt)) %>%
  pivot_wider(names_from = year, values_from = dropout) %>%
  mutate(
    diff     = .data[[recent_col]] - .data[[comparison_col]],
    diff_txt = case_when(
      diff == 1  ~ str_glue(get_text2("txt_general_dropout_diff_higher_1", text_vars)),
      diff > 0   ~ str_glue(get_text2("txt_general_dropout_diff_higher_pp", text_vars)),
      diff == -1 ~ str_glue(get_text2("txt_general_dropout_diff_lower_1", text_vars)),
      diff < -1  ~ str_glue(get_text2("txt_general_dropout_diff_lower_pp", text_vars)),
      TRUE       ~ str_glue(get_text2("txt_general_dropout_diff_same", text_vars))))

###### first bullet ######
dir_comp <- case_when(
  ctry_cvg_rev > ctry_cvg_comp ~ "inc",
  ctry_cvg_rev < ctry_cvg_comp ~ "dec",
  TRUE                         ~ "sam"
)
dir_minusone <- case_when(
  ctry_cvg_rev > ctry_cvg_minusone ~ "inc",
  ctry_cvg_rev < ctry_cvg_minusone ~ "dec",
  TRUE                             ~ "same"
)
target_text_key <- paste0("txt_vax_", dir_comp, "_", dir_minusone)
bullet_dtp1_1 <- str_glue(get_text2(target_text_key, text_vars), vax = dtp1_translated)
#####

##### bullet 2 - IA2030 zero-dose comparison + zero-dose counts vs 2019 and 2024
dtp1_dropout_row <- dropout_dtp %>% filter(year == rev_yr)

ia2030_key <- case_when(
  zd_pcnt_diff > 0 ~ "txt_dtp1_zd_ia2030_higher",
  zd_pcnt_diff < 0 ~ "txt_dtp1_zd_ia2030_lower",
  TRUE             ~ "txt_dtp1_zd_ia2030_same"
)
sentence_zd_ia2030 <- str_glue(get_text2(ia2030_key, text_vars),
  abs_zd_pcnt_diff = abs(zd_pcnt_diff),
  rev_yr           = rev_yr)

zd_rev      <- unvac_v %>% pull(recent_col)
zd_comp     <- unvac_v %>% pull(comparison_col)
zd_minusone <- unvac_v %>% pull(minusone_col)

zd_tgt <- tibble(rev = zd_rev, comp = zd_comp, minusone = zd_minusone) %>%
  label_vals_millions(rev, 'rev_lbl') %>%
  label_vals_millions(comp, 'comp_lbl') %>%
  label_vals_millions(minusone, 'minusone_lbl')

dir_comp_zd <- case_when(
  zd_rev > zd_comp ~ "inc",
  zd_rev < zd_comp ~ "dec",
  TRUE             ~ "sam"
)
dir_minusone_zd <- case_when(
  zd_rev > zd_minusone ~ "inc",
  zd_rev < zd_minusone ~ "dec",
  TRUE                 ~ "same"
)
target_zd_key <- paste0("txt_dtp1_zd_count_", dir_comp_zd, "_", dir_minusone_zd)

sentence_zd_count <- str_glue(get_text2(target_zd_key, text_vars),
  zd_rev_lbl      = zd_tgt$rev_lbl,
  zd_comp_lbl     = zd_tgt$comp_lbl,
  zd_minusone_lbl = zd_tgt$minusone_lbl,
  rev_yr          = rev_yr,
  comp_yr         = comp_yr,
  minusone_yr     = minusone_yr)

bullet_dtp1_2 <- paste(sentence_zd_ia2030, sentence_zd_count)

##### bullet 3: comparison to regional average coverage
# get regional average DTP1 coverage
reg_avg_cvg <- wuenic_dta %>% 
  filter(Region == region, lvl_1 == "region") %>% 
  filter(vaccine == v, year == rev_yr) %>%
  pull(coverage)
regional_avg_direction <- case_when(
  ctry_cvg_rev > reg_avg_cvg ~ "txt_cvg_higher_than_regional_avg",
  ctry_cvg_rev < reg_avg_cvg ~ "txt_cvg_lower_than_regional_avg",
  TRUE                       ~ "txt_cvg_same_as_regional_avg"
)
bullet_dtp1_3 <- str_glue(get_text2(regional_avg_direction, text_vars), vax = dtp1_translated)

##### bullet 4: dropout
if (dtp1_dropout_row$dropout < 0) {
  dropout_sentence <- get_text2("txt_dtp1_negative_dropout", text_vars)
  bullet_dtp1_4 <- str_glue(
    dropout_sentence,
    abs_dropout_pct  = abs(dtp1_dropout_row$dropout), # Converted to absolute positive value
    dropout_diff_txt = dropout_comp_dtp %>% pull(diff_txt),
    rev_yr           = rev_yr
  )
} else {
  dropout_sentence <- get_text2("txt_dtp1_dropout", text_vars)
  bullet_dtp1_4 <- str_glue(
    dropout_sentence,
    dropout_pct      = dtp1_dropout_row$dropout,
    dropout_diff_txt = dropout_comp_dtp %>% pull(diff_txt),
    dropout_txt      = dtp1_dropout_row$dropout_txt,
    rev_yr           = rev_yr
  )
}
```

```{r dtp1-line-plot, echo=FALSE, warning=FALSE, message=FALSE}
plot_dtp1  <- df %>% filter(vaccine == "dtp1") %>% select(year, coverage)
plot_zd_dtp <- df %>% filter(vaccine == "dtp1") %>% select(year, unvaccinated)

scale_factor <- max(plot_dtp1$coverage, na.rm = TRUE) /
                max(plot_zd_dtp$unvaccinated, na.rm = TRUE)

dtp_plot <- ggplot() +
  geom_col(data = plot_zd_dtp, aes(x = year, y = unvaccinated, fill = t_lookup("txt_plt_dtp1_legend_zd", language)), alpha = 0.4, width = 0.6) +
  geom_line(data = plot_dtp1, aes(x = year, y = coverage / scale_factor, color = t_lookup("txt_plt_dtp1_legend_cvg", language)), size = 1.5) +
  geom_point(data = plot_dtp1, aes(x = year, y = coverage / scale_factor, color = t_lookup("txt_plt_dtp1_legend_cvg", language)), size = 3) +
  geom_hline(yintercept = 90 / scale_factor, linetype = "dashed") +
  scale_color_manual(name = NULL, values = setNames("#1CABE2", t_lookup("txt_plt_dtp1_legend_cvg", language))) +
  scale_fill_manual(name = NULL, values = setNames("#E2231A", t_lookup("txt_plt_dtp1_legend_zd", language))) +
  scale_x_continuous(breaks = seq(min(plot_dtp1$year), max(plot_dtp1$year), 1)) +
  scale_y_continuous(name = t_lookup("txt_plt_zd_children", language), labels = scales::label_comma(),
    sec.axis = sec_axis(~ . * scale_factor, name = t_lookup("txt_plt_cvg_pct", language), breaks = seq(0, 100, 25))) +
  labs(title = str_glue(get_text2("txt_plt_dtp1_title", plot_vars)),
       caption = paste0(t_lookup("txt_plt_ia2030_caption", language), "\n", str_glue(get_text2("txt_wuenic_source", plot_vars))),
       x = t_lookup("txt_plt_year", language)) +
  guides(color = guide_legend(order = 1), fill = guide_legend(order = 2)) +
  theme_unicef() +
  theme(panel.grid.minor = element_blank(), panel.grid.major.x = element_line(),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
        plot.title = element_text(size = 14, hjust = 0.5, face = "bold"), legend.position = "bottom")
```

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{`r t_lookup("txt_sec_dtp1", language)`}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:dtp1}{}")
cat(paste0("\\addcontentsline{toc}{section}{", t_lookup("txt_sec_dtp1", language), "}"))
```

- `r bullet_dtp1_1`

- `r bullet_dtp1_2`

- `r bullet_dtp1_3`

- `r bullet_dtp1_4`

\vspace{1em}

```{r dtp1, echo=FALSE, fig.width=10, fig.height=6.2}
print(dtp_plot)
```

<!-- ===================== PAGE 3: DTP3 ===================== -->

```{r setup-dtp3, include=FALSE}
v       <- "dtp3"
unvac_v <- unvac %>% filter(vaccine == v)
cvg_v   <- cvg   %>% filter(vaccine == v)
vacc_v  <- vacc  %>% filter(vaccine == v)

ctry_cvg_rev  <- ctry_cvg_rev_dtp3 <- cvg_v %>% pull(recent_col)
ctry_cvg_comp <- ctry_cvg_comp_dtp3 <- cvg_v %>% pull(comparison_col)
ctry_cvg_minusone <- ctry_cvg_minusone_dtp3 <- cvg_v %>% pull(minusone_col)

reg_rank_dtp3 <- reg_peers %>%
  filter(vaccine == v, year == rev_yr) %>%
  arrange(desc(coverage)) %>%
  mutate(rank = row_number()) %>%
  filter(iso3c == ctry_iso)

unvac_dtp3_rev <- unvac_v %>% pull(recent_col)
unvac_dtp3_lbl <- label_vals_millions_num(unvac_dtp3_rev)

tgt_chng <- tgt %>%
  mutate(pop_grwth = round((.data[[recent_col]] / .data[[comparison_col]] - 1) * 100, 0)) %>%
  label_vals_millions(.[[recent_col]], "tgt_rev_yr") %>%
  label_vals_millions(.[[comparison_col]], "tgt_comp_yr")

pop_chng_txt <- if (ctry_cvg_rev_dtp3 < ctry_cvg_comp_dtp3 & tgt[[comparison_col]] > tgt[[recent_col]]) {
  str_glue(get_text2("txt_pop_growth_context", text_vars),
           pop_grwth = tgt_chng$pop_grwth)
} else { "" }

###### first bullet ######
dir_comp <- case_when(
  ctry_cvg_rev > ctry_cvg_comp ~ "inc",
  ctry_cvg_rev < ctry_cvg_comp ~ "dec",
  TRUE                         ~ "sam"
)
dir_minusone <- case_when(
  ctry_cvg_rev > ctry_cvg_minusone ~ "inc",
  ctry_cvg_rev < ctry_cvg_minusone ~ "dec",
  TRUE                             ~ "same"
)
target_text_key <- paste0("txt_vax_", dir_comp, "_", dir_minusone)
bullet_dtp3_1 <- str_glue(get_text2(target_text_key, text_vars), vax = dtp3_translated)
#####

# get perc. change and mark if increasing, decreasing, or steady
raw_pchng_value <- vacc_v %>% pull(pchng) 
base_sentence_key <- case_when(
  raw_pchng_value > 0 ~ "txt_dtp3_vaccinated_increased",
  raw_pchng_value < 0 ~ "txt_dtp3_vaccinated_decreased",
  TRUE                ~ "txt_dtp3_vaccinated_unchanged"
)

# get translation
dtp3_sentence <- get_text2(base_sentence_key, text_vars)

bullet_dtp3_2 <- str_glue(
  dtp3_sentence,
  vacc_comp_yr_lbl   = vacc_v %>% pull(vacc_comp_yr_lbl),
  vacc_rev_yr_lbl    = vacc_v %>% pull(vacc_rev_yr_lbl),
  comp_yr            = comp_yr,
  rev_yr             = rev_yr,
  the_ctry_name      = the_ctry_name,
  ctry_cvg_rev_dtp3  = ctry_cvg_rev_dtp3,
  ctry_cvg_comp_dtp3 = ctry_cvg_comp_dtp3,
  pop_chng_txt       = pop_chng_txt
)

##### bullet 3: number unvaccinated
bullet_dtp3_3 <- str_glue(get_text2("txt_dtp3_unvac", text_vars),
  ctry_cvg_rev_dtp3 = ctry_cvg_rev_dtp3,
  unvac_dtp3_lbl    = unvac_dtp3_lbl)

##### bullet 4: dropout
dtp3_dropout_row      <- dropout_dtp %>% filter(year == rev_yr)
dtp3_dropout_comp     <- dropout_dtp %>% filter(year == comp_yr)
dtp3_dropout_minusone <- dropout_dtp %>% filter(year == minusone_col)
current_dropout  <- dtp3_dropout_row$dropout
comp_dropout     <- dtp3_dropout_comp$dropout
minusone_dropout <- dtp3_dropout_minusone$dropout

# get direction of change
trend_comp <- case_when(
  current_dropout > comp_dropout ~ "higher",
  current_dropout < comp_dropout ~ "lower",
  TRUE                           ~ "same"
)
trend_2024 <- case_when(
  current_dropout > minusone_dropout ~ "higher",
  current_dropout < minusone_dropout ~ "lower",
  TRUE                               ~ "same"
)

# generate text and bullet
target_text_key <- paste0("txt_dropout_", trend_comp, "_", trend_2024)
bullet_dtp3_4 <- str_glue(
  get_text2(target_text_key, text_vars),
  vax1 = dtp1_translated, vax2 = dtp3_translated,
  dropout_pct      = current_dropout,
  comp_pct         = comp_dropout,
  minusone_pct     = minusone_dropout,
  diff_comp        = abs(current_dropout - comp_dropout),
  diff_minusone    = abs(current_dropout - minusone_dropout)
)

##### bullet 5: comparison to regional average coverage
# get regional average DTP3 coverage
reg_avg_cvg <- wuenic_dta %>% 
  filter(Region == region, lvl_1 == "region") %>% 
  filter(vaccine == v, year == rev_yr) %>%
  pull(coverage)
regional_avg_direction <- case_when(
  ctry_cvg_rev > reg_avg_cvg ~ "txt_cvg_higher_than_regional_avg",
  ctry_cvg_rev < reg_avg_cvg ~ "txt_cvg_lower_than_regional_avg",
  TRUE                       ~ "txt_cvg_same_as_regional_avg"
)
bullet_dtp3_5 <- str_glue(get_text2(regional_avg_direction, text_vars), vax = dtp3_translated)
```

```{r dtp3-line-plot, echo=FALSE, warning=FALSE, message=FALSE}
plot_dtp3   <- df %>% filter(vaccine == "dtp3") %>% select(year, coverage)
plot_zd_dtp3 <- df %>% filter(vaccine == "dtp3") %>% select(year, unvaccinated)

scale_factor <- max(plot_zd_dtp3$unvaccinated, na.rm = TRUE) / 100

dtp3_plot <- ggplot() +
  geom_col(data = plot_zd_dtp3, aes(x = year, y = unvaccinated, 
                                   fill = t_lookup("txt_plt_dtp3_legend_unvac", language)), alpha = 0.4, width = 0.6) +
  geom_line(data = plot_dtp3, aes(x = year, y = coverage * scale_factor, 
                                 color = t_lookup("txt_plt_dtp3_legend_cvg", language)), size = 1.5) +
  geom_point(data = plot_dtp3, aes(x = year, y = coverage * scale_factor, 
                                   color = t_lookup("txt_plt_dtp3_legend_cvg", language)), size = 3) +
  geom_hline(yintercept = 90 * scale_factor, linetype = "dashed") +
  scale_color_manual(name = NULL, values = setNames("#1CABE2", t_lookup("txt_plt_dtp3_legend_cvg", language))) +
  scale_fill_manual(name = NULL, values = setNames("#E2231A", t_lookup("txt_plt_dtp3_legend_unvac", language))) +
  scale_x_continuous(breaks = seq(min(plot_dtp3$year), max(plot_dtp3$year), 1)) +
  scale_y_continuous(name = t_lookup("txt_plt_unvac_undervac", language), 
                     labels = scales::label_comma(),
                     sec.axis = sec_axis(~ . / scale_factor, 
                                         name = t_lookup("txt_plt_cvg_pct", language),
                                         breaks = seq(0, 100, 25))) +
  labs(title = str_glue(get_text2("txt_plt_dtp3_title", plot_vars)),
       caption = paste0(t_lookup("txt_plt_ia2030_caption", language), "\n", str_glue(get_text2("txt_wuenic_source", plot_vars))),
       x = t_lookup("txt_plt_year", language), color = NULL) +
  theme_unicef() +
  theme(panel.grid.minor = element_blank(), 
        panel.grid.major.x = element_line(),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
        plot.title = element_text(size = 14, hjust = 0.5, face = "bold"), 
        legend.position = "bottom")
```

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{`r t_lookup("txt_sec_dtp3", language)`}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:dtp3}{}")
cat(paste0("\\addcontentsline{toc}{section}{", t_lookup("txt_sec_dtp3", language), "}"))
```

- `r bullet_dtp3_1`

- `r bullet_dtp3_2`

- `r bullet_dtp3_3`

- `r bullet_dtp3_4`

- `r bullet_dtp3_5`

\vspace{1em}

```{r dtp3, echo=FALSE, fig.width=10, fig.height=6.2}
print(dtp3_plot)
```

<!-- ===================== PAGE 4: MCV ===================== -->

```{r setup-mcv, include=FALSE}
v <- "mcv1"
v_mcv   <- if ("mcv2" %in% available_vaccines) c("mcv1", "mcv2") else "mcv1"
unvac_v <- unvac %>% filter(vaccine %in% v_mcv)
cvg_v   <- cvg   %>% filter(vaccine %in% v_mcv)

ctry_cvg_rev  <- cvg_v %>% filter(vaccine == v) %>% pull(recent_col)
ctry_cvg_comp <- cvg_v %>% filter(vaccine == v) %>% pull(comparison_col)
ctry_cvg_minusone <- cvg_v %>% filter(vaccine == v) %>% pull(minusone_col)

cvg_mcv1 <- cvg_v %>%
  filter(vaccine == v) %>%
  mutate(
    diff         = .[[recent_col]] - .[[comparison_col]],
    cvg_chng_txt = case_when(
      diff %in% c(-1:1) ~ str_glue(get_text2("txt_cvg_stable", text_vars)),
      diff < -1         ~ str_glue(get_text2("txt_cvg_decreased", text_vars),
                                   cvg_comp = .[[comparison_col]], cvg_rev = .[[recent_col]]),
      diff > 1          ~ str_glue(get_text2("txt_cvg_increased", text_vars),
                                   cvg_comp = .[[comparison_col]], cvg_rev = .[[recent_col]])))

if ("mcv2" %in% available_vaccines) {
  cvg_mcv2 <- cvg_v %>%
    filter(vaccine == "mcv2") %>%
    mutate(
      diff         = .[[recent_col]] - .[[comparison_col]],
      cvg_chng_txt = case_when(
        diff %in% c(-1:1) ~ str_glue(get_text2("txt_cvg_stable",    text_vars)),
        diff < -1         ~ str_glue(get_text2("txt_cvg_decreased",  text_vars),
                                     cvg_comp = .[[comparison_col]], cvg_rev = .[[recent_col]]),
        diff > 1          ~ str_glue(get_text2("txt_cvg_increased",  text_vars),
                                     cvg_comp = .[[comparison_col]], cvg_rev = .[[recent_col]])))

  unvac_mcv <- unvac_v %>%
    select(iso3c, country, vaccine, all_of(recent_col)) %>% 
    pivot_wider(names_from = vaccine, values_from = last_col()) %>%
    mutate(diff = mcv2 - mcv1) %>%
    label_vals_millions(mcv1, "mcv1_lbl") %>%
    label_vals_millions(mcv2, "mcv2_lbl") %>%
    label_vals_millions(diff, "diff_lbl")
} else {
  cvg_mcv2  <- cvg_v %>%
    mutate(cvg_chng_txt = str_glue(
      get_text2("txt_mcv2_no_data", text_vars)))
  unvac_mcv <- NULL
}

reg_rank_mcv1 <- reg_peers %>%
  filter(vaccine == "mcv1", year == rev_yr) %>%
  arrange(desc(coverage)) %>%
  mutate(rank = row_number()) %>%
  filter(iso3c == ctry_iso)

dropout_mcv <- df %>%
  filter(year %in% c(comp_yr, rev_yr, minusone_col), vaccine %in% c("dtp1", "mcv1")) %>%
  select(iso3c, country, year, vaccine, vaccinated) %>%
  pivot_wider(names_from = vaccine, values_from = vaccinated) %>%
  mutate(
    dropout     = round(((dtp1 - mcv1) / dtp1) * 100, 0),
    dropout_txt = case_when(
      dropout < 5   ~ get_text2("txt_dropout_low_mcv",    text_vars),
      dropout <= 10 ~ get_text2("txt_dropout_medium_mcv", text_vars),
      dropout > 10  ~ get_text2("txt_dropout_high_mcv",   text_vars)))

dropout_comp_mcv <- dropout_mcv %>%
  select(-c(dtp1, mcv1, dropout_txt)) %>%
  pivot_wider(names_from = year, values_from = dropout) %>%
  mutate(
    diff     = .[[recent_col]] - .[[comparison_col]],
    diff_txt = case_when(
      diff == 1  ~ str_glue(get_text2("txt_general_dropout_diff_higher_1",  text_vars)),
      diff > 0   ~ str_glue(get_text2("txt_general_dropout_diff_higher_pp", text_vars)),
      diff == -1 ~ str_glue(get_text2("txt_general_dropout_diff_lower_1",   text_vars)),
      diff < -1  ~ str_glue(get_text2("txt_general_dropout_diff_lower_pp",  text_vars)),
      TRUE       ~ str_glue(get_text2("txt_general_dropout_diff_same",      text_vars))))

nyrs     <- 3; thrshold <- 80
cvg_trend_mcv1 <- df %>%
  filter(vaccine == "mcv1", year %in% seq(rev_yr - nyrs + 1, rev_yr)) %>%
  mutate(ltx = ifelse(coverage < thrshold, 1, 0)) %>%
  summarise(sum = sum(ltx), mean_cvg = mean(coverage, na.rm = TRUE))

capital_the_name <- if_else(str_starts(the_ctry_name, "the "),
                             str_replace(the_ctry_name, "^the ", "The "),
                             the_ctry_name)

persistent_low_txt <- if (cvg_trend_mcv1$sum == nyrs) {
  str_glue(get_text2("txt_mcv_persistent_low_yes", text_vars),
           capital_the_name = capital_the_name,
           nyrs = nyrs, thrshold = thrshold)
} else {
  str_glue(get_text2("txt_mcv_persistent_low_no",  text_vars),
           capital_the_name = capital_the_name,
           nyrs = nyrs, thrshold = thrshold)
}

###### first bullet ######
dir_comp <- case_when(
  ctry_cvg_rev > ctry_cvg_comp ~ "inc",
  ctry_cvg_rev < ctry_cvg_comp ~ "dec",
  TRUE                         ~ "sam"
)
dir_minusone <- case_when(
  ctry_cvg_rev > ctry_cvg_minusone ~ "inc",
  ctry_cvg_rev < ctry_cvg_minusone ~ "dec",
  TRUE                             ~ "same"
)
target_text_key <- paste0("txt_vax_", dir_comp, "_", dir_minusone)
bullet_mcv_1 <- str_glue(get_text2(target_text_key, text_vars), vax = mcv1_translated)
#####

direction_of_mcv_change <- if (cvg_mcv1$diff > 0) {
  "increased"
} else if (cvg_mcv1$diff < 0) {
  "decreased"
} else {
  "stable"
}
bullet2_type <- ifelse(direction_of_mcv_change %in% c("increased", "stable"), "txt_mcv2_cvg_change_increased", "txt_mcv2_cvg_change_decreased")
                       
bullet_mcv_2 <- str_glue(get_text2(bullet2_type, text_vars),
  cvg_chng_txt  = cvg_mcv2 %>% pull(cvg_chng_txt),
  mcv1_unvac_lbl = unvac_mcv$mcv1_lbl,
  mcv_diff_lbl   = unvac_mcv$diff_lbl)

##### bullet 3: dropout
mcv_dropout_row      <- dropout_mcv %>% filter(year == rev_yr)
mcv_dropout_comp     <- dropout_mcv %>% filter(year == comp_yr)
mcv_dropout_minusone <- dropout_mcv %>% filter(year == minusone_col)
current_dropout  <- mcv_dropout_row$dropout
comp_dropout     <- mcv_dropout_comp$dropout
minusone_dropout <- mcv_dropout_minusone$dropout

# get direction of change
trend_comp <- case_when(
  current_dropout > comp_dropout ~ "higher",
  current_dropout < comp_dropout ~ "lower",
  TRUE                           ~ "same"
)
trend_2024 <- case_when(
  current_dropout > minusone_dropout ~ "higher",
  current_dropout < minusone_dropout ~ "lower",
  TRUE                               ~ "same"
)

# generate text and bullet
target_text_key <- paste0("txt_dropout_", trend_comp, "_", trend_2024)
bullet_mcv_3 <- str_glue(
  get_text2(target_text_key, text_vars),
  vax1 = dtp1_translated, vax2 = mcv1_translated,
  dropout_pct      = current_dropout,
  comp_pct         = comp_dropout,
  minusone_pct     = minusone_dropout,
  diff_comp        = abs(current_dropout - comp_dropout),
  diff_minusone    = abs(current_dropout - minusone_dropout)
)

bullet_mcv_4 <- persistent_low_txt

##### bullet 5: comparison to regional average coverage
# get regional average MCV coverage
reg_avg_cvg <- wuenic_dta %>% 
  filter(Region == region, lvl_1 == "region") %>% 
  filter(vaccine == v, year == rev_yr) %>%
  pull(coverage)
regional_avg_direction <- case_when(
  ctry_cvg_rev > reg_avg_cvg ~ "txt_cvg_higher_than_regional_avg",
  ctry_cvg_rev < reg_avg_cvg ~ "txt_cvg_lower_than_regional_avg",
  TRUE                       ~ "txt_cvg_same_as_regional_avg"
)
bullet_mcv_5 <- str_glue(get_text2(regional_avg_direction, text_vars), vax = mcv1_translated)
```

```{r mcv-line-plot, echo=FALSE, warning=FALSE, message=FALSE}
plot_mcv1  <- df %>% filter(vaccine == "mcv1") %>% select(year, coverage)
if ("mcv2" %in% available_vaccines)
  plot_mcv2 <- df %>% filter(vaccine == "mcv2") %>% select(year, coverage)
plot_zd_mcv <- df %>% filter(vaccine == "mcv1") %>% select(year, unvaccinated)

scale_factor <- max(plot_mcv1$coverage, na.rm = TRUE) /
                max(plot_zd_mcv$unvaccinated, na.rm = TRUE)

mcv_plot <- ggplot() +
  geom_col(data = plot_zd_mcv, aes(x = year, y = unvaccinated, fill = t_lookup("txt_plt_mcv_legend_zd", language)), alpha = 0.4, width = 0.6) +
  geom_line(data = plot_mcv1, aes(x = year, y = coverage / scale_factor, color = t_lookup("txt_plt_mcv1_legend_cvg", language)), size = 1.5) +
  geom_point(data = plot_mcv1, aes(x = year, y = coverage / scale_factor, color = t_lookup("txt_plt_mcv1_legend_cvg", language)), size = 3) +
  { if ("mcv2" %in% available_vaccines) list(
      geom_line(data = plot_mcv2, aes(x = year, y = coverage / scale_factor, color = t_lookup("txt_plt_mcv2_legend_cvg", language)), size = 1.5),
      geom_point(data = plot_mcv2, aes(x = year, y = coverage / scale_factor, color = t_lookup("txt_plt_mcv2_legend_cvg", language)), size = 3)) } +
  geom_hline(yintercept = 90 / scale_factor, linetype = "dashed") +
  scale_color_manual(name = NULL, values = setNames(c("#1CABE2", "#6A1E74"),
    c(t_lookup("txt_plt_mcv1_legend_cvg", language), t_lookup("txt_plt_mcv2_legend_cvg", language)))) +
  scale_fill_manual(name = NULL, values = setNames("#E2231A", t_lookup("txt_plt_mcv_legend_zd", language))) +
  scale_x_continuous(breaks = seq(min(plot_mcv1$year), max(plot_mcv1$year), 1)) +
  scale_y_continuous(name = t_lookup("txt_plt_mcv_legend_zd", language), labels = scales::label_comma(),
    sec.axis = sec_axis(~ . * scale_factor, name = t_lookup("txt_plt_cvg_pct", language), breaks = seq(0, 100, 25))) +
  guides(color = guide_legend(order = 1), fill = guide_legend(order = 2)) +
  labs(title = str_glue(get_text2("txt_plt_mcv_title", plot_vars)),
       caption = paste0(t_lookup("txt_plt_ia2030_caption", language), "\n", str_glue(get_text2("txt_wuenic_source", plot_vars))),
       x = t_lookup("txt_plt_year", language)) +
  theme_unicef() +
  theme(panel.grid.minor = element_blank(), panel.grid.major.x = element_line(),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
        plot.title = element_text(size = 14, hjust = 0.5, face = "bold"), legend.position = "bottom")
```

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{`r t_lookup("txt_sec_mcv", language)`}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:mcv}{}")
cat(paste0("\\addcontentsline{toc}{section}{", t_lookup("txt_sec_mcv", language), "}"))
```

- `r bullet_mcv_1`

- `r bullet_mcv_2`

- `r bullet_mcv_3`

- `r bullet_mcv_4`

- `r bullet_mcv_5`

\vspace{1em}

```{r mcv, echo=FALSE, fig.width=10, fig.height=6.2}
print(mcv_plot)
```

<!-- ===================== PAGE 5: HPV ===================== -->

```{r setup-hpv, include=FALSE}
# helper: ensure a column exists after pivot_wider; if missing, add as all-NA
ensure_col <- function(data, col_name) {
  if (!col_name %in% names(data)) {
    data[[col_name]] <- NA_real_
  }
  data
}

if (has_hpv_data && nrow(hpv_df) > 0) {
  available_hpv <- unique(hpv_df$vaccine_code)
  has_hpvc      <- "PRHPVC_F" %in% available_hpv
  has_hpv1      <- "PRHPV1_F" %in% available_hpv

  hpvc_yrs <- hpv_df %>% filter(vaccine_code == "PRHPVC_F") %>% pull(year)
  hpv1_yrs <- hpv_df %>% filter(vaccine_code == "PRHPV1_F") %>% pull(year)

  # --- DYNAMIC YEAR MANAGEMENT FALLBACK ---
  # Check if the global evaluation year exists in this country's data
  if (hpv_rev_yr %in% hpv_df$year) {
    local_rev_yr       <- hpv_rev_yr
    local_recent_col   <- recent_col
    local_minusone_col <- minusone_col
  } else {
    # Fallback to the country's most recent available data year (e.g., 2024)
    local_rev_yr       <- max(hpv_df$year, na.rm = TRUE)
    local_recent_col   <- as.character(local_rev_yr)
    # If 2019 exists, prioritize it for historic long-term comparison
    local_minusone_col <- if (2019 %in% hpv_df$year) "2019" else as.character(local_rev_yr - 1)
  }

  has_hpvc_comparison <- has_hpvc && all(c(as.numeric(local_recent_col), as.numeric(local_minusone_col)) %in% hpvc_yrs)
  has_hpv1_comparison <- has_hpv1 && all(c(as.numeric(local_recent_col), as.numeric(local_minusone_col)) %in% hpv1_yrs)

  primary_hpv_code  <- if (has_hpvc_comparison) "PRHPVC_F" else "PRHPV1_F"
  primary_hpv_label <- if (has_hpvc_comparison) t_lookup("hpvc", language) else t_lookup("hpv1", language)
  
  hpvc_label <- t_lookup("hpvc", language)
  has_hpvc_for_rev_yr <- has_hpvc && local_rev_yr %in% hpvc_yrs
  
  # coverage change strings helper
  make_hpv_diff_txt <- function(df_row) {
    d <- df_row$diff
    if (is.na(d)) {
      str_glue(get_text2("txt_hpv_single_year_hpv1", text_vars))
    } else if (d > 0) {
      str_glue(t_lookup("txt_vacc_increased", language),
               cvg_prev = df_row[[local_minusone_col]], cvg_rev = df_row[[local_recent_col]])
    } else if (d < 0) {
      str_glue(t_lookup("txt_vacc_decreased", language),
               cvg_prev = df_row[[local_minusone_col]], cvg_rev = df_row[[local_recent_col]])
    } else {
      str_glue(t_lookup("txt_vacc_constant", language),
               cvg_rev = df_row[[local_recent_col]])
    }
  }
  
  hpv1_rev_cvg <- hpv_df %>%
    filter(vaccine_code == "PRHPV1_F", year == local_rev_yr) %>%
    pull(coverage)
  if (length(hpv1_rev_cvg) == 0) hpv1_rev_cvg <- NA_real_

  # coverage change strings
  hpv <- hpv_df %>%
    filter(vaccine_code == primary_hpv_code) %>%
    select(iso3c, country, vaccine_code, year, coverage) %>%
    filter(year %in% c(as.numeric(local_recent_col), as.numeric(local_minusone_col))) %>%
    pivot_wider(names_from = year, values_from = coverage) %>%
    ensure_col(local_recent_col) %>%
    ensure_col(local_minusone_col) %>%
    mutate(
      diff = .data[[local_recent_col]] - .data[[local_minusone_col]],
      chng = case_when(
        is.na(diff) ~ NA_character_,
        diff > 0 ~ "increased", 
        diff < 0 ~ "decreased", 
        TRUE ~ "constant"
      ),
      rev_val = .data[[local_recent_col]],
      minusone_val = .data[[local_minusone_col]],
      diff_txt = case_when(
        is.na(diff) ~ str_glue(get_text2("txt_hpv_single_year_hpv1", text_vars)),
        chng %in% c("increased", "decreased") ~ 
          str_glue(get_text2("txt_hpv_changed", text_vars)),
        TRUE ~ 
          str_glue(get_text2("txt_hpv_constant", text_vars))
      ),
      
      # Target threshold evaluation text
      gt90 = case_when(
        is.na(.data[[local_recent_col]]) ~ NA_character_,
        .data[[local_recent_col]] < 80  & chng == "increased" ~ 
          str_glue(t_lookup("txt_hpv_gt90_below80_up", language), primary_hpv_label = primary_hpv_label),
        .data[[local_recent_col]] < 80  & chng != "increased" ~ 
          str_glue(t_lookup("txt_hpv_gt90_below80", language), primary_hpv_label = primary_hpv_label),
        .data[[local_recent_col]] %in% 80:89 & chng == "increased" ~ 
          str_glue(t_lookup("txt_hpv_gt90_near_up", language), primary_hpv_label = primary_hpv_label),
        .data[[local_recent_col]] %in% 80:89 & chng != "increased" ~ 
          str_glue(t_lookup("txt_hpv_gt90_near", language), primary_hpv_label = primary_hpv_label),
        .data[[local_recent_col]] == 90 ~ 
          str_glue(t_lookup("txt_hpv_gt90_achieved", language), primary_hpv_label = primary_hpv_label),
        .data[[local_recent_col]] > 90  ~ 
          str_glue(t_lookup("txt_hpv_gt90_exceeded", language), primary_hpv_label = primary_hpv_label),
        TRUE ~ NA_character_
      )
    )

  # Check if the reference year exists in the dataset
if (rev_yr %in% hpv_df$year) {
  
  hpv_unvac <- hpv_df %>%
    filter(year == rev_yr) %>%
    select(iso3c, country, vaccine_code, unvaccinated) %>%
    distinct(vaccine_code, .keep_all = TRUE) %>% 
    pivot_wider(names_from = vaccine_code, values_from = unvaccinated)

  # Safe generation if pivot_wider returns an empty dataframe structure
  if (nrow(hpv_unvac) == 0) {
    hpv_unvac <- tibble(PRHPVC_F = NA_real_, PRHPV1_F = NA_real_)
  } else {
    if (!"PRHPVC_F" %in% names(hpv_unvac)) hpv_unvac$PRHPVC_F <- NA_real_
    if (!"PRHPV1_F" %in% names(hpv_unvac)) hpv_unvac$PRHPV1_F <- NA_real_
  }

  hpv_unvac <- hpv_unvac %>%
    mutate(diff = PRHPVC_F - PRHPV1_F) %>%
    label_vals_millions(PRHPVC_F, "PRHPVC_F_lbl") %>%
    label_vals_millions(PRHPV1_F, "PRHPV1_F_lbl") %>%
    label_vals_millions(diff,     "diff_lbl")

  # Set flag based on actual data presence
  has_hpv_num_unvac <- !is.na(hpv_unvac$PRHPV1_F) | !is.na(hpv_unvac$PRHPVC_F)

} else {
  # If rev_yr isn't in the data, turn the flag off so the bullet is suppressed
  has_hpv_num_unvac <- FALSE
}
  
  first_hpv_year <- NA_integer_
  if (exists("hpv_intro") && nrow(hpv_intro) > 0) {
    first_hpv_year <- hpv_intro %>% filter(introduced == 1) %>% pull(year) %>% min(na.rm = TRUE)
  }

  reg_rank_hpv <- hpv_peers %>%
    filter(vaccine_code == primary_hpv_code, year == rev_yr) %>%
    arrange(desc(coverage)) %>%
    mutate(rank = row_number()) %>%
    filter(iso3c == ctry_iso)

  n_peers_hpv <- hpv_peers %>%
    filter(vaccine_code == primary_hpv_code, year == rev_yr) %>%
    distinct(iso3c) %>% 
    nrow()
  
  hpv_years_wide <- hpv_df %>% 
    select(iso3c, country, vaccine_code, year, coverage) %>%
    pivot_wider(names_from = year, values_from = coverage)

  # get hpv1 diff_txt for the "both" bullet
  if (has_hpv1 && has_hpvc) {
    hpv1_row <- hpv_df %>%
      filter(vaccine_code == "PRHPV1_F") %>%
      select(iso3c, country, vaccine_code, year, coverage) %>%
      filter(year %in% c(as.numeric(local_recent_col), as.numeric(local_minusone_col))) %>%
      pivot_wider(names_from = year, values_from = coverage) %>%
      ensure_col(local_recent_col) %>%
      ensure_col(local_minusone_col) %>%
      mutate(diff = .data[[local_recent_col]] - .data[[local_minusone_col]])
    hpv1_diff_txt <- make_hpv_diff_txt(hpv1_row)
    hpvc_diff_txt <- hpv %>% pull(diff_txt)
  } else {
    hpv1_diff_txt <- hpv %>% pull(diff_txt)
    hpvc_diff_txt <- NULL
  }
}
```

```{r hpv-line-plot, echo=FALSE, warning=FALSE, message=FALSE}
if (has_hpv_data) {
  plot_hpv1 <- if (has_hpv1)
    hpv_df %>% filter(vaccine_code == "PRHPV1_F", year >= first_hpv_year) %>%
    select(year, coverage) else NULL
  plot_hpvc <- if (has_hpvc)
    hpv_df %>% filter(vaccine_code == "PRHPVC_F", year >= first_hpv_year) %>%
    select(year, coverage) else NULL
  plot_hpv_ref <- if (has_hpvc) plot_hpvc else plot_hpv1
  
  min_hpv_year <- min(c(plot_hpv1$year, plot_hpvc$year))
  max_hpv_year <- max(c(plot_hpv1$year, plot_hpvc$year))
  
  has_hpv_num_unvac <- sum(!is.na(hpv_df$unvaccinated)) > 1
    
  if (has_hpv_num_unvac) {
    plot_zd_hpv <- hpv_df %>%
      filter(vaccine_code == primary_hpv_code, year >= first_hpv_year) %>%
      select(year, unvaccinated)
    hpv_scale_factor <- 100 / max(plot_zd_hpv$unvaccinated, na.rm = TRUE)
    hpv_unvac <- hpv_df %>%
    filter(year == rev_yr) %>%
    select(iso3c, country, vaccine_code, unvaccinated) %>%
    distinct(vaccine_code, .keep_all = TRUE) %>% 
    pivot_wider(names_from = vaccine_code, values_from = unvaccinated)

  # Safe generation if pivot_wider returns an empty dataframe structure
  if (nrow(hpv_unvac) == 0) {
    hpv_unvac <- tibble(PRHPVC_F = NA_real_, PRHPV1_F = NA_real_)
  } else {
    if (!"PRHPVC_F" %in% names(hpv_unvac)) hpv_unvac$PRHPVC_F <- NA_real_
    if (!"PRHPV1_F" %in% names(hpv_unvac)) hpv_unvac$PRHPV1_F <- NA_real_
  }

  hpv_unvac <- hpv_unvac %>%
    mutate(diff = PRHPVC_F - PRHPV1_F) %>%
    label_vals_millions(PRHPVC_F, "PRHPVC_F_lbl") %>%
    label_vals_millions(PRHPV1_F, "PRHPV1_F_lbl") %>%
    label_vals_millions(diff,     "diff_lbl")
  }
  
  title_key <- case_when(
    has_hpvc & has_hpv_num_unvac  ~ "txt_plt_hpv_both_unvac",
    has_hpvc & !has_hpv_num_unvac ~ "txt_plt_hpv_both_only",
    !has_hpvc & has_hpv_num_unvac ~ "txt_plt_hpv1_unvac",
    TRUE                          ~ "txt_plt_hpv1_only"
  )
  
  hpv_plot <- ggplot() +
    { if (has_hpv_num_unvac && nrow(plot_zd_hpv) > 1)
      geom_col(data = plot_zd_hpv,
               aes(x = year, y = unvaccinated,
                   fill = t_lookup("txt_plt_unvac_undervac_fem", language)),
               alpha = 0.4, width = 0.6) } +
    { if (has_hpv1 && nrow(plot_hpv1) > 1) list(
      geom_line(data = plot_hpv1,
                aes(x = year, y = coverage / hpv_scale_factor,
                    color = t_lookup("txt_plt_hpv1_legend_cvg", language)),
                size = 1.5),
      geom_point(data = plot_hpv1,
                 aes(x = year, y = coverage / hpv_scale_factor,
                     color = t_lookup("txt_plt_hpv1_legend_cvg", language)),
                 size = 3))
      else if (has_hpv1 && nrow(plot_hpv1) == 1)
        geom_point(data = plot_hpv1,
                   aes(x = year, y = coverage / hpv_scale_factor,
                       color = t_lookup("txt_plt_hpv1_legend_cvg", language)),
                   size = 3) } +
    { if (has_hpvc && nrow(plot_hpvc) > 1) list(
      geom_line(data = plot_hpvc,
                aes(x = year, y = coverage / hpv_scale_factor,
                    color = t_lookup("txt_plt_hpvc_legend_cvg", language)),
                size = 1.5),
      geom_point(data = plot_hpvc,
                 aes(x = year, y = coverage / hpv_scale_factor,
                     color = t_lookup("txt_plt_hpvc_legend_cvg", language)),
                 size = 3))
      else if (has_hpvc && nrow(plot_hpvc) == 1)
        geom_point(data = plot_hpvc,
                   aes(x = year, y = coverage / hpv_scale_factor,
                       color = t_lookup("txt_plt_hpvc_legend_cvg", language)),
                   size = 3) } +
    geom_hline(yintercept = 90 / hpv_scale_factor, linetype = "dashed") +
    scale_color_manual(name   = NULL,
                       values = setNames(
                         c("#1CABE2", "#6A1E72"),
                         c(t_lookup("txt_plt_hpv1_legend_cvg", language),
                           t_lookup("txt_plt_hpvc_legend_cvg", language)))) +
    scale_fill_manual(name   = NULL,
                      values = setNames("#E2231A",
                                        t_lookup("txt_plt_unvac_undervac_fem", language))) +
    scale_x_continuous(breaks = seq(min_hpv_year, max_hpv_year, 1)) +
    { if (has_hpv_num_unvac)
      scale_y_continuous(
        name   = t_lookup("txt_plt_un_under_vac_females", language),
        labels = scales::label_comma(),
        sec.axis = sec_axis(
          ~ . * hpv_scale_factor,
          name   = t_lookup("txt_plt_cvg_pct", language),
          breaks = seq(0, 100, 25)
        )
      )
      else
        scale_y_continuous(limits = c(0, 100),
                           name   = t_lookup("txt_plt_cvg_pct", language)) } +
    { if (has_hpv_num_unvac)
      guides(color = guide_legend(order = 1), fill = guide_legend(order = 2)) } +
    labs(title = str_glue(t_lookup(title_key, language), ctry_name = ctry_name, first_hpv_year = first_hpv_year, hpv_rev_yr = hpv_rev_yr),
         caption = paste0(
           t_lookup("txt_plt_ia2030_caption", language), "\n",
           if (!has_hpvc)         paste0(get_text2("txt_hpv_no_hpvc_caption",  plot_vars), "\n") else "",
           if (!has_hpv_num_unvac) paste0(get_text2("txt_hpv_no_unvac_caption", plot_vars), "\n") else "",
           str_glue(get_text2("txt_wuenic_source", plot_vars))),
         x = t_lookup("txt_plt_year", language),
         color = NULL) +
    theme_unicef() +
    theme(panel.grid.minor   = element_blank(),
          panel.grid.major.x = element_line(),
          axis.text.x        = element_text(angle = 45, hjust = 1, size = 9),
          plot.title         = element_text(size = 14, hjust = 0.5, face = "bold"),
          legend.position    = "bottom")
}
```

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{`r t_lookup("txt_sec_hpv", language)`}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:hpv}{}")
cat(paste0("\\addcontentsline{toc}{section}{", t_lookup("txt_sec_hpv", language), "}"))
```

```{r hpv-bullets, results='asis', echo=FALSE}
# helper: safely pull a single value, returning NA if no matching row/column exists
safe_pull_hpv <- function(data, code, col) {
  val <- data %>% filter(vaccine_code == code) %>% pull({{ col }})
  if (length(val) == 0) NA_real_ else val
}

# helper: ensure a column exists in hpv_years_wide before selecting it;
ensure_col <- function(data, col_name) {
  if (!col_name %in% names(data)) {
    data[[col_name]] <- NA_real_
  }
  data
}

if (hpv_intro_flag == 0) {
  cat(paste0("- ", str_glue(get_text2("txt_hpv_not_introduced", text_vars)), "\n\n"))
} else if (hpv_intro_flag == 1 & !has_hpv_data) {
  cat(paste0("- ", str_glue(get_text2("txt_hpv_intro_no_data", text_vars)), "\n\n"))
} else {

  # Guard against missing year-columns using our local structural variables
  hpv_years_wide <- hpv_years_wide %>%
    ensure_col(local_minusone_col) %>%
    ensure_col(local_recent_col)

  if (has_hpvc) {

    hpv_metrics <- hpv_years_wide %>%
      select(vaccine_code, all_of(local_minusone_col), all_of(local_recent_col)) %>%
      rename(comp_val = !!sym(local_minusone_col), rev_val = !!sym(local_recent_col))

    hpv1_comp_cvg     <- safe_pull_hpv(hpv_metrics, "PRHPV1_F", comp_val)
    hpv1_rev_cvg      <- safe_pull_hpv(hpv_metrics, "PRHPV1_F", rev_val)
    hpvc_comp_cvg     <- safe_pull_hpv(hpv_metrics, "PRHPVC_F", comp_val)
    hpvc_rev_cvg      <- safe_pull_hpv(hpv_metrics, "PRHPVC_F", rev_val)

    if (is.na(hpv1_comp_cvg) | is.na(hpvc_comp_cvg)) {
      target_key <- "txt_hpv_single_year_both"
    } else {
      hpv1_diff <- hpv1_rev_cvg - hpv1_comp_cvg
      hpvc_diff <- hpvc_rev_cvg - hpvc_comp_cvg

      target_key <- case_when(
        hpv1_diff == 0 & hpvc_diff == 0 ~ "txt_hpv_both_constant",
        hpv1_diff == 0 & hpvc_diff > 0  ~ "txt_hpv_hpv1_const_hpvc_inc",
        hpv1_diff == 0 & hpvc_diff < 0  ~ "txt_hpv_hpv1_const_hpvc_dec",
        hpv1_diff > 0  & hpvc_diff == 0 ~ "txt_hpv_hpv1_inc_hpvc_const",
        hpv1_diff > 0  & hpvc_diff > 0  ~ "txt_hpv_hpv1_inc_hpvc_inc",
        hpv1_diff > 0  & hpvc_diff < 0  ~ "txt_hpv_hpv1_inc_hpvc_dec",
        hpv1_diff < 0  & hpvc_diff == 0 ~ "txt_hpv_hpv1_dec_hpvc_const",
        hpv1_diff < 0  & hpvc_diff > 0  ~ "txt_hpv_hpv1_dec_hpvc_inc",
        hpv1_diff < 0  & hpvc_diff < 0  ~ "txt_hpv_hpv1_dec_hpvc_dec"
      )
    }
    hpv_cov_sentence <- get_text2(target_key, text_vars)
    bullet_text <- str_glue(
      hpv_cov_sentence,
      comp_yr       = local_minusone_col, # Maps to 2019 baseline
      rev_yr        = local_rev_yr,       # Maps to 2024 baseline
      hpv1_rev_cvg  = hpv1_rev_cvg,
      hpvc_rev_cvg  = hpvc_rev_cvg
    )

    cat("- ", bullet_text, "\n\n", sep = "")
    cat("- ", hpv %>% pull(gt90), "\n\n", sep = "")

  } else {
    hpv_metrics <- hpv_years_wide %>%
      select(vaccine_code, all_of(local_minusone_col), all_of(local_recent_col)) %>%
      rename(comp_val = !!sym(local_minusone_col), rev_val = !!sym(local_recent_col))

    hpv1_comp_cvg     <- safe_pull_hpv(hpv_metrics, "PRHPV1_F", comp_val)
    hpv1_rev_cvg      <- safe_pull_hpv(hpv_metrics, "PRHPV1_F", rev_val)

    if (is.na(hpv1_comp_cvg)) {
      bullet_text <- str_glue(get_text2("txt_hpv_single_year_hpv1", text_vars))
    } else {
      hpv1_diff <- hpv1_rev_cvg - hpv1_comp_cvg
      hpv1_diff_txt <- case_when(
        hpv1_diff > 0  ~ str_glue(get_text2("txt_hpv_diff_inc", text_vars)),
        hpv1_diff < 0  ~ str_glue(get_text2("txt_hpv_diff_dec", text_vars)),
        TRUE           ~ str_glue(get_text2("txt_hpv_diff_const", text_vars))
      )
      bullet_text <- str_glue(get_text2("txt_hpv_cvg_hpv1_only", text_vars), hpv1_diff_txt = hpv1_diff_txt)
    }

    # Format the root historical change bullet point
    bullet_text <- str_glue(bullet_text, comp_yr = local_minusone_col, rev_yr = local_rev_yr, hpv1_rev_cvg = hpv1_rev_cvg)
    cat("- ", bullet_text, "\n\n", sep = "")

    cat("- ", hpv %>% pull(gt90), "\n\n", sep = "")

    if (exists("has_hpvc_comparison") && !has_hpvc_comparison) {
      cat("- ", str_glue(get_text2("txt_hpv_no_hpvc_one_yr", text_vars)), "\n\n", sep = "")
    } else {
      cat("- ", str_glue(get_text2("txt_hpv_no_hpvc_data",   text_vars)), "\n\n", sep = "")
    }
  }

  # ── 1. CHECK UNVACCINATED DATA AVAILABILITY ──────────────────────────────────
hpv_unvac_years_avail <- hpv_df %>% filter(!is.na(unvaccinated)) %>% pull(year) %>% unique()
has_hpv_unvac_data   <- length(hpv_unvac_years_avail) > 0
hpv_unvac_is_rev_yr  <- rev_yr %in% hpv_unvac_years_avail

# Use rev_yr if available, otherwise fall back to most recent year with data
hpv_unvac_year <- if (has_hpv_unvac_data) {
  if (hpv_unvac_is_rev_yr) rev_yr else max(hpv_unvac_years_avail)
} else {
  NA_real_
}

# data from target year
hpv_unvac <- hpv_df %>%
  filter(year == hpv_unvac_year) %>%
  select(iso3c, country, vaccine_code, unvaccinated) %>%
  distinct(vaccine_code, .keep_all = TRUE) %>%
  pivot_wider(names_from = vaccine_code, values_from = unvaccinated)

# see if data is empty
if (nrow(hpv_unvac) == 0) {
  hpv_unvac <- tibble(PRHPVC_F = NA_real_, PRHPV1_F = NA_real_)
} else {
  if (!"PRHPVC_F" %in% names(hpv_unvac)) hpv_unvac$PRHPVC_F <- NA_real_
  if (!"PRHPV1_F" %in% names(hpv_unvac)) hpv_unvac$PRHPV1_F <- NA_real_
}

hpv_unvac <- hpv_unvac %>%
  mutate(diff = PRHPVC_F - PRHPV1_F) %>%
  label_vals_millions(PRHPVC_F, "PRHPVC_F_lbl") %>%
  label_vals_millions(PRHPV1_F, "PRHPV1_F_lbl") %>%
  label_vals_millions(diff,     "diff_lbl")

# lookups
hpv1_unvac_lbl <- if (has_hpv_unvac_data) hpv_unvac$PRHPV1_F_lbl[[1]] else NA_character_
hpv_rev_yr     <- rev_yr

# build bullets
hpv_unvac_bullet <- if (!has_hpv_unvac_data) {
  str_glue(get_text2("txt_hpv_no_unvac_data", text_vars))
} else if (hpv_unvac_is_rev_yr) {
  str_glue(get_text2("txt_hpv_unvac", text_vars))
} else {
  str_glue(get_text2("txt_hpv_unvac_diff_fallback_yr", text_vars))
}

if (!is.na(hpv_unvac_bullet)) {
  cat("- ", hpv_unvac_bullet, "\n\n", sep = "")
}
}

if (exists("reg_rank_hpv") && nrow(reg_rank_hpv) > 0) {
  cat("- ", str_glue(get_text2("txt_hpv_reg_rank", text_vars),
    rank              = reg_rank_hpv$rank,
    n_peers_hpv       = n_peers_hpv,
    primary_hpv_label = primary_hpv_label,
    rev_yr            = local_rev_yr), "\n\n", sep = "")
}
invisible(NULL)
```

```{r hpv, echo=FALSE, warning=FALSE, fig.width=10, fig.height=6.2}
if (exists("hpv_plot")) print(hpv_plot)
```

<!-- ===================== PAGE 6: REGIONAL COMPARISON ===================== -->

```{r setup-regional-comparison, include=FALSE, warning=FALSE, message=FALSE}
reg_comp_vax <- reg_peers %>%
  filter(vaccine %in% c("dtp1", "dtp3", "mcv1"), year == rev_yr) %>%
  select(iso3c, country, vaccine, coverage)

reg_comp_hpv <- hpv_peers %>%
  filter(vaccine_code == "PRHPVC_F", year == hpv_rev_yr) %>%
  select(iso3c, country, coverage) %>%
  mutate(vaccine = "hpvc")

if (has_hpv_data && has_hpvc_for_rev_yr) {
  reg_comp <- bind_rows(reg_comp_vax, reg_comp_hpv) %>%
    mutate(highlight_type = ifelse(iso3c == ctry_iso, "focal", "peer"),
           vaccine        = toupper(vaccine),
           vaccine        = factor(vaccine, levels = c("DTP1","DTP3","MCV1","HPVC")))
} else {
  reg_comp <- reg_comp_vax %>%
    mutate(highlight_type = ifelse(iso3c == ctry_iso, "focal", "peer"),
           vaccine        = toupper(vaccine),
           vaccine        = factor(vaccine, levels = c("DTP1","DTP3","MCV1")))
}

reg_avg <- reg_comp %>%
  group_by(vaccine) %>%
  summarise(avg = round(mean(coverage, na.rm = TRUE), 0))

hpvc_suffix <- if (has_hpv_data && has_hpvc_for_rev_yr)
  paste0(", ", get_text2("txt_reg_comp_hpvc_suffix", text_vars)) else ""

# translated legend text
legend_labels <- c("focal" = ctry_name, "peer" = t_lookup("txt_plt_reg_comp_peers", language))

reg_comp <- translate_vaccine_names(reg_comp)
reg_avg <- translate_vaccine_names(reg_avg)

plt_reg_comp <- reg_comp %>%
  mutate(country_label = if_else(
    highlight_type == "focal",
    paste0("<span style='color:#E2231A;'><b>", country, "</b></span>"),
    country
  )) %>%
  mutate(country_label = reorder_within(country_label, coverage, vaccine)) %>%
  ggplot(aes(x = coverage, y = country_label, colour = highlight_type, size = highlight_type)) +
  geom_point() +
  geom_vline(data = reg_avg, aes(xintercept = avg),
             linetype = "dashed", colour = "grey50", linewidth = 1.5) +
  facet_wrap(~vaccine, ncol = 2, scales = "free_y") +
  scale_y_reordered() +
  scale_colour_manual(
    values = c("focal" = "#E2231A", "peer" = "#A8D5E2"),
    labels = legend_labels,
    name = ""
  ) +
  scale_size_manual(
    values = c("focal" = 4, "peer" = 2.5),
    labels = legend_labels,
    name = ""
  ) +
  labs(x        = t_lookup("txt_plt_cvg_pct", language),
       y        = "",
       title    = str_glue(get_text2("txt_plt_reg_comp_title", plot_vars)),
       subtitle = t_lookup("txt_plt_reg_comp_subtitle", language),
       caption  = str_glue(get_text2("txt_wuenic_source", plot_vars))) +
  theme_unicef() +
  theme(legend.position  = "bottom",
        strip.background   = element_rect(fill = "#0058AB", color = NA),
        strip.text.x        = element_text(color = "white"),
        strip.text       = element_text(face = "bold"),
        panel.grid.minor = element_blank(),
        axis.text.y      = ggtext::element_markdown(size = 9))
```

```{r calc-height, echo=FALSE}
n_countries    <- length(unique(reg_comp$country))
dynamic_height <- max(4, min(2.5 + (n_countries * 0.28), 12))

# regional comparison text string (see if has HPVC or not)
region_comparison_string <- ifelse(has_hpvc, str_glue(get_text2("txt_reg_comp_intro_has_hpvc", text_vars)), str_glue(get_text2("txt_reg_comp_intro_no_hpvc", text_vars)))
```

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{`r t_lookup("txt_sec_reg_comp", language)`}}
\end{tcolorbox}

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:regcomp}{}")
cat(paste0("\\addcontentsline{toc}{section}{",
           t_lookup("txt_sec_reg_comp", language), "}"))
```

\vspace{1em}

`r region_comparison_string`

\vspace{1em}

```{r reg-comp-hpv-note, echo=FALSE, results='asis'}
if (has_hpv_data) {
  if (has_hpvc && !has_hpvc_for_rev_yr) {
    cat(str_glue(get_text2("txt_reg_comp_no_hpvc_yr", text_vars)))
    cat("\\vspace{0.6em}\n")
  } else if (!has_hpvc) {
    cat(str_glue(get_text2("txt_reg_comp_no_hpvc", text_vars)))
    cat("\\vspace{0.6em}\n")
  }
}
```

```{r ggplot-reg-comp, echo=FALSE, fig.width=10, fig.height=dynamic_height}
print(plt_reg_comp)
```

\newpage

<!-- ===================== PAGE 7: ADDITIONAL VACCINES ===================== -->

```{r setup-additional-vax, include=FALSE}
add_vax <- c("bcg","hib3","hepb3","ipv1","pcvc","rotac")

ipv1_translated <- ifelse(language %in% c("fr", "es", "pt"), "VPI1", "IPV1")

add_vax_df <- df %>%
  filter(
    vaccine %in% add_vax,
    year %in% c(comp_yr, rev_yr, minusone_col)
  ) %>%
  select(iso3c, country, vaccine, year, coverage) %>%
  pivot_wider(
    names_from = year,
    values_from = coverage
  ) %>%
  mutate(
    vaccine_label = case_when(
      vaccine == "bcg"   ~ "BCG",
      vaccine == "hib3"  ~ "Hib3",
      vaccine == "hepb3" ~ "HepB3",
      vaccine == "ipv1"  ~ ipv1_translated,
      vaccine == "pcvc"  ~ "PCVC",
      vaccine == "rotac" ~ "RotaC",
      TRUE ~ toupper(vaccine)
    ),
    diff    = .data[[recent_col]] - .data[[comparison_col]],
    diff_txt = case_when(
      diff > 0  ~ str_glue(get_text2("txt_add_vax_increased", text_vars), comp_val = .data[[comparison_col]], rev_val = .data[[recent_col]]),
      diff < 0  ~ str_glue(get_text2("txt_add_vax_decreased", text_vars), comp_val = .data[[comparison_col]], rev_val = .data[[recent_col]]),
      diff == 0 ~ str_glue(get_text2("txt_add_vax_same", text_vars), comp_val = .data[[comparison_col]], rev_val = .data[[recent_col]]),
      TRUE      ~ NA_character_
    )
  )

add_vax_heatmap_df <- df %>%
  filter(vaccine %in% add_vax, year >= 2000) %>%
  mutate(vaccine_label = case_when(
    vaccine == "bcg"   ~ "BCG",   vaccine == "hib3"  ~ "Hib3",
    vaccine == "hepb3" ~ "HepB3", vaccine == "ipv1"  ~ ipv1_translated,
    vaccine == "pcvc"  ~ "PCVC",  vaccine == "rotac" ~ "RotaC",
    TRUE ~ toupper(vaccine)))

plt_add_vax <- add_vax_heatmap_df %>%
  ggplot(aes(x = as.factor(year), y = vaccine_label, fill = coverage)) +
  geom_tile(color = "white") +
  geom_text(aes(label = coverage), colour = "white", size = 3) +
  labs(x       = "",
       y       = "",
       title   = str_glue(get_text2("txt_plt_add_vax_title", plot_vars)),
       caption = str_glue(get_text2("txt_wuenic_source", plot_vars))) +
  scale_x_discrete(position = "top") +
  scale_y_discrete(limits = rev) +
  theme_minimal() +
  theme(panel.grid       = element_blank(),
        legend.position  = "bottom",
        legend.key.width = unit(1.2, "cm"),
        axis.text.x      = element_text(size = 9),
        plot.title       = element_text(size = 14, hjust = 0.5, face = "bold")) +
  scale_fill_gradientn(
    colours = c('#B50800','#E2231A','#F26A21','#FFC20E','#80BD41','#00833D'),
    values = scales::rescale(c(0, 59, 69, 79, 88, 90, 100)),
    guide = "colorbar",
    limits = c(0, 100),
    name    = t_lookup("txt_plt_cvg_pct_legend", language),
    breaks = c(0, 60, 70, 80, 90, 100),
    labels = c("0", "60", "70", "80", "90", "100"))
```

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{`r t_lookup("txt_sec_add_vax", language)`}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:addvax}{}")
cat(paste0("\\addcontentsline{toc}{section}{",
           t_lookup("txt_sec_add_vax", language), "}"))
```

``` {r additional-vax-bullets, results='asis', echo=FALSE}
# helper: classify direction of change between two coverage values
classify_dir <- function(new_val, old_val, same_label = "same") {
  case_when(
    new_val > old_val ~ "inc",
    new_val < old_val ~ "dec",
    TRUE               ~ same_label
  )
}

for (i in seq_len(nrow(add_vax_df))) {
  row <- add_vax_df[i, ] %>% 
    rename(comp_val = !!sym(comparison_col), rev_val = !!sym(recent_col))
  
  # ensure minusone column exists and safely pull value
  minusone_val <- if (as.character(minusone_col) %in% names(row)) row[[as.character(minusone_col)]] else NA
  
  has_comp     <- !is.na(row$comp_val)
  has_minusone <- !is.na(minusone_val)
  has_rev      <- !is.na(row$rev_val)
  
  # --- CASE 1: core recent year data is missing completely -> print "no data" and skip ---
  if (!has_rev) {
    raw_sentence <- get_text2("txt_add_vax_no_data", text_vars)
    bullet_text <- str_glue(raw_sentence,
                             vaccine_label = coalesce(as.character(row$vaccine_label), "Vaccine"),
                             rev_yr        = coalesce(as.character(rev_yr), "NA"))
    cat("- ", bullet_text, "\n\n")
    next
  }
  
  # --- CASE 2: neither comp_yr nor minusone available -> single year only ---
  if (!has_comp && !has_minusone) {
    raw_sentence <- get_text2("txt_add_vax_single_year", text_vars)
    
  # --- CASE 3: only minusone available (comp_yr missing) -> compare vs minusone only ---
  } else if (!has_comp && has_minusone) {
    dir_minusone <- classify_dir(row$rev_val, minusone_val, same_label = "sam")
    target_key <- paste0("txt_add_vax_vsminusone_", dir_minusone)
    raw_sentence <- get_text2(target_key, text_vars)
    
  # --- CASE 4: only comp_yr available (minusone missing) -> compare vs comp_yr only ---
  } else if (has_comp && !has_minusone) {
    dir_comp <- classify_dir(row$rev_val, row$comp_val, same_label = "sam")
    target_key <- paste0("txt_add_vax_vscomp_", dir_comp)
    raw_sentence <- get_text2(target_key, text_vars)
    
  # --- CASE 5: both comp_yr and minusone available -> full three-way comparison ---
  } else {
    dir_comp <- case_when(
      row$rev_val > row$comp_val ~ "inc",
      row$rev_val < row$comp_val ~ "dec",
      TRUE                       ~ "sam"
    )
    dir_minusone <- case_when(
      row$rev_val > minusone_val ~ "inc",
      row$rev_val < minusone_val ~ "dec",
      TRUE                       ~ "same"
    )
    target_key <- paste0("txt_add_vax_", dir_comp, "_", dir_minusone)
    raw_sentence <- get_text2(target_key, text_vars)
  }
  
  # safe string glue substitution (coerces any floating NA fields cleanly to text)
  bullet_text <- str_glue(
    raw_sentence,
    vaccine_label = coalesce(as.character(row$vaccine_label), "Vaccine"),
    rev_cvg       = coalesce(as.character(row$rev_val), "NA"),
    rev_yr        = coalesce(as.character(rev_yr), "NA"),
    comp_yr       = coalesce(as.character(comp_yr), "NA"),
    comp_cvg      = coalesce(as.character(row$comp_val), "NA"),
    minusone_col  = coalesce(as.character(minusone_col), "NA"),
    minusone_cvg  = coalesce(as.character(minusone_val), "NA")
  )
  
  cat("- ", bullet_text, "\n\n")
}
```


```{r ggplot-add-vax, echo=FALSE, fig.width=10, fig.height=4}
print(plt_add_vax)
```

<!-- ===================== APPENDIX ===================== -->
```{r appendix-fontsize, include=FALSE}
appendix_fontsize <- switch(language,
  "ar" = "\\fontsize{12}{15}\\selectfont",
  "fr" = "\\fontsize{10}{13}\\selectfont",
  "es" = "\\fontsize{10}{13}\\selectfont",
  "pt" = "\\fontsize{10}{13}\\selectfont",
       "\\fontsize{11}{14}\\selectfont"
)
```
\newpage
```{r, echo=FALSE, results='asis'}
cat(appendix_fontsize)
```

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{`r t_lookup("txt_sec_appendix", language)`}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:overview}{}")
cat(paste0("\\addcontentsline{toc}{section}{", t_lookup("txt_sec_appendix", language), "}"))
```

\vspace{0.8em}

\textcolor{cyan}{\large `r t_lookup("txt_appendix_wuenic_header", language)`}

\vspace{0.8em}

`r t_lookup("txt_appendix_wuenic_body", language)`

\vspace{0.3em}
`r t_lookup("txt_appendix_methods_header", language)`:

\begin{itemize}
  \item \begin{sloppypar}
        \textcolor{blue}{\href{https://pubmed.ncbi.nlm.nih.gov/19649368/}{Burton et al. 2009. WHO and UNICEF estimates of national infant immunization coverage: methods and processes}}
        \end{sloppypar}
  \item \begin{sloppypar}
        \textcolor{blue}{\href{https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0047806}{Burton et al. 2012. A formal representation of the WHO and UNICEF estimates of national immunization coverage: a computational logic approach}}
        \end{sloppypar}
  \item \begin{sloppypar}
        \textcolor{blue}{\href{https://openpublichealthjournal.com/VOLUME/6/PAGE/73/ABSTRACT/}{Brown et al. 2013. An introduction to the grade of confidence used to characterize uncertainty around the WHO and UNICEF estimates of national immunization coverage}}
        \end{sloppypar}
\end{itemize}

\vspace{0.3em}

`r str_glue(get_text2("txt_appendix_wuenic_revision", text_vars))`

\vspace{1.3em}

\textcolor{cyan}{\fontsize{14}{16}\selectfont `r t_lookup("txt_appendix_definitions_header", language)`}

\vspace{0.3em}

```{r appendix-definitions, results='asis', echo=FALSE}
# definitions are stored as a named list in txt_vars_talking_points.R
# and retrieved here so they translate properly
defs <- list(
  t_lookup("txt_def_bcg",   language),
  t_lookup("txt_def_hepbb", language),
  t_lookup("txt_def_dtp",   language),
  t_lookup("txt_def_hepb3", language),
  t_lookup("txt_def_hib3",  language),
  t_lookup("txt_def_ipv",   language),
  t_lookup("txt_def_mcv",   language),
  t_lookup("txt_def_rotac", language),
  t_lookup("txt_def_pcv3",  language),
  t_lookup("txt_def_yfv",   language),
  t_lookup("txt_def_menga", language),
  t_lookup("txt_def_hpv",   language))

for (d in defs) {
  cat(paste0("-   ", d, "\n"))
  cat("\\vspace{0.5em}\n")
}
```

\vspace{1.3em}
\textcolor{cyan}{\fontsize{14}{16}\selectfont `r t_lookup("txt_appendix_resources_header", language)`} \textcolor{blue}{\href{https://data.unicef.org/resources/immunization/}{\underline{https://data.unicef.org/resources/immunization/}}}