---
title: ""
output:
  pdf_document:
    latex_engine: xelatex
    toc: false
    toc_depth: 3
geometry: "left=1cm,right=1cm,top=1.1cm,bottom=1.1cm"
fontsize: 12pt
params:
  country: "Default"
  language: "en"
header-includes:
  - \usepackage{enumitem}
  - \geometry{a4paper, left=15mm, right=15mm, top=15mm, bottom=20mm}
  - \setlist[itemize,1]{label=\textbullet}
  - \setlist[itemize,2]{label=\(\circ\)}
  - \setlist[itemize,3]{label=\(\infty\)}
  - \setmainfont{Arial}
  - \usepackage{hyperref}
  - \hypersetup{colorlinks=true, linkcolor=blue, urlcolor=blue, breaklinks=true}
  - \usepackage{tcolorbox}
  - \usepackage{fancyhdr}
  - \setlength{\headheight}{15pt}
  - \addtolength{\topmargin}{-3pt}
  - \definecolor{unicefblue}{RGB}{28, 171, 226}
  - \pagestyle{fancy}
  - \fancyhf{}
  - \fancyhead[L]{}
  - \fancyhead[C]{}
  - \fancyhead[R]{\hyperlink{top}{\small \color{black} $\uparrow$ Back to top}}
  - \fancyfoot[C]{\thepage}
  - \usepackage{tocloft}
  - \renewcommand{\cftsecfont}{\normalfont}
  - \renewcommand{\cftsecpagefont}{\normalfont}
  - \renewcommand{\cftsecleader}{\cftdotfill{\cftdotsep}}
---

\hypertarget{top}{}

```{r setup, include=FALSE}
options(scipen = 999)

# ── TRANSLATION SETUP ─────────────────────────────────────────────────────────
source("txt_vars_talking_points.R")        # loads text_vars_en
source(file.path(utils, "R/narrative-translation_deepl.R"))    # loads text_vars (translated)

current_country <- current_country
language <- params$language

# ── COUNTRY DATA ──────────────────────────────────────────────────────────────
df <- wuenic_dta %>% filter(country == current_country) %>%
  filter(!vaccine == "pol3")
hpv_df <- hpv_dta %>% filter(country == current_country)
has_hpv_data <- nrow(hpv_df) > 0
has_hpvc     <- has_hpv_data && any(hpv_df$vaccine_code == "PRHPVC_F")

add_region_names <- function(df) {
  df %>%
    mutate(Region_Name = case_match(Region,
      "ROSA"          ~ "South Asia",
      "ESAR"          ~ "Eastern and Southern Africa",
      "ECAR"          ~ "Europe and Central Asia",
      "MENA"          ~ "Middle East and North Africa",
      "LACR"          ~ "Latin America and the Caribbean",
      "WCAR"          ~ "West and Central Africa",
      "EAPR"          ~ "East Asia and the Pacific",
      "Non-programme" ~ "Non-programme countries",
      .default        = Region))
}
df     <- add_region_names(df)
hpv_df <- add_region_names(hpv_df)

ctry_iso  <- unique(df$iso3c)
region    <- unique(df$Region)
ctry_regn <- df %>%
  distinct(Region_Name, Region) %>%
  mutate(formatted = str_glue("{Region_Name} ({Region})")) %>%
  pull(formatted)

reg_peers <- wuenic_dta %>% filter(Region == region, lvl_1 == "country")
hpv_peers <- hpv_dta    %>% filter(Region == region, lvl_1 == "country")

available_vaccines <- unique(df$vaccine)

sub_df <- wuenic_dta %>% filter(iso3c == ctry_iso)

wuenic_src <- str_glue("Source: WHO/UNICEF Estimates of National Immunization Coverage, {rev_yr} revision")

the_countries <- c(
  "United Arab Emirates", "Bahamas", "Central African Republic", "DRC", "Congo",
  "Dominican Republic", "Gambia", "Maldives", "Marshall Islands",
  "Netherlands (Kingdom of the)", "Philippines", "Republic of Korea",
  "Republic of Moldova", "Russian Federation", "Solomon Islands",
  "State of Palestine", "United Kingdom", "United States")

df <- df %>%
  mutate(the_country = ifelse(country %in% the_countries, paste("the", country), country))

ctry_name     <- unique(df$country)
the_ctry_name <- unique(df$the_country)

hpv_intro <- wiise_hpv_intro_yrs %>%
  filter(iso3c == ctry_iso) %>%
  mutate(introduced = case_when(nationwide == "yes" | partially == "yes" ~ 1, TRUE ~ 0))

hpv_intro_flag <- ifelse(nrow(hpv_intro) > 0 && any(hpv_intro$introduced == 1, na.rm = TRUE), 1, 0)

if (!require("countrycode", quietly = TRUE)) install.packages("countrycode")
library(countrycode)
iso2 <- tolower(countrycode(ctry_name, origin = "country.name", destination = "iso2c"))
if (ctry_name == "Micronesia") iso2 <- "fm"

flag_path <- paste0("flags/", iso2, ".png")
if (!file.exists(flag_path)) {
  dir.create("flags", showWarnings = FALSE)
  download.file(paste0("https://flagcdn.com/160x120/", iso2, ".png"),
                destfile = flag_path, mode = "wb")
}
```


```{r banner, echo=FALSE, results='asis'}
cat(paste0('
\\begin{tcolorbox}[
  colback=unicefblue,
  colframe=unicefblue,
  arc=6pt,
  boxrule=0pt,
  left=4mm,
  right=4mm,
  top=3mm,
  bottom=3mm,
  halign=center,
  valign=center
]
\\begin{center}
  {\\LARGE \\color{white} \\textbf{', get_text2("txt_sec_banner_title", text_vars), ': ', t_lookup(x, language), '}} \\hspace{0.2cm}\\includegraphics[height=0.6cm]{', flag_path, '}\\\\[0.4em]
  {\\large \\color{white} ', get_text2("txt_sec_banner_revision", text_vars), ': ', rev_yr, '}
\\end{center}
\\end{tcolorbox}
\\vspace{0.4in}
'))
```

```{r toc, echo=FALSE, results='asis'}
cat("\\tableofcontents")
```


```{r country-map, echo=FALSE, warning=FALSE, message=FALSE}
region_for_country <- base_map_df %>%
  filter(iso3c == ctry_iso) %>%
  pull(map_region)

map_df <- base_map_df %>% filter(map_region == region_for_country)

if (!is.na(region_for_country) && region_for_country == "Oceania") {
  map_df <- st_shift_longitude(map_df)
}

map_df <- map_df %>%
  mutate(highlight = ifelse(iso3c == ctry_iso,
                            get_text2("txt_map_country", text_vars),
                            get_text2("txt_map_region",  text_vars)))

if (!is.na(region_for_country)) {
  limits <- switch(region_for_country,
    "Africa"   = list(x = c(-24, 56),   y = c(-40, 46)),
    "Americas" = list(x = c(-170, -23), y = c(-60, 85)),
    "Asia"     = list(x = c(25, 150),   y = c(-15, 60)),
    "Europe"   = list(x = c(-25, 100),  y = c(34, 75)),
    "Oceania"  = list(x = c(110, 240),  y = c(-50, 5)),
    list(x = c(-180, 180), y = c(-90, 90)))
  xlim_val <- limits$x
  ylim_val <- limits$y
}

if (!is.na(region_for_country) && region_for_country == "Oceania") {
  pacific_crs <- st_crs("+proj=longlat +lon_0=180")
  map_df      <- st_transform(map_df, pacific_crs)
  xlim_val    <- c(-70, 15)
  ylim_val    <- c(-50, 3)
}

if (current_country == "Russian Federation") {
  xlim_val <- c(-25, 200)
  ylim_val <- c(34, 85)
}

target_country_geo <- map_df %>% filter(highlight == get_text2("txt_map_country", text_vars))
is_small_country   <- as.numeric(st_area(target_country_geo)) < 5e9
if (is_small_country) country_centroid <- st_centroid(target_country_geo)

country_map <- ggplot(map_df) +
  geom_sf(aes(fill = highlight), color = "white", size = 0.2) +
  { if (is_small_country)
      geom_sf(data = country_centroid, inherit.aes = FALSE,
              size = 6, shape = 21, color = "#1CABE2", stroke = 1.5) } +
  scale_fill_manual(
    values = setNames(c("#1CABE2", "#d7d7d7"),
                      c(get_text2("txt_map_country", text_vars),
                        get_text2("txt_map_region",  text_vars))),
    name = NULL) +
  coord_sf(xlim = xlim_val, ylim = ylim_val, expand = FALSE) +
  theme_void() +
  theme(legend.position = "none",
        plot.title  = element_markdown(hjust = 0.5, size = 14, face = "bold"),
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.5)) +
  labs(title = str_glue(
    get_text2("txt_map_title_full", text_vars),
    ctry_name          = ctry_name,
    region_for_country = region_for_country
  ))
```

\vspace{2em}
```{r ggplot-country-map, echo=FALSE, warning=FALSE, fig.align='center'}
suppressMessages(print(country_map))
```

```{r setup-vax, include=FALSE}
vax <- df %>%
  filter(vaccine %in% c("dtp1", "dtp3", "mcv1", "mcv2"),
         year %in% c(comp_yr, rev_yr - 1, rev_yr))

unvac_lbl <- vax %>%
  select(Region, iso3c, country, year, vaccine, unvaccinated_lbl) %>%
  pivot_wider(names_from = year, values_from = unvaccinated_lbl) %>%
  rename_with(~paste0(., "_lbl"), .cols = (ncol(.) - 2):ncol(.))

unvac <- vax %>%
  select(Region, iso3c, country, year, vaccine, unvaccinated) %>%
  pivot_wider(names_from = year, values_from = unvaccinated) %>%
  left_join(., unvac_lbl) %>%
  distinct() %>%
  mutate(
    pchng     = round((.[[7]] / .[[5]] - 1) * 100, 0),
    pchng_txt = case_when(
      .[[8]] == .[[10]] ~ str_glue(get_text2("txt_unvac_stable", text_vars)),
      pchng < 0         ~ str_glue(get_text2("txt_unvac_declined", text_vars)),
      pchng > 0         ~ str_glue(get_text2("txt_unvac_increased", text_vars)),
      TRUE              ~ NA_character_))

vacc <- vax %>%
  select(Region, iso3c, country, year, vaccine, vaccinated) %>%
  pivot_wider(names_from = year, values_from = vaccinated) %>%
  mutate(
    pchng     = round((.[[7]] / .[[5]] - 1) * 100, 1),
    pchng_txt = case_when(
      pchng > 0 ~ get_text2("txt_vacc_increased", text_vars),
      pchng < 0 ~ get_text2("txt_vacc_decreased", text_vars),
      TRUE      ~ get_text2("txt_vacc_same", text_vars))) %>%
  label_vals_millions(.[[5]], "vacc_comp_yr_lbl") %>%
  label_vals_millions(.[[7]], "vacc_rev_yr_lbl")

cvg <- vax %>%
  select(Region, iso3c, country, year, vaccine, coverage) %>%
  pivot_wider(names_from = year, values_from = coverage)

tgt <- vax %>%
  filter(vaccine == "dtp1") %>%
  select(Region, iso3c, country, year, vaccine, target) %>%
  pivot_wider(names_from = year, values_from = target) %>%
  mutate(
    diff     = .[[7]] - .[[5]],
    moreless = case_when(
      diff < 0 ~ get_text2("txt_moreless_fewer", text_vars),
      diff > 0 ~ get_text2("txt_moreless_more",  text_vars),
      TRUE ~ get_text2("txt_moreless_same", text_vars)),
    to_vax = case_when(
      diff < 0 ~ get_text2("txt_surv_inf_vax_3", text_vars),
      diff > 0 & diff < 1000000 ~ get_text2("txt_surv_inf_vax_1", text_vars),
      diff > 0 & diff >= 1000000 ~ get_text2("txt_surv_inf_vax_2", text_vars))) %>%
  label_vals(diff, "diff_lbl") %>%
  mutate(diff_abs = abs(diff)) %>%
  label_vals_millions(diff_abs, "diff_abs_lbl")
```

```{r setup-overview, include=FALSE}
gt90 <- df %>%
  select(-target_grp) %>%
  distinct() %>%
  filter(year == rev_yr) %>%
  mutate(gt90 = ifelse(coverage >= 90, 1, 0))

n_vax_total <- n_distinct(gt90$vaccine)
n_vax_gt90  <- sum(gt90$gt90, na.rm = TRUE)

cvg_chng <- vax %>%
  filter(year %in% c(rev_yr, rev_yr - 1)) %>%
  select(iso3c, country, vaccine, year, coverage) %>%
  pivot_wider(names_from = year, values_from = coverage) %>%
  mutate(
    diff     = .[[5]] - .[[4]],
    diff_txt = case_when(
      diff > 0 ~ str_glue(get_text2("txt_cvg_increased_short", text_vars),
                          cvg_prev = .[[4]], cvg_rev = .[[5]]),
      diff < 0 ~ str_glue(get_text2("txt_cvg_decreased_short", text_vars),
                          cvg_prev = .[[4]], cvg_rev = .[[5]]),
      TRUE     ~ str_glue(get_text2("txt_cvg_same_short",      text_vars),
                          cvg_rev = .[[5]])))

# pre-build overview bullets
bullet_ovw_1 <- if (n_vax_gt90 == 0) {
  str_glue(get_text2("txt_ovw_vax_none_gt90", text_vars))
} else {
  str_glue(get_text2("txt_ovw_vax_gt90", text_vars))
}

bullet_ovw_2 <- str_glue(
  get_text2("txt_ovw_dtp1_cvg", text_vars),
  cvg_chng_dtp1 = cvg_chng %>% filter(vaccine == "dtp1") %>% pull(diff_txt),
  cvg_chng_dtp3 = cvg_chng %>% filter(vaccine == "dtp3") %>% pull(diff_txt),
  cvg_chng_mcv1 = cvg_chng %>% filter(vaccine == "mcv1") %>% pull(diff_txt))
```

```{r setup-progress-ia2030-zd-target, include=FALSE}
zd <- df %>%
  filter(vaccine == "dtp1", year >= 2000) %>%
  select(iso3c, country, year, unvaccinated)

zd_2019            <- zd %>% filter(year == 2019) %>% pull(unvaccinated)
ia2030_annual_target <- zd_2019 / 2

target <- zd %>%
  filter(year == 2019) %>%
  mutate(target_2030 = unvaccinated / 2,
         factor      = (unvaccinated - target_2030) / (2030 - 2019)) %>%
  crossing(year_new = 2020:2030) %>%
  mutate(unvaccinated_new = unvaccinated - factor * (year_new - 2019)) %>%
  select(iso3c, country, year = year_new, unvaccinated_new)

zd_rev_yr         <- zd     %>% filter(year == rev_yr) %>% pull(unvaccinated)
ia2030_rev_yr     <- target %>% filter(year == rev_yr) %>% pull(unvaccinated_new)
zd_pcnt_diff      <- round((zd_rev_yr / ia2030_rev_yr - 1) * 100, 0)
zd_rev_yr_lbl     <- label_vals_millions_num(zd_rev_yr)
ia2030_rev_yr_lbl <- label_vals_millions_num(ia2030_rev_yr)

bullet_ovw_3 <- str_glue(
  get_text2("txt_ovw_zd_ia2030", text_vars),
  abs_zd_pcnt_diff = abs(zd_pcnt_diff),
  zd_diff_txt      = get_text2(case_when(
    zd_pcnt_diff > 0 ~ "txt_zd_higher_than",
    zd_pcnt_diff < 0 ~ "txt_zd_lower_than",
    TRUE             ~ "txt_zd_same_as"), text_vars))
```

<!-- ===================== PAGE 1: OVERVIEW ===================== -->

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{`r get_text2("txt_sec_overview", text_vars)`}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:overview}{}")
cat(paste0("\\addcontentsline{toc}{section}{", get_text2("txt_sec_overview", text_vars), "}"))
```

- `r bullet_ovw_1`

- `r bullet_ovw_2`

- `r bullet_ovw_3`


```{r setup-heatmap-overview, include=FALSE}
if (has_hpvc) {
  hpv_merge <- hpv_df %>%
    filter(vaccine_code == "PRHPVC_F") %>%
    select(-vaccine) %>%
    rename(vaccine = vaccine_code)
  shared_cols <- intersect(names(df), names(hpv_merge))
  df <- bind_rows(df %>% select(all_of(shared_cols)),
                  hpv_merge %>% select(all_of(shared_cols)))
}

heatmap_ovw_df <- df %>%
  mutate(
    year    = as.character(year),
    vaccine = case_when(
      vaccine == "hepbb"    ~ "HepBB", vaccine == "hepb3"    ~ "HepB3",
      vaccine == "hib3"     ~ "Hib3",  vaccine == "rotac"    ~ "RotaC",
      vaccine == "menga"    ~ "MengA", vaccine == "PRHPVC_F" ~ "HPVc",
      TRUE ~ toupper(vaccine)),
    cvg_cat = case_when(
      coverage < 60              ~ "<60",
      coverage %in% 60:69        ~ "60-69",
      coverage %in% 70:79        ~ "70-79",
      coverage %in% 80:89        ~ "80-89",
      coverage %in% 90:94        ~ "90-94",
      coverage >= 95             ~ ">=95",
      TRUE                       ~ NA_character_))

heatmap_ovw_df$vaccine <- factor(heatmap_ovw_df$vaccine,
  levels = c("BCG","HepBB","DTP1","DTP3","Hib3","HepB3",
             "PCVC","RotaC","IPV1","IPVC","MCV1","RCV1",
             "MCV2","YFV","MengA","HPVc"))

heatmap_overview <- heatmap_ovw_df %>%
  ggplot(aes(x = year, y = vaccine, fill = coverage)) +
  geom_tile(color = "white") +
  geom_text(aes(label = coverage), colour = "white", vjust = 0.5, size = 3) +
  labs(x = "", y = "",
       title   = str_glue(get_text2("txt_plt_ovw_heatmap_title", text_vars)),
       caption = wuenic_src) +
  scale_x_discrete(position = "top") +
  scale_y_discrete(limits = rev) +
  theme_minimal() +
  theme(panel.grid       = element_blank(),
        legend.position  = "bottom",
        legend.key.width = unit(1.2, "cm"),
        plot.title       = element_text(size = 14, hjust = 0.5, face = "bold")) +
  scale_fill_gradientn(
    colours = c("#B50800","#E2231A","#F26A21","#FFC20E","#80BD41","#00833D"),
    values  = scales::rescale(c(0, 59, 69, 79, 89, 94, 100)),
    guide   = "colorbar", limits = c(0, 100),
    name    = get_text2("txt_plt_cvg_pct_legend", text_vars),
    breaks  = c(0, 59, 69, 79, 89, 100),
    labels  = c("0", "60", "70", "80", "90", "100"))
```

```{r overview-line-plot, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center'}
all_vax_plot <- heatmap_ovw_df %>%
  mutate(year = as.numeric(year)) %>%
  filter(!is.na(coverage))

vaccine_colors <- c(
  "BCG"   = "#1CABE2", "HepBB" = "#0058AB", "DTP1"  = "#00833D",
  "DTP3"  = "#80BD41", "Hib3"  = "#6A1E74", "HepB3" = "#961A49",
  "PCVC"  = "#E2231A", "RotaC" = "#F26A21", "IPV1"  = "#B50800",
  "IPVC"  = "#FF8C00", "MCV1"  = "#2E7EBB", "RCV1"  = "#9E9E9E",
  "MCV2"  = "#FFB6C1", "YFV"   = "#5B4A9C", "MengA" = "#C8A900",
  "HPVc"  = "#D45F9E")

first_points <- all_vax_plot %>%
  group_by(vaccine) %>%
  filter(year == min(year)) %>%
  ungroup()

num_vaccines <- n_distinct(all_vax_plot$vaccine)
min_coverage <- min(all_vax_plot$coverage, na.rm = TRUE)

all_vax_line_plot <- ggplot(all_vax_plot,
    aes(x = year, y = coverage, color = vaccine, group = vaccine)) +
  geom_line(size = 0.8, alpha = 0.7) +
  geom_point(size = 1.5, alpha = 0.7) +
  geom_hline(yintercept = 90, linetype = "dashed") +
  geom_text(data = first_points,
            aes(label = vaccine, color = vaccine),
            hjust = 1.2, size = 2.5, show.legend = FALSE) +
  scale_color_manual(values = vaccine_colors) +
  scale_x_continuous(breaks    = seq(2000, rev_yr, 1),
                     expand    = expansion(add = c(2, 0.5))) +
  scale_y_continuous(limits = c(min_coverage - 5, 100),
                     breaks = seq(0, 100, by = 10),
                     name   = get_text2("txt_plt_cvg_pct", text_vars)) +
  labs(title   = str_glue(get_text2("txt_plt_ovw_line_title", text_vars)),
       caption = paste0(get_text2("txt_plt_ia2030_caption", text_vars),
                        "\n", wuenic_src),
       x       = get_text2("txt_plt_year", text_vars),
       color   = NULL) +
  theme_unicef() +
  theme(panel.grid.minor  = element_blank(),
        panel.grid.major.x = element_line(),
        axis.text.x       = element_text(angle = 45, hjust = 1, size = 9),
        plot.title        = element_text(size = 14, hjust = 0.5, face = "bold"),
        legend.position   = "right",
        legend.text       = element_text(size = 8),
        legend.key.height = unit(0.6, "cm")) +
  guides(color = guide_legend(nrow = num_vaccines))
```

```{r overview-plot, echo=FALSE, fig.width=10, fig.height=5.5}
print(heatmap_overview)
```

```{r all_vax_plot, echo=FALSE, fig.width=10, fig.height=5.5}
print(all_vax_line_plot)
```

\newpage

<!-- ===================== PAGE 2: DTP1 ===================== -->

```{r setup-dtp1, include=FALSE}
v       <- "dtp1"
unvac_v <- unvac %>% filter(vaccine == v)
cvg_v   <- cvg   %>% filter(vaccine == v)
vacc_v  <- vacc  %>% filter(vaccine == v)

ctry_cvg_rev  <- cvg_v %>% pull(7)
ctry_cvg_comp <- cvg_v %>% pull(5)

reg_rank <- reg_peers %>%
  filter(vaccine == v, year == rev_yr) %>%
  arrange(desc(coverage)) %>%
  mutate(rank = row_number()) %>%
  filter(iso3c == ctry_iso)

n_peers <- reg_peers %>% filter(vaccine == v, year == rev_yr) %>% n_distinct("iso3c")

dropout_dtp <- df %>%
  filter(year %in% c(comp_yr, rev_yr), vaccine %in% c("dtp1", "dtp3")) %>%
  select(iso3c, country, year, vaccine, vaccinated) %>%
  pivot_wider(names_from = vaccine, values_from = vaccinated) %>%
  mutate(
    dropout     = round(((dtp1 - dtp3) / dtp1) * 100, 0),
    dropout_txt = case_when(
      dropout < 5  ~ str_glue(get_text2("txt_dropout_low", text_vars), vax_pair = "DTP"),
      dropout <= 10 ~ str_glue(get_text2("txt_dropout_medium", text_vars), vax_pair = "DTP"),
      dropout > 10 ~ str_glue(get_text2("txt_dropout_high", text_vars), vax_pair = "DTP")
    )
  )

dropout_comp_dtp <- dropout_dtp %>%
  select(-c(dtp1, dtp3, dropout_txt)) %>%
  pivot_wider(names_from = year, values_from = dropout) %>%
  mutate(
    diff     = .[[4]] - .[[3]],
    diff_txt = case_when(
      diff == 1  ~ str_glue(get_text2("txt_dropout_diff_higher_1", text_vars)),
      diff > 0   ~ str_glue(get_text2("txt_dropout_diff_higher_pp", text_vars)),
      diff == -1 ~ str_glue(get_text2("txt_dropout_diff_lower_1", text_vars)),
      diff < -1  ~ str_glue(get_text2("txt_dropout_diff_lower_pp", text_vars)),
      TRUE       ~ str_glue(get_text2("txt_dropout_diff_same", text_vars))))

# pre-build DTP1 bullets
dtp1_dropout_row <- dropout_dtp %>% filter(year == rev_yr)

bullet_dtp1_1 <- str_glue(get_text2("txt_dtp1_cvg_change", text_vars),
  ctry_cvg_rev  = ctry_cvg_rev,
  ctry_cvg_comp = ctry_cvg_comp)

bullet_dtp1_2 <- str_glue(get_text2("txt_dtp1_zd_ia2030", text_vars),
  abs_zd_pcnt_diff = abs(zd_pcnt_diff),
  zd_diff_txt      = get_text2(case_when(
    zd_pcnt_diff > 0 ~ "txt_zd_higher_than",
    zd_pcnt_diff < 0 ~ "txt_zd_lower_than",
    TRUE             ~ "txt_zd_same_as"), text_vars))

bullet_dtp1_3 <- str_glue(get_text2("txt_dtp1_surv_infants", text_vars),
  diff_abs_lbl = tgt$diff_abs_lbl,
  moreless     = tgt$moreless,
  to_vax       = tgt$to_vax)

bullet_dtp1_4 <- str_glue(get_text2("txt_dtp1_reg_rank", text_vars),
  rank         = reg_rank$rank,
  n_peers      = n_peers,
  ctry_cvg_rev = ctry_cvg_rev)

bullet_dtp1_5 <- str_glue(get_text2("txt_dtp1_dropout", text_vars),
  dropout_pct      = dtp1_dropout_row$dropout,
  dropout_diff_txt = dropout_comp_dtp %>% pull(diff_txt),
  dropout_txt      = dtp1_dropout_row$dropout_txt)
```

```{r dtp1-line-plot, echo=FALSE, warning=FALSE, message=FALSE}
plot_dtp1  <- df %>% filter(vaccine == "dtp1") %>% select(year, coverage)
plot_zd_dtp <- df %>% filter(vaccine == "dtp1") %>% select(year, unvaccinated)

scale_factor <- max(plot_dtp1$coverage, na.rm = TRUE) /
                max(plot_zd_dtp$unvaccinated, na.rm = TRUE)

dtp_plot <- ggplot() +
  geom_col(data = plot_zd_dtp,
           aes(x = year, y = unvaccinated * scale_factor,
               fill = get_text2("txt_plt_dtp1_legend_zd", text_vars)),
           alpha = 0.4, width = 0.6) +
  geom_line(data  = plot_dtp1,
            aes(x = year, y = coverage,
                color = get_text2("txt_plt_dtp1_legend_cvg", text_vars)),
            size = 0.8) +
  geom_point(data = plot_dtp1,
             aes(x = year, y = coverage,
                 color = get_text2("txt_plt_dtp1_legend_cvg", text_vars)),
             size = 1.5) +
  geom_hline(yintercept = 90, linetype = "dashed") +
  scale_color_manual(name   = NULL,
                     values = setNames("#1CABE2",
                       get_text2("txt_plt_dtp1_legend_cvg", text_vars))) +
  scale_fill_manual(name   = NULL,
                    values = setNames("#E2231A",
                      get_text2("txt_plt_dtp1_legend_zd", text_vars))) +
  scale_x_continuous(breaks = seq(min(plot_dtp1$year), max(plot_dtp1$year), 1)) +
  scale_y_continuous(
    limits = c(0, 100),
    name   = get_text2("txt_plt_cvg_pct", text_vars),
    sec.axis = sec_axis(~ . / scale_factor,
                        name   = get_text2("txt_plt_zd_children", text_vars),
                        labels = scales::label_comma())) +
  labs(title   = str_glue(get_text2("txt_plt_dtp1_title",      text_vars)),
       caption = paste0(get_text2("txt_plt_ia2030_caption",     text_vars),
                        "\n", wuenic_src),
       x       = get_text2("txt_plt_year", text_vars)) +
  guides(color = guide_legend(order = 1), fill = guide_legend(order = 2)) +
  theme_unicef() +
  theme(panel.grid.minor   = element_blank(),
        panel.grid.major.x = element_line(),
        axis.text.x        = element_text(angle = 45, hjust = 1, size = 9),
        plot.title         = element_text(size = 14, hjust = 0.5, face = "bold"),
        legend.position    = "bottom")
```

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{`r get_text2("txt_sec_dtp1", text_vars)`}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:dtp1}{}")
cat(paste0("\\addcontentsline{toc}{section}{", get_text2("txt_sec_dtp1", text_vars), "}"))
```

- `r bullet_dtp1_1`

- `r bullet_dtp1_2`

- `r bullet_dtp1_3`

- `r bullet_dtp1_4`

- `r bullet_dtp1_5`

\vspace{1em}

```{r dtp1, echo=FALSE, fig.width=10, fig.height=6.2}
print(dtp_plot)
```

<!-- ===================== PAGE 3: DTP3 ===================== -->

```{r setup-dtp3, include=FALSE}
v       <- "dtp3"
unvac_v <- unvac %>% filter(vaccine == v)
cvg_v   <- cvg   %>% filter(vaccine == v)
vacc_v  <- vacc  %>% filter(vaccine == v)

ctry_cvg_rev_dtp3  <- cvg_v %>% pull(7)
ctry_cvg_comp_dtp3 <- cvg_v %>% pull(5)

reg_rank_dtp3 <- reg_peers %>%
  filter(vaccine == v, year == rev_yr) %>%
  arrange(desc(coverage)) %>%
  mutate(rank = row_number()) %>%
  filter(iso3c == ctry_iso)

unvac_dtp3_rev <- unvac_v %>% pull(7)
unvac_dtp3_lbl <- label_vals_millions_num(unvac_dtp3_rev)

tgt_chng <- tgt %>%
  mutate(pop_grwth = round((.[[7]] / .[[5]] - 1) * 100, 0)) %>%
  label_vals_millions(.[[7]], "tgt_rev_yr") %>%
  label_vals_millions(.[[5]], "tgt_comp_yr")

pop_chng_txt <- if (ctry_cvg_rev_dtp3 < ctry_cvg_comp_dtp3 & tgt[[7]] > tgt[[5]]) {
  str_glue(get_text2("txt_pop_growth_context", text_vars),
           pop_grwth = tgt_chng$pop_grwth)
} else { "" }

# pre-build DTP3 bullets
dtp3_dropout_row <- dropout_dtp %>% filter(year == rev_yr)

bullet_dtp3_1 <- str_glue(get_text2("txt_dtp3_vaccinated", text_vars),
  vacc_pchng_txt   = vacc_v %>% pull(pchng_txt),
  vacc_comp_yr_lbl = vacc_v %>% pull(vacc_comp_yr_lbl),
  vacc_rev_yr_lbl  = vacc_v %>% pull(vacc_rev_yr_lbl),
  pop_chng_txt     = pop_chng_txt)

bullet_dtp3_2 <- str_glue(get_text2("txt_dtp3_dropout", text_vars),
  dropout_pct      = dtp3_dropout_row$dropout,
  dropout_diff_txt = dropout_comp_dtp %>% pull(diff_txt),
  dropout_txt      = dtp3_dropout_row$dropout_txt)

bullet_dtp3_3 <- str_glue(get_text2("txt_dtp3_unvac", text_vars),
  ctry_cvg_rev_dtp3 = ctry_cvg_rev_dtp3,
  unvac_dtp3_lbl    = unvac_dtp3_lbl)

bullet_dtp3_4 <- str_glue(get_text2("txt_dtp3_reg_rank", text_vars),
  rank    = reg_rank_dtp3$rank,
  n_peers = n_peers)
```

```{r dtp3-line-plot, echo=FALSE, warning=FALSE, message=FALSE}
plot_dtp3   <- df %>% filter(vaccine == "dtp3") %>% select(year, coverage)
plot_zd_dtp3 <- df %>% filter(vaccine == "dtp3") %>% select(year, unvaccinated)

scale_factor <- max(plot_dtp3$coverage, na.rm = TRUE) /
                max(plot_zd_dtp3$unvaccinated, na.rm = TRUE)

dtp3_plot <- ggplot() +
  geom_col(data = plot_zd_dtp,
           aes(x = year, y = unvaccinated * scale_factor,
               fill = get_text2("txt_plt_dtp3_legend_unvac", text_vars)),
           alpha = 0.4, width = 0.6) +
  geom_line(data  = plot_dtp3,
            aes(x = year, y = coverage,
                color = get_text2("txt_plt_dtp3_legend_cvg", text_vars)),
            size = 0.8) +
  geom_point(data = plot_dtp3,
             aes(x = year, y = coverage,
                 color = get_text2("txt_plt_dtp3_legend_cvg", text_vars)),
             size = 1.5) +
  geom_hline(yintercept = 90, linetype = "dashed") +
  scale_color_manual(name   = NULL,
                     values = setNames("#1CABE2",
                       get_text2("txt_plt_dtp3_legend_cvg", text_vars))) +
  scale_fill_manual(name   = NULL,
                    values = setNames("#E2231A",
                      get_text2("txt_plt_dtp3_legend_unvac", text_vars))) +
  scale_x_continuous(breaks = seq(min(plot_dtp3$year), max(plot_dtp3$year), 1)) +
  scale_y_continuous(
    limits   = c(0, 100),
    name     = get_text2("txt_plt_cvg_pct", text_vars),
    sec.axis = sec_axis(~ . / scale_factor,
                        name   = get_text2("txt_plt_unvac_undervac", text_vars),
                        labels = scales::label_comma())) +
  labs(title   = str_glue(get_text2("txt_plt_dtp3_title",  text_vars)),
       caption = paste0(get_text2("txt_plt_ia2030_caption", text_vars),
                        "\n", wuenic_src),
       x       = get_text2("txt_plt_year", text_vars),
       color   = NULL) +
  theme_unicef() +
  theme(panel.grid.minor   = element_blank(),
        panel.grid.major.x = element_line(),
        axis.text.x        = element_text(angle = 45, hjust = 1, size = 9),
        plot.title         = element_text(size = 14, hjust = 0.5, face = "bold"),
        legend.position    = "bottom")
```

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{`r get_text2("txt_sec_dtp3", text_vars)`}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:dtp3}{}")
cat(paste0("\\addcontentsline{toc}{section}{", get_text2("txt_sec_dtp3", text_vars), "}"))
```

- `r bullet_dtp3_1`

- `r bullet_dtp3_2`

- `r bullet_dtp3_3`

- `r bullet_dtp3_4`

\vspace{1em}

```{r dtp3, echo=FALSE, fig.width=10, fig.height=6.2}
print(dtp3_plot)
```

<!-- ===================== PAGE 4: MCV ===================== -->

```{r setup-mcv, include=FALSE}
v_mcv   <- if ("mcv2" %in% available_vaccines) c("mcv1", "mcv2") else "mcv1"
unvac_v <- unvac %>% filter(vaccine %in% v_mcv)
cvg_v   <- cvg   %>% filter(vaccine %in% v_mcv)

cvg_mcv1 <- cvg_v %>%
  filter(vaccine == "mcv1") %>%
  mutate(
    diff         = .[[7]] - .[[5]],
    cvg_chng_txt = case_when(
      diff %in% c(-1:1) ~ str_glue(get_text2("txt_cvg_stable", text_vars)),
      diff < -1         ~ str_glue(get_text2("txt_cvg_decreased", text_vars),
                                   cvg_comp = .[[5]], cvg_rev = .[[7]]),
      diff > 1          ~ str_glue(get_text2("txt_cvg_increased", text_vars),
                                   cvg_comp = .[[5]], cvg_rev = .[[7]])))

if ("mcv2" %in% available_vaccines) {
  cvg_mcv2 <- cvg_v %>%
    filter(vaccine == "mcv2") %>%
    mutate(
      diff         = .[[7]] - .[[5]],
      cvg_chng_txt = case_when(
        diff %in% c(-1:1) ~ str_glue(get_text2("txt_cvg_stable",    text_vars)),
        diff < -1         ~ str_glue(get_text2("txt_cvg_decreased",  text_vars),
                                     cvg_comp = .[[5]], cvg_rev = .[[7]]),
        diff > 1          ~ str_glue(get_text2("txt_cvg_increased",  text_vars),
                                     cvg_comp = .[[5]], cvg_rev = .[[7]])))

  unvac_mcv <- unvac_v %>%
    select(iso3c, country, vaccine, 7) %>%
    pivot_wider(names_from = vaccine, values_from = last_col()) %>%
    mutate(diff = mcv2 - mcv1) %>%
    label_vals_millions(mcv1, "mcv1_lbl") %>%
    label_vals_millions(mcv2, "mcv2_lbl") %>%
    label_vals_millions(diff, "diff_lbl")
} else {
  cvg_mcv2  <- cvg_v %>%
    mutate(cvg_chng_txt = str_glue(
      get_text2("txt_mcv2_no_data", text_vars)))
  unvac_mcv <- NULL
}

reg_rank_mcv1 <- reg_peers %>%
  filter(vaccine == "mcv1", year == rev_yr) %>%
  arrange(desc(coverage)) %>%
  mutate(rank = row_number()) %>%
  filter(iso3c == ctry_iso)

dropout_mcv <- df %>%
  filter(year %in% c(comp_yr, rev_yr), vaccine %in% c("dtp1", "mcv1")) %>%
  select(iso3c, country, year, vaccine, vaccinated) %>%
  pivot_wider(names_from = vaccine, values_from = vaccinated) %>%
  mutate(
    dropout     = round(((dtp1 - mcv1) / dtp1) * 100, 0),
    dropout_txt = case_when(
      dropout < 5   ~ get_text2("txt_dropout_low_mcv",    text_vars),
      dropout <= 10 ~ get_text2("txt_dropout_medium_mcv", text_vars),
      dropout > 10  ~ get_text2("txt_dropout_high_mcv",   text_vars)))

dropout_comp_mcv <- dropout_mcv %>%
  select(-c(dtp1, mcv1, dropout_txt)) %>%
  pivot_wider(names_from = year, values_from = dropout) %>%
  mutate(
    diff     = .[[4]] - .[[3]],
    diff_txt = case_when(
      diff == 1  ~ str_glue(get_text2("txt_dropout_diff_higher_1",  text_vars)),
      diff > 0   ~ str_glue(get_text2("txt_dropout_diff_higher_pp", text_vars)),
      diff == -1 ~ str_glue(get_text2("txt_dropout_diff_lower_1",   text_vars)),
      diff < -1  ~ str_glue(get_text2("txt_dropout_diff_lower_pp",  text_vars)),
      TRUE       ~ str_glue(get_text2("txt_dropout_diff_same",      text_vars))))

nyrs     <- 3; thrshold <- 80
cvg_trend_mcv1 <- df %>%
  filter(vaccine == "mcv1", year %in% seq(rev_yr - nyrs + 1, rev_yr)) %>%
  mutate(ltx = ifelse(coverage < thrshold, 1, 0)) %>%
  summarise(sum = sum(ltx), mean_cvg = mean(coverage, na.rm = TRUE))

capital_the_name <- if_else(str_starts(the_ctry_name, "the "),
                             str_replace(the_ctry_name, "^the ", "The "),
                             the_ctry_name)

persistent_low_txt <- if (cvg_trend_mcv1$sum == nyrs) {
  str_glue(get_text2("txt_mcv_persistent_low_yes", text_vars),
           capital_the_name = capital_the_name,
           nyrs = nyrs, thrshold = thrshold)
} else {
  str_glue(get_text2("txt_mcv_persistent_low_no",  text_vars),
           capital_the_name = capital_the_name,
           nyrs = nyrs, thrshold = thrshold)
}

# pre-build MCV bullets
mcv_dropout_row <- dropout_mcv %>% filter(year == rev_yr)

bullet_mcv_1 <- str_glue(get_text2("txt_mcv1_cvg_change", text_vars),
  mcv1_cvg_rev = cvg_mcv1 %>% pull(7),
  cvg_chng_txt = cvg_mcv1 %>% pull(cvg_chng_txt))

bullet_mcv_2 <- str_glue(get_text2("txt_mcv2_cvg_change", text_vars),
  cvg_chng_txt  = cvg_mcv2 %>% pull(cvg_chng_txt),
  mcv1_unvac_lbl = unvac_mcv$mcv1_lbl,
  mcv_diff_lbl   = unvac_mcv$diff_lbl)

bullet_mcv_3 <- str_glue(get_text2("txt_mcv_dropout", text_vars),
  dropout_pct      = mcv_dropout_row$dropout,
  dropout_diff_txt = dropout_comp_mcv %>% pull(diff_txt),
  dropout_txt      = mcv_dropout_row$dropout_txt)

bullet_mcv_4 <- persistent_low_txt

bullet_mcv_5 <- str_glue(get_text2("txt_mcv_reg_rank", text_vars),
  rank    = reg_rank_mcv1$rank,
  n_peers = n_peers)
```

```{r mcv-line-plot, echo=FALSE, warning=FALSE, message=FALSE}
plot_mcv1  <- df %>% filter(vaccine == "mcv1") %>% select(year, coverage)
if ("mcv2" %in% available_vaccines)
  plot_mcv2 <- df %>% filter(vaccine == "mcv2") %>% select(year, coverage)
plot_zd_mcv <- df %>% filter(vaccine == "mcv1") %>% select(year, unvaccinated)

scale_factor <- max(plot_mcv1$coverage, na.rm = TRUE) /
                max(plot_zd_mcv$unvaccinated, na.rm = TRUE)

mcv_plot <- ggplot() +
  geom_col(data = plot_zd_mcv,
           aes(x = year, y = unvaccinated * scale_factor,
               fill = get_text2("txt_plt_mcv_legend_zd", text_vars)),
           alpha = 0.4, width = 0.6) +
  geom_line(data  = plot_mcv1,
            aes(x = year, y = coverage,
                color = get_text2("txt_plt_mcv1_legend_cvg", text_vars)),
            size = 0.8) +
  geom_point(data = plot_mcv1,
             aes(x = year, y = coverage,
                 color = get_text2("txt_plt_mcv1_legend_cvg", text_vars)),
             size = 1.5) +
  { if ("mcv2" %in% available_vaccines) list(
      geom_line(data  = plot_mcv2,
                aes(x = year, y = coverage,
                    color = get_text2("txt_plt_mcv2_legend_cvg", text_vars)),
                size = 0.8),
      geom_point(data = plot_mcv2,
                 aes(x = year, y = coverage,
                     color = get_text2("txt_plt_mcv2_legend_cvg", text_vars)),
                 size = 1.5)) } +
  geom_hline(yintercept = 90, linetype = "dashed") +
  scale_color_manual(name   = NULL,
                     values = setNames(
                       c("#1CABE2", "#6A1E74"),
                       c(get_text2("txt_plt_mcv1_legend_cvg", text_vars),
                         get_text2("txt_plt_mcv2_legend_cvg", text_vars)))) +
  scale_fill_manual(name   = NULL,
                    values = setNames("#E2231A",
                      get_text2("txt_plt_mcv_legend_zd", text_vars))) +
  scale_x_continuous(breaks = seq(min(plot_mcv1$year), max(plot_mcv1$year), 1)) +
  scale_y_continuous(
    limits   = c(0, 100),
    name     = get_text2("txt_plt_cvg_pct", text_vars),
    sec.axis = sec_axis(~ . / scale_factor,
                        name   = get_text2("txt_plt_mcv_legend_zd", text_vars),
                        labels = scales::label_comma())) +
  guides(color = guide_legend(order = 1), fill = guide_legend(order = 2)) +
  labs(title   = str_glue(get_text2("txt_plt_mcv_title", text_vars)),
       caption = paste0(get_text2("txt_plt_ia2030_caption", text_vars),
                        "\n", wuenic_src),
       x = get_text2("txt_plt_year", text_vars)) +
  theme_unicef() +
  theme(panel.grid.minor   = element_blank(),
        panel.grid.major.x = element_line(),
        axis.text.x        = element_text(angle = 45, hjust = 1, size = 9),
        plot.title         = element_text(size = 14, hjust = 0.5, face = "bold"),
        legend.position    = "bottom")
```

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{`r get_text2("txt_sec_mcv", text_vars)`}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:mcv}{}")
cat(paste0("\\addcontentsline{toc}{section}{", get_text2("txt_sec_mcv", text_vars), "}"))
```

\vspace{1em}

- `r bullet_mcv_1`

- `r bullet_mcv_2`

- `r bullet_mcv_3`

- `r bullet_mcv_4`

- `r bullet_mcv_5`

\vspace{1em}

```{r mcv, echo=FALSE, fig.width=10, fig.height=6.2}
print(mcv_plot)
```

<!-- ===================== PAGE 5: HPV ===================== -->

```{r setup-hpv, include=FALSE}
if (has_hpv_data) {
  available_hpv <- unique(hpv_df$vaccine_code)
  has_hpvc      <- "PRHPVC_F" %in% available_hpv
  has_hpv1      <- "PRHPV1_F" %in% available_hpv

  hpvc_yrs <- hpv_df %>% filter(vaccine_code == "PRHPVC_F") %>% pull(year)
  hpv1_yrs <- hpv_df %>% filter(vaccine_code == "PRHPV1_F") %>% pull(year)

  has_hpvc_comparison <- has_hpvc && all(c(hpv_rev_yr, hpv_rev_yr - 1) %in% hpvc_yrs)
  has_hpv1_comparison <- has_hpv1 && all(c(hpv_rev_yr, hpv_rev_yr - 1) %in% hpv1_yrs)

  primary_hpv_code  <- if (has_hpvc_comparison) "PRHPVC_F" else "PRHPV1_F"
  primary_hpv_label <- if (has_hpvc_comparison) t_lookup("hpvc", language) else t_lookup("hpv1", language)
  
  hpvc_label <- t_lookup("hpvc", language)

  has_hpvc_for_rev_yr <- has_hpvc && hpv_rev_yr %in% hpvc_yrs
  
  # coverage change strings
  make_hpv_diff_txt <- function(df_row) {
    d <- df_row$diff
    if (d > 0)       str_glue(get_text2("txt_vacc_increased", text_vars),
                               cvg_prev = df_row[[4]], cvg_rev = df_row[[5]])
    else if (d < 0)  str_glue(get_text2("txt_vacc_decreased", text_vars),
                               cvg_prev = df_row[[4]], cvg_rev = df_row[[5]])
    else             str_glue(get_text2("txt_vacc_constant",    text_vars),
                               cvg_rev = df_row[[5]])
  }

  # coverage change strings
  hpv <- hpv_df %>%
    filter(vaccine_code == primary_hpv_code) %>%
    select(iso3c, country, vaccine_code, year, coverage) %>%
    filter(year %in% c(hpv_rev_yr, hpv_rev_yr - 1)) %>%
    pivot_wider(names_from = year, values_from = coverage) %>%
    mutate(
      diff = .[[5]] - .[[4]],
      chng = case_when(
        diff > 0 ~ get_text2("txt_vacc_increased", text_vars), 
        diff < 0 ~ get_text2("txt_vacc_decreased", text_vars), 
        TRUE ~ get_text2("txt_vacc_constant", text_vars)
      ),
      diff_txt = case_when(
        chng %in% c("increased", "decreased") ~ 
          str_glue(get_text2("txt_hpv_changed", text_vars), 
                   chng = chng, prev_val = .[[4]], prev_yr = hpv_rev_yr - 1, cur_val = .[[5]], cur_yr = hpv_rev_yr),
        TRUE ~ 
          str_glue(get_text2("txt_hpv_constant", text_vars), 
                   chng = chng, cur_val = .[[5]], prev_yr = hpv_rev_yr - 1, cur_yr = hpv_rev_yr)
      ),
      
      # Target threshold evaluation text
      gt90 = case_when(
        .[[5]] < 80  & chng == "increased" ~ 
          str_glue(get_text2("txt_hpv_gt90_below80_up", text_vars), primary_hpv_label = primary_hpv_label),
        .[[5]] < 80  & chng != "increased" ~ 
          str_glue(get_text2("txt_hpv_gt90_below80", text_vars), primary_hpv_label = primary_hpv_label),
        .[[5]] %in% 80:89 & chng == "increased" ~ 
          str_glue(get_text2("txt_hpv_gt90_near_up", text_vars), primary_hpv_label = primary_hpv_label),
        .[[5]] %in% 80:89 & chng != "increased" ~ 
          str_glue(get_text2("txt_hpv_gt90_near", text_vars), primary_hpv_label = primary_hpv_label),
        .[[5]] == 90 ~ 
          str_glue(get_text2("txt_hpv_gt90_achieved", text_vars), primary_hpv_label = primary_hpv_label),
        .[[5]] > 90  ~ 
          str_glue(get_text2("txt_hpv_gt90_exceeded", text_vars), primary_hpv_label = primary_hpv_label),
        TRUE ~ NA_character_
      )
    )

  hpv_unvac <- hpv_df %>%
    filter(year == hpv_rev_yr) %>%
    select(iso3c, country, vaccine_code, unvaccinated) %>%
    pivot_wider(names_from = vaccine_code, values_from = unvaccinated)

  if (!"PRHPVC_F" %in% names(hpv_unvac)) hpv_unvac$PRHPVC_F <- NA
  if (!"PRHPV1_F" %in% names(hpv_unvac)) hpv_unvac$PRHPV1_F <- NA

  hpv_unvac <- hpv_unvac %>%
    mutate(diff = PRHPVC_F - PRHPV1_F) %>%
    label_vals_millions(PRHPVC_F, "PRHPVC_F_lbl") %>%
    label_vals_millions(PRHPV1_F, "PRHPV1_F_lbl") %>%
    label_vals_millions(diff,     "diff_lbl")

  has_hpv_num_unvac <- !is.na(hpv_unvac$PRHPV1_F) | !is.na(hpv_unvac$PRHPVC_F)
  first_hpv_year    <- hpv_intro %>% filter(introduced == 1) %>% pull(year) %>% min(na.rm = TRUE)

  reg_rank_hpv <- hpv_peers %>%
    filter(vaccine_code == primary_hpv_code, year == hpv_rev_yr) %>%
    arrange(desc(coverage)) %>%
    mutate(rank = row_number()) %>%
    filter(iso3c == ctry_iso)

  n_peers_hpv <- hpv_peers %>%
    filter(vaccine_code == primary_hpv_code, year == hpv_rev_yr) %>%
    n_distinct("iso3c")

  # also get hpv1 diff_txt for the "both" bullet
  if (has_hpv1 && has_hpvc) {
    hpv1_row <- hpv_df %>%
      filter(vaccine_code == "PRHPV1_F") %>%
      select(iso3c, country, vaccine_code, year, coverage) %>%
      filter(year %in% c(hpv_rev_yr, hpv_rev_yr - 1)) %>%
      pivot_wider(names_from = year, values_from = coverage) %>%
      mutate(diff = .[[5]] - .[[4]])
    hpv1_diff_txt <- make_hpv_diff_txt(hpv1_row)
    hpvc_diff_txt <- hpv %>% pull(diff_txt)
  } else {
    hpv1_diff_txt <- hpv %>% pull(diff_txt)
    hpvc_diff_txt <- NULL
  }
}
```

```{r hpv-line-plot, echo=FALSE, warning=FALSE, message=FALSE}
if (has_hpv_data) {
  plot_hpv1 <- if (has_hpv1)
    hpv_df %>% filter(vaccine_code == "PRHPV1_F", year >= first_hpv_year) %>%
               select(year, coverage) else NULL
  plot_hpvc <- if (has_hpvc)
    hpv_df %>% filter(vaccine_code == "PRHPVC_F", year >= first_hpv_year) %>%
               select(year, coverage) else NULL
  plot_hpv_ref <- if (has_hpvc) plot_hpvc else plot_hpv1

  if (has_hpv_num_unvac) {
    plot_zd_hpv     <- hpv_df %>%
      filter(vaccine_code == primary_hpv_code, year >= first_hpv_year) %>%
      select(year, unvaccinated)
    hpv_scale_factor <- 100 / max(plot_zd_hpv$unvaccinated, na.rm = TRUE)
  }
  
  title_key <- case_when(
    has_hpvc & has_hpv_num_unvac  ~ "txt_plt_hpv_both_unvac",
    has_hpvc & !has_hpv_num_unvac ~ "txt_plt_hpv_both_only",
    !has_hpvc & has_hpv_num_unvac ~ "txt_plt_hpv1_unvac",
    TRUE                          ~ "txt_plt_hpv1_only"
  )

  hpv_plot <- ggplot() +
    { if (has_hpv1 && nrow(plot_hpv1) > 1) list(
        geom_line(data = plot_hpv1,
                  aes(x = year, y = coverage,
                      color = get_text2("txt_plt_hpv1_legend_cvg", text_vars)),
                  size = 0.8),
        geom_point(data = plot_hpv1,
                   aes(x = year, y = coverage,
                       color = get_text2("txt_plt_hpv1_legend_cvg", text_vars)),
                   size = 1.5))
      else if (has_hpv1 && nrow(plot_hpv1) == 1)
        geom_point(data = plot_hpv1,
                   aes(x = year, y = coverage,
                       color = get_text2("txt_plt_hpv1_legend_cvg", text_vars)),
                   size = 1.5) } +
    { if (has_hpvc && nrow(plot_hpvc) > 1) list(
        geom_line(data = plot_hpvc,
                  aes(x = year, y = coverage,
                      color = get_text2("txt_plt_hpvc_legend_cvg", text_vars)),
                  size = 0.8),
        geom_point(data = plot_hpvc,
                   aes(x = year, y = coverage,
                       color = get_text2("txt_plt_hpvc_legend_cvg", text_vars)),
                   size = 1.5))
      else if (has_hpvc && nrow(plot_hpvc) == 1)
        geom_point(data = plot_hpvc,
                   aes(x = year, y = coverage,
                       color = get_text2("txt_plt_hpvc_legend_cvg", text_vars)),
                   size = 1.5) } +
    { if (has_hpv_num_unvac && nrow(plot_zd_hpv) > 1)
        geom_col(data = plot_zd_hpv,
                 aes(x = year, y = unvaccinated * hpv_scale_factor,
                     fill = get_text2("txt_plt_hpv_legend_unvac", text_vars)),
                 alpha = 0.4, width = 0.6) } +
    geom_hline(yintercept = 90, linetype = "dashed") +
    scale_color_manual(name   = NULL,
                       values = setNames(
                         c("#1CABE2", "#6A1E72"),
                         c(get_text2("txt_plt_hpv1_legend_cvg", text_vars),
                           get_text2("txt_plt_hpvc_legend_cvg", text_vars)))) +
    scale_fill_manual(name   = NULL,
                      values = setNames("#E2231A",
                        get_text2("txt_plt_hpv_legend_unvac", text_vars))) +
    scale_x_continuous(breaks = seq(min(plot_hpv_ref$year),
                                    max(plot_hpv_ref$year), 1)) +
    { if (has_hpv_num_unvac)
        scale_y_continuous(limits = c(0, 100),
                           name   = get_text2("txt_plt_cvg_pct", text_vars),
                           sec.axis = sec_axis(~ . / hpv_scale_factor,
                             name   = get_text2("txt_plt_unvac_females", text_vars),
                             labels = scales::label_comma()))
      else
        scale_y_continuous(limits = c(0, 100),
                           name   = get_text2("txt_plt_cvg_pct", text_vars)) } +
    { if (has_hpv_num_unvac)
        guides(color = guide_legend(order = 1), fill = guide_legend(order = 2)) } +
    labs(title = str_glue(get_text2(title_key, text_vars), ctry_name = ctry_name, first_hpv_year = first_hpv_year, hpv_rev_yr = hpv_rev_yr),
         caption = paste0(
           get_text2("txt_plt_ia2030_caption", text_vars), "\n",
           if (!has_hpvc)         paste0(get_text2("txt_hpv_no_hpvc_caption",  text_vars), "\n") else "",
           if (!has_hpv_num_unvac) paste0(get_text2("txt_hpv_no_unvac_caption", text_vars), "\n") else "",
           wuenic_src),
         x = get_text2("txt_plt_year", text_vars),
         color = NULL) +
    theme_unicef() +
    theme(panel.grid.minor   = element_blank(),
          panel.grid.major.x = element_line(),
          axis.text.x        = element_text(angle = 45, hjust = 1, size = 9),
          plot.title         = element_text(size = 14, hjust = 0.5, face = "bold"),
          legend.position    = "bottom")
}
```

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{`r get_text2("txt_sec_hpv", text_vars)`}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:hpv}{}")
cat(paste0("\\addcontentsline{toc}{section}{", get_text2("txt_sec_hpv", text_vars), "}"))
```

\vspace{1em}

```{r hpv-bullets, results='asis', echo=FALSE}
if (hpv_intro_flag == 0) {
  cat(paste0("- ", str_glue(get_text2("txt_hpv_not_introduced", text_vars)), "\n"))
  cat("\\vspace{0.6em}\n")
} else if (hpv_intro_flag == 1 & !has_hpv_data) {
  cat(paste0("- ", str_glue(get_text2("txt_hpv_intro_no_data", text_vars)), "\n"))
  cat("\\vspace{0.6em}\n")
} else {
  if (has_hpvc) {
    cat(paste0("- ", str_glue(get_text2("txt_hpv_cvg_both", text_vars),
      hpv1_diff_txt = hpv1_diff_txt,
      hpvc_diff_txt = hpvc_diff_txt), "\n"))
    cat("\\vspace{0.6em}\n")
    cat(paste0("- ", hpv %>% pull(gt90), "\n"))
    cat("\\vspace{0.6em}\n")
  } else {
    cat(paste0("- ", str_glue(get_text2("txt_hpv_cvg_hpv1_only", text_vars),
      hpv1_diff_txt = hpv1_diff_txt), "\n"))
    cat("\\vspace{0.6em}\n")
    cat(paste0("- ", hpv %>% pull(gt90), "\n"))
    cat("\\vspace{0.6em}\n")
    if (exists("has_hpvc_comparison") && !has_hpvc_comparison) {
      cat(paste0("- ", str_glue(get_text2("txt_hpv_no_hpvc_one_yr", text_vars)), "\n"))
    } else {
      cat(paste0("- ", str_glue(get_text2("txt_hpv_no_hpvc_data",   text_vars)), "\n"))
    }
    cat("\\vspace{0.6em}\n")
  }
  if (has_hpv_num_unvac) {
    cat(paste0("- ", str_glue(get_text2("txt_hpv_unvac", text_vars),
      hpv1_unvac_lbl = hpv_unvac$PRHPV1_F_lbl), "\n"))
    cat("\\vspace{0.6em}\n")
  } else {
    cat(paste0("- ", str_glue(get_text2("txt_hpv_no_unvac_data", text_vars)), "\n"))
    cat("\\vspace{0.6em}\n")
  }
}
if (exists("reg_rank_hpv") && nrow(reg_rank_hpv) > 0) {
  cat(paste0("- ", str_glue(get_text2("txt_hpv_reg_rank", text_vars),
    rank            = reg_rank_hpv$rank,
    n_peers_hpv     = n_peers_hpv,
    primary_hpv_label = primary_hpv_label), "\n"))
  cat("\\vspace{0.6em}\n")
}
invisible(NULL)
```

```{r hpv, echo=FALSE, warning=FALSE, fig.width=10, fig.height=6.2}
if (exists("hpv_plot")) print(hpv_plot)
```

<!-- ===================== PAGE 6: REGIONAL COMPARISON ===================== -->

```{r setup-regional-comparison, include=FALSE}
reg_comp_vax <- reg_peers %>%
  filter(vaccine %in% c("dtp1", "dtp3", "mcv1"), year == rev_yr) %>%
  select(iso3c, country, vaccine, coverage)

reg_comp_hpv <- hpv_peers %>%
  filter(vaccine_code == "PRHPVC_F", year == hpv_rev_yr) %>%
  select(iso3c, country, coverage) %>%
  mutate(vaccine = "hpvc")

if (has_hpv_data && has_hpvc_for_rev_yr) {
  reg_comp <- bind_rows(reg_comp_vax, reg_comp_hpv) %>%
    mutate(highlight = ifelse(iso3c == ctry_iso, ctry_name,
                              get_text2("txt_plt_reg_comp_peers", text_vars)),
           vaccine   = toupper(vaccine),
           vaccine   = factor(vaccine, levels = c("DTP1","DTP3","MCV1","HPVC")))
} else {
  reg_comp <- reg_comp_vax %>%
    mutate(highlight = ifelse(iso3c == ctry_iso, ctry_name,
                              get_text2("txt_plt_reg_comp_peers", text_vars)),
           vaccine   = toupper(vaccine),
           vaccine   = factor(vaccine, levels = c("DTP1","DTP3","MCV1")))
}

reg_avg <- reg_comp %>%
  group_by(vaccine) %>%
  summarise(avg = round(mean(coverage, na.rm = TRUE), 0))

hpvc_suffix <- if (has_hpv_data && has_hpvc_for_rev_yr)
  paste0(", ", get_text2("txt_reg_comp_hpvc_suffix", text_vars)) else ""

plt_reg_comp <- reg_comp %>%
  mutate(country = reorder_within(country, coverage, vaccine)) %>%
  ggplot(aes(x = coverage, y = country, colour = highlight, size = highlight)) +
  geom_point() +
  geom_vline(data = reg_avg, aes(xintercept = avg),
             linetype = "dashed", colour = "grey50", linewidth = 0.5) +
  facet_wrap(~vaccine, ncol = 2, scales = "free_y") +
  scale_y_reordered() +
  scale_colour_manual(
    values = setNames(c("#E2231A", "#A8D5E2"),
                      c(ctry_name, get_text2("txt_plt_reg_comp_peers", text_vars))),
    name = "") +
  scale_size_manual(
    values = setNames(c(4, 2.5),
                      c(ctry_name, get_text2("txt_plt_reg_comp_peers", text_vars))),
    name = "") +
  labs(x        = get_text2("txt_plt_cvg_pct", text_vars),
       y        = "",
       title    = str_glue(get_text2("txt_plt_reg_comp_title", text_vars)),
       subtitle = get_text2("txt_plt_reg_comp_subtitle", text_vars),
       caption  = wuenic_src) +
  theme_unicef() +
  theme(legend.position = "bottom",
        strip.text      = element_text(face = "bold"),
        panel.grid.minor = element_blank(),
        axis.text.y     = element_text(size = 9))
```

```{r calc-height, echo=FALSE}
n_countries    <- length(unique(reg_comp$country))
dynamic_height <- max(4, min(2.5 + (n_countries * 0.28), 12))
```

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{`r get_text2("txt_sec_reg_comp", text_vars)`}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:regcomp}{}")
cat(paste0("\\addcontentsline{toc}{section}{",
           get_text2("txt_sec_reg_comp", text_vars), "}"))
```

\vspace{1em}

`r str_glue(get_text2("txt_reg_comp_intro", text_vars), hpvc_suffix = hpvc_suffix)`

\vspace{1em}

```{r reg-comp-hpv-note, echo=FALSE, results='asis'}
if (has_hpv_data) {
  if (has_hpvc && !has_hpvc_for_rev_yr) {
    cat(str_glue(get_text2("txt_reg_comp_no_hpvc_yr", text_vars)))
    cat("\\vspace{0.6em}\n")
  } else if (!has_hpvc) {
    cat(str_glue(get_text2("txt_reg_comp_no_hpvc", text_vars)))
    cat("\\vspace{0.6em}\n")
  }
}
```

```{r ggplot-reg-comp, echo=FALSE, fig.width=10, fig.height=dynamic_height}
print(plt_reg_comp)
```

\newpage

<!-- ===================== PAGE 7: ADDITIONAL VACCINES ===================== -->

```{r setup-additional-vax, include=FALSE}
add_vax <- c("bcg","hib3","hepb3","ipv1","pcvc","rotac")

add_vax_df <- df %>%
  filter(vaccine %in% add_vax, year %in% c(comp_yr, rev_yr)) %>%
  select(iso3c, country, vaccine, year, coverage) %>%
  pivot_wider(names_from = year, values_from = coverage) %>%
  filter(!is.na(.[[4]]) | !is.na(.[[5]])) %>%
  mutate(
    diff      = .[[5]] - .[[4]],
    diff_txt  = case_when(
      diff > 0  ~ str_glue(get_text2("txt_add_vax_increased", text_vars),
                            cvg_comp = .[[4]], cvg_rev = .[[5]]),
      diff < 0  ~ str_glue(get_text2("txt_add_vax_decreased", text_vars),
                            cvg_comp = .[[4]], cvg_rev = .[[5]]),
      diff == 0 ~ str_glue(get_text2("txt_add_vax_same",      text_vars),
                            cvg_comp = .[[4]]),
      TRUE      ~ NA_character_),
    vaccine_label = case_when(
      vaccine == "bcg"   ~ "BCG",   vaccine == "hib3"  ~ "Hib3",
      vaccine == "hepb3" ~ "HepB3", vaccine == "ipv1"  ~ "IPV1",
      vaccine == "pcvc"  ~ "PCV3",  vaccine == "rotac" ~ "RotaC",
      TRUE ~ toupper(vaccine)),
    gt90 = ifelse(.[[5]] >= 90,
                  get_text2("txt_add_vax_above90",    text_vars),
                  get_text2("txt_add_vax_not_above90", text_vars)))

add_vax_heatmap_df <- df %>%
  filter(vaccine %in% add_vax, year >= 2000) %>%
  mutate(vaccine_label = case_when(
    vaccine == "bcg"   ~ "BCG",   vaccine == "hib3"  ~ "Hib3",
    vaccine == "hepb3" ~ "HepB3", vaccine == "ipv1"  ~ "IPV1",
    vaccine == "pcvc"  ~ "PCV3",  vaccine == "rotac" ~ "RotaC",
    TRUE ~ toupper(vaccine)))

plt_add_vax <- add_vax_heatmap_df %>%
  ggplot(aes(x = as.factor(year), y = vaccine_label, fill = coverage)) +
  geom_tile(color = "white") +
  geom_text(aes(label = coverage), colour = "white", size = 3) +
  labs(x       = "",
       y       = "",
       title   = str_glue(get_text2("txt_plt_add_vax_title", text_vars)),
       caption = wuenic_src) +
  scale_x_discrete(position = "top") +
  scale_y_discrete(limits = rev) +
  theme_minimal() +
  theme(panel.grid       = element_blank(),
        legend.position  = "bottom",
        legend.key.width = unit(1.2, "cm"),
        axis.text.x      = element_text(size = 9),
        plot.title       = element_text(size = 14, hjust = 0.5, face = "bold")) +
  scale_fill_gradientn(
    colours = c("#B50800","#E2231A","#F26A21","#FFC20E","#80BD41","#00833D"),
    values  = scales::rescale(c(0, 59, 69, 79, 89, 94, 100)),
    guide   = "colorbar", limits = c(0, 100),
    name    = get_text2("txt_plt_cvg_pct_legend", text_vars),
    breaks  = c(0, 59, 69, 79, 89, 100),
    labels  = c("0", "60", "70", "80", "90", "100"))
```

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{`r get_text2("txt_sec_add_vax", text_vars)`}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:addvax}{}")
cat(paste0("\\addcontentsline{toc}{section}{",
           get_text2("txt_sec_add_vax", text_vars), "}"))
```

\vspace{1em}

```{r additional-vax-bullets, results='asis', echo=FALSE}
for (i in seq_len(nrow(add_vax_df))) {
  row <- add_vax_df[i, ]
  cat(str_glue(get_text2("txt_add_vax_bullet", text_vars),
               vaccine_label = row$vaccine_label,
               diff_txt      = row$diff_txt,
               gt90          = row$gt90), "\n\n")
}
```

```{r ggplot-add-vax, echo=FALSE, fig.width=10, fig.height=4}
print(plt_add_vax)
```

<!-- ===================== APPENDIX ===================== -->

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{`r get_text2("txt_sec_appendix", text_vars)`}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:appendix}{}")
cat(paste0("\\addcontentsline{toc}{section}{",
           get_text2("txt_sec_appendix", text_vars), "}"))
```

\vspace{0.8em}

\textcolor{cyan}{\large `r get_text2("txt_appendix_wuenic_header", text_vars)`}

\vspace{0.8em}

`r get_text2("txt_appendix_wuenic_body", text_vars)`

\vspace{0.3em}
`r get_text2("txt_appendix_methods_header", text_vars)`:

\begin{itemize}
  \item \begin{sloppypar}
        \textcolor{blue}{\href{https://pubmed.ncbi.nlm.nih.gov/19649368/}{Burton et al. 2009. WHO and UNICEF estimates of national infant immunization coverage: methods and processes}}
        \end{sloppypar}
  \item \begin{sloppypar}
        \textcolor{blue}{\href{https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0047806}{Burton et al. 2012. A formal representation of the WHO and UNICEF estimates of national immunization coverage: a computational logic approach}}
        \end{sloppypar}
  \item \begin{sloppypar}
        \textcolor{blue}{\href{https://openpublichealthjournal.com/VOLUME/6/PAGE/73/ABSTRACT/}{Brown et al. 2013. An introduction to the grade of confidence used to characterize uncertainty around the WHO and UNICEF estimates of national immunization coverage}}
        \end{sloppypar}
\end{itemize}

\vspace{0.3em}

`r str_glue(get_text2("txt_appendix_wuenic_revision", text_vars))`

\vspace{1.3em}

\textcolor{cyan}{\fontsize{14}{16}\selectfont `r get_text2("txt_appendix_definitions_header", text_vars)`}

\vspace{0.3em}

```{r appendix-definitions, results='asis', echo=FALSE}
# definitions are stored as a named list in txt_vars_talking_points.R
# and retrieved here so they translate properly
defs <- list(
  get_text2("txt_def_bcg",   text_vars),
  get_text2("txt_def_hepbb", text_vars),
  get_text2("txt_def_dtp",   text_vars),
  get_text2("txt_def_hepb3", text_vars),
  get_text2("txt_def_hib3",  text_vars),
  get_text2("txt_def_pol3",  text_vars),
  get_text2("txt_def_ipv",   text_vars),
  get_text2("txt_def_mcv",   text_vars),
  get_text2("txt_def_rotac", text_vars),
  get_text2("txt_def_pcv3",  text_vars),
  get_text2("txt_def_yfv",   text_vars),
  get_text2("txt_def_menga", text_vars),
  get_text2("txt_def_hpv",   text_vars))

for (d in defs) {
  cat(paste0("-   ", d, "\n"))
  cat("\\vspace{0.5em}\n")
}
```

\vspace{1.3em}

\textcolor{cyan}{\fontsize{14}{16}\selectfont `r get_text2("txt_appendix_resources_header", text_vars)`} \textcolor{blue}{\href{https://data.unicef.org/resources/immunization/}{\underline{https://data.unicef.org/resources/immunization/}}}
