---
title: ''
output:
  word_document:
    reference_docx: TP_word_template.dotx
    toc_depth: 3
  pdf_document:
    keep_tex: false
    latex_engine: xelatex
    toc: false
    toc_depth: 3
geometry: left=1cm,right=1cm,top=1.1cm,bottom=1.1cm
fontsize: 12pt
params:
  region: Default
  language: en
  plot_language: en
header-includes:
- \usepackage{enumitem}
- \geometry{a4paper, left=15mm, right=15mm, top=15mm, bottom=20mm}
- \setlist[itemize,1]{label=\textbullet}
- \setlist[itemize,2]{label=\(\circ\)}
- \setlist[itemize,3]{label=\(\infty\)}
- \setmainfont{Arial}
- \usepackage{hyperref}
- \hypersetup{colorlinks=true, linkcolor=blue, urlcolor=blue, breaklinks=true}
- \usepackage{tcolorbox}
- \usepackage{fancyhdr}
- \setlength{\headheight}{15pt}
- "\\addtolength{\\topmargin}{-3pt}"
- \definecolor{unicefblue}{RGB}{28, 171, 226}
- \pagestyle{fancy}
- \fancyhf{}
- \fancyhead[L]{}
- \fancyhead[C]{}
- \fancyhead[R]{\hyperlink{top}{\small \color{black} $\uparrow$ Back to top}}
- \fancyfoot[C]{\thepage}
- \usepackage{tocloft}
- \renewcommand{\cftsecfont}{\normalfont}
- \renewcommand{\cftsecpagefont}{\normalfont}
- \renewcommand{\cftsecleader}{\cftdotfill{\cftdotsep}}
---

\hypertarget{top}{}

```{r setup, include=FALSE}
options(scipen=999)

# ── TRANSLATION SETUP ─────────────────────────────────────────────────────────
source(file.path(RevDir, str_glue("/unicef-products/{type}/talking-points/region-specific-talking-points/txt_vars_talking_points.R")))        # loads text_vars_en
source(file.path(utils, "R/narrative-translation_deepl_grace.R"))    # loads text_vars (translated)

language <- params$language

comparison_col <- as.character(comp_yr)
recent_col  <- as.character(rev_yr)
minusone_col <- as.character(rev_yr - 1)

# filter based on selected region
df <- wuenic_dta %>% filter(Region == current_region) %>% 
  label_vals_millions(target, "target_lbl") %>%
  label_vals_millions(vaccinated, "vaccinated_lbl") %>%
  label_vals_millions(unvaccinated, "unvaccinated_lbl") %>%
  filter(vaccine != "pol3") %>% 
  mutate(region_name = case_when(
    Region == "EAPR" ~ "East Asia and Pacific",
    Region == "ECAR" ~ "Eastern Europe and Central Asia",
    Region == "ESAR" ~ "Eastern and Southern Africa",
    Region == "LACR" ~ "Latin America and the Caribbean",
    Region == "MENA" ~ "Middle East and North Africa",
    Region == "Non-programme" ~ "Non-programme",
    Region == "ROSA" ~ "South Asia",
    Region == "WCAR" ~ "West and Central Africa",
    Region == "APRO" ~ "Asia-Pacific Regional Office (APRO)",
    Region == "Gavi 57" ~ "Gavi 57",
    Region == "Never Gavi MICs" ~ "Never Gavi Middle Income Countries",
    Region == "Gavi HICs" ~ "Gavi High-Income Countries",
    Region == "Gavi transition" ~ "Gavi Transition Countries",
    Region == "LICs" ~ "World Bank Low-Income Countries",
    Region == "UMICs" ~ "World Bank Upper-Middle-Income Countries",
    Region == "WB HICs" ~ "World Bank High-Income Countries",
    Region == "LMICs" ~ "World Bank Lower-Middle-Income Countries",
    Region == "World Bank FCV" ~ "World Bank Fragile, Conflict, and Violence-Affected (FCV) Countries",
    Region == "World Bank non-FCV" ~ "World Bank Non-FCV Countries (Non-Fragile, Conflict, and Violence-Affected)",
    Region == "EUR" ~ "WHO European",
    Region == "AFR" ~ "WHO African",
    Region == "AMR" ~ "WHO Region of the Americas",
    Region == "WPR" ~ "WHO Western Pacific",
    Region == "SEAR" ~ "WHO South-East Asia",
    Region == "EMR" ~ "WHO Eastern Mediterranean",
    Region == "WHO FCV" ~ "WHO Fragile, Conflict, and Vulnerable Settings",
    Region == "WHO non-FCV" ~ "WHO Non-FCV (Non-Fragile, Conflict, and Vulnerable) Settings",
    Region == "Northern Africa" ~ "Northern Africa",
    Region == "Southern Africa" ~ "Southern Africa",
    Region == "Western Africa" ~ "Western Africa",
    Region == "Central Africa" ~ "Central Africa",
    Region == "Eastern Africa" ~ "Eastern Africa",
    TRUE ~ Region
  ))

hpv_df <- hpv_dta %>% filter(Region == current_region) %>% 
  label_vals_millions(target, "target_lbl") %>%
  label_vals_millions(vaccinated, "vaccinated_lbl") %>%
  label_vals_millions(unvaccinated, "unvaccinated_lbl")

## vars
regn       <- unique(df$Region)
region <- tolower(regn)
regn_name  <- unique(df$region_name)
wuenic_src <- str_glue("Source: WHO/UNICEF Estimates of National Immunization Coverage, {rev_yr} revision")
reg_type   <- "region"


if(org %in% c("au", "gavi", "wb", "wb_fcv", "who", "who_fcv", "european_union")) {
  region_translated_plots <- regn
  region_translated <- regn
} else {
  region_translated_plots <- t_lookup(region, plot_lang)
  region_translated <- t_lookup(region, language)
}

if(region %in% c("Western Africa", "Central Africa")) {
  region_translated_plots <- t_lookup(region, plot_lang)
}

region_translated <- case_when(
  regn == "LMICs (Low- and Middle-Income Countries)" ~ "LMICs",
  regn == "WB HICs (World Bank High-Income Countries)" ~ "WB HICs",
  regn == "UMICs (World Bank Upper-Middle-Income Countries)" ~ "UMICs",
  regn == "LICs (World Bank Low-Income Countries)" ~ "LICs",
  regn == "World Bank non-FCV" ~ "World Bank Non-FCV Countries (Non-Fragile, Conflict, and Violence-Affected)",
  regn == "WHO non-FCV" ~ "WHO Non-FCV (Non-Fragile, Conflict, and Vulnerable) Settings",
  regn == "World Bank FCV" ~ "World Bank Fragile, Conflict, and Violence-Affected (FCV) Countries",
  regn == "WHO FCV" ~ "WHO Fragile, Conflict, and Vulnerable (FCV) Settings",
  regn == "EUR" ~ "WHO European",
  regn == "AFR" ~ "WHO African",
  regn == "AMR" ~ "WHO Region of the Americas",
  regn == "WPR" ~ "WHO Western Pacific",
  regn == "SEAR" ~ "WHO South-East Asia",
  regn == "EMR" ~ "WHO Eastern Mediterranean",
  regn == "Never Gavi MICs" ~ "Never Gavi Middle Income Countries",
  regn == "Gavi HICs" ~ "Gavi High-Income Countries",
  regn == "Gavi transition" ~ "Gavi Transition Countries",
  regn == "LICs" ~ "World Bank Low-Income Countries",
  regn == "UMICs" ~ "World Bank Upper-Middle-Income Countries",
  regn == "WB HICs" ~ "World Bank High-Income Countries",
  regn == "LMICs" ~ "World Bank Lower-Middle-Income Countries",
  regn == "APRO" ~ "Asia-Pacific Regional Office (APRO)",
  TRUE ~ region_translated_plots
)

region_translated_plots<- case_when(
  regn == "LMICs (Low- and Middle-Income Countries)" ~ "LMICs",
  regn == "WB HICs (World Bank High-Income Countries)" ~ "WB HICs",
  regn == "UMICs (World Bank Upper-Middle-Income Countries)" ~ "UMICs",
  regn == "LICs (World Bank Low-Income Countries)" ~ "LICs",
  regn == "World Bank non-FCV" ~ "World Bank Non-FCV Countries",
  regn == "WHO non-FCV" ~ "WHO Non-FCV Settings",
  regn == "World Bank FCV" ~ "World Bank FCV Countries",
  regn == "WHO FCV" ~ "WHO FCV Settings",
  regn == "EUR" ~ "WHO European",
  regn == "AFR" ~ "WHO African",
  #regn == "AMR" ~ "WHO Region of the Americas",
  regn == "WPR" ~ "WHO Western Pacific",
  regn == "SEAR" ~ "WHO South-East Asia",
  regn == "Never Gavi MICs" ~ "Never Gavi Middle Income Countries",
  regn == "Gavi HICs" ~ "Gavi High-Income Countries",
  regn == "Gavi transition" ~ "Gavi Transition Countries",
  regn == "LICs" ~ "World Bank Low-Income Countries",
  regn == "UMICs" ~ "World Bank Upper-Middle-Income Countries",
  regn == "WB HICs" ~ "World Bank High-Income Countries",
  regn == "LMICs" ~ "World Bank Lower-Middle-Income Countries",
  TRUE ~ regn
)

report_title <- str_glue(get_text2("txt_region_banner_title", text_vars), region_translated = region_translated)
if(reg %in% c("European Union", "Central Africa", "Western Africa", "Eastern Africa", "Northern Africa", "Southern Africa",
              "World Bank FCV", "World Bank non-FCV", "WHO FCV", "WHO non-FCV",
              "Gavi transition", "Gavi 57", "Non-Programme", "APRO", "Non-programme", "AFRO-West", "AFRO-Central", "AFRO-East and South")) {
  report_title <- str_glue(get_text2("txt_region_banner_title_no_parentheses", text_vars))
}
```

```{r, results='asis', echo=FALSE}
cat(paste0(
"# ", report_title, "\n",
"# ", get_text2("txt_sec_banner_revision", text_vars), " ", rev_yr, "\n"
))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}

# list of countries in this region
ctry_list <- df %>% 
  filter(lvl_1 == "country") %>%
  select(country) %>%
  distinct() %>%
  arrange(country) %>%
  pull(country)

map_df <- base_map_df %>%
  mutate(admin = case_match(admin,
    "United States of America" ~ "United States",
    "Vietnam"                  ~ "Viet Nam",
    "Ivory Coast"              ~ "Cote d'Ivoire",
    "Democratic Republic of the Congo" ~ "DRC",
    "Republic of the Congo"    ~ "Congo",
    "United Republic of Tanzania" ~ "Tanzania",
    "North Korea"              ~ "DPRK",
    "South Korea"              ~ "Republic of Korea",
    "Republic of Serbia"       ~ "Serbia",
    "Moldova"                  ~ "Republic of Moldova",
    "Palestine"                ~ "State of Palestine",
    "Turkey"                   ~ "Turkiye",
    "Russia"                   ~ "Russian Federation",
    "East Timor"               ~ "Timor-Leste",
    "Federated States of Micronesia" ~ "Micronesia",
    "Papua New Guinea"         ~ "PNG",
    "The Bahamas"              ~ "Bahamas",
    "Brunei"                   ~ "Brunei Darussalam",
    "eSwatini"                 ~ "Eswatini",
    "São Tomé and Principe"    ~ "Sao Tome and Principe",
    "Netherlands"              ~ "Netherlands (Kingdom of the)",
    .default = admin
  ))

# flag region
map_df <- map_df %>%
  mutate(highlight = ifelse(admin %in% ctry_list, "Region", "World"))

# plot
world_map <- ggplot(map_df) +
  geom_sf(aes(fill = highlight), color = "white", size = 0.2) +
  scale_fill_manual(values = c("Region" = "#1CABE2", "World" = "#d7d7d7"), name = NULL) +
  theme_void() +
  theme(legend.position = "none", 
        plot.title = element_markdown(hjust = 0.5, size = 12, face = "bold", color = "#1CABE2"),
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.5)) +
  labs(title = str_glue(get_text2("txt_region_title", plot_vars)))
```
\vspace{-40pt}
```{r ggplot-country-map, echo=FALSE, warning=FALSE, fig.align='center', dpi=300}
suppressMessages(print(world_map))
```
\vspace{-60pt}

```{r region-includes, echo=FALSE, results='asis'}
cat("\n\n")

cat(
  "- ",
  str_glue(
    get_text2("txt_region_includes_list", text_vars),
    region_translated = region_translated,
    n_ctry = length(ctry_list),
    ctry_list_txt = knitr::combine_words(ctry_list)
  ),
  "\n\n",
  sep = ""
)
```

```{r setup-vax, include=FALSE}
# number of countries in region
nctry <- n_distinct(df$iso3c) - 1

# key vaccines and years
vax <- df %>% 
  filter(vaccine %in% c("dtp1","dtp3","mcv1","mcv2"),
         year %in% c(comp_yr, rev_yr-1, rev_yr)) %>%
    arrange(desc(lvl_1), country, vaccine, year)

# unvaccinated data labels
unvac_lbl <- vax %>%
    select(lvl_1, lvl_2, Region, iso3c, country, year, vaccine, unvaccinated_lbl) %>%
  pivot_wider(., names_from = year, values_from = unvaccinated_lbl) %>%
  rename_with(~ paste0(., "_lbl"), .cols = (ncol(.)-2):ncol(.))

# unvaccinated raw values
unvac <- vax %>%
    select(lvl_1, lvl_2, Region, iso3c, country, year, vaccine, unvaccinated) %>%
  pivot_wider(., names_from = year, values_from = unvaccinated) %>%
  left_join(., unvac_lbl) %>%
  distinct() %>%
  mutate(pchng = round((.[[9]]/.[[7]] - 1) * 100, 0),
         pchng_txt = case_when(.[[10]] == .[[12]] ~ str_glue(get_text2("txt_unvac_stable", text_vars)),
                               pchng < 0 ~ str_glue(get_text2("txt_unvac_declined", text_vars)),
                               pchng > 0 ~ str_glue(get_text2("txt_unvac_increased", text_vars)),
                               TRUE ~ NA_character_)) %>%
  mutate(unvac_rgn = ifelse(lvl_1 == "region", .[[9]], NA_real_)) %>%
  group_by(Region, vaccine) %>%
  fill(unvac_rgn, .direction = "updown") %>%
  ungroup() %>%
  mutate(unvac_ctry_pcnt = round(.[[9]] / unvac_rgn * 100, 0)) %>%
  arrange(desc(lvl_1), vaccine, desc(unvac_ctry_pcnt)) %>%
  group_by(lvl_1, Region, vaccine) %>%
  mutate(unvac_ctry_pcnt_cumsum = cumsum(unvac_ctry_pcnt)) %>%
  ungroup() 

# vaccinated
vacc <- vax %>%
  select(lvl_1, lvl_2, Region, iso3c, country, year, vaccine, vaccinated) %>%
  pivot_wider(., names_from = year, values_from = vaccinated) %>%
  mutate(pchng = round((.[[9]] / .[[7]] - 1) * 100, 1),
         pchng_txt = case_when(pchng > 0 ~ t_lookup("txt_vacc_increased", language),
                               pchng < 0 ~ t_lookup("txt_vacc_decreased", language),
                               TRUE ~ t_lookup("txt_vacc_same", language))) %>%
  label_vals_millions(.[[7]], 'vacc_comp_yr_lbl') %>%
  label_vals_millions(.[[9]], 'vacc_rev_yr_lbl')

# coverage   
cvg <- vax %>%
  select(lvl_1, lvl_2, Region, iso3c, country, year, vaccine, coverage) %>%
  pivot_wider(., names_from = year, values_from = coverage)

# target pop (si)
tgt <- vax %>%
  filter(vaccine == "dtp1") %>%
    select(lvl_1, lvl_2, Region, iso3c, country, year, vaccine, target) %>%
  pivot_wider(., names_from = year, values_from = target) %>%
  mutate(diff = .[[9]] - .[[7]],
         moreless = case_when(diff < 0 ~ t_lookup("txt_moreless_fewer", language),
                              diff > 0 ~ t_lookup("txt_moreless_more", language),
                              TRUE ~ t_lookup("txt_moreless_same", language)),
         to_vax = case_when(diff < 0 ~ t_lookup("txt_surv_inf_vax_3", language),
                              diff > 0 & diff < 1000000 ~ t_lookup("txt_surv_inf_vax_1", language),
                              diff > 0 & diff >= 1000000 ~ t_lookup("txt_surv_inf_vax_2", language))) %>%
    label_vals(diff, 'diff_lbl') %>%
  mutate(diff_abs = abs(diff)) %>%
  label_vals_millions(diff_abs, 'diff_abs_lbl')
```

```{r setup-overview, include=FALSE}
# coverage >90% [region, any vaccine]
gt90_reg <- df %>%
  select(-target_grp) %>%
  distinct() %>%
  filter(lvl_1 == "region",
         year == rev_yr) %>%
  mutate(gt90 = case_when(coverage >= 90 ~ 1, 
                          TRUE ~ 0),
         gt90_sum = sum(gt90))

# coverage >90% [country, any vaccine]
gt90_ctry <- df %>%
  select(-target_grp) %>%
  distinct() %>%
  filter(lvl_1 == "country",
         year == rev_yr) %>%
  group_by(iso3c) %>%
  mutate(gt90 = case_when(coverage >= 90 ~ 1, 
                          TRUE ~ 0),
         gt90_sum = sum(gt90)) %>%
  ungroup() %>%
  mutate(ctry_anyvax_gt90 = ifelse(gt90_sum > 0, 1, 0)) %>%
  select(ctry_anyvax_gt90, iso3c, country) %>%
  distinct() %>%
  arrange(ctry_anyvax_gt90)

# number of countries that did not achieve 90% coverage in a single vaccine
nctry_lt90 <- gt90_ctry %>%
  #filter(country != "Nicaragua") %>% 
  filter(ctry_anyvax_gt90 == 0) %>%
  summarise(count = n()) %>%
  pull(count) %>%
  replace_na(0)

# Generate the list of failed countries ahead of time
failed_countries_list <- gt90_ctry %>% 
  #filter(country != "Nicaragua") %>% 
  filter(ctry_anyvax_gt90 == 0) %>% 
  pull(country) %>% 
  knitr::combine_words()

if (nctry_lt90 == 1) {
  nctry_lt90_txt <- str_glue(
    get_text2("txt_ovw_ctry_level_failed_one", text_vars), 
    nctry_lt90 = nctry_lt90, 
    failed_list = failed_countries_list
  )
} else if (nctry_lt90 > 1) {
  nctry_lt90_txt <- str_glue(
    get_text2("txt_ovw_ctry_level_failed_multiple", text_vars), 
    nctry_lt90 = nctry_lt90, 
    failed_list = failed_countries_list
  )
} else {
  nctry_lt90_txt <- get_text2("txt_ovw_ctry_level_failed_none", text_vars)
}

# summary
gt90_ctry_summary <- gt90_ctry %>%
  group_by(ctry_anyvax_gt90) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  mutate(total = sum(n),
         pcnt = round(n / total * 100, 0))

# coverage change [prev_yr, rev_yr]
cvg_chng <- vax %>%
  filter(lvl_1 == "region",
         year %in% c(rev_yr, rev_yr-1)) %>%
  select(lvl_1:vaccine, coverage) %>%
  pivot_wider(., names_from = year, values_from = coverage) %>%
    mutate(diff = .[[8]] - .[[7]]) %>%
  mutate(diff_txt = case_when(diff > 0 ~ str_glue(get_text2("txt_cvg_increased_short", text_vars)),
                              diff < 0 ~ str_glue(get_text2("txt_cvg_decreased_short", text_vars)),
                              TRUE ~ str_glue(get_text2("txt_cvg_same_short", text_vars))))
```

```{r setup-progress-ia2030-zd-target, include=FALSE}
zd <- df %>%
  filter(vaccine == 'dtp1', 
         lvl_1 =='region',
         year >= 2000) %>%
  mutate(measure = 'unvaccinated') %>%
  select(lvl_1:vaccine, measure, unvaccinated)

zd_target <- zd %>%
  filter(year == 2019) %>%
  mutate(unvaccinated = unvaccinated/2,
         measure = 'IA2030 zero-dose target',
         year = 2030)

zd2 <- zd %>%
  mutate(measure = 'Estimate') %>%
  rbind(zd_target) %>%
  mutate(chart = 'bar',
         unvaccinated = round(unvaccinated, 0))

target <- zd %>%
  filter(vaccine == 'dtp1',
         year %in% c(2019)) %>%
  select(lvl_1:country, vaccine, year, unvaccinated) %>%
  pivot_wider(., names_from = year, values_from = unvaccinated) %>%
  mutate(target = `2019`/2) %>%
  mutate(`2020` = `2019`) %>%
  mutate(factor = (`2020`-target)/(2030-2020)) %>%
  mutate(`2021` = `2020` - factor) %>%
  mutate(`2022` = `2021` - factor) %>%
  mutate(`2023` = `2022` - factor) %>%
  mutate(`2024` = `2023` - factor) %>%
  mutate(`2025` = `2024` - factor) %>%
  mutate(`2026` = `2025` - factor) %>%
  mutate(`2027` = `2026` - factor) %>%
  mutate(`2028` = `2027` - factor) %>%
  mutate(`2029` = `2028` - factor) %>%
  mutate(`2030` = `2029` - factor) %>%
  pivot_longer(-c(lvl_1:vaccine), names_to = 'year', values_to = 'zerodose_mt') %>%
  filter(year != 'factor', year != 'target') %>%
  mutate(year = as.numeric(year)) %>%
  select(lvl_1:year, unvaccinated_new = zerodose_mt) %>%
  left_join(., zd) %>%
  filter(year %in% c(2020:2030)) %>%
  mutate(diff = unvaccinated - unvaccinated_new,
         measure = "unvaccinated")  %>%
  select(lvl_1:year, measure, unvaccinated_new, unvaccinated, diff)

target_lines <- target %>%
  select(lvl_1:unvaccinated_new) %>%
  rename(unvaccinated = unvaccinated_new) %>%
  mutate(chart = 'line',
         measure = 'IA2030')

ia2030_2019 <- zd2 %>%
  filter(year == 2019) %>%
  mutate(chart = 'line',
         measure = 'IA2030')

zd2 <- rbind(zd2, target_lines, ia2030_2019) %>%
  filter(!(year == 2019 & chart == 'line'))

data_text <- zd2 %>%
  filter(year == rev_yr) %>%
  select(Region, measure, unvaccinated) %>%
  pivot_wider(., names_from = measure, values_from = unvaccinated) %>%
  mutate(diff = Estimate - IA2030,
         diff_txt = case_when(diff > 0 ~ get_text2("txt_zd_higher_than", text_vars),
                              diff < 0 ~ get_text2("txt_zd_lower_than", text_vars),
                              diff == 0 ~ get_text2("txt_zd_same_as", text_vars)),
         pcnt_diff = round((Estimate / IA2030 - 1) * 100,0)) %>%
  label_vals_millions(Estimate, "Estimate_lbl") %>%
  label_vals_millions(IA2030, "IA2030_lbl")
```

<!-- ===================== HEATMAP & LINE PLOT FOR EACH VACCINE ===================== -->

```{r setup-heatmap, include=FALSE}
if (regn == "Africa") {
  by <- get_text2("txt_map_region", text_vars)
  heatmap_df <- wuenic_dta %>%
    filter(lvl_1 == "region", lvl_2 %in% c("region_au", "region_au_africa"), vaccine %in% c("dtp1", "dtp3", "mcv1"), year >= 2000) %>%
    select(lvl_1:vaccine, coverage)
} else if (regn == "Global") {
  by <- get_text2("txt_map_region", text_vars)
  heatmap_df <- wuenic_dta %>%
    filter(lvl_1 == "region", lvl_2 %in% c("region_unicef_ops", "region_unicef_global_old"), 
           vaccine %in% c("dtp1", "dtp3", "mcv1"), year >= 2000) %>%
    select(lvl_1:vaccine, coverage)
} else {
  by <- get_text2("txt_map_country", text_vars)
  heatmap_df <- df %>% filter(vaccine %in% c("dtp1", "dtp3", "mcv1"), year >= 2000) %>% select(lvl_1:vaccine, coverage)
}

heatmap_df <- heatmap_df %>%
  mutate(cvg_cat = case_when(coverage < 60 ~ '<60',
                             coverage %in% c(60:69) ~ '60-69',
                             coverage %in% c(70:79) ~ '70-79',
                             coverage %in% c(80:89) ~ '80-89',
                             coverage %in% c(90:94) ~ '90-94',
                             coverage >=95 ~ '>=95',
                             TRUE ~ NA_character_)) %>%
  group_by(country, vaccine) %>%
  mutate(rev_yr_cvg = ifelse(year == rev_yr, coverage, NA_real_)) %>%
  fill(rev_yr_cvg, .direction = "updown") %>%
  ungroup()

n_ctry <- heatmap_df %>% mutate(n = n_distinct(country)) %>% pull(n) %>% unique

position <- heatmap_df %>% 
  filter(year == rev_yr) %>%
  arrange(vaccine, coverage, Region == regn) %>%
  group_by(vaccine) %>%
  mutate(cvg_position = seq(n())) %>%
  ungroup()

if (regn == "Africa") {
  position <- position %>% filter(Region == regn) %>% select(vaccine, cvg_position)
} else {
  position <- position %>% filter(lvl_1 == "region") %>% select(vaccine, cvg_position)
}

if (regn == "LACR") {
  yaxis_txt <- 10
# } else if (n_ctry > 35) {
#   yaxis_txt <- 7
} else if (regn == "ROSA") {
  yaxis_txt <- 13
} else if (regn %in% c("LMICs", "WB HICs", "UMICs", "Gavi HICs", "Never Gavi MICs", "Gavi 57", "Non-programme")) {
  yaxis_txt <- 12
} else {
  yaxis_txt <- 11
}

# heatmap for each vaccine
for (v in c('dtp1', 'dtp3', 'mcv1')) {
  plt_df <- heatmap_df %>% 
    filter(vaccine == v) %>%
    mutate(country = ifelse(country == filter_regn, paste0("**", country, "**"), country)) %>%
    left_join(position) %>%
    distinct()# %>% 
    #filter(country != "Nicaragua")
  
  if(v == "dtp1") {
    heatmap_title <- str_glue(get_text2("txt_dtp1_heatmap_title", plot_vars))
  } else if(v == "dtp3") {
    heatmap_title <- str_glue(get_text2("txt_dtp3_heatmap_title", plot_vars))
  } else if(v == "mcv1") {
    heatmap_title <- str_glue(get_text2("txt_mcv1_heatmap_title", plot_vars))
  }
  
  gg <- plt_df %>%
    ggplot(aes(x = year, y = reorder(country, -rev_yr_cvg), fill = coverage)) +
    geom_tile(color = "white") +
    # geom_text(data = plt_df %>% filter(coverage > 80 | coverage < 95), aes(label = paste0(as.character(coverage))), colour = 'black', vjust = 0.5, size = 3.5) +
    # geom_text(data = plt_df %>% filter(coverage <= 80 | coverage >= 95), aes(label = paste0(as.character(coverage))), colour = 'white', vjust = 0.5, size = 3.5) +
    geom_text(aes(label = paste0(as.character(coverage))), color = 'white', vjust = 0.5, size = 3.5) +
    geom_rect(aes(xmin = 2000 - 0.5, xmax = rev_yr + 0.5, ymin = unique(cvg_position) - 0.5, ymax = unique(cvg_position) + 0.5), 
              fill = "transparent", color = "black", size = 1) +
    scale_x_continuous(breaks = seq(2000, rev_yr, by = 1), position = "top", expand = c(0, 0)) +
    scale_y_discrete(limits = rev, expand = c(0, 0)) +
    scale_fill_gradientn(
      colours = c('#B50800','#E2231A','#F26A21','#FFC20E','#80BD41','#00833D'),
      values = scales::rescale(c(0, 59, 69, 79, 88, 90, 100)),
      guide = "colorbar",
      limits = c(0, 100),
      name    = get_text2("txt_plt_cvg_pct_legend", plot_vars),
      breaks = c(0, 60, 70, 80, 90, 100),
      labels = c("0", "60", "70", "80", "90", "100")) +
    theme_minimal() +
    theme(
      strip.placement = "outside", panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
      axis.text.y = ggtext::element_markdown(size = yaxis_txt), 
      axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, hjust = 1),
      strip.text.x = element_text(size = 12), legend.position = "bottom", legend.key.width = unit(1.2, "cm"),
      plot.title = element_text(size = 14, hjust = 0.5, face = "bold")
    ) +
    labs(
      x = '', y = '',
      title = heatmap_title,
      caption = str_glue(get_text2("txt_wuenic_source", plot_vars), "\n", 
                         get_text2("txt_plt_ovw_heatmap_caption_vax", plot_vars), v_upper = toupper(v), rev_yr = rev_yr)
    )
  assign(paste0("heatmap_", plt_df$vaccine, sep = ""), gg)
}

# line plot for each vaccine
for (v in c('dtp1', 'dtp3', 'mcv1')) {
  local({
    v <- v
    plt_vax_df <- df %>% filter(country == filter_regn, vaccine == v) %>% select(year, coverage, unvaccinated)

    if (v == "dtp1") {
      vax_label  <- t_lookup("txt_plt_dtp1_legend_cvg", language)
      count_label <- t_lookup("txt_plt_dtp1_legend_zd", language)
      title_tpl   <- get_text2("txt_plt_dtp1_title", plot_vars)
    } else if (v == "dtp3") {
      vax_label  <- t_lookup("txt_plt_dtp3_legend_cvg", language)
      count_label <- t_lookup("txt_plt_unvac_undervac", language)
      title_tpl   <- get_text2("txt_plt_dtp3_title", plot_vars)
    } else if (v == "mcv1") {
      vax_label  <- t_lookup("txt_plt_mcv1_legend_cvg", language)
      count_label <- t_lookup("txt_plt_mcv_legend_zd", language)
      title_tpl   <- get_text2("txt_plt_mcv_title", plot_vars)
    }

    max_cvg <- max(plt_vax_df$coverage, na.rm = TRUE)
    max_zd  <- max(plt_vax_df$unvaccinated, na.rm = TRUE)
    scale_factor <- ifelse(max_zd > 0, max_cvg / max_zd, 1)

    vax_values  <- setNames("#1CABE2", vax_label)
    fill_values <- setNames("#E2231A", count_label)

    gg_line <- ggplot(plt_vax_df) +
      geom_col(aes(x = year, y = unvaccinated, fill = count_label), alpha = 0.4, width = 0.6) +
      geom_line(aes(x = year, y = coverage / scale_factor, color = vax_label), size = 0.8, linewidth = 1.5) +
      geom_point(aes(x = year, y = coverage / scale_factor, color = vax_label), size = 3) +
      geom_hline(yintercept = 90 / scale_factor, linetype = "dashed", alpha = 0.7) +
      scale_color_manual(name = NULL, values = vax_values, breaks = vax_label) +
      scale_fill_manual(name = NULL, values = fill_values, breaks = count_label) +
      guides(color = guide_legend(order = 1), fill = guide_legend(order = 2)) +
      scale_x_continuous(breaks = seq(2000, rev_yr, by = 1)) +
      scale_y_continuous(name = count_label, labels = scales::label_comma(),
        sec.axis = sec_axis(~ . * scale_factor, name = get_text2("txt_plt_cvg_pct", plot_vars), breaks = seq(0, 100, 25))) +
      labs(
        title = str_glue(title_tpl, ctry_name = regn, rev_yr = rev_yr),
        caption = paste0(t_lookup("txt_plt_ia2030_caption", language), "\n", str_glue(get_text2("txt_wuenic_source", plot_vars))),
        x = get_text2("txt_plt_year", plot_vars)
      ) +
      theme_unicef() +
      theme(
        panel.grid.minor = element_blank(), panel.grid.major.x = element_line(color = "grey90"),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 9), plot.title = element_text(size = 14, hjust = 0.5, face = "bold"), legend.position = "bottom"
      )
    assign(paste0("line_plt_", v), gg_line, envir = .GlobalEnv)
  })
}
```

<!-- ===================== PAGE 1: OVERVIEW ===================== -->

\newpage

## `r get_text2("txt_sec_overview", text_vars)`

\vspace{0.2em}

```{r, results='asis', echo=FALSE}
cat(paste0("-   ", str_glue(
  get_text2("txt_ovw_vax_gt90_summary", text_vars),
  n_achieved_txt = if (gt90_reg$gt90_sum[1] == 0) get_text2("txt_word_none", text_vars) else paste(gt90_reg$gt90_sum[1], get_text2("txt_word_out", text_vars)),
  n_vax = n_distinct(gt90_reg$vaccine),
  regn = regn,
  rev_yr = rev_yr
), "\n"))
```

```{r, results='asis', echo=FALSE}
cat(paste0("-   ", str_glue(
  get_text2("txt_ovw_coverage_change_summary", text_vars),
  dtp1_diff = cvg_chng %>% filter(vaccine == 'dtp1') %>% pull(diff_txt),
  dtp3_diff = cvg_chng %>% filter(vaccine == 'dtp3') %>% pull(diff_txt),
  mcv1_diff = cvg_chng %>% filter(vaccine == 'mcv1') %>% pull(diff_txt)
), "\n"))
```

```{r, results='asis', echo=FALSE}
n_achieved   <- gt90_ctry_summary %>% filter(ctry_anyvax_gt90 == 1) %>% pull(n)
total_ctry   <- unique(gt90_ctry_summary$total)
pct_achieved <- gt90_ctry_summary %>% filter(ctry_anyvax_gt90 == 1) %>% pull(pcnt)
n_vax        <- n_distinct(gt90_reg$vaccine)

if (nctry_lt90 == 0) {
  cat(str_glue(
    "    -   ", get_text2("txt_ovw_ctry_level_achieved", text_vars),
    "\n\\vspace{{1em}}\n"
  ))
} else {
  cat(str_glue(
    "-   ", get_text2("txt_ovw_ctry_level_achieved", text_vars), "\n",
    "-   {nctry_lt90_txt}\n"
  ))
}
```

```{r, results='asis', echo=FALSE}
cat(paste0("-   ", str_glue(
  get_text2("txt_ovw_zero_dose_summary", text_vars),
  rev_yr = rev_yr,
  estimate_lbl = data_text$Estimate_lbl,
  pcnt_diff = data_text %>% pull(pcnt_diff) %>% abs(),
  diff_txt = data_text %>% pull(diff_txt),
  regn = regn,
  ia2030_lbl = data_text$IA2030_lbl
), "\n"))
```

```{r setup-heatmap-reg, include=FALSE}
df_to_merge <- df %>% select(-region_name)
heatmap_reg_df <- df_to_merge %>%
  rbind(hpv_df %>% filter(lvl_1 == "region", vaccine_code == "PRHPVC_F") %>% select(-vaccine_code)) %>%
  filter(lvl_1 == "region") %>%
  mutate(year = as.character(year),
         cvg_cat = case_when(coverage < 60 ~ '<60',
                             coverage %in% c(60:69) ~ '60-69',
                             coverage %in% c(70:79) ~ '70-79',
                             coverage %in% c(80:89) ~ '80-89',
                             coverage %in% c(90:94) ~ '90-94',
                             coverage >=95 ~ '>=95',
                             TRUE ~ NA_character_)) %>%
  mutate(vaccine = case_when(vaccine == "hepbb" ~ "HepBB", 
                             vaccine == "hepb3" ~ "HepB3", 
                             vaccine == "hib3" ~ "Hib3", 
                             vaccine == "rotac" ~ "RotaC", 
                             vaccine == "menga" ~ "MengA", 
                             vaccine == "HPVc Females" ~ "HPVc", 
                             TRUE ~ toupper(vaccine))) %>%
  group_by(country, vaccine) %>% 
  mutate(mean = mean(coverage)) %>% 
  ungroup()

heatmap_reg_df$vaccine <- factor(heatmap_reg_df$vaccine, levels = c("BCG", "HepBB", "DTP1", "DTP3", "Hib3", "HepB3", "PCVC", "RotaC", "IPV1", "IPVC", "MCV1", "RCV1", "MCV2", "YFV", "MengA", "HPVc"))

heatmap_reg <- heatmap_reg_df %>%
  ggplot(aes(x = year, y = vaccine, fill = coverage)) +
  geom_tile(color = "white") +
  theme_minimal() +
  # geom_text(data = heatmap_reg_df %>% filter(coverage > 80 | coverage < 95), 
  #           aes(label = paste0(as.character(coverage))), colour = 'black', vjust = 0.5, size = 3.5) +
  # geom_text(data = heatmap_reg_df %>% filter(coverage <= 80 | coverage >= 95), 
  #           aes(label = paste0(as.character(coverage))), colour = 'white', vjust = 0.5, size = 3.5) +
  geom_text(data = heatmap_reg_df, aes(label = paste0(as.character(coverage))), colour = 'white', vjust = 0.5, size = 3.5) +
  labs(x = '', y = '', 
       title = str_glue(get_text2("txt_plt_ovw_heatmap_title_reg", plot_vars), regn = regn, min_yr = min(heatmap_reg_df$year), rev_yr = rev_yr), 
       caption = str_glue(get_text2("txt_wuenic_source", plot_vars))) +
  scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) +
  theme(strip.placement = "outside", 
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        strip.text.x = element_text(size = 12), 
        legend.position="bottom", legend.key.width = unit(1.2, "cm"), 
        plot.title = element_text(size = 14, hjust = 0.5, face = "bold")) +
  scale_fill_gradientn(
      colours = c('#B50800','#E2231A','#F26A21','#FFC20E','#80BD41','#00833D'),
      values = scales::rescale(c(0, 59, 69, 79, 88, 90, 100)),
      guide = "colorbar",
      limits = c(0, 100),
      name    = get_text2("txt_plt_cvg_pct_legend", plot_vars),
      breaks = c(0, 60, 70, 80, 90, 100),
      labels = c("0", "60", "70", "80", "90", "100"))
```

```{r zd-plot, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center'}
# chart showing the progress towards the IA2030 goal of reducing ZD children by half (bar chart with line showing linear trajectory annual targets)

###### plt_ia2030_zd_proj_bar_line :: progress towards zd target :: bar-chart, linechart ######
zd <- df %>%
  filter(lvl_1 == "region") %>% 
  filter(vaccine == 'dtp1', year >= 2000) %>%
  mutate(measure = 'unvaccinated') %>%
  select(lvl_1:vaccine, measure, unvaccinated)

zd_target <- zd %>%
  filter(year == 2019) %>%
  mutate(unvaccinated = unvaccinated/2,
         measure = 'IA2030 zero-dose target',
         year = 2030)

# data for bars
zd2 <- zd %>%
  mutate(measure = 'Estimate') %>%
  rbind(zd_target) %>%
  mutate(chart = 'bar',
         unvaccinated = round(unvaccinated, 0))

# target 
target <- zd %>%
  filter(vaccine == 'dtp1',
         year %in% c(2019)) %>%
  select(lvl_1:country, vaccine, year, unvaccinated) %>%
  pivot_wider(., names_from = year, values_from = unvaccinated) %>%
  mutate(target = `2019`/2) %>%
  mutate(`2020` = `2019`) %>%
  mutate(factor = (`2020`-target)/(2030-2020)) %>%
  mutate(`2021` = `2020` - factor) %>%
  mutate(`2022` = `2021` - factor) %>%
  mutate(`2023` = `2022` - factor) %>%
  mutate(`2024` = `2023` - factor) %>%
  mutate(`2025` = `2024` - factor) %>%
  mutate(`2026` = `2025` - factor) %>%
  mutate(`2027` = `2026` - factor) %>%
  mutate(`2028` = `2027` - factor) %>%
  mutate(`2029` = `2028` - factor) %>%
  mutate(`2030` = `2029` - factor) %>%
  pivot_longer(-c(lvl_1:vaccine), names_to = 'year', values_to = 'zerodose_mt') %>%
  filter(year != 'factor', year != 'target') %>%
  mutate(year = as.numeric(year)) %>%
  select(lvl_1:year, unvaccinated_new = zerodose_mt) %>%
  left_join(., zd) %>%
  filter(year %in% c(2020:2030)) %>%
  mutate(diff = unvaccinated - unvaccinated_new,
         measure = "unvaccinated")  %>%
  select(lvl_1:year, measure, unvaccinated_new, unvaccinated, diff)

# data for lines
target_lines <- target %>%
  select(lvl_1:unvaccinated_new) %>%
  rename(unvaccinated = unvaccinated_new) %>%
  mutate(chart = 'line',
         measure = 'IA2030')

ia2030_2019 <- zd2 %>%
  filter(year == 2019) %>%
  mutate(chart = 'line',
         measure = 'IA2030')

# append
zd2 <- rbind(zd2, target_lines, ia2030_2019)  %>%
  filter(!(year == 2019 & chart == 'line')) %>%
  distinct()

plt_ia2030_zd_proj_bar_line <- gg <-
  ggplot() + 
  geom_col(data = zd2 %>% filter(chart=='bar'), aes(y=unvaccinated, x=year, fill=measure)) +
  geom_line(data = zd2 %>% filter(chart=='line'), aes(y=unvaccinated, x=year)) +
  geom_point(data = zd2 %>% filter(chart=='line'), aes(y=unvaccinated, x=year)) +
  # geom_text(data = zd2 %>% filter(chart=='bar'), aes(y=unvaccinated, x=year, label = format(unvaccinated, big.mark=",")), vjust = -0.5, size = 3) +
  labs(y = get_text2("txt_ia2030_zd_proj_bar_line_y", plot_vars),
       x = "",
       title = str_glue(get_text2("txt_ia2030_zd_proj_bar_line_title", plot_vars)),
       caption = paste0(str_glue(get_text2("txt_wuenic_source", plot_vars)), "\n", 
                        str_glue(get_text2("txt_ia2030_zd_proj_bar_line_cpt", plot_vars)))) +
  theme_minimal()+
  theme_unicef()+
  theme(legend.position="bottom",
        legend.title = element_blank(),
        #axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=8.5),
        axis.text.y = element_text(size=9),
        legend.text = element_text(size=11),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor   = element_blank(),
        panel.grid.major.x = element_line(),
        axis.text.x        = element_text(angle = 45, hjust = 1, size = 9),
        plot.title         = element_text(size = 14, hjust = 0.5, face = "bold")) +
  scale_fill_manual(values = c("#0058AB","#69DBFF"),name = "") +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = c(seq(2000,rev_yr,1), 2025, 2030))
```

```{r ggplot-allvax-reg, echo=FALSE, fig.width=11, fig.height=7, dpi=300}
print(heatmap_reg)
```

```{r ggplot-zd_plot, echo=FALSE, fig.width=10, fig.height=5.5, dpi=300}
print(plt_ia2030_zd_proj_bar_line)
```

```{r setup-map-africa, include=FALSE, warning=FALSE}
map_dta <- df %>% 
  filter(year == rev_yr, vaccine %in% c('dtp1','dtp3','mcv1'), lvl_1 == 'country') %>%
  mutate(coverage = case_when(Region != regn ~ NA_real_, TRUE ~ coverage),
         cvg_cat = case_when(coverage < 60 ~ '<60', coverage %in% c(60:69) ~ '60-69', coverage %in% c(70:79) ~ '70-79', coverage %in% c(80:89) ~ '80-89', coverage %in% c(90:94) ~ '90-94', coverage >=95 ~ '>=95'))

map_df <- left_join(base_map_df, map_dta, by = "iso3c") %>% filter(!is.na(coverage))

for (v in c("dtp1","dtp3","mcv1")) {
  plt_df <- map_df %>% filter(vaccine == v)
  if (regn == "EAPR") {
    pacific_crs <- st_crs("+proj=robin +lon_0=160 +x_0=0 +y_0=0 +datum=WGS84 +units=m")
    plt_df <- plt_df %>% st_wrap_dateline(options = c("WRAPDATELINE=YES", "DATELINEOFFSET=10")) %>% st_transform(pacific_crs)
    coord_layer <- coord_sf(crs = pacific_crs, expand = FALSE)
  } else {
    coord_layer  <- coord_sf(expand = FALSE)
  }
  
  plt_df <- plt_df %>%
    #mutate(cvg_cat = if_else(regn == "LACR" & admin == "Nicaragua", get_text2("txt_map_other_countries", text_vars), as.character(cvg_cat))) %>%
    mutate(cvg_cat = factor(cvg_cat, levels = c('>=95', '90-94', '80-89', '70-79', '60-69', '<60', get_text2("txt_map_other_countries", plot_vars))))
  
  if(v == "dtp1") {
    title_text <- str_glue(get_text2("txt_map_dtp1_title", plot_vars))
  } else if(v == "dtp3") {
    title_text <- str_glue(get_text2("txt_map_dtp3_title", plot_vars))
  } else if(v == "mcv1") {
    title_text <- str_glue(get_text2("txt_map_mcv1_title", plot_vars))
  }
  
  gg <- ggplot() +
    geom_sf(data = plt_df, aes(fill = cvg_cat), alpha = 0.75, size = 0.5, color = "black") +
    scale_fill_manual(name = '', breaks = levels(plt_df$cvg_cat), 
                      values = c('<60'='#B50800','60-69'='#E2231A','70-79'='#F26A21',
                                 '80-89'='#FFC20E','90-94'='#80BD41','>=95'='#00833D','Other countries'='lightgrey')) +
    coord_layer + theme_minimal(base_size = 12) +
    theme(axis.title = element_blank(), 
          axis.text = element_blank(), 
          axis.line = element_blank(), 
          axis.ticks = element_blank(), 
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          plot.title = element_text(size = 14, hjust = 0.5, face = "bold"), 
          caption = element_text(size = 8, hjust = 0, vjust = 1)) +
    labs(title = title_text, 
         caption = str_glue(get_text2("txt_map_disclaimer", plot_vars)))
  assign(paste0("map_", v, sep = ""), gg)
}
```

<!-- ===================== DTP1 ===================== -->

```{r setup-dtp1, include=FALSE}
v <- "dtp1"

unvac_v <- unvac %>% filter(vaccine == v) %>%
  mutate(ia2030tgt = round(.[[7]]/2, 0),
         ia2030tgt_pdiff = round((.[[9]]/ia2030tgt - 1) * 100),
         ia2030tgt_pdiff_hilo = case_when(ia2030tgt_pdiff > 0 ~ "higher", ia2030tgt_pdiff < 0 ~ "lower", TRUE ~ "same")) %>%
  label_vals(ia2030tgt, "ia2030tgt_lbl")

vaccinated_v <- vacc %>% 
  filter(vaccine == v, country == filter_regn)

cvg_v <- cvg %>% filter(vaccine == v)

zd_ctry <- unvac_v %>%
  #filter(country != "Nicaragua") %>%
  arrange(desc(.[[9]])) %>%
  filter(lvl_1 == "country") %>%
  mutate(rank = seq(n())) %>%
  slice_max(order_by = .[[9]], n = 2)

cvg_v <- cvg %>%
  filter(vaccine == v, lvl_1 == "country") %>%
  arrange(.[[9]]) %>%
  mutate(rank = dense_rank(.[[9]])) %>%
  mutate(gt90 = ifelse(.[[9]] >= 90, 1, 0), n_gt90 = sum(gt90)) %>%
  mutate(gt90_txt = ifelse(
    n_gt90 == 0, get_text2("txt_dtp1_none_achieved", text_vars),
    ifelse(n_gt90 < 10, str_glue(" - {knitr::combine_words(unique(country[.[[9]] >= 90]))}"), "")
  )) %>%
  mutate(n_gt90_txt = ifelse(
    n_gt90 == 1,
    str_glue(get_text2("txt_was_1_country", text_vars)),
    ifelse(n_gt90 > 1,
           str_glue(get_text2("txt_were_n_countries_short", text_vars), n_gt90 = n_gt90),
           "")
  ))

#cvg_v_no_nicaragua <- cvg_v %>% filter(country != "Nicaragua")
```

\newpage

## `r t_lookup("txt_sec_dtp1", language)`

```{r, results='asis', echo=FALSE}
cat(paste0("-   ", str_glue(
  get_text2("txt_dtp1_change_summary", text_vars),
  pchng = vaccinated_v %>% filter(lvl_1 == 'region') %>% pull(pchng) %>% abs(),
  direction = vaccinated_v %>% filter(lvl_1 == 'region') %>% pull(pchng_txt),
  comp_yr = comp_yr,
  comp_val = vaccinated_v %>% filter(lvl_1 == 'region') %>% pull(vacc_comp_yr_lbl),
  rev_yr = rev_yr,
  rev_val = vaccinated_v %>% filter(lvl_1 == 'region') %>% pull(vacc_rev_yr_lbl)
), "\n"))
```

```{r, results='asis', echo=FALSE}
cvg_dtp1   <- cvg %>% filter(vaccine == v)

dtp1_rev_cvg <- cvg_dtp1 %>% filter(lvl_1 == "region") %>% pull(as.character(rev_yr))
dtp1_comp_cvg <- cvg_dtp1 %>% filter(lvl_1 == "region") %>% pull(as.character(comp_yr))
dtp1_minusone_cvg <- cvg_dtp1 %>% filter(lvl_1 == "region") %>% pull(minusone_col)

dtp1_comp_dir <- if (dtp1_rev_cvg > dtp1_comp_cvg) {
  "higher"
} else if (dtp1_rev_cvg < dtp1_comp_cvg) {
  "lower"
} else {
  "same"
}

dtp1_minusone_dir <- if (dtp1_rev_cvg > dtp1_minusone_cvg) {
  "higher"
} else if (dtp1_rev_cvg < dtp1_minusone_cvg) {
  "lower"
} else {
  "same"
}

dtp1_cvg_key <- paste0("txt_dtp1_cvg_", dtp1_comp_dir, "_", dtp1_minusone_dir)

cat(paste0("-   ", str_glue(
  get_text2(dtp1_cvg_key, text_vars),
  rev_yr       = rev_yr,
  rev_pct      = dtp1_rev_cvg,
  comp_yr      = comp_yr,
  comp_pct     = dtp1_comp_cvg,
  minusone_yr  = minusone_col,
  minusone_pct = dtp1_minusone_cvg
), "\n"))
```

```{r, results='asis', echo=FALSE}
cat(paste0("-   ", str_glue(
  get_text2("txt_dtp1_zd_ia2030_summary", text_vars),
  rev_yr = rev_yr,
  estimate_lbl = data_text$Estimate_lbl,
  pcnt_diff = data_text %>% pull(pcnt_diff) %>% abs(),
  diff_txt = data_text %>% pull(diff_txt),
  regn = regn,
  ia2030_lbl = data_text$IA2030_lbl
), "\n"))
```

```{r, results='asis', echo=FALSE}
# moreless_val <- tgt %>% filter(lvl_1 == 'region') %>% pull(moreless)
# 
# txt_key <- if (moreless_val == "more") "txt_dtp1_surviving_infants_more" else "txt_dtp1_surviving_infants_fewer"
# 
# cat(paste0("-   ", str_glue(
#   get_text2(txt_key, text_vars),
#   diff_abs_lbl = tgt %>% filter(lvl_1 == 'region') %>% pull(diff_abs_lbl),
#   comp_yr = comp_yr,
#   to_vax = tgt %>% filter(lvl_1 == 'region') %>% pull(to_vax)
# ), "\n"))
```

```{r, results='asis', echo=FALSE}
cat(paste0("-   ", str_glue(
  get_text2("txt_dtp1_top2_zd_ctry_summary", text_vars),
  rev_yr = rev_yr,
  n_ctry = n_distinct(zd_ctry$iso3c),
  ctry_list = unique(zd_ctry$country) %>% knitr::combine_words(),
  pct1 = zd_ctry %>% filter(rank == 1) %>% pull(unvac_ctry_pcnt),
  pct2 = zd_ctry %>% filter(rank == 2) %>% pull(unvac_ctry_pcnt),
  regn = regn,
  total_zd = unvac_v %>% filter(lvl_1 == 'region') %>% pull(12),
  ctry1 = zd_ctry %>% filter(rank == 1) %>% pull(country),
  n1 = zd_ctry %>% filter(rank == 1) %>% pull(12),
  ctry2 = zd_ctry %>% filter(rank == 2) %>% pull(country),
  n2 = zd_ctry %>% filter(rank == 2) %>% pull(12)
), "\n"))
```

```{r, results='asis', echo=FALSE}
cat(paste0("-   ", str_glue(
  get_text2("txt_dtp1_lowest_coverage_summary", text_vars),
  ctry1 = cvg_v %>% filter(rank == 1) %>% pull(country) %>% knitr::combine_words(),
  pct1 = cvg_v %>% filter(rank == 1) %>% pull(9) %>% unique(),
  ctry2 = cvg_v %>% filter(rank == 2) %>% pull(country) %>% knitr::combine_words(),
  pct2 = cvg_v %>% filter(rank == 2) %>% pull(9) %>% unique()
), "\n"))
```

```{r, results='asis', echo=FALSE}
if (unique(cvg_v$n_gt90) == 0) {
  cat(paste0("- ", str_glue(get_text2("txt_dtp1_none_achieved", text_vars), nctry = nctry, regn = regn, rev_yr = rev_yr)))
} else if (unique(cvg_v$n_gt90 == 1)) {
  cat(paste0("- ", str_glue(get_text2("txt_was_1_country", text_vars), gt90_txt = unique(cvg_v$gt90_txt))))
} else {
  cat(paste0("- ", str_glue(get_text2("txt_dtp1_share_achieved", text_vars), n_gt90_txt = unique(cvg_v$n_gt90_txt), pct = round(unique(cvg_v$n_gt90)/nctry*100,0), nctry = nctry, gt90_txt = cvg_v %>% pull(gt90_txt) %>% unique()), "\\vspace{1em}\n\n"))
}
```

\vspace{1em}

```{r, results='asis', echo=FALSE}
if (regn %in% c("Africa", "Global")) {
  reg_zd_txt <- heatmap_df
  cat(paste0("    -   ", "", "\n"))
}

# dynamic heatmap height
heatmap_height <- min(0.4 * n_ctry, 22)
if(reg %in% c("Non-programme", "Gavi 57")) { 
  heatmap_height <- 22
} else if (reg %in% c("ROSA", "Central Africa", "Southern Africa", "AFRO-Central", "SEAR")) {
  heatmap_height <- 6
} else if (reg == "Northern Africa") {
  heatmap_height <- 4
}

```

```{r heatmap-dtp1, echo=FALSE, warning=FALSE, fig.width=10, fig.height=heatmap_height, dpi=300}
print(heatmap_dtp1)
```

```{r map_dtp1, echo=FALSE, warning=FALSE, fig.width=10, fig.height=6.5, dpi=300}
print(map_dtp1)
```

```{r line_dtp1, echo=FALSE, warning=FALSE, fig.width=10, fig.height=6.5, dpi=300}
print(line_plt_dtp1)
```

<!-- ===================== DTP3 ===================== -->

```{r setup-dtp3, include=FALSE}
v <- "dtp3"
unvac_v <- unvac %>% filter(vaccine == v) 
cvg_v   <- cvg %>% filter(vaccine == v)
vaccinated_v  <- vacc %>% filter(vaccine == v)

reg_cvg_rev_yr  <- cvg_v %>% filter(lvl_1 == "region") %>% pull(9)   
reg_cvg_comp_yr <- cvg_v %>% filter(lvl_1 == "region") %>% pull(7) 

unvac_ctry <- unvac_v %>% 
  arrange(desc(.[[9]])) %>% 
  filter(lvl_1 == "country") %>% 
  mutate(rank = seq(n())) %>% 
  slice_max(order_by = .[[9]], n = 2)
cvg_v <- cvg_v %>% 
  filter(vaccine == v, lvl_1 == "country") %>% 
  arrange(.[[9]]) %>% mutate(rank = dense_rank(.[[9]])) %>% 
  mutate(gt90 = ifelse(.[[9]] >= 90, 1, 0), n_gt90 = sum(gt90)) 
#cvg_v_no_nicaragua <- cvg_v %>% filter(country != "Nicaragua")

tgt_chng <- tgt %>% filter(lvl_1 == "region") %>% mutate(pop_grwth = round((.[[9]] / .[[7]] - 1) * 100, 0)) %>% label_vals_millions(.[[9]], "tgt_rev_yr") %>% label_vals_millions(.[[7]], "tgt_comp_yr")

if((reg_cvg_rev_yr < reg_cvg_comp_yr) & (tgt_chng[[9]] > tgt_chng[[7]])) {
  pop_chng_txt <- str_glue(get_text2("txt_pop_growth_context", text_vars), pop_grwth = tgt_chng$pop_grwth, comp_yr = comp_yr, rev_yr = rev_yr, ctry_cvg_rev_dtp3 = reg_cvg_rev_yr, ctry_cvg_comp_dtp3 = reg_cvg_comp_yr)
} else {
  pop_chng_txt <- ""
}

dropout_df <- df %>%
  filter(lvl_1 == "region", year %in% c(comp_yr, rev_yr), vaccine %in% c("dtp1","dtp3")) %>% select(lvl_1, lvl_2, Region, iso3c, country, year, vaccine, vaccinated) %>% pivot_wider(., names_from = vaccine, values_from = vaccinated) %>%
  mutate(dropout = round(((dtp1-dtp3)/dtp1) * 100, 0),
         dropout_txt = case_when(dropout < 5 ~ str_glue(get_text2("txt_dropout_low", text_vars), vax_pair="DTP"),
                                 dropout %in% c(5:10) ~ str_glue(get_text2("txt_dropout_medium", text_vars), vax_pair="DTP"),
                                 dropout > 5 ~ str_glue(get_text2("txt_dropout_high", text_vars), vax_pair="DTP")))

dropout_comp <- dropout_df %>%
  select(-c(dtp1, dtp3, dropout_txt)) %>%
  pivot_wider(names_from = year, values_from = dropout) %>%
  mutate(
    comp_pct = .[[6]],
    diff     = .[[7]] - .[[6]],
    diff_key = case_when(
      diff == 1  ~ "txt_dtp3_diff_higher_1",
      diff > 1   ~ "txt_dtp3_diff_higher_pp",
      diff == -1 ~ "txt_dtp3_diff_lower_1",
      diff < -1  ~ "txt_dtp3_diff_lower_pp",
      TRUE       ~ "txt_dtp3_diff_same"
    )
  )

dropout_diff_txt <- str_glue(
  get_text2(dropout_comp$diff_key, text_vars),
  diff     = abs(dropout_comp$diff),
  comp_yr  = comp_yr,
  comp_pct = dropout_comp$comp_pct
)

unvac_cvg <- unvac_v %>% arrange(desc(.[[9]])) %>% filter(lvl_1 == "country") %>% mutate(rank = seq(n())) %>% select(lvl_1:vaccine, rank, 9, 12) %>%
  rename_with(~"unvaccinated", .cols = 8) %>% rename_with(~"unvaccinated_lbl", .cols = 9) %>%
  left_join(., cvg_v %>% select(iso3c, vaccine, 9), by = c("iso3c", "vaccine")) %>% rename_with(~"coverage", .cols = 10)# %>% filter(country != "Nicaragua")
```

## `r t_lookup("txt_sec_dtp3", language)`

\vspace{0.2em}

```{r, results='asis', echo=FALSE}
cat(paste0("-   ", str_glue(
  get_text2("txt_dtp3_change_summary", text_vars),
  direction  = vaccinated_v %>% filter(lvl_1 == 'region') %>% pull(pchng_txt),
  pchng      = vaccinated_v %>% filter(lvl_1 == 'region') %>% pull(pchng) %>% abs(),
  comp_val   = vaccinated_v %>% filter(lvl_1 == 'region') %>% pull(vacc_comp_yr_lbl),
  comp_yr    = comp_yr,
  rev_val    = vaccinated_v %>% filter(lvl_1 == 'region') %>% pull(vacc_rev_yr_lbl),
  rev_yr     = rev_yr,
  pop_chng_txt = pop_chng_txt
), "\n"))
```

```{r, results='asis', echo=FALSE}
cvg_v   <- cvg %>% filter(vaccine == v)

dtp3_rev_cvg <- reg_cvg_rev_yr
dtp3_comp_cvg <- reg_cvg_comp_yr
dtp3_minusone_cvg <- cvg_v %>% filter(lvl_1 == "region") %>% pull(minusone_col)

dtp3_comp_dir <- if (dtp3_rev_cvg > dtp3_comp_cvg) {
  "higher"
} else if (dtp3_rev_cvg < dtp3_comp_cvg) {
  "lower"
} else {
  "same"
}

dtp3_minusone_dir <- if (dtp3_rev_cvg > dtp3_minusone_cvg) {
  "higher"
} else if (dtp3_rev_cvg < dtp3_minusone_cvg) {
  "lower"
} else {
  "same"
}

dtp3_cvg_key <- paste0("txt_dtp3_cvg_", dtp3_comp_dir, "_", dtp3_minusone_dir)

cat(paste0("-   ", str_glue(
  get_text2(dtp3_cvg_key, text_vars),
  rev_yr       = rev_yr,
  rev_pct      = dtp3_rev_cvg,
  comp_yr      = comp_yr,
  comp_pct     = dtp3_comp_cvg,
  minusone_yr  = minusone_col,
  minusone_pct = dtp3_minusone_cvg
), "\n"))
```

```{r, results='asis', echo=FALSE}
dropout_level_val <- dropout_df %>% filter(year == rev_yr) %>% pull(dropout)

txt_key <- if (dropout_level_val < 5) {
  "txt_dtp3_dropout_low"
} else if (dropout_level_val <= 10) {
  "txt_dtp3_dropout_medium"
} else {
  "txt_dtp3_dropout_high"
}

cat(paste0("-   ", str_glue(
  get_text2(txt_key, text_vars),
  rev_yr           = rev_yr,
  dropout_pct      = dropout_level_val,
  dropout_diff_txt = dropout_diff_txt
), "\n"))
```

```{r, results='asis', echo=FALSE}
cat(paste0("-   ", str_glue(
  get_text2("txt_dtp3_coverage_range", text_vars),
  rev_yr   = rev_yr,
  min_pct  = min(cvg_v[[9]], na.rm = TRUE),
  min_ctry = cvg_v %>% filter(.[[9]] == min(.[[9]], na.rm = TRUE)) %>% pull(country) %>% knitr::combine_words(),
  max_pct  = max(cvg_v[[9]], na.rm = TRUE),
  max_ctry = cvg_v %>% filter(.[[9]] == max(.[[9]], na.rm = TRUE)) %>% pull(country) %>% knitr::combine_words()
), "\n"))
```

```{r, results='asis', echo=FALSE}
cat(paste0("-   ", str_glue(
  get_text2("txt_dtp3_top2_unvac_summary", text_vars),
  ctry1  = unvac_cvg %>% filter(rank == 1) %>% pull(country),
  pct1   = unvac_cvg %>% filter(rank == 1) %>% pull(coverage) %>% unique(),
  unvac1 = unvac_cvg %>% filter(rank == 1) %>% pull(unvaccinated_lbl) %>% unique(),
  ctry2  = unvac_cvg %>% filter(rank == 2) %>% pull(country),
  pct2   = unvac_cvg %>% filter(rank == 2) %>% pull(coverage),
  unvac2 = unvac_cvg %>% filter(rank == 2) %>% pull(unvaccinated_lbl)
), "\n"))
```

\vspace{1em}

```{r ggplot-dtp3, echo=FALSE, warning=FALSE, fig.width=10, fig.height=heatmap_height, dpi=300}
print(heatmap_dtp3)
```

```{r ggplot_dtp3, echo=FALSE, warning=FALSE, fig.width=10, fig.height=4.5, dpi=300}
print(map_dtp3)
```

```{r line_dtp3, echo=FALSE, warning=FALSE, fig.width=10, fig.height=6.5, dpi=300}
print(line_plt_dtp3)
```

<!-- ===================== MCV ===================== -->

```{r setup-mcv, include=FALSE}
v <- c("mcv1", "mcv2")
unvac_v <- unvac %>% filter(vaccine %in% c(v))
cvg_v   <- cvg %>% filter(vaccine %in% c(v))

cvg_mcv1 <- cvg_v %>% filter(vaccine == "mcv1") %>%
  mutate(diff = .[[9]] - .[[7]],
         cvg_chng_txt = case_when(
           diff %in% c(-1:1) ~ get_text2("txt_cvg_stable", text_vars),
           diff < -1          ~ get_text2("txt_cvg_lower_than_baseline", text_vars),
           diff > 1           ~ get_text2("txt_cvg_higher_than_baseline", text_vars)
         ))

cvg_mcv1_reg <- cvg_mcv1 %>% filter(lvl_1 == "region")

cvg_mcv2 <- cvg_v %>% filter(lvl_1 == "region", vaccine == "mcv2") %>%
  mutate(diff = .[[9]] - .[[7]],
         cvg_chng_txt = case_when(
           diff %in% c(-1:1) ~ get_text2("txt_cvg_stable", text_vars),
           diff < -1          ~ get_text2("txt_cvg_decreased", text_vars),
           diff > 1           ~ get_text2("txt_cvg_increased", text_vars)
         ))

unvac_mcv <- unvac_v %>%
  filter(lvl_1 == "region") %>%
  select(lvl_1:vaccine, 9) %>%
  pivot_wider(names_from = vaccine, values_from = last_col()) %>%
  mutate(diff = mcv2 - mcv1) %>%
  label_vals_millions(mcv1, 'mcv1_lbl') %>%
  label_vals_millions(mcv2, 'mcv2_lbl') %>%
  label_vals_millions(diff, 'diff_lbl')

unvac_ctry <- unvac_v %>%
  #filter(country != "Nicaragua") %>% 
  filter(lvl_1 == "country", vaccine == "mcv1") %>%
  arrange(desc(.[[9]])) %>%
  mutate(rank = seq(n())) %>%
  slice_max(order_by = .[[9]], n = 1)

nyrs <- 3; thrshold <- 80

cvg_trends <- df %>%
  filter(vaccine == "mcv1", year %in% seq((rev_yr - nyrs + 1), rev_yr), lvl_1 == "country") %>%
  select(lvl_1, lvl_2, Region, iso3c, country, year, vaccine, coverage) %>%
  arrange(country) %>%
  mutate(ltx = ifelse(coverage < thrshold, 1, 0)) %>%
  group_by(country) %>%
  mutate(sum = sum(ltx)) %>%
  ungroup() %>%
  select(iso3c, country, sum) %>%
  distinct()# %>%
  #filter(country != "Nicaragua")

n_ctry_lower_x <- nrow(cvg_trends %>% filter(sum == nyrs))

dropout <- df %>%
  filter(lvl_1 == "region", year %in% c(comp_yr, rev_yr, minusone_col), vaccine %in% c("dtp1","mcv1")) %>%
  select(lvl_1, lvl_2, Region, iso3c, country, year, vaccine, vaccinated) %>%
  pivot_wider(names_from = vaccine, values_from = vaccinated) %>%
  mutate(dropout = round(((dtp1 - mcv1) / dtp1) * 100, 0),
         dropout_level = case_when(
           dropout < 5  ~ "low",
           dropout <= 10 ~ "medium",
           TRUE          ~ "high"
         ))

dropout_comp <- dropout %>%
  select(-c(dtp1, mcv1, dropout_level)) %>%
  pivot_wider(names_from = year, values_from = dropout) %>%
  mutate(
    diff     = .[[7]] - .[[6]],
    comp_pct = .[[6]],
    diff_key = case_when(
      diff == 1  ~ "txt_dropout_diff_higher_1",
      diff > 1   ~ "txt_dropout_diff_higher_pp",
      diff == -1 ~ "txt_dropout_diff_lower_1",
      diff < -1  ~ "txt_dropout_diff_lower_pp",
      TRUE       ~ "txt_dropout_diff_same"
    )
  )

dropout_diff_txt <- str_glue(
  get_text2(dropout_comp$diff_key, text_vars),
  diff     = abs(dropout_comp$diff),
  comp_yr  = comp_yr,
  comp_pct = dropout_comp$comp_pct
)

dropout_level_val <- dropout %>% filter(year == rev_yr) %>% pull(dropout_level)
dropout_txt_key <- paste0("txt_dropout_", dropout_level_val, "_mcv")
```

\newpage

## `r t_lookup("txt_sec_mcv", language)`

\vspace{0.2em}

```{r, results='asis', echo=FALSE}
mcv1_diff <- cvg_mcv1_reg[[9]] - cvg_mcv1_reg[[7]]
comp_dir <- if (mcv1_diff > 1) "higher" else if (mcv1_diff < -1) "lower" else "same"

mcv1_minusone_pct  <- cvg_mcv1_reg[[minusone_col]]
mcv1_minusone_diff <- cvg_mcv1_reg[[9]] - mcv1_minusone_pct
minusone_dir <- if (mcv1_minusone_diff > 1) "higher" else if (mcv1_minusone_diff < -1) "lower" else "same"

txt_key <- paste0("txt_mcv1_cvg_", comp_dir, "_", minusone_dir)

sentence <- str_glue(
  get_text2(txt_key, text_vars),
  rev_yr       = rev_yr,
  rev_pct      = cvg_mcv1_reg[[9]],
  comp_yr      = comp_yr,
  comp_pct     = cvg_mcv1_reg[[7]],
  minusone_yr  = minusone_col,
  minusone_pct = mcv1_minusone_pct
)

cat(paste0("-   ", sentence, "\n"))
```

```{r, results='asis', echo=FALSE}
mcv2_rev_cvg <- cvg_mcv2[[as.character(rev_yr)]]
mcv2_comp_cvg <- cvg_mcv2[[as.character(comp_yr)]]
mcv2_minusone_pct  <- cvg_mcv2[[minusone_col]]
mcv2_comp_dir <- if (mcv2_rev_cvg > mcv2_comp_cvg) {
  "higher"
} else if (mcv2_rev_cvg < mcv2_comp_cvg) {
  "lower"
} else {
  "same"
}
mcv2_minusone_dir <- if (mcv2_rev_cvg > mcv2_minusone_pct) {
  "higher"
} else if (mcv2_rev_cvg < mcv2_minusone_pct) {
  "lower"
} else {
  "same"
}
mcv2_txt_key <- paste0("txt_mcv2_cvg_", mcv2_comp_dir, "_", mcv2_minusone_dir)
sentence <- str_glue(
  get_text2(mcv2_txt_key, text_vars),
  mcv1_lbl     = unvac_mcv$mcv1_lbl,
  diff_lbl     = unvac_mcv$diff_lbl
)
cat(paste0("-   ", sentence, "\n"))

dropout_rev  <- dropout %>% filter(year == rev_yr) %>% pull(dropout)
dropout_comp <- dropout %>% filter(year == comp_yr) %>% pull(dropout)
dropout_minusone <- dropout %>% filter(year == minusone_col) %>% pull(dropout)

dropout_comp_diff <- dropout_rev - dropout_comp
comp_dir_drop <- if (dropout_comp_diff > 0) "higher" else if (dropout_comp_diff < 0) "lower" else "same"

dropout_minusone_diff <- dropout_rev - dropout_minusone
minusone_dir_drop <- if (dropout_minusone_diff > 0) "higher" else if (dropout_minusone_diff < 0) "lower" else "same"

diff_pp_abs <- abs(round(dropout_comp_diff, 0))
pp_suffix <- ifelse(diff_pp_abs == 1, "_1pp", "")
dropout_key <- paste0("txt_mcv_dropout_", comp_dir_drop, "_", minusone_dir_drop, pp_suffix)

dropout_sentence <- str_glue(
  get_text2(dropout_key, text_vars),
  rev_yr       = rev_yr,
  dropout_pct  = dropout_rev,
  comp_yr      = comp_yr,
  comp_pct     = dropout_comp,
  diff_pp      = diff_pp_abs,
  minusone_yr  = minusone_col,
  minusone_pct = dropout_minusone
)
cat(paste0("-   ", dropout_sentence, "\n"))
cat(paste0("-   ", get_text2(dropout_txt_key, text_vars), "\n"))
```

```{r, results='asis', echo=FALSE}
if (n_ctry_lower_x == 1) {
  cat(paste0("- ", str_glue(get_text2("txt_mcv_persistent_low_single", text_vars),
    nyrs = nyrs, list = knitr::combine_words(cvg_trends %>% filter(sum == nyrs) %>% pull(country)), thrshold = thrshold), "\n\n"))
} else if (n_ctry_lower_x > 1) {
  cat(paste0("- ", str_glue(get_text2("txt_mcv_persistent_low_multi", text_vars),
    nyrs = nyrs, count = n_ctry_lower_x, list = knitr::combine_words(cvg_trends %>% filter(sum == nyrs) %>% pull(country)), thrshold = thrshold), "\n\n"))
} else {
  cat(paste0("- ", str_glue(get_text2("txt_mcv_persistent_low_none", text_vars),
    nyrs = nyrs, thrshold = thrshold), "\n\n"))
}
```

```{r, results='asis', echo=FALSE}
cat(paste0("-   ", str_glue(
  get_text2("txt_mcv_top_unvac_ctry", text_vars),
  country  = unvac_ctry %>% pull(country),
  unvac_n  = unvac_ctry[[12]],
  unvac_pct = unvac_ctry %>% pull(unvac_ctry_pcnt),
  region   = unique(df$Region),
  mcv1_lbl = unvac_mcv %>% pull(mcv1_lbl)
), "\n"))
```

\vspace{1em}

```{r ggplot-mcv1, echo=FALSE, warning=FALSE, fig.width=10, fig.height=heatmap_height, dpi=300}
print(heatmap_mcv1)
```

```{r ggplot_mcv1, echo=FALSE, warning=FALSE, fig.width=10, fig.height=4.5, dpi=300}
print(map_mcv1)
```

```{r line_mcv1, echo=FALSE, warning=FALSE, fig.width=10, fig.height=6.5, dpi=300}
print(line_plt_mcv1)
```

<!-- ===================== HPV ===================== -->

```{r setup-hpv, include=FALSE}
if(nrow(hpv_df > 0)) {
  hpv <- hpv_df %>%
  select(lvl_1:vaccine_code, coverage) %>%
  filter(year %in% c(hpv_rev_yr, hpv_rev_yr - 1), lvl_1 == "region") %>%
  pivot_wider(names_from = year, values_from = coverage) %>%
  mutate(
    diff = .[[8]] - .[[7]],
    chng = case_when(diff > 0 ~ "increased", diff < 0 ~ "decreased", TRUE ~ "constant"),
    gt90 = case_when(
      .[[8]] < 80  & chng == "increased" ~ t_lookup("txt_hpv_gt90_below80_up", language),
      .[[8]] < 80  & chng != "increased" ~ t_lookup("txt_hpv_gt90_below80", language),
      .[[8]] >= 80 & .[[8]] <= 89 & chng == "increased" ~ t_lookup("txt_hpv_gt90_near_up", language),
      .[[8]] >= 80 & .[[8]] <= 89 & chng != "increased" ~ t_lookup("txt_hpv_gt90_near", language),
      .[[8]] == 90 ~ t_lookup("txt_hpv_gt90_achieved", language),
      .[[8]] > 90  ~ t_lookup("txt_hpv_gt90_exceeded", language),
      TRUE ~ NA_character_
    )
  )

hpv1  <- hpv %>% filter(vaccine_code == 'PRHPV1_F')
hpvc  <- hpv %>% filter(vaccine_code == 'PRHPVC_F')

hpv1_txt_key <- paste0("txt_hpv_hpv1_", hpv1$chng)
hpvc_txt_key <- paste0("txt_hpv_hpvc_", hpvc$chng)

hpv_unvac <- hpv_df %>%
  select(lvl_1:vaccine_code, unvaccinated) %>%
  filter(year %in% c(hpv_rev_yr), lvl_1 == "region") %>%
  pivot_wider(names_from = vaccine_code, values_from = unvaccinated) %>%
  mutate(diff = PRHPVC_F - PRHPV1_F) %>%
  label_vals_millions(PRHPVC_F, "PRHPVC_F_lbl") %>%
  label_vals_millions(PRHPV1_F, "PRHPV1_F_lbl") %>%
  label_vals_millions(diff, "diff_lbl")

# binary vars for if region has HPVc and HPV unvaccinated data
has_hpv <- ifelse(nrow(hpv) > 0, 1, 0)
has_hpvc <- ifelse(nrow(hpvc) > 0, 1, 0)
has_hpv_unvac <- ifelse(nrow(hpv_unvac) > 0, 1, 0)

n_ctry_in_reg <- df %>% filter(lvl_1 == "country") %>% pull(iso3c) %>% unique()
hpv_intro_y <- wiise_hpv_intro_yrs %>%
  filter(iso3c %in% c(n_ctry_in_reg)) %>%
  filter(nationwide == "yes" | partially == "yes") %>%
  filter(year <= rev_yr) %>%
  group_by(iso3c) %>%
  slice_max(year, n = 1, with_ties = FALSE) %>%
  ungroup()

n_ctry_intro       <- n_distinct(hpv_intro_y$iso3c)
n_ctry_total       <- length(n_ctry_in_reg)
pct_ctry_intro     <- round(n_ctry_intro / n_ctry_total * 100, 0)
hpv_missing_ctry   <- df %>% filter(lvl_1 == "country") %>% select(iso3c, country) %>% distinct() %>%
  anti_join(hpv_intro_y, by = "iso3c") %>% pull(country)
hpv_missing_flag   <- ifelse(length(hpv_missing_ctry) > 0, 1, 0)
n_ctry_wuenic_hpv  <- hpv_df %>% filter(lvl_1 == "country", year == hpv_rev_yr, vaccine_code == "PRHPVC_F") %>%
  summarise(n = n_distinct(iso3c)) %>% pull(n)
pct_ctry_wuenic_hpv <- round(n_ctry_wuenic_hpv / n_ctry_total * 100, 0)
}
```

\newpage

## `r t_lookup("txt_sec_hpv", language)`

\vspace{0.2em}

```{r, results='asis', echo=FALSE}
if(nrow(hpv_df) > 0) {
cat(paste0("-   ", str_glue(
  get_text2(hpv1_txt_key, text_vars),
  hpv1_comp_pct = hpv1[[7]],
  hpv1_rev_pct  = hpv1[[8]],
  hpv_comp_yr   = hpv_rev_yr - 1,
  hpv_rev_yr    = hpv_rev_yr
), "\n"))

cat(paste0("-   ", str_glue(
  get_text2(hpvc_txt_key, text_vars),
  hpvc_comp_pct = hpvc[[7]],
  hpvc_rev_pct  = hpvc[[8]],
  hpv_comp_yr   = hpv_rev_yr - 1,
  hpv_rev_yr    = hpv_rev_yr
), "\n"))

cat(paste0("-   ", hpvc$gt90, "\n"))

cat(paste0("-   ", str_glue(
  get_text2("txt_hpv_unvac_girls", text_vars),
  hpv_rev_yr = hpv_rev_yr,
  hpv1_lbl   = hpv_unvac$PRHPV1_F_lbl
), "\n"))
} else {
  cat(paste0("-   ", str_glue(get_text2("txt_hpv_no_data", text_vars)), "\n"))
}
```

```{r, results='asis', echo=FALSE}
if(nrow(hpv_df) > 0) {
if (hpv_missing_flag == 1) {
  bullet_1 <- str_glue(get_text2("txt_hpv_intro_partial", text_vars),
                       n_ctry_intro = n_ctry_intro, n_ctry_total = n_ctry_total, 
                       pct_ctry_intro = pct_ctry_intro, regn = regn, rev_yr = rev_yr)
  string_key <- ifelse(n_ctry_wuenic_hpv == n_ctry_intro, "txt_hpv_est_all_intro", "txt_hpv_est_partial")
  bullet_2 <- str_glue(get_text2(string_key, text_vars),
                       n_ctry_wuenic_hpv = n_ctry_wuenic_hpv, 
                       hpv_rev_yr = hpv_rev_yr)
  cat(paste0("- ", bullet_1, "\n\t- ", bullet_2, "\n\n"))

} else {
  bullet_1 <- str_glue(get_text2("txt_hpv_intro_all", text_vars),
                       n_ctry_total = n_ctry_total, regn = regn, rev_yr = rev_yr)
  bullet_2 <- str_glue(get_text2("txt_hpv_est_all", text_vars),
                       n_ctry_wuenic_hpv = n_ctry_wuenic_hpv, 
                       pct_ctry_wuenic_hpv = pct_ctry_wuenic_hpv, hpv_rev_yr = hpv_rev_yr)
  cat(paste0("- ", bullet_1, "\n\t- ", bullet_2, "\n\n"))
}

if (n_ctry_intro > n_ctry_wuenic_hpv) {
  hpv_country_df <- hpv_df %>% filter(country != regn)
  hpv_intro_no_wuenic <- hpv_intro_y %>%
    filter(!iso3c %in% (hpv_country_df %>% filter(lvl_1 == "country", year == hpv_rev_yr) %>% pull(iso3c))) %>%
    left_join(df %>% filter(lvl_1 == "country") %>% select(iso3c, country) %>% distinct(), by = "iso3c") %>%
    pull(country)
  hpv_intro_no_wuenic <- hpv_intro_no_wuenic[!is.na(hpv_intro_no_wuenic)]

  if (length(hpv_intro_no_wuenic) > 0) {
    if (length(hpv_intro_no_wuenic) == 1) {
      cat(paste0("    - ", str_glue(get_text2("txt_hpv_intro_no_data_single", text_vars),
        hpv_rev_yr = hpv_rev_yr, list = knitr::combine_words(hpv_intro_no_wuenic)), "\n\n"))
    } else {
      cat(paste0("    - ", str_glue(get_text2("txt_hpv_intro_no_data_multi", text_vars),
        count = length(hpv_intro_no_wuenic), hpv_rev_yr = hpv_rev_yr,
        list = knitr::combine_words(hpv_intro_no_wuenic)), "\n\n"))
    }
  }
}

if (hpv_missing_flag == 1) {
  if (length(hpv_missing_ctry) == 1) {
    cat(paste0("- ", str_glue(get_text2("txt_hpv_not_introduced_single", text_vars),
      list = knitr::combine_words(hpv_missing_ctry)), "\n\n"))
  } else {
    cat(paste0("- ", str_glue(get_text2("txt_hpv_not_introduced_multi", text_vars),
      count = length(hpv_missing_ctry), list = knitr::combine_words(hpv_missing_ctry)), "\n\n"))
  }
}
}  
```

```{r hpv-region-line-plot, echo=FALSE, warning=FALSE, message=FALSE}
if(nrow(hpv_df) > 0) {
plot_hpv1 <- hpv_df %>% filter(vaccine_code == "PRHPV1_F", country == filter_regn) %>% select(year, coverage)
plot_hpvc <- hpv_df %>% filter(vaccine_code == "PRHPVC_F", country == filter_regn) %>% select(year, coverage)
plot_zd_hpv <- hpv_df %>% filter(vaccine_code == "PRHPV1_F", country == filter_regn) %>% select(year, unvaccinated)
first_hpv_year <- min(c(plot_hpv1$year, plot_hpvc$year), na.rm = TRUE)
hpv_scale_factor <- 100 / max(plot_zd_hpv$unvaccinated, na.rm = TRUE)

if(has_hpvc == 0 & has_hpv_unvac == 0) {
  hpv_line_plot_title <- str_glue(get_text2("txt_hpv_line_title_no_hpvc_no_unvac", plot_vars))
} else if (has_hpvc == 0 & has_hpv_unvac == 1) {
  hpv_line_plot_title <- str_glue(get_text2("txt_hpv_line_title_no_hpvc_unvac", plot_vars))
} else if (has_hpvc == 1 & has_hpv_unvac == 0) {
  hpv_line_plot_title <- str_glue(get_text2("txt_hpv_line_title_hpvc_no_unvac", plot_vars))
} else if (has_hpvc == 1 & has_hpv_unvac == 1) {
  hpv_line_plot_title <- str_glue(get_text2("txt_hpv_line_title_hpvc_unvac", plot_vars))
}

if (has_hpv == 1) {
  hpv_region_plot <- ggplot() +
  geom_col(data = plot_zd_hpv, aes(x = year, y = unvaccinated, fill = get_text2("txt_plt_unvac_females", plot_vars)), alpha = 0.4, width = 0.6) +
  {if (nrow(plot_hpv1) > 1)
    list(
      geom_line(data=plot_hpv1, aes(x=year, y=coverage / hpv_scale_factor, 
                                    color=get_text2("txt_plt_hpv1_legend_cvg", plot_vars)), size=1.5, alpha = 0.7),
      geom_point(data=plot_hpv1, aes(x=year, y=coverage / hpv_scale_factor, 
                                     color=get_text2("txt_plt_hpv1_legend_cvg", plot_vars)), size=3, alpha = 0.7))} +
  {if (nrow(plot_hpvc) > 1)
    list(geom_line(data=plot_hpvc, aes(x=year, y=coverage / hpv_scale_factor, 
                                       color=get_text2("txt_plt_hpvc_legend_cvg", plot_vars)), size=1.5, alpha = 0.7),
         geom_point(data=plot_hpvc, aes(x=year, y=coverage / hpv_scale_factor, 
                                        color=get_text2("txt_plt_hpvc_legend_cvg", plot_vars)), size=3, alpha = 0.7))
    else list(geom_point(data=plot_hpvc, aes(x=year, y=coverage / hpv_scale_factor, 
                                             color=get_text2("txt_plt_hpvc_legend_cvg", plot_vars)), size=3, alpha = 0.7))} +
    geom_hline(yintercept = 90 / hpv_scale_factor, linetype = "dashed", alpha = 0.7) +
  scale_color_manual(name = NULL, values = setNames(c("#1CABE2", "#6A1E72"), 
                                                    c(get_text2("txt_plt_hpv1_legend_cvg", plot_vars), 
                                                      get_text2("txt_plt_hpvc_legend_cvg", plot_vars)))) +
  scale_fill_manual(name = NULL, values = setNames("#E2231A", get_text2("txt_plt_unvac_females", plot_vars))) +
  scale_x_continuous(breaks = seq(min(plot_hpv1$year), max(plot_hpv1$year), by = 1)) +
  scale_y_continuous(name = get_text2("txt_plt_unvac_females", plot_vars), labels = scales::label_comma(),
                     sec.axis = sec_axis(~ . * hpv_scale_factor, name = get_text2("txt_plt_cvg_pct", plot_vars), breaks = seq(0, 100, 25))) +
  guides(color = guide_legend(order = 1, nrow = 1),
         fill = guide_legend(order = 2, nrow = 1)) +
  labs(title = hpv_line_plot_title,
       caption = paste0(t_lookup("txt_plt_ia2030_caption", language), "\n", str_glue(get_text2("txt_wuenic_source", plot_vars))),
       x = get_text2("txt_plt_year", plot_vars)) +
  theme_unicef() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_line(color = "grey90"),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
        plot.title = element_text(size = 14, hjust = 0.5, face = "bold"),
        legend.position = "bottom",
        legend.box = "vertical")
}
}
```

```{r line_hpv, echo=FALSE, warning=FALSE, fig.width=10, fig.height=6.5, dpi=300}
if(nrow(hpv_df) > 0) {
if (has_hpv == 1) {
  print(hpv_region_plot)
}
}
```

\vspace{1em}

<!-- ===================== BIG CATCH-UP ===================== -->

```{r setup-bcu-ctry, include=FALSE}
# new methodology:
  # pandemic_impact is based only on the immediate pandemic years (2020, 2021) rather than all years through rev_yr
  # recovery status is computed relative to rev_yr

bcu_ctry <- c("afg","ben","bfa","bdi","cmr","caf","tcd","tgo","com","civ","cod","eth","gmb","gin","gnb","ken","kgz",
              "mdg","mli","mrt","moz","mmr","npl","ner","nga","pak","slb","som","ssd","sdn","syr","tjk","tza","yem","zmb")

n_bcu_ctry <- df %>% select(iso3c) %>% distinct() %>% summarise(count = sum(iso3c %in% bcu_ctry, na.rm = TRUE)) %>% pull(count)
if (n_bcu_ctry > 0) {
  rev_yr_chr <- as.character(rev_yr)
  recovery <- df %>%
    filter(lvl_1 == 'country', iso3c %in% bcu_ctry, vaccine == 'dtp3', year %in% c(2019:2021, rev_yr)) %>%
    select(iso3c, country, vaccine, year, coverage) %>%
    arrange(iso3c, vaccine, year) %>%
    pivot_wider(names_from = year, values_from = coverage) %>%
    # calculate difference in coverage 2020/2021 vs 2019, and rev_yr vs 2019
    mutate(
      diff_2019_2020    = `2020` - `2019`,
      diff_2019_2021    = `2021` - `2019`,
      diff_2019_revyr    = .data[[rev_yr_chr]] - `2019`
    ) %>%
    # identify largest decline (i.e. lowest difference) during the pandemic years only
    rowwise() %>%
    mutate(pandemic_impact = min(diff_2019_2020, diff_2019_2021)) %>%
    # assign pandemic impact in terms of decline
    mutate(pandemic_effect = case_when(
      pandemic_impact <= -5 ~ 'substantial decline',
      pandemic_impact < 0   ~ 'small decline',
      pandemic_impact >= 0  ~ 'no decline'
    )) %>%
    # determine recovery status relative to rev_yr
    mutate(recovery_status = case_when(
      diff_2019_revyr >= 0 ~ 'recovered or strengthened',
      diff_2019_revyr >= pandemic_impact + 3 ~ 'partially recovered',
      TRUE ~ 'lower than 2019 levels'
    )) %>%
    ungroup()
  
  recovery_summary <- recovery %>%
    group_by(pandemic_effect) %>%
    mutate(n = n_distinct(iso3c)) %>%
    ungroup() %>%
    mutate(pandemic_effect = paste0(pandemic_effect, ' (n=', n, ')')) %>%
    group_by(pandemic_effect, recovery_status) %>%
    summarise(n_countries = n(), .groups = "drop")
  
  n_no_decline <- recovery_summary %>% 
    filter(recovery_status == "recovered or strengthened", grepl("no decline", pandemic_effect)) %>% 
    summarise(total = sum(n_countries)) %>% pull(total)
  
  n_recovered  <- recovery_summary %>% 
    filter(recovery_status == "recovered or strengthened", !grepl("no decline", pandemic_effect)) %>% 
    summarise(total = sum(n_countries)) %>% pull(total)
  
  n_recovering <- recovery_summary %>% 
    filter(recovery_status == "partially recovered") %>% summarise(total = sum(n_countries)) %>% pull(total)
  
  n_no <- recovery_summary %>% 
    filter(recovery_status == "lower than 2019 levels") %>% summarise(total = sum(n_countries)) %>% pull(total)
  
  # function to get text for recovery status
  build_text <- function(count, key_single, key_multi, countries_vec) {
    if (count == 0) return(NULL)
    names_txt <- knitr::combine_words(countries_vec, and = get_text2("txt_word_and", text_vars))
    if (count == 1) {
      str_glue(get_text2(key_single, text_vars), country = names_txt)
    } else {
      str_glue(get_text2(key_multi, text_vars), count = count, countries_list = names_txt)
    }
  }
  recovery_txt1 <- build_text(n_recovered, "txt_bcu_recovered_single", "txt_bcu_recovered_multi",
                     recovery %>% filter(recovery_status == 'recovered or strengthened', grepl('decline', pandemic_effect)) %>% pull(country))
  recovery_txt2 <- build_text(n_recovering, "txt_bcu_recovering_single", "txt_bcu_recovering_multi",
                     recovery %>% filter(recovery_status == 'partially recovered') %>% pull(country))
  recovery_txt3 <- build_text(n_no, "txt_bcu_not_recovered_single", "txt_bcu_not_recovered_multi",
                     recovery %>% filter(recovery_status == 'lower than 2019 levels') %>% pull(country))
  recovery_txt4 <- build_text(n_no_decline, "txt_bcu_no_decline_single", "txt_bcu_no_decline_multi",
                     recovery %>% filter(recovery_status == 'recovered or strengthened', pandemic_effect == 'no decline') %>% pull(country))
}
```

```{r, results='asis', echo=FALSE}
if (n_bcu_ctry > 0) {

  cat("\\newpage\n\n")

  cat("## ", get_text2("txt_sec_bcu_header", text_vars), "\n\n", sep = "")

}
```

```{r, results='asis', echo=FALSE}
if (n_bcu_ctry > 0) {
  cat(paste0("-   ", str_glue(get_text2("txt_bcu_intro_summary", text_vars), regn = regn, n_bcu_ctry = n_bcu_ctry), "\n"))
}
```

```{r, results='asis', echo=FALSE}
if (n_bcu_ctry > 0 && !is.null(recovery_txt1) && nzchar(recovery_txt1)) {
  cat(paste0("    -   ", recovery_txt1, "\n"))
}
```

```{r, results='asis', echo=FALSE}
if (n_bcu_ctry > 0 && !is.null(recovery_txt2) && nzchar(recovery_txt2)) {
  cat(paste0("    -   ", recovery_txt2, "\n"))
}
```

```{r, results='asis', echo=FALSE}
if (n_bcu_ctry > 0 && !is.null(recovery_txt3) && nzchar(recovery_txt3)) {
  cat(paste0("    -   ", recovery_txt3, "\n"))
}
```

```{r, results='asis', echo=FALSE}
if (n_bcu_ctry > 0 && !is.null(recovery_txt4) && nzchar(recovery_txt4)) {
  cat(paste0("    -   ", recovery_txt4, "\n"))
}
```

```{r bcu-recovery-plot, echo=FALSE, warning=FALSE, fig.width=9, fig.height=max(4, n_bcu_ctry * 0.4 + 1.5), out.width="100%", dpi=300}
if (n_bcu_ctry > 0) {
  
  # ggh4x library - using this to dynamically adjust facet heights based on number of each countries with each status
  if (!requireNamespace("ggh4x", quietly = TRUE)) install.packages("ggh4x")
  
  bcu_plot_df <- recovery %>%
    select(country, iso3c, `2019`, all_of(rev_yr_chr), pandemic_effect, recovery_status) %>%
    rename(rev_yr_cvg = all_of(rev_yr_chr)) %>%
    mutate(
      recovery_label = case_when(
        recovery_status == "recovered or strengthened" ~ get_text2("txt_bcu_lbl_recovered", plot_vars),
        recovery_status == "partially recovered"        ~ get_text2("txt_bcu_lbl_recovering", plot_vars),
        recovery_status == "lower than 2019 levels"     ~ get_text2("txt_bcu_lbl_not_recovered", plot_vars)
      ),
      recovery_label = fct_relevel(recovery_label,
                        get_text2("txt_bcu_lbl_recovered", plot_vars),
                        get_text2("txt_bcu_lbl_recovering", plot_vars),
                        get_text2("txt_bcu_lbl_not_recovered", plot_vars)),
      country = fct_reorder(country, rev_yr_cvg)
    )

  # country counts in each recovery category
  facet_counts <- bcu_plot_df %>% 
    count(recovery_label, .drop = FALSE) %>% 
    arrange(recovery_label) %>% 
    pull(n)
  
  # avoid 0-weight issues if a category has no countries
  facet_height_ratios <- ifelse(facet_counts == 0, 0.1, facet_counts)

  status_colors <- setNames(
    c("#00833D", "#FFaf00", "#E2231A"),
    c(get_text2("txt_bcu_lbl_recovered", plot_vars), get_text2("txt_bcu_lbl_recovering", plot_vars), get_text2("txt_bcu_lbl_not_recovered", plot_vars))
  )

  gg_bcu <- ggplot(bcu_plot_df) +
    geom_segment(aes(x = `2019`, xend = rev_yr_cvg, y = country, yend = country, color = recovery_label), linewidth = 0.7, alpha = 0.5) +
    geom_point(aes(x = `2019`, y = country), shape = 21, fill = "white", color = "grey40", size = 2.5) +
    geom_point(aes(x = rev_yr_cvg, y = country, color = recovery_label), size = 3) +
    geom_vline(xintercept = 90, linetype = "dashed", color = "grey40", alpha = 0.7) +
    scale_color_manual(name = get_text2("txt_bcu_plot_status_name", plot_vars), values = status_colors) +
    scale_x_continuous(limits = c(0, 100), breaks = seq(0, 100, 20), labels = function(x) paste0(x, "%")) +
    facet_wrap(
      ~ recovery_label,
      ncol   = 1,
      scales = "free_y",
      labeller = labeller(recovery_label = label_wrap_gen(width = 20))
    ) +
    # apply dynamic heights here
    ggh4x::force_panelsizes(rows = facet_height_ratios) +
    labs(
      title    = str_glue(get_text2("txt_plt_bcu_title", plot_vars)),
      subtitle = str_glue(get_text2("txt_plt_bcu_subtitle", plot_vars)),
      x        = get_text2("txt_plt_bcu_xaxis", plot_vars),
      y        = NULL,
      caption = paste0(str_glue(get_text2("txt_plt_bcu_caption", text_vars), regn = regn), "\n", str_glue(get_text2("txt_wuenic_source", text_vars)))
    ) +
    theme_unicef() +
    theme(
      legend.position    = "none",
      strip.text = element_text(
        face = "bold",
        size = 10,
        margin = margin(t = 8, b = 8, l = 5, r = 5)
      ),
      strip.background   = element_rect(fill = "#0058AB", color = NA),
      strip.text.x        = element_text(color = "white"),
      panel.spacing      = unit(1.2, "lines"),
      panel.grid.major.y = element_blank(),
      panel.grid.major.x = element_line(color = "grey90"),
      panel.grid.minor   = element_blank(),
      axis.text.y        = element_text(size = 9),
      plot.subtitle      = element_text(size = 9, color = "grey40"),
      plot.title         = element_text(size = 13, face = "bold", hjust = 0.5)
    )

  print(gg_bcu)
}
```

```{r recovery-methodology, echo=FALSE, results='asis'}
if (n_bcu_ctry > 0) {

cat(
  paste0(
"### ", get_text2("txt_bcu_methodology_title", text_vars), "\n\n",
str_glue(get_text2("txt_bcu_methodology_intro", text_vars), rev_yr = rev_yr), "\n\n",
"- **", get_text2("txt_bcu_methodology_recovered_lbl", text_vars), ":** ",
str_glue(get_text2("txt_bcu_methodology_recovered_desc", text_vars), rev_yr = rev_yr), "\n",
"- **", get_text2("txt_bcu_methodology_recovering_lbl", text_vars), ":** ",
str_glue(get_text2("txt_bcu_methodology_recovering_desc", text_vars), rev_yr = rev_yr), "\n",
"- **", get_text2("txt_bcu_methodology_not_recovered_lbl", text_vars), ":** ",
str_glue(get_text2("txt_bcu_methodology_not_recovered_desc", text_vars), rev_yr = rev_yr), "\n\n"
  )
)

}
```

<!-- ===================== APPENDIX ===================== -->

\newpage

## `r t_lookup("txt_sec_appendix", language)`

### `r t_lookup("txt_appendix_wuenic_header", language)`

`r t_lookup("txt_appendix_wuenic_body", language)`

### `r t_lookup("txt_appendix_methods_header", language)`

- [Burton et al. 2009. WHO and UNICEF estimates of national infant immunization coverage: methods and processes](https://pubmed.ncbi.nlm.nih.gov/19649368/)
- [Burton et al. 2012. A formal representation of the WHO and UNICEF estimates of national immunization coverage: a computational logic approach](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0047806)
- [Brown et al. 2013. An introduction to the grade of confidence used to characterize uncertainty around the WHO and UNICEF estimates of national immunization coverage](https://openpublichealthjournal.com/VOLUME/6/PAGE/73/ABSTRACT/)

`r str_glue(get_text2("txt_appendix_wuenic_revision", text_vars))`

### `r t_lookup("txt_appendix_definitions_header", language)`

```{r appendix-definitions, results='asis', echo=FALSE}
# definitions are stored as a named list in txt_vars_talking_points.R
# and retrieved here so they translate properly
defs <- list(
  t_lookup("txt_def_bcg",   language),
  t_lookup("txt_def_hepbb", language),
  t_lookup("txt_def_dtp",   language),
  t_lookup("txt_def_hepb3", language),
  t_lookup("txt_def_hib3",  language),
  t_lookup("txt_def_ipv",   language),
  t_lookup("txt_def_mcv",   language),
  t_lookup("txt_def_rotac", language),
  t_lookup("txt_def_pcv3",  language),
  t_lookup("txt_def_yfv",   language),
  t_lookup("txt_def_menga", language),
  t_lookup("txt_def_hpv",   language)
)

for (d in defs) {
  cat("- ", d, "\n", sep = "")
}
```

### `r t_lookup("txt_appendix_resources_header", language)`

<https://data.unicef.org/resources/immunization/>