---
title: "WUENIC Talking Points: `r params$region`"
output:
  pdf_document:
    latex_engine: xelatex
geometry: "left=1cm,right=1cm,top=1.1cm,bottom=1.1cm"
fontsize: 12pt
params:
  region: "Default"
header-includes:
  - \usepackage{enumitem}
  - \setlist[itemize,1]{label=\textbullet}  # lines 12-15 are to specify point type when using tiered bullets
  - \setlist[itemize,2]{label=\(\circ\)}   
  - \setlist[itemize,3]{label=\(\infty\)} 
  - \setmainfont{Arial}
  - \usepackage{hyperref}
  - \hypersetup{colorlinks=true, linkcolor=blue, urlcolor=blue, breaklinks=true}
---


```{r setup, include=FALSE}
options(scipen=999)

# Filter based on selected region
df <- wuenic_dta %>% filter(Region == params$region)
hpv_df <- hpv_dta %>% filter(Region == params$region)
# df <- wuenic_dta %>% filter(Region == "MENA")
# hpv_df <- hpv_dta %>% filter(Region == "MENA")

## vars
regn       <- unique(df$Region)
wuenic_src <- str_glue("Source: WHO/UNICEF Estimates of National Immunization Coverage, {rev_yr} revision")
reg_type   <- "region"
```

```{r setup-vax, include=FALSE}
# number of countries in region
nctry <- n_distinct(df$iso3c) - 1

# key vaccines and years
vax <- df %>% 
  filter(vaccine %in% c("dtp1","dtp3","mcv1","mcv2"),
         year %in% c(comp_yr, rev_yr-1, rev_yr)) %>%
    arrange(desc(lvl_1), country, vaccine, year)

# unvaccinated data labels
unvac_lbl <- vax %>%
    select(lvl_1, lvl_2, Region, iso3c, country, year, vaccine, unvaccinated_lbl) %>%
  pivot_wider(., names_from = year, values_from = unvaccinated_lbl) %>%
  rename_with(~ paste0(., "_lbl"), .cols = (ncol(.)-2):ncol(.))  # add "_lbl" suffix to last three columns

# unvaccinated raw values
unvac <- vax %>%
    select(lvl_1, lvl_2, Region, iso3c, country, year, vaccine, unvaccinated) %>%
  pivot_wider(., names_from = year, values_from = unvaccinated) %>%
  # join rounded labels
  left_join(., unvac_lbl) %>%
  distinct() %>%
  mutate(pchng = round((.[[9]]/.[[7]] - 1) * 100, 0),
         pchng_txt = case_when(.[[10]] == .[[12]] ~ str_glue("remained the relatively stable between {comp_yr} and {rev_yr} at {.[[12]]}"),
                               pchng < 0 ~ str_glue("declined {abs(pchng)}% from {.[[10]]} in {comp_yr} to {.[[12]]} in {rev_yr}"),
                               pchng > 0 ~ str_glue("increased {pchng}% from {.[[10]]} in {comp_yr} to {.[[12]]} in {rev_yr}"),
                               TRUE ~ NA_character_)) %>%
  # unvaccinated in region
  mutate(unvac_rgn = ifelse(lvl_1 == "region", .[[9]], NA_real_)) %>%
  group_by(Region, vaccine) %>%
  fill(unvac_rgn, .direction = "updown") %>%
  ungroup() %>%
  # pcnt of unvacc in each country
  mutate(unvac_ctry_pcnt = round(.[[9]] / unvac_rgn * 100, 0)) %>%
  arrange(desc(lvl_1), vaccine, desc(unvac_ctry_pcnt)) %>%
  group_by(lvl_1, Region, vaccine) %>%
  # cumulative sum of pct of regional unvac
  mutate(unvac_ctry_pcnt_cumsum = cumsum(unvac_ctry_pcnt)) %>%
  ungroup() 

# vaccinated
vacc <- vax %>%
  select(lvl_1, lvl_2, Region, iso3c, country, year, vaccine, vaccinated) %>%
  pivot_wider(., names_from = year, values_from = vaccinated) %>%
  mutate(pchng = round((.[[9]] / .[[7]] - 1) * 100, 1),
         pchng_txt = case_when(pchng > 0 ~ "increased",
                               pchng < 0 ~ "decreased",
                               TRUE ~ "same")) %>%
  label_vals_millions(.[[7]], 'vacc_comp_yr_lbl') %>%
  label_vals_millions(.[[9]], 'vacc_rev_yr_lbl')


# coverage   
cvg <- vax %>%
  select(lvl_1, lvl_2, Region, iso3c, country, year, vaccine, coverage) %>%
  pivot_wider(., names_from = year, values_from = coverage)

# target pop (si)
tgt <- vax %>%
  filter(vaccine == "dtp1") %>%
    select(lvl_1, lvl_2, Region, iso3c, country, year, vaccine, target) %>%
  pivot_wider(., names_from = year, values_from = target) %>%
  mutate(diff = .[[9]] - .[[7]],
         moreless = case_when(diff < 0 ~ "fewer",
                              diff > 0 ~ "more",
                              TRUE ~ "same"),
         to_vax = case_when(diff < 0 ~ "for vaccine coverage to increase, the number of children vaccinated needs to either increase or at decline at a slower rate than the decline in surviving infant target population",
                              diff > 0 & diff < 1000000 ~ "vaccinating approximately the same number of children as currently reached will be sufficient due to the anticipated decrease or relatively small increases in the birth cohorts, and thus the number of children requiring vaccination.",
                              diff > 0 & diff >= 1000000 ~ "maintaining current coverage requires vaccinating an increasing number of children. To improve coverage and reach the 2030 ZD target will require substantial increases in immunization programme and health system capacity")) %>%
    label_vals(diff, 'diff_lbl') %>%
  mutate(diff_abs = abs(diff)) %>%
  label_vals_millions(diff_abs, 'diff_abs_lbl')
```

```{r setup-overview, include=FALSE}
# coverage >90% [region, any vaccine]
gt90_reg <- df %>%
  select(-target_grp) %>%
  distinct() %>%
  filter(lvl_1 == "region",
         year == rev_yr) %>%
  mutate(gt90 = case_when(coverage >= 90 ~ 1, 
                          TRUE ~ 0),
         gt90_sum = sum(gt90))

# coverage >90% [country, any vaccine]
gt90_ctry <- df %>%
  select(-target_grp) %>%
  distinct() %>%
  filter(lvl_1 == "country",
         year == rev_yr) %>%
  group_by(iso3c) %>%
  mutate(gt90 = case_when(coverage >= 90 ~ 1, 
                          TRUE ~ 0),
         gt90_sum = sum(gt90)) %>%
  ungroup() %>%
  mutate(ctry_anyvax_gt90 = ifelse(gt90_sum > 0, 1, 0)) %>%
  select(ctry_anyvax_gt90, iso3c, country) %>%
  distinct() %>%
  arrange(ctry_anyvax_gt90)

# number of countries that did not achieve 90% coverage in a single vaccine (returns 0 if all countries did)
nctry_lt90 <- gt90_ctry %>%
  filter(ctry_anyvax_gt90 == 0) %>%
  summarise(count = n()) %>%
  pull(count) %>%
  replace_na(0)

if (nctry_lt90 == 1) {
  nctry_lt90_txt <- str_glue("was {nctry_lt90} country")
} else if (nctry_lt90 > 1) {
    nctry_lt90_txt <- str_glue("were {nctry_lt90} countries")
} else {
  nctry_lt90_txt <- "were no countries"
}

# summary
gt90_ctry_summary <- gt90_ctry %>%
  group_by(ctry_anyvax_gt90) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  mutate(total = sum(n),
         pcnt = round(n / total * 100, 0))

# coverage change [prev_yr, rev_yr]
cvg_chng <- vax %>%
  filter(lvl_1 == "region",
         year %in% c(rev_yr, rev_yr-1)) %>%
  select(lvl_1:vaccine, coverage) %>%
  pivot_wider(., names_from = year, values_from = coverage) %>%
    mutate(diff = .[[8]] - .[[7]]) %>%
  mutate(diff_txt = case_when(diff > 0 ~ str_glue("increased from {.[[7]]}% in {rev_yr-1} to {.[[8]]}% in {rev_yr}"),
                              diff < 0 ~ str_glue("declined from {.[[7]]}% in {rev_yr-1} to {.[[8]]}% in {rev_yr}"),
                              TRUE ~ str_glue("was the same in {rev_yr} as it was in {rev_yr-1} ({.[[8]]}%)")))
```

```{r setup-progress-ia2030-zd-target, include=FALSE}
zd <- df %>%
  filter(vaccine == 'dtp1', 
         lvl_1=='region',
         year >= 2000) %>%
  mutate(measure = 'unvaccinated') %>%
  select(lvl_1:vaccine, measure, unvaccinated)

zd_target <- zd %>%
  filter(year == 2019) %>%
  mutate(unvaccinated = unvaccinated/2,
         measure = 'IA2030 zero-dose target',
         year = 2030)

# data for bars
zd2 <- zd %>%
  mutate(measure = 'Estimate') %>%
  rbind(zd_target) %>%
  mutate(chart = 'bar',
         unvaccinated = round(unvaccinated, 0))

# calculate annual target
target <- zd %>%
  filter(vaccine == 'dtp1',
         year %in% c(2019)) %>%
  select(lvl_1:country, vaccine, year, unvaccinated) %>%
  pivot_wider(., names_from = year, values_from = unvaccinated) %>%
  mutate(target = `2019`/2) %>%
  mutate(`2020` = `2019`) %>%
  mutate(factor = (`2020`-target)/(2030-2020)) %>%
  mutate(`2021` = `2020` - factor) %>%
  mutate(`2022` = `2021` - factor) %>%
  mutate(`2023` = `2022` - factor) %>%
  mutate(`2024` = `2023` - factor) %>%
  mutate(`2025` = `2024` - factor) %>%
  mutate(`2026` = `2025` - factor) %>%
  mutate(`2027` = `2026` - factor) %>%
  mutate(`2028` = `2027` - factor) %>%
  mutate(`2029` = `2028` - factor) %>%
  mutate(`2030` = `2029` - factor) %>%
  pivot_longer(-c(lvl_1:vaccine), names_to = 'year', values_to = 'zerodose_mt') %>%
  filter(year != 'factor', year != 'target') %>%
  mutate(year = as.numeric(year)) %>%
  select(lvl_1:year, unvaccinated_new = zerodose_mt) %>%
  left_join(., zd) %>%
  filter(year %in% c(2020:2030)) %>%
  mutate(diff = unvaccinated - unvaccinated_new,
         measure = "unvaccinated")  %>%
  select(lvl_1:year, measure, unvaccinated_new, unvaccinated, diff)

# data for lines
target_lines <- target %>%
  select(lvl_1:unvaccinated_new) %>%
  rename(unvaccinated = unvaccinated_new) %>%
  mutate(chart = 'line',
         measure = 'IA2030')

ia2030_2019 <- zd2 %>%
  filter(year == 2019) %>%
  mutate(chart = 'line',
         measure = 'IA2030')

# append
zd2 <- rbind(zd2, target_lines, ia2030_2019)  %>%
  filter(!(year == 2019 & chart == 'line'))

txt_dta <- zd2 %>%
  # rbind(target_lines) %>%
  filter(year == rev_yr) %>%
  select(Region, measure, unvaccinated) %>%
  pivot_wider(., names_from = measure, values_from = unvaccinated) %>%
  mutate(diff = Estimate - IA2030,
         diff_txt = case_when(diff > 0 ~ "higher than",
                              diff < 0 ~ "lower than",
                              diff == 0 ~ "same as"),
         pcnt_diff = round((Estimate / IA2030 - 1) * 100,0)) %>%
  label_vals_millions(Estimate, "Estimate_lbl") %>%
  label_vals_millions(IA2030, "IA2030_lbl") 
```

\textcolor{cyan}{\Large Overview}

\vspace{1em}

-   While noting variations across countries and vaccines, `r if (gt90_reg$gt90_sum[1] == 0) "none" else paste(gt90_reg$gt90_sum[1], "out")` of the `r n_distinct(gt90_reg$vaccine)` childhood vaccines achieved an average coverage of 90% or higher within the `r regn` region in `r rev_yr`.

\vspace{1em}

-   Coverage of the first dose of diphtheria-tetanus-pertussis-containing vaccine (DTP1) `r cvg_chng %>% filter(vaccine == 'dtp1') %>% pull(diff_txt)`, and the third dose of DTP-containing vaccine (DTP3) `r cvg_chng %>% filter(vaccine == 'dtp3') %>% pull(diff_txt)`. Coverage of the first dose of the vaccine against measles (MCV1) `r cvg_chng %>% filter(vaccine == 'mcv1') %>% pull(diff_txt)`.

\vspace{1em}

```{r, results='asis', echo=FALSE}
if (nctry_lt90 == 0) {
  cat(str_glue(
    "    -   At the country level, {gt90_ctry_summary %>% filter(ctry_anyvax_gt90 == 1) %>% pull(n)} out of {unique(gt90_ctry_summary$total)} ({gt90_ctry_summary %>% filter(ctry_anyvax_gt90 == 1) %>% pull(pcnt)}%) countries achieved coverage of 90% or higher in at least one of the {n_distinct(gt90_reg$vaccine)} childhood vaccines administered in the region.",
    "\n\\vspace{{1em}}\n"
  ))
} else {
cat(str_glue(
  "-   At the country level, {gt90_ctry_summary %>% filter(ctry_anyvax_gt90 == 1) %>% pull(n)} out of {unique(gt90_ctry_summary$total)} ({gt90_ctry_summary %>% filter(ctry_anyvax_gt90 == 1) %>% pull(pcnt)}%) countries achieved coverage of 90% or higher in at least one of the {n_distinct(gt90_reg$vaccine)} childhood vaccines administered in the region.",
  "\n\\vspace{{1em}}\n",
  "-   There {nctry_lt90_txt} that did not achieve at least 90% coverage in any vaccine ({gt90_ctry %>% filter(ctry_anyvax_gt90 == 0) %>% pull(country) %>% knitr::combine_words()}).",
  "\n\\vspace{{1em}}\n"
))
}
```

-   In `r rev_yr`, the number of zero-dose children (`r txt_dta$Estimate_lbl`) was approximately `r txt_dta %>% pull(pcnt_diff) %>% abs()`% `r txt_dta %>% pull(diff_txt)` the annual goal to achieve the Immunization Agenda 2030 target of reducing the number of zero-dose children by half by 2030 in `r regn` region (`r txt_dta$IA2030_lbl`).

\vspace{1em}


```{r setup-heatmap-reg, include=FALSE}
heatmap_reg_df <- df %>%
  rbind(hpv_df %>% filter(lvl_1 == "region", vaccine_code == "PRHPVC_F") %>% select(-vaccine_code)) %>%
  filter(lvl_1 == "region") %>%
  mutate(year = as.character(year),
         cvg_cat = case_when(coverage < 60 ~ '<60',
                             coverage %in% c(60:69) ~ '60-69',
                             coverage %in% c(70:79) ~ '70-79',
                             coverage %in% c(80:89) ~ '80-89',
                             coverage %in% c(90:94) ~ '90-94',
                             coverage >=95 ~ '>=95',
                             TRUE ~ NA_character_)) %>%
  mutate(vaccine = case_when(vaccine == "hepbb" ~ "HepBB",
                             vaccine == "hepb3" ~ "HepB3",
                             vaccine == "hib3" ~ "Hib3",
                             vaccine == "rotac" ~ "RotaC",
                             vaccine == "menga" ~ "MengA",
                             vaccine == "HPVc Females" ~ "HPVc",
                             TRUE ~ toupper(vaccine))) %>%
  group_by(country, vaccine) %>%
  mutate(mean = mean(coverage)) %>%
  ungroup()

heatmap_reg_df$vaccine <- factor(heatmap_reg_df$vaccine, levels = c("BCG", "HepBB", "DTP1", "DTP3", "Hib3", "HepB3", "PCVC", "RotaC", "IPV1", "IPVC", "MCV1", "RCV1", "MCV2", "YFV", "MengA","HPVc"))


plt_df <- heatmap_reg_df


heatmap_reg <-
  plt_df %>%
  ggplot(aes(x = year, y = vaccine, fill = coverage)) +
  geom_tile(color = "white") +
  theme_minimal() +
  geom_text(data = plt_df %>% filter(coverage>=60), aes(label = paste0(as.character(coverage))), colour='black', vjust = 0.5, size=3) +
  geom_text(data = plt_df %>% filter(coverage<60), aes(label = paste0(as.character(coverage))), colour='white', vjust = 0.5, size=3) +
  labs(x = '', y = '',
       title = str_glue("Vaccine coverage (%), by year, {regn}, {min(plt_df$year)}-{rev_yr}"),
       caption = str_glue("{wuenic_src}")) +
  scale_x_discrete(position = "top") +
  scale_y_discrete(limits=rev) +  # reverse y-axis
  theme(strip.placement = "outside",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text.x = element_text(size = 12),
        legend.position="bottom",
        legend.key.width = unit(1.2, "cm")) +
  scale_fill_gradientn(
    colours = c('#B50800', '#E2231A', '#F26A21', '#FFC20E', '#80BD41', '#00833D') ,
    values = scales::rescale(c(0,59,69,79,89,94, 100)),
    guide = "colorbar", limits=c(0,100),
    name = str_glue("Vaccine coverage (%)  "),
    breaks = c(0,59,69,79,89,100),
    labels = c("0", "60", "70", "80", "90", "100")
  )

```

```{r ggplot-allvax-reg, echo=FALSE, fig.width=10, fig.height=6.2}
print(heatmap_reg)
```

```{r setup-map-africa, include=FALSE}
map_dta <- df %>% 
  filter(year == rev_yr,
         vaccine %in% c('dtp1','dtp3','mcv1'),
         lvl_1 == 'country'
  ) %>%
  mutate(coverage = case_when(Region != regn ~ NA_real_,
                              TRUE ~ coverage),
         cvg_cat = case_when(coverage < 60 ~ '<60',
                             coverage %in% c(60:69) ~ '60-69',
                             coverage %in% c(70:79) ~ '70-79',
                             coverage %in% c(80:89) ~ '80-89',
                             coverage %in% c(90:94) ~ '90-94',
                             coverage >=95 ~ '>=95'))

# # for gavi and wb regions, add in missing countries in grey
# if (org %in% c('wb', 'wb_short', 'gavi')) {
#   # vaccines
#   other_options <- data.frame(vaccine = c("dtp1", "dtp3", "mcv1")) %>% mutate(join = 1)
#   
#   # other countries not in the region
#   other_countries <- wueniclatestrev %>%
#     filter(lvl_1 == "country") %>%
#     select(iso3c, country) %>%
#     distinct() %>%
#     anti_join(., map_dta) %>%
#     mutate(join = 1)
#   
#   other_countries <- left_join(other_countries, other_options) %>% distinct() %>% select(-join)
#   
#   map_dta <- bind_rows(map_dta, other_countries) %>%
#     mutate(
#       coverage = ifelse(is.na(coverage), 0, coverage),
#       cvg_cat = ifelse(coverage == 0, "Other countries", cvg_cat))
#   
# } else {
#   map_dta <- map_dta
# }


# join data to map df
map_df <- left_join(base_map_df, map_dta, by = "iso3c") %>%
  filter(!is.na(coverage))

for (v in c("dtp1","dtp3","mcv1")) {
  
  plt_df <- map_df %>% filter(vaccine == v)
  
  # plot
  gg <- ggplot() +
    geom_sf(data = plt_df, aes(fill = cvg_cat), alpha = 0.75, size = 0.5, color = "black") +
    coord_sf() +
    # set legend name, category order and colours
    scale_fill_manual(
      name = str_glue(''),
      breaks = c('>=95', '90-94', '80-89', '70-79', '60-69', '<60', 'Other countries'),
      values = c('<60' = '#B50800', 
                 '60-69' = '#E2231A',
                 '70-79' = '#F26A21',
                 '80-89' = '#FFC20E',
                 '90-94' = '#80BD41',
                 '>=95' = '#00833D',
                 'Other countries' = 'lightgrey')
    ) +
    theme_minimal(base_size = 12) +
    theme(axis.title = element_blank(),
          axis.text = element_blank(),
          axis.line = element_blank(),
          axis.ticks = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          plot.title = element_text(size=10)
          # legend.position = "bottom"
    ) +
    labs(title = str_glue("        {toupper(v)} coverage (%), {regn}, {rev_yr}"),
         caption = str_glue("Source: WHO/UNICEF estimates of national immunization coverage, {rev_yr} revision
                  Note: This map is stylized and based on an approximate scale.This map does not reflect a position\nby UNICEF or WHO on the legal status of any country or territory or the delimitation of any frontiers.")) 
  # guides(fill = guide_legend(title.position = "left", title.hjust = 0.5, nrow = 1)) 
  
  # assign the ggplot chart with a name based on the vaccine
  name <- paste0("map_", v, sep = "")
  assign(name , gg)
  
  
}
```
\vspace{1em} 
\vspace{1em} 

```{r ggplot_dtp3_africa, echo=FALSE, fig.width=10, fig.height=4.5}
if (regn %in% c("Africa", "Global")) {
  print(map_dtp3)
}
```

```{r setup-heatmap, include=FALSE}
## heatmap :: [dtp1, dtp3, mcv1, by year] ----
if (regn == "Africa") {
    by <- "region"
    
  heatmap_df <- wuenic_dta %>%
    filter(lvl_1 == "region",
           lvl_2 %in% c("region_au", "region_au_africa"),
           vaccine %in% c("dtp1", "dtp3", "mcv1"),
           year >= 2000) %>%
    select(lvl_1:vaccine, coverage)
  
} else if (regn == "Global") {
    by <- "region"
    
  heatmap_df <- wuenic_dta %>%
    filter(lvl_1 == "region",
           lvl_2 %in% c("region_unicef_ops", "region_unicef_global_old"),
           vaccine %in% c("dtp1", "dtp3", "mcv1"),
           year >= 2000) %>%
    select(lvl_1:vaccine, coverage)
  
} else {
  by <- "country"
  
  heatmap_df <- df %>%
  filter(vaccine %in% c("dtp1", "dtp3", "mcv1"),
         year >= 2000) %>%
  select(lvl_1:vaccine, coverage)
}


# set data labels and coverage category
heatmap_df <- heatmap_df %>%
  mutate(cvg_cat = case_when(coverage < 60 ~ '<60',
                             coverage %in% c(60:69) ~ '60-69',
                             coverage %in% c(70:79) ~ '70-79',
                             coverage %in% c(80:89) ~ '80-89',
                             coverage %in% c(90:94) ~ '90-94',
                             coverage >=95 ~ '>=95',
                             TRUE ~ NA_character_)) %>%
  group_by(country, vaccine) %>%
  mutate(rev_yr_cvg = ifelse(year == rev_yr, coverage, NA_real_)) %>%
  fill(rev_yr_cvg, .direction = "updown") %>%
  ungroup()

# text size
n_ctry <- heatmap_df %>% mutate(n = n_distinct(country)) %>% pull(n) %>% unique

if (regn == 'LACR') {
  txt_size <- 2.7
  yaxis_txt <- 9
} else if (n_ctry > 35) {
  txt_size <- 2
  yaxis_txt <- 7
} else if (regn == "ROSA") {
  txt_size <- 4
  yaxis_txt <- 13
} else {
  txt_size <- 3
  yaxis_txt <- 11
}

# region position
position <- heatmap_df %>% 
  filter(year == rev_yr) %>%
  arrange(vaccine, coverage, Region == regn) %>%
  group_by(vaccine) %>%
  mutate(cvg_position = seq(n())) %>%
  ungroup()

if (regn == "Africa") {
  position <- position %>%
    filter(Region == regn) %>%
    select(vaccine, cvg_position)
} else {
  position <- position %>%
    filter(lvl_1 == "region") %>%
    select(vaccine, cvg_position)
}

for (v in c('dtp1', 'dtp3', 'mcv1')) {
  
  plt_df <- heatmap_df %>% filter(vaccine == v) %>%
    mutate(country = ifelse(country == regn, str_glue("<b>{country}</b>"), country)) %>%
    # join on position of region
    left_join(position) %>%
    distinct()
  
  gg <- plt_df %>%
    ggplot(aes(x = year, y = reorder(country, -rev_yr_cvg), fill = coverage)) +
    geom_tile(color = "white") +
    labs(caption = str_glue("{wuenic_src}\nNote: Countries ordered based on descending {toupper(v)} coverage in {rev_yr}.")) +
    theme_minimal() +
    geom_text(data = plt_df %>% filter(coverage >= 60), aes(label = paste0(as.character(coverage))), 
              colour = 'black', vjust = 0.5, size = txt_size) +
    geom_text(data = plt_df %>% filter(coverage < 60), aes(label = paste0(as.character(coverage))), 
              colour = 'white', vjust = 0.5, size = txt_size) +  # Add data labels
    labs(x = '', y = '',
         title = str_glue('{toupper(v)} coverage, by {by}, {regn}, 2000-{rev_yr}')) +
    scale_x_continuous(breaks = seq(2000, rev_yr, by = 1), position = "top") +
    scale_y_discrete(limits = rev) +  # reverse y-axis
    theme(
      strip.placement = "outside",
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.text.y = ggtext::element_markdown(size = yaxis_txt),
      strip.text.x = element_text(size = 12),
      legend.position = "bottom",
      legend.key.width = unit(1.2, "cm"),
      axis.text.x = element_text(angle = 45, hjust = 0, vjust = 0.5, size = 11)
    ) +
    scale_fill_gradientn(
      colours = c('#B50800', '#E2231A', '#F26A21', '#FFC20E', '#80BD41', '#00833D'),
      values = scales::rescale(c(0, 59, 69, 79, 89, 94, 100)),
      guide = "colorbar", limits = c(0, 100),
      name = str_glue("Vaccine coverage (%)  "),
      breaks = c(0, 59, 69, 79, 89, 100),  # Specify where the labels should appear
      labels = c("0", "60", "70", "80", "90", "100")
    ) +
    geom_rect(aes(xmin = 2000 - 0.5, xmax = rev_yr + 0.5, ymin = unique(cvg_position) - 0.5, ymax = unique(cvg_position) + 0.5),
              fill = "transparent", color = "black", size = 1)
  
  # assign the ggplot chart with a name based on the vaccine
  name <- paste0("heatmap_", plt_df$vaccine, sep = "")
  assign(name , gg)
}
```

```{r setup-dtp1, include=FALSE}
v <- "dtp1"

unvac_v <- unvac %>% filter(vaccine == v) %>%
  mutate(ia2030tgt = round(.[[7]]/2, 0),
         ia2030tgt_pdiff = round((.[[9]]/ia2030tgt - 1) * 100),
         ia2030tgt_pdiff_hilo = case_when(ia2030tgt_pdiff > 0 ~ "higher",
                                          ia2030tgt_pdiff < 0 ~ "lower",
                                          TRUE ~ "same")) %>%
  label_vals(ia2030tgt, "ia2030tgt_lbl")

cvg_v <- cvg %>% filter(vaccine == v) 

# countries with most zd
zd_ctry <- unvac_v %>%
  arrange(desc(.[[9]])) %>%
  filter(lvl_1 == "country") %>%
  mutate(rank = seq(n())) %>%
  slice_max(order_by = .[[9]], n = 2)

cvg_v <- cvg %>%
  filter(vaccine == v, lvl_1 == "country") %>%
  arrange(.[[9]]) %>%
  mutate(rank = dense_rank(.[[9]])) %>%
  # Identify countries with cvg >= 90%
  mutate(gt90 = ifelse(.[[9]] >= 90, 1, 0),
         n_gt90 = sum(gt90)) %>%
  # Ensure `gt90_txt` has a consistent length
  mutate(gt90_txt = ifelse(
    n_gt90 == 0, 
    "There were no", 
    ifelse(n_gt90 < 10, 
           str_glue(" - {knitr::combine_words(unique(country[.[[9]] >= 90]))}"), 
           "")
  )) %>%
  mutate(n_gt90_txt = ifelse(
    n_gt90 == 1, 
    "There was 1", 
    ifelse(n_gt90 > 1, 
           str_glue("There were {n_gt90}"), 
           "")
  ))
  
```

\newpage 
\textcolor{cyan}{\Large DTP1: Reaching zero-dose children}

\vspace{1em} 

-   The number of children vaccinated with DTP1 `r unvac_v %>% filter(lvl_1 == 'region') %>% pull(pchng_txt)`. 

\vspace{1em} 

-   In `r rev_yr`, the number of zero-dose children (`r txt_dta$Estimate_lbl`) was approximately `r txt_dta %>% pull(pcnt_diff) %>% abs()`% `r txt_dta %>% pull(diff_txt)` the annual goal to achieve the Immunization Agenda 2030 target in `r regn` region (`r txt_dta$IA2030_lbl`).

\vspace{1em} 

-   With `r tgt %>% filter(lvl_1 == 'region') %>% pull(diff_abs_lbl)` `r tgt %>% filter(lvl_1 == 'region') %>% pull(moreless)` surviving infants in the region compared to in `r comp_yr`, `r tgt %>% filter(lvl_1 == 'region') %>% pull(to_vax)`. 

\vspace{1em} 

-   In `r rev_yr`, `r n_distinct(zd_ctry$iso3c)` countries, `r unique(zd_ctry$country) %>% knitr::combine_words()`, represented approximately `r zd_ctry %>% filter(rank == 1) %>% pull(unvac_ctry_pcnt)`% and `r zd_ctry %>% filter(rank == 2) %>% pull(unvac_ctry_pcnt)`% of all zero-dose children in `r params$region` (n = `r unvac_v %>% filter(lvl_1 == 'region') %>% pull(12)`), respectively. `r zd_ctry %>% filter(rank == 1) %>% pull(country)` had the highest number of zero-dose children at `r zd_ctry %>% filter(rank == 1) %>% pull(12)`, and `r zd_ctry %>% filter(rank == 2) %>% pull(country)` had the second highest at `r zd_ctry %>% filter(rank == 2) %>% pull(12)`. 

\vspace{1em} 

-   The lowest DTP1 coverage was observed in `r cvg_v %>% filter(rank == 1) %>% pull(country) %>% knitr::combine_words()` at `r cvg_v %>% filter(rank == 1) %>% pull(9) %>% unique()`%, followed by `r cvg_v %>% filter(rank == 2) %>% pull(country) %>% knitr::combine_words()` at `r cvg_v %>% filter(rank == 2) %>% pull(9) %>% unique()`%. 

\vspace{1em} 

```{r, results='asis', echo=FALSE}
# text for if no countries achieved at least 90% coverage
if (unique(cvg_v$n_gt90) == 0) {
cat(paste0("-   None of the ", nctry, " countries in ", regn, " achieved DTP1 coverage of 90% or higher in ", rev_yr, "."))
  
  # text for when at least one country achieved at least 90% coverage
} else {
  cat(paste0("-   ", unique(cvg_v$n_gt90_txt), " (", round(unique(cvg_v$n_gt90)/nctry*100,0), "%) out of ", nctry, " countries in the region that achieved DTP1 coverage of 90% or higher", cvg_v %>% pull(gt90_txt) %>% unique() , ".", "\\vspace{1em}\n\n"))
}
```

\vspace{1em} 

```{r, results='asis', echo=FALSE}
if (regn %in% c("Africa", "Global")) {
  reg_zd_txt <- heatmap_df
  cat(paste0("    -   ", "", "\n\\vspace{1em}\n"))
}
```

\vspace{1em} 
\vspace{1em} 

```{r ggplot-dtp1, echo=FALSE, fig.width=10, fig.height=6.5}
print(heatmap_dtp1)

```

```{r setup-dtp3, include=FALSE}
v <- "dtp3"
unvac_v <- unvac %>% filter(vaccine == v) 
cvg_v   <- cvg %>% filter(vaccine == v)
vacc_v  <- vacc %>% filter(vaccine == v)

# region coverage [rev_yr, comp_yr]
reg_cvg_rev_yr <- cvg_v %>% filter(lvl_1 == "region") %>% pull(9)   
reg_cvg_comp_yr <- cvg_v %>% filter(lvl_1 == "region") %>% pull(7) 

# countries with most unvacc
unvac_ctry <- unvac_v %>%
  arrange(desc(.[[9]])) %>%
  filter(lvl_1 == "country") %>%
  mutate(rank = seq(n())) %>%
  slice_max(order_by = .[[9]], n = 2)

cvg_v <- cvg_v %>%
  filter(vaccine == v,
         lvl_1 == "country") %>%
  arrange(.[[9]]) %>%
  mutate(rank = dense_rank(.[[9]])) %>%
  # countries with cvg >= 90%
  mutate(gt90 = ifelse(.[[9]] >= 90, 1, 0),
         n_gt90 = sum(gt90)) 

# text for if coverage declined but number vaccinated increased
tgt_chng <- tgt %>%
  filter(lvl_1 == "region") %>%
  mutate(pop_grwth = round((.[[9]] / .[[7]] - 1) * 100, 0)) %>%
  label_vals_millions(.[[9]], "tgt_rev_yr") %>%
  label_vals_millions(.[[7]], "tgt_comp_yr")

if((reg_cvg_rev_yr < reg_cvg_comp_yr) & (tgt_chng[[9]] > tgt_chng[[7]])) {
  pop_chng_txt <- str_glue("However, due to population growth - approximately a {tgt_chng$pop_grwth}% increase between {comp_yr} ({tgt_chng$tgt_comp_yr}) and {rev_yr} ({tgt_chng$tgt_rev_yr}) - DTP3 coverage in {rev_yr} ({reg_cvg_rev_yr}%) was lower than in {comp_yr} ({reg_cvg_comp_yr}%)")
} else {
  pop_chng_txt <- ""
}

# dtp dropout
dropout <- df %>%
  filter(lvl_1 == "region",
         year %in% c(comp_yr, rev_yr),
         vaccine %in% c("dtp1","dtp3")) %>%
  select(lvl_1, lvl_2, Region, iso3c, country, year, vaccine, vaccinated) %>%
  pivot_wider(., names_from = vaccine, values_from = vaccinated) %>%
  mutate(dropout = round(((dtp1-dtp3)/dtp1) * 100, 0),
         dropout_txt = case_when(dropout < 5 ~ "low DTP drop-out rates imply good ability to provide a complete series of vaccines early in life",
                                 dropout %in% c(5:10) ~ "medium DTP drop-out rates imply moderate ability to provide a complete series of vaccines early in life",
                                 dropout > 5 ~ "high DTP drop-out rates imply poor ability to provide a complete series of vaccines early in life"))

# dropout comparison to comp_yr
dropout_comp <- dropout %>%
  select(-c(dtp1, dtp3, dropout_txt)) %>%
  pivot_wider(., names_from = year, values_from = dropout) %>%
  mutate(diff = .[[7]] - .[[6]],
         diff_txt  = case_when(diff > 0 ~ str_glue("{diff} percentage points higher than in {comp_yr} ({.[[6]]}%)"),
                               diff < 0 ~ str_glue("{abs(diff)} percentage points lower than in {comp_yr} ({.[[6]]}%)"),
                               TRUE ~ str_glue("the same as in {comp_yr}")))

# combine unvaccinated and coverage
unvac_cvg <- unvac_v %>%
  arrange(desc(.[[9]])) %>%
  filter(lvl_1 == "country") %>%
  mutate(rank = seq(n())) %>%
  select(lvl_1:vaccine, rank, 9, 12) %>%
  # rename vars
  rename_with(~"unvaccinated", .cols = 8) %>%
  rename_with(~"unvaccinated_lbl", .cols = 9) %>%
  left_join(., cvg_v %>% select(iso3c, vaccine, 9), by = c("iso3c", "vaccine")) %>%
  rename_with(~"coverage", .cols = 10)

```
\newpage  
\textcolor{cyan}{\Large DTP3: A marker of routine immunization service delivery to children}

\vspace{1em} 

-   The number of children vaccinated with DTP3 `r vacc_v %>% filter(lvl_1 == 'region') %>% pull(pchng_txt)` from `r vacc_v %>% filter(lvl_1 == 'region') %>% pull(vacc_comp_yr_lbl)` in `r comp_yr` to `r vacc_v %>% filter(lvl_1 == 'region') %>% pull(vacc_rev_yr_lbl)` in `r rev_yr`. `r {pop_chng_txt}`. 

\vspace{1em} 

-   In `r rev_yr`, `r dropout %>% filter(year == rev_yr) %>% pull(dropout)`% of children who received DTP1 did not receive DTP3. DTP drop-out was `r dropout_comp %>% pull(diff_txt)`. The `r dropout %>% filter(year == rev_yr) %>% pull(dropout_txt)`. 

\vspace{1em} 

-   In `r rev_yr`, DTP3 coverage ranged from `r min(cvg_v[[9]], na.rm = TRUE)`% in `r cvg_v %>% filter(.[[9]] == min(.[[9]])) %>% pull(country) %>% knitr::combine_words()` to `r max(cvg_v[[9]], na.rm = TRUE)`% in `r cvg_v %>% filter(.[[9]] == max(.[[9]])) %>% pull(country) %>% knitr::combine_words()`. 

\vspace{1em} 

-   `r unvac_cvg %>% filter(rank == 1) %>% pull(country)` had DTP3 coverage of `r unvac_cvg %>% filter(rank == 1) %>% pull(coverage) %>% unique()`% and the highest absolute number of un- and undervaccinated children in the region (`r unvac_cvg %>% filter(rank == 1) %>% pull(unvaccinated_lbl) %>% unique()`), followed by `r unvac_cvg %>% filter(rank == 2) %>% pull(country)` with `r unvac_cvg %>% filter(rank == 2) %>% pull(coverage)`% coverage and `r unvac_cvg %>% filter(rank == 2) %>% pull(unvaccinated_lbl)` un- and undervaccinated children. 

\vspace{1em} 
\vspace{1em} 

```{r ggplot-dtp3, echo=FALSE, fig.width=10, fig.height=7}
print(heatmap_dtp3)

```

```{r setup-mcv, include=FALSE}
v <- c("mcv1", "mcv2")
unvac_v <- unvac %>% filter(vaccine %in% c(v)) 
cvg_v   <- cvg %>% filter(vaccine %in% c(v)) 

# mcv1 coverage [rev_yr, comp_yr]
cvg_mcv1 <- cvg_v %>%
  filter(vaccine == "mcv1") %>%
  mutate(diff = .[[9]] - .[[7]],
         cvg_chng_txt = case_when(diff %in% c(-1:1) ~ str_glue("remained relatively stable, within 1% of the {comp_yr} level"),
                                  diff < -1 ~ str_glue("was lower than in {comp_yr}"),
                                  diff > 1 ~ str_glue("was higher than in {comp_yr}")))

# mcv1 region coverage
cvg_mcv1_reg <- cvg_mcv1 %>% filter(lvl_1 == "region")

# mcv2 region coverage and unvacc
cvg_mcv2 <- cvg_v %>%
  filter(lvl_1== "region",
         vaccine == "mcv2") %>%
  mutate(diff = .[[9]] - .[[7]],
         cvg_chng_txt = case_when(diff %in% c(-1:1) ~ str_glue("remained relatively stable, within 1% of the {comp_yr} level"),
                                  diff < -1 ~ str_glue("decreased from {.[[7]]}% in {comp_yr} to {.[[9]]}% in {rev_yr}"),
                                  diff > 1 ~ str_glue("increased from {.[[7]]}% in {comp_yr} to {.[[9]]}% in {rev_yr}")))

# number unvaccinated in rev_yr [mcv1, mcv2]
unvac_mcv <- unvac_v %>%
  filter(lvl_1 == "region") %>%
  select(lvl_1:vaccine, 9) %>%
  pivot_wider(., names_from = vaccine, values_from = last_col()) %>%
  mutate(diff = mcv2 - mcv1) %>%
  label_vals_millions(mcv1, 'mcv1_lbl') %>%
  label_vals_millions(mcv2, 'mcv2_lbl') %>%
  label_vals_millions(diff, 'diff_lbl') 

# mcv2 coverage
unvac_cvg_mcv2 <- unvac %>%
  filter(lvl_1 == "region",
         vaccine %in% c("mcv1","mcv2")) %>%
  select(lvl_1:vaccine, 10, 12) %>%
  # rename vars
  rename_with(~"rev_yr_unvaccinated_lbl", .cols = 8) %>%
    rename_with(~"comp_yr_unvaccinated_lbl", .cols = 7) %>%
  left_join(., cvg_v %>% select(iso3c, vaccine, 9), by = c("iso3c", "vaccine")) %>%
  rename_with(~"coverage", .cols = 9)

# countries with most unvacc
unvac_ctry <- unvac_v %>%
    filter(lvl_1 == "country",
           vaccine == "mcv1") %>%
  arrange(desc(.[[9]])) %>%
  mutate(rank = seq(n())) %>%
  slice_max(order_by = .[[9]], n = 1)

# countries with coverage persistently below x%
nyrs <- 3

if (unique(df$Region) %in% c("ECAR", "LACR")) {
  thrshold <- 80
} else {
thrshold <- 80
}

cvg_trends <- df %>% 
  filter(vaccine == "mcv1",
         year %in% seq((rev_yr - nyrs + 1), rev_yr),
         lvl_1 == "country") %>%
    select(lvl_1, lvl_2, Region, iso3c, country, year, vaccine, coverage) %>%
  arrange(country) %>%
  # countries with cvg < x%
  mutate(ltx = ifelse(coverage < thrshold, 1, 0)) %>%
  group_by(country) %>%
  mutate(sum = sum(ltx)) %>%
  ungroup() %>%
  select(iso3c, country, sum) %>%
  distinct()
  # filter(sum == nyrs)

# number of countries with coverage below the threshold
n_ctry_lower_x <- max(cvg_trends$sum)


# dtp-mcv dropout
dropout <- df %>%
  filter(lvl_1 == "region",
         year %in% c(comp_yr, rev_yr),
         vaccine %in% c("dtp1","mcv1")) %>%
  select(lvl_1, lvl_2, Region, iso3c, country, year, vaccine, vaccinated) %>%
  pivot_wider(., names_from = vaccine, values_from = vaccinated) %>%
  mutate(dropout = round(((dtp1-mcv1)/dtp1) * 100, 0),
         dropout_txt = case_when(dropout < 5 ~ "low DTP-MCV drop-out rates imply good retention in immunization programmes and ability to provide a full course of vaccines in infancy (up to one year)",
                                 dropout %in% c(5:10) ~ "medium DTP-MCV drop-out rates imply moderate retention in immunization programmes and ability to provide a full course of vaccines in infancy (up to one year)",
                                 dropout > 5 ~ "high DTP-MCV drop-out rates imply poor retention in immunization programmes and ability to provide a full course of vaccines in infancy (up to one year)"))

# dtp-mcv dropout comparison to comp_yr
dropout_comp <- dropout %>%
  select(-c(dtp1, mcv1, dropout_txt)) %>%
  pivot_wider(., names_from = year, values_from = dropout) %>%
  mutate(diff = .[[7]] - .[[6]],
         diff_txt  = case_when(diff > 0 ~ str_glue("{diff} percentage points higher than in {comp_yr} ({.[[6]]}%)"),
                               diff < 0 ~ str_glue("{abs(diff)} percentage points lower than in {comp_yr} ({.[[6]]}%)"),
                               TRUE ~ str_glue("the same as in {comp_yr}")))


  
```
\newpage  
\textcolor{cyan}{\Large MCV: Canary in the coal mine}

\vspace{1em} 

-   MCV1 coverage in `r rev_yr` (`r cvg_mcv1_reg[[9]]`%) `r cvg_mcv1_reg$cvg_chng_txt` (`r cvg_mcv1_reg[[7]]`%). 

\vspace{1em} 

-   MCV2 coverage `r cvg_mcv2$cvg_chng_txt`. This still leaves `r unvac_mcv$mcv1_lbl` children without any protection and another `r unvac_mcv$diff_lbl` with only partial protection against the measles virus. 

\vspace{1em} 

-   In `r rev_yr`, `r dropout %>% filter(year == rev_yr) %>% pull(dropout)`% of children who received DTP1 did not receive MCV1. DTP-MCV drop-out was `r dropout_comp %>% pull(diff_txt)`. The `r dropout %>% filter(year == rev_yr) %>% pull(dropout_txt)`.

\vspace{1em} 

```{r, results='asis', echo=FALSE}
if (n_ctry_lower_x == 1) {
    n_ctry_low_txt <- "one country"
    cvg_trends <- cvg_trends %>% filter(sum == nyrs)
cat(str_glue(paste0("-   Over the last {nyrs} years, {n_ctry_low_txt} ({knitr::combine_words(cvg_trends$country)}) has consistently had MCV1 coverage below {thrshold}%", ".\n", "\\vspace{{1em}}\n")))
} else if (n_ctry_lower_x > 1) {
   n_ctry_low_txt <- str_glue("{n_ctry_lower_x} countries") 
   cvg_trends <- cvg_trends %>% filter(sum == nyrs)
   cat(str_glue(paste0("-   Over the last {nyrs} years, {n_ctry_low_txt} ({knitr::combine_words(cvg_trends$country)}) has consistently had MCV1 coverage below {thrshold}%", ".\n", "\\vspace{{1em}}\n")))
} else {
     cat(str_glue(paste0("-   Over the last {nyrs} years, no countries have had MCV1 coverage below {thrshold}%", ".\n", "\\vspace{{1em}}\n")))
}
```

-   `r unvac_ctry %>% pull(country)` had the highest number of children who did not receive the measles vaccine (`r unvac_ctry[[12]]`), representing approximately `r unvac_ctry %>% pull(unvac_ctry_pcnt)`% of all children unprotected against measles in `r unique(df$Region)` (`r unvac_mcv %>% pull(mcv1_lbl)`). 

\vspace{1em} 
\vspace{1em}

```{r ggplot-mcv1, echo=FALSE, fig.width=10, fig.height=7}
print(heatmap_mcv1)

```

```{r setup-hpv, include=FALSE}
hpv <- hpv_df %>%
  select(lvl_1:vaccine_code, coverage) %>%
  filter(year %in% c(hpv_rev_yr, hpv_rev_yr - 1),
         lvl_1 == "region") %>%
  pivot_wider(names_from = year, values_from = coverage) %>%
  mutate(diff = .[[8]] - .[[7]],
         chng = case_when(
           diff > 0 ~ "increased",
           diff < 0 ~ "decreased",
           TRUE ~ "remained constant"
         ),
         diff_txt = case_when(
           chng %in% c("increased", "decreased") ~ str_glue("{chng} from {.[[7]]}% in {hpv_rev_yr-1} to {.[[8]]}% in {hpv_rev_yr}"),
           TRUE ~ str_glue("{chng} at {.[[8]]}% between {hpv_rev_yr-1} and {hpv_rev_yr}")
         ),
         gt90 = case_when(
           .[[8]] < 80 & chng == "increased" ~ "Despite the important increases, HPVc coverage remains below the 90% target",
           .[[8]] < 80 & chng != "increased" ~ "HPVc coverage remains below the 90% target",
           .[[8]] >= 80 & .[[8]] <= 89  & chng == "increased" ~ "Despite the increase, HPVc coverage has not yet reached the 90% target",
           .[[8]] >= 80 & .[[8]] <= 89  & chng != "increased" ~ "HPVc coverage has not yet reached the 90% target",
           .[[8]] == 90 ~ "Coverage of the last dose of HPV vaccine achieved the 90% target",
           .[[8]] > 90 ~ "Coverage of the last dose of HPV vaccine has exceeded the 90% target",
           TRUE ~ NA_character_
         ))

# number girls unvaccinated with HPV1 and HPVc
hpv_unvac <- hpv_df %>%
  select(lvl_1:vaccine_code, unvaccinated) %>%
  filter(year %in% c(hpv_rev_yr),
         lvl_1 == "region") %>%
  pivot_wider(., names_from = vaccine_code, values_from = unvaccinated) %>%
  # calculate difference between HPVc and HPV1
  mutate(diff = PRHPVC_F - PRHPV1_F) %>%
  label_vals_millions(PRHPVC_F, "PRHPVC_F_lbl") %>%
  label_vals_millions(PRHPV1_F, "PRHPV1_F_lbl") %>%
  label_vals_millions(diff, "diff_lbl") 

# country iso codes in the region
n_ctry_in_reg <- df %>% filter(lvl_1 == "country") %>% pull(iso3c) %>% unique()

# filter hpv intro data for these countries
n_ctry_intro_f <- wiise_hpv_intro_yrs %>%
  filter(iso3c %in% c(n_ctry_in_reg) | iso3c == "nic") %>%
  distinct() %>%
  # count number of countries with nationwide or partial intro, by year
  filter(nationwide == "yes" | partially == "yes") %>%
  group_by(year) %>%
  mutate(n_ctry = n_distinct(iso3c)) %>%
  ungroup() %>%
  select(year, n_ctry) %>%
  distinct() %>%
  mutate(sex = "female")

# countries that have introduced hpv
hpv_intro_y <- wiise_hpv_intro_yrs %>%
  filter(year == rev_yr) %>%
  filter(nationwide == "yes" | partially == "yes") %>%
  filter(iso3c %in% c(n_ctry_in_reg) | iso3c == "nic")

# countries without hpv
hpv_missing_ctry <- anti_join(df %>% filter(lvl_1 == "country") %>% select(iso3c, country) %>% distinct(), hpv_intro_y[c("iso3c")]) %>% pull(country)

# flag for if there is at least one country in the region without hpv vaccine
hpv_missing_flag <- ifelse(length(hpv_missing_ctry) > 0, 1, 0)

```

\newpage  
\textcolor{cyan}{\Large HPV: }

\vspace{1em} 

-   Programme coverage of the first dose of HPV vaccine (HPV1) among girls `r hpv %>% filter(vaccine_code == 'PRHPV1_F') %>% pull(diff_txt)` and coverage of the last dose (HPVc) `r hpv %>% filter(vaccine_code == 'PRHPVC_F') %>% pull(diff_txt)`.

\vspace{1em}

-   `r hpv %>% filter(vaccine_code == 'PRHPVC_F') %>% pull(gt90)`.

\vspace{1em}

-   In `r rev_yr`, there were `r hpv_unvac$PRHPV1_F_lbl` girls that did not receive any HPV vaccination.

\vspace{1em}

```{r, results='asis', echo=FALSE}
if (hpv_missing_flag == 1) {
cat(paste0("-   There were ", n_ctry_intro_f %>% filter(year == rev_yr) %>% pull(n_ctry), " (", round(n_ctry_intro_f %>% filter(year == rev_yr) %>% pull(n_ctry) / length(n_ctry_in_reg) * 100, 0), "%) countries in ", regn , " that had introduced HPV vaccination by ", rev_yr, ".",  "\\vspace{1em}\n", 
           "-   The ", length(hpv_missing_ctry),"  countries that had not introduced HPV vaccination include: ", knitr::combine_words(hpv_missing_ctry), "\\vspace{1em}\n\n"))
} else {
  cat(paste0("-   All countries in ", regn , " had introduced HPV vaccination by ", rev_yr, "\\vspace{1em}\n\n"))
}
```

\vspace{1em}
\vspace{1em}
\vspace{1em}

```{r setup-bcu-ctry, include=FALSE}
bcu_ctry <- c("afg","ben","bfa","bdi","cmr","caf","tcd","tgo","com","civ","cod","eth","gmb","gin","gnb","ken","kgz","mdg","mli","mrt","moz","mmr","npl","ner","nga","pak","slb","som","ssd","sdn","syr","tjk","tza","yem","zmb")

# number of bcu countries in region
n_bcu_ctry <- df %>%
  select(iso3c) %>%
  distinct() %>%
  summarise(count = sum(iso3c %in% bcu_ctry, na.rm = TRUE)) %>%
  pull(count)

if (n_bcu_ctry > 0) {
recovery <- df %>%
  filter(lvl_1 == 'country',
         iso3c %in% c(bcu_ctry),
         vaccine == 'dtp3',
         year %in% c(2019:rev_yr)) %>%
  select(iso3c:vaccine, year, coverage) %>%
  arrange(iso3c, vaccine, year) %>%
  pivot_wider(., names_from = year, values_from = coverage) %>%
  # calculate difference in coverage 2020/2021 vs 2019
  mutate(diff_2019_2020 = `2020`-`2019`,
         diff_2019_2021 = `2021`-`2019`,
         diff_2019_2022 = `2022`-`2019`,
         diff_2019_2023 = `2023`-`2019`,
         diff_2019_2024 = `2024`-`2019`) %>%
  # identify largest decline (ie. lowest difference)
  rowwise() %>%
  mutate(pandemic_impact = min(diff_2019_2020, diff_2019_2021, diff_2019_2022, diff_2019_2023, diff_2019_2024)) %>%
  # assign pandemic impact in terms of decline
  mutate(pandemic_effect = case_when(pandemic_impact <= -5 ~ 'substantial decline',
                                     pandemic_impact < 0 ~ 'small decline',
                                     pandemic_impact >=0 ~ 'no decline')) %>%
  # determine coverage status
  mutate(recovery_status = case_when(diff_2019_2024 >= 0 ~ 'recovered',
                                     diff_2019_2024 >= pandemic_impact+3 ~ 'recovering',
                                     TRUE ~ 'not recovered'))

# recovery_summary <- recovery %>%
#   mutate(pandemic_effect = case_when(grepl("small|substantial", pandemic_effect) ~ "decline",
#                                      TRUE ~ "no decline")) %>%
#   # number of countries in each pandemic_effect group
#   group_by(pandemic_effect) %>%
#   mutate(n = n_distinct(iso3c)) %>%
#   ungroup() %>%
#   # mutate(pandemic_effect = paste0(pandemic_effect, ' (n=', n, ')')) %>%
#   # number of countries in each recovery_status group by pandemic_effect
#   group_by(pandemic_effect, recovery_status) %>%
#   summarise(n_countries = n()) %>%
#   # add in any missing categories and set number of countries to 0
#     complete(pandemic_effect = c("yes", "no"), fill = list(n_countries = 0)) %>%
#   complete(recovery_status = c("recovered", "recovering", "no"), fill = list(n_countries = 0))

recovery_summary <- recovery %>%
  mutate(pandemic_effect = case_when(
    grepl("small|substantial", pandemic_effect) ~ "decline",
    TRUE ~ "no decline"
  )) %>%
  # number of countries in each pandemic_effect group
  group_by(pandemic_effect, recovery_status) %>%
  summarise(n_countries = n(), .groups = "drop") %>%
  # add in any missing categories and set number of countries to 0
  complete(pandemic_effect = c("decline", "no decline"), 
           recovery_status = c("recovered", "recovering", "not recovered"), 
           fill = list(n_countries = 0)) 

# n countries in each recovery group
n_no_decline <- recovery_summary %>% filter(recovery_status == "recovered", grepl("no decline", pandemic_effect)) %>% pull(n_countries)
n_recovered <- recovery_summary %>% filter(recovery_status == "recovered", !grepl("no decline", pandemic_effect)) %>% pull(n_countries)
n_recovering <- recovery_summary %>% filter(recovery_status == "recovering", !grepl("no decline", pandemic_effect)) %>% pull(n_countries)
n_no <- recovery_summary %>% filter(recovery_status == "not recovered", !grepl("no decline", pandemic_effect)) %>% pull(n_countries)

# text for countries that recovered
if (n_recovered == 1) {
  recovery_txt1 <- str_glue("One country ({recovery %>% filter(recovery_status == 'recovered', !grepl('no decline', pandemic_effect)) %>% pull(country) %>% knitr::combine_words()}) has fully recovered to or exceeded their 2019 DTP3 levels.")
} else if (n_recovered > 1) {
  recovery_txt1 <- str_glue("{n_recovered} countries ({recovery %>% filter(recovery_status == 'recovered', !grepl('no decline', pandemic_effect)) %>% pull(country) %>% knitr::combine_words()}) have fully recovered to or exceeded their 2019 DTP3 levels.")
} else {
  recovery_txt1 <- ""
}

# text for countries that are recovering
if (n_recovering == 1) {
  recovery_txt2 <- str_glue("One country ({recovery %>% filter(recovery_status == 'recovering') %>% pull(country) %>% knitr::combine_words()}) is on the path to recovery to 2019 DTP3 levels.")
} else if (n_recovering > 1) {
  recovery_txt2 <- str_glue("{n_recovering} countries ({recovery %>% filter(recovery_status == 'recovering') %>% pull(country) %>% knitr::combine_words()}) are on the path to recovery to 2019 DTP3 levels.")
} else {
  recovery_txt2 <- ""
}

# text for countries that have not recovered
if (n_no == 1) {
  recovery_txt3 <- str_glue("One country ({recovery %>% filter(recovery_status == 'not recovered') %>% pull(country) %>% knitr::combine_words()}) has not yet recovered, with stagnation or decline from 2019 DTP3 levels.")
} else if (n_no > 1) {
  recovery_txt3 <- str_glue("{n_no} countries ({recovery %>% filter(recovery_status == 'not recovered') %>% pull(country) %>% knitr::combine_words()}) have not yet recovered, with stagnation or decline from 2019 DTP3 levels.")
} else {
  recovery_txt3 <- ""
}

# text for countries that had no decline during pandemic
if (n_no_decline == 1) {
  recovery_txt4 <- str_glue("The remaining country ({recovery %>% filter(recovery_status == 'recovered', grepl('no decline', pandemic_effect)) %>% pull(country) %>% knitr::combine_words()}) did not experience a decline in DTP3 levels compared to 2019.")
} else if (n_no_decline > 1) {
  recovery_txt4 <- str_glue("The remaining {n_no_decline} countries ({recovery %>% filter(recovery_status == 'recovered', grepl('no decline', pandemic_effect)) %>% pull(country) %>% knitr::combine_words()}) did not experience a decline in DTP3 coverage levels compared to 2019.")
} else {
  recovery_txt4 <- ""
}


recovery_txt_final <- str_glue("{recovery_txt1}\n\n{recovery_txt2}\n\n{recovery_txt3}\n\n{recovery_txt4}")
}
  
```


```{r, results='asis', echo=FALSE}
# title
if (n_bcu_ctry > 0) {
cat(paste0("\\textcolor{cyan}{\\Large Priority countries supported to recover from the COVID-19 pandemic}\n", "\\vspace{1em}\n"))
}
```

```{r, results='asis', echo=FALSE}
if (n_bcu_ctry > 0) {
cat(paste0("-   In ", regn, ", there are ", n_bcu_ctry, " out of the 35 countries globally supported for the Big Catch-up. Of these:", "\\vspace{1em}\n"))
}
```

```{r, results='asis', echo=FALSE}
# ensure that recovery_txt1 is only checked if it exists
if (exists("recovery_txt1") && n_bcu_ctry > 0 && recovery_txt1 != "") {
  cat(paste0("    -   ", recovery_txt1, "\n\\vspace{1em}\n"))
}
```
```{r, results='asis', echo=FALSE}
if (exists("recovery_txt2") && n_bcu_ctry > 0 && recovery_txt2 != "") {
  cat(paste0("    -   ", recovery_txt2, "\n\\vspace{1em}\n"))
}
```
```{r, results='asis', echo=FALSE}
if (exists("recovery_txt3") && n_bcu_ctry > 0 && recovery_txt3 != "") {
  cat(paste0("    -   ", recovery_txt3, "\n\\vspace{1em}\n"))
}
```

```{r, results='asis', echo=FALSE}
if (exists("recovery_txt4") && n_bcu_ctry > 0 && recovery_txt4 != "") {
  cat(paste0("    -   ", recovery_txt4, "\n\\vspace{1em}\n"))
}
```

\newpage  

\textcolor{cyan}{\Large Appendix }

\vspace{1em} 

\textcolor{cyan}{\Large WHO/UNICEF Estimates of National Immunization Coverage (WUENIC) }

\vspace{1em} 

Each year, WHO and UNICEF review national immunization data from Member States, including administrative and official coverage, survey estimates and contextual information, and use rules-based data triangulation approach to assess coverage. Estimates are produced independently, without borrowing data from other countries, and are not based on ad hoc adjustments. In some cases, they rely on a single source, typically nationally reported coverage. When data for a specific country-vaccine-year are unavailable, values from adjacent years are interpolated or extrapolated, and conflicting sources are assessed to determine the most reliable estimate, considering potential biases and expert opinion.

\vspace{0.5em} 
Methods:

\vspace{0.5em}  

\begin{itemize}
  \item \begin{sloppypar} 
        \textcolor{blue}{\href{https://pubmed.ncbi.nlm.nih.gov/19649368/}{Burton et al. 2009. WHO and UNICEF estimates of national infant immunization coverage: methods and processes}}
        \end{sloppypar}
  \item \begin{sloppypar} 
        \textcolor{blue}{\href{https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0047806}{Burton et al. 2012. A formal representation of the WHO and UNICEF estimates of national immunization coverage: a computational logic approach}}
        \end{sloppypar}
  \item \begin{sloppypar} 
        \textcolor{blue}{\href{https://openpublichealthjournal.com/VOLUME/6/PAGE/73/ABSTRACT/}{Brown et al. 2013. An introduction to the grade of confidence used to characterize uncertainty around the WHO and UNICEF estimates of national immunization coverage}}
        \end{sloppypar}
\end{itemize}

\vspace{0.5em}

These talking points present the latest WUENIC estimates (published 15 July `r rev_yr+1`).

\vspace{1.5em} 

\textcolor{cyan}{\fontsize{14}{16}\selectfont Definitions of immunization terms: }

\vspace{0.5em} 

-   Bacillus Calmette-Guerin (BCG): vaccine against tuberculosis
\vspace{0.5em}
-   Hepatitis B birth dose, given within 24 hours after birth (HepBB)
\vspace{0.5em}
-   Diphtheria, tetanus, and pertussis vaccine, first dose (DTP1) and third dose (DTP3)
\vspace{0.5em}
-   Hepatitis B vaccine, third dose (HepB3)
\vspace{0.5em}
-   Haemophilus influenzae type B vaccine, third dose (Hib3)
\vspace{0.5em}
-   Poliomyelitis vaccine, third dose (Polio3)
\vspace{0.5em}
-   Inactivated polio vaccine, first dose (IPV1) and second dose (IPV2): second dose is only shown for oral polio vaccine (OPV) using countries
\vspace{0.5em}
-   Measles containing vaccine, first dose (MCV1) and second dose (MCV2)
\vspace{0.5em}
-   Rotavirus vaccine, last dose (RotaC)
\vspace{0.5em}
-   Pneumococcal vaccine, third dose (PCV3)
\vspace{0.5em}
-   Yellow Fever vaccine (YFV) 
\vspace{0.5em}
-   Meningococcal A vaccine (MengA) 
\vspace{0.5em}
-   Human Papillomavirus vaccine, first dose (HPV1) and last dose (HPVc)

\vspace{1.5em} 

\textcolor{cyan}{\fontsize{14}{16}\selectfont Additional resources: } \textcolor{blue}{\href{https://data.unicef.org/resources/immunization/}{\underline{https://data.unicef.org/resources/immunization/}}}







