---
title: ""
output:
  pdf_document:
    latex_engine: xelatex
    toc: false
    toc_depth: 3
geometry: "left=1cm,right=1cm,top=1.1cm,bottom=1.1cm"
fontsize: 12pt
params:
  region: "Default"
header-includes:
  - \usepackage{enumitem}
  - \geometry{a4paper, left=15mm, right=15mm, top=15mm, bottom=20mm}
  - \setlist[itemize,1]{label=\textbullet}
  - \setlist[itemize,2]{label=\(\circ\)}
  - \setlist[itemize,3]{label=\(\infty\)}
  - \setmainfont{Arial}
  - \usepackage{hyperref}
  - \hypersetup{colorlinks=true, linkcolor=blue, urlcolor=blue, breaklinks=true}
  - \usepackage{tcolorbox}
  - \usepackage{fancyhdr}
  - \setlength{\headheight}{15pt}
  - \addtolength{\topmargin}{-3pt}
  - \definecolor{unicefblue}{RGB}{28, 171, 226}
  - \pagestyle{fancy}
  - \fancyhf{}
  - \fancyhead[L]{}
  - \fancyhead[C]{}
  - \fancyhead[R]{\hyperlink{top}{\small \color{black} $\uparrow$ Back to top}}
  - \fancyfoot[C]{\thepage}
  - \usepackage{tocloft}
  - \renewcommand{\cftsecfont}{\normalfont}
  - \renewcommand{\cftsecpagefont}{\normalfont}
  - \renewcommand{\cftsecleader}{\cftdotfill{\cftdotsep}}
---

\hypertarget{top}{}

```{r setup, include=FALSE}
options(scipen=999)

# Filter based on selected region
df <- wuenic_dta %>% filter(Region == params$region) %>% 
  filter(vaccine != "pol3") %>% 
  mutate(region_name = case_when(
    Region == "EAPR" ~ "East Asia and Pacific",
    Region == "ECAR" ~ "Eastern Europe and Central Asia",
    Region == "ESAR" ~ "Eastern and Southern Africa",
    Region == "LACR" ~ "Latin America and the Caribbean",
    Region == "MENA" ~ "Middle East and North Africa",
    Region == "Non-programme" ~ "Non-programme",
    Region == "ROSA" ~ "South Asia",
    Region == "WCAR" ~ "West and Central Africa",
    TRUE ~ Region
  ))

hpv_df <- hpv_dta %>% filter(Region == params$region)

## vars
regn       <- unique(df$Region)
regn_name  <- unique(df$region_name)
wuenic_src <- str_glue("Source: WHO/UNICEF Estimates of National Immunization Coverage, {rev_yr} revision")
reg_type   <- "region"
```

```{r banner, echo=FALSE, results='asis'}
cat(paste0('
\\begin{tcolorbox}[
  colback=unicefblue,
  colframe=unicefblue,
  arc=6pt,
  boxrule=0pt,
  left=4mm,
  right=4mm,
  top=3mm,
  bottom=3mm,
  halign=center,
  valign=center
]
\\begin{center}
  {\\LARGE \\color{white} \\textbf{WUENIC Talking Points: ', regn_name, ' (', regn, ') Region}}\\\\[0.4em]
  {\\large \\color{white} Revision Year: ', rev_yr, '}
\\end{center}
\\end{tcolorbox}
\\vspace{0.4in}
'))
```

```{r toc, echo=FALSE, results='asis'}
cat("\\tableofcontents")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}

# list of countries in this region
ctry_list <- df %>% 
  filter(lvl_1 == "country") %>%
  select(country) %>%
  distinct() %>%
  arrange(country) %>%
  pull(country)

map_df <- base_map_df %>%
  mutate(admin = case_match(admin,
    "United States of America" ~ "United States",
    "Vietnam"                  ~ "Viet Nam",
    "Ivory Coast"              ~ "Cote d'Ivoire",
    "Democratic Republic of the Congo" ~ "DRC",
    "Republic of the Congo"    ~ "Congo",
    "United Republic of Tanzania" ~ "Tanzania",
    "North Korea"              ~ "DPRK",
    "South Korea"              ~ "Republic of Korea",
    "Republic of Serbia"       ~ "Serbia",
    "Moldova"                  ~ "Republic of Moldova",
    "Palestine"                ~ "State of Palestine",
    "Turkey"                   ~ "Turkiye",
    "Russia"                   ~ "Russian Federation",
    "East Timor"               ~ "Timor-Leste",
    "Federated States of Micronesia" ~ "Micronesia",
    "Papua New Guinea"         ~ "PNG",
    "The Bahamas"              ~ "Bahamas",
    "Brunei"                   ~ "Brunei Darussalam",
    "eSwatini"                 ~ "Eswatini",
    "São Tomé and Principe"    ~ "Sao Tome and Principe",
    "Netherlands"              ~ "Netherlands (Kingdom of the)",
    .default = admin
  ))

# flag region
map_df <- map_df %>%
  mutate(highlight = ifelse(admin %in% ctry_list, "Region", "World"))

# plot
world_map <- ggplot(map_df) +
  geom_sf(aes(fill = highlight), color = "white", size = 0.2) +
  scale_fill_manual(values = c("Region" = "#1CABE2", "World" = "#d7d7d7"), name = NULL) +
  theme_void() +
  theme(legend.position = "none", 
        plot.title = element_markdown(hjust = 0.5, size = 14, face = "bold"),
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.5)) +
  labs(title = paste0("<span style='color:#1CABE2;'>", regn_name, " Region"))
```
\vspace{-40pt}
```{r ggplot-country-map, echo=FALSE, warning=FALSE, fig.align='center'}
suppressMessages(print(world_map))
```
\vspace{-60pt}
- The `r regn` region includes `r length(ctry_list)` countries (`r knitr::combine_words(ctry_list)`).

```{r nicaragua-note, echo=FALSE, results='asis'}
if(regn == "LACR") {
  cat("- Note: Data from Nicaragua is not shown individually, but it is included in aggregate measures in this report.")
}
```

```{r setup-vax, include=FALSE}
# number of countries in region
nctry <- n_distinct(df$iso3c) - 1

# key vaccines and years
vax <- df %>% 
  filter(vaccine %in% c("dtp1","dtp3","mcv1","mcv2"),
         year %in% c(comp_yr, rev_yr-1, rev_yr)) %>%
    arrange(desc(lvl_1), country, vaccine, year)

# unvaccinated data labels
unvac_lbl <- vax %>%
    select(lvl_1, lvl_2, Region, iso3c, country, year, vaccine, unvaccinated_lbl) %>%
  pivot_wider(., names_from = year, values_from = unvaccinated_lbl) %>%
  rename_with(~ paste0(., "_lbl"), .cols = (ncol(.)-2):ncol(.))

# unvaccinated raw values
unvac <- vax %>%
    select(lvl_1, lvl_2, Region, iso3c, country, year, vaccine, unvaccinated) %>%
  pivot_wider(., names_from = year, values_from = unvaccinated) %>%
  left_join(., unvac_lbl) %>%
  distinct() %>%
  mutate(pchng = round((.[[9]]/.[[7]] - 1) * 100, 0),
         pchng_txt = case_when(.[[10]] == .[[12]] ~ str_glue("remained the relatively stable between {comp_yr} and {rev_yr} at {.[[12]]}"),
                               pchng < 0 ~ str_glue("declined {abs(pchng)}% from {.[[10]]} in {comp_yr} to {.[[12]]} in {rev_yr}"),
                               pchng > 0 ~ str_glue("increased {pchng}% from {.[[10]]} in {comp_yr} to {.[[12]]} in {rev_yr}"),
                               TRUE ~ NA_character_)) %>%
  mutate(unvac_rgn = ifelse(lvl_1 == "region", .[[9]], NA_real_)) %>%
  group_by(Region, vaccine) %>%
  fill(unvac_rgn, .direction = "updown") %>%
  ungroup() %>%
  mutate(unvac_ctry_pcnt = round(.[[9]] / unvac_rgn * 100, 0)) %>%
  arrange(desc(lvl_1), vaccine, desc(unvac_ctry_pcnt)) %>%
  group_by(lvl_1, Region, vaccine) %>%
  mutate(unvac_ctry_pcnt_cumsum = cumsum(unvac_ctry_pcnt)) %>%
  ungroup() 

# vaccinated
vacc <- vax %>%
  select(lvl_1, lvl_2, Region, iso3c, country, year, vaccine, vaccinated) %>%
  pivot_wider(., names_from = year, values_from = vaccinated) %>%
  mutate(pchng = round((.[[9]] / .[[7]] - 1) * 100, 1),
         pchng_txt = case_when(pchng > 0 ~ "increased",
                               pchng < 0 ~ "decreased",
                               TRUE ~ "same")) %>%
  label_vals_millions(.[[7]], 'vacc_comp_yr_lbl') %>%
  label_vals_millions(.[[9]], 'vacc_rev_yr_lbl')

# coverage   
cvg <- vax %>%
  select(lvl_1, lvl_2, Region, iso3c, country, year, vaccine, coverage) %>%
  pivot_wider(., names_from = year, values_from = coverage)

# target pop (si)
tgt <- vax %>%
  filter(vaccine == "dtp1") %>%
    select(lvl_1, lvl_2, Region, iso3c, country, year, vaccine, target) %>%
  pivot_wider(., names_from = year, values_from = target) %>%
  mutate(diff = .[[9]] - .[[7]],
         moreless = case_when(diff < 0 ~ "fewer",
                              diff > 0 ~ "more",
                              TRUE ~ "same"),
         to_vax = case_when(diff < 0 ~ "for vaccine coverage to increase, the number of children vaccinated needs to either increase or at decline at a slower rate than the decline in surviving infant target population",
                              diff > 0 & diff < 1000000 ~ "vaccinating approximately the same number of children as currently reached will be sufficient due to the anticipated decrease or relatively small increases in the birth cohorts, and thus the number of children requiring vaccination",
                              diff > 0 & diff >= 1000000 ~ "maintaining current coverage requires vaccinating an increasing number of children. To improve coverage and reach the 2030 ZD target will require substantial increases in immunization programme and health system capacity")) %>%
    label_vals(diff, 'diff_lbl') %>%
  mutate(diff_abs = abs(diff)) %>%
  label_vals_millions(diff_abs, 'diff_abs_lbl')
```

```{r setup-overview, include=FALSE}
# coverage >90% [region, any vaccine]
gt90_reg <- df %>%
  select(-target_grp) %>%
  distinct() %>%
  filter(lvl_1 == "region",
         year == rev_yr) %>%
  mutate(gt90 = case_when(coverage >= 90 ~ 1, 
                          TRUE ~ 0),
         gt90_sum = sum(gt90))

# coverage >90% [country, any vaccine]
gt90_ctry <- df %>%
  select(-target_grp) %>%
  distinct() %>%
  filter(lvl_1 == "country",
         year == rev_yr) %>%
  group_by(iso3c) %>%
  mutate(gt90 = case_when(coverage >= 90 ~ 1, 
                          TRUE ~ 0),
         gt90_sum = sum(gt90)) %>%
  ungroup() %>%
  mutate(ctry_anyvax_gt90 = ifelse(gt90_sum > 0, 1, 0)) %>%
  select(ctry_anyvax_gt90, iso3c, country) %>%
  distinct() %>%
  arrange(ctry_anyvax_gt90)

# number of countries that did not achieve 90% coverage in a single vaccine
nctry_lt90 <- gt90_ctry %>%
  filter(country != "Nicaragua") %>%  # exclude Nicaragua from this count since it is not shown individually in the report
  filter(ctry_anyvax_gt90 == 0) %>%
  summarise(count = n()) %>%
  pull(count) %>%
  replace_na(0)

if (nctry_lt90 == 1) {
  nctry_lt90_txt <- str_glue("was {nctry_lt90} country")
} else if (nctry_lt90 > 1) {
    nctry_lt90_txt <- str_glue("were {nctry_lt90} countries")
} else {
  nctry_lt90_txt <- "were no countries"
}

# summary
gt90_ctry_summary <- gt90_ctry %>%
  group_by(ctry_anyvax_gt90) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  mutate(total = sum(n),
         pcnt = round(n / total * 100, 0))

# coverage change [prev_yr, rev_yr]
cvg_chng <- vax %>%
  filter(lvl_1 == "region",
         year %in% c(rev_yr, rev_yr-1)) %>%
  select(lvl_1:vaccine, coverage) %>%
  pivot_wider(., names_from = year, values_from = coverage) %>%
    mutate(diff = .[[8]] - .[[7]]) %>%
  mutate(diff_txt = case_when(diff > 0 ~ str_glue("increased from {.[[7]]}% in {rev_yr-1} to {.[[8]]}% in {rev_yr}"),
                              diff < 0 ~ str_glue("declined from {.[[7]]}% in {rev_yr-1} to {.[[8]]}% in {rev_yr}"),
                              TRUE ~ str_glue("was the same in {rev_yr} as it was in {rev_yr-1} ({.[[8]]}%)")))
```

```{r setup-progress-ia2030-zd-target, include=FALSE}
zd <- df %>%
  filter(vaccine == 'dtp1', 
         lvl_1 =='region',
         year >= 2000) %>%
  mutate(measure = 'unvaccinated') %>%
  select(lvl_1:vaccine, measure, unvaccinated)

zd_target <- zd %>%
  filter(year == 2019) %>%
  mutate(unvaccinated = unvaccinated/2,
         measure = 'IA2030 zero-dose target',
         year = 2030)

zd2 <- zd %>%
  mutate(measure = 'Estimate') %>%
  rbind(zd_target) %>%
  mutate(chart = 'bar',
         unvaccinated = round(unvaccinated, 0))

target <- zd %>%
  filter(vaccine == 'dtp1',
         year %in% c(2019)) %>%
  select(lvl_1:country, vaccine, year, unvaccinated) %>%
  pivot_wider(., names_from = year, values_from = unvaccinated) %>%
  mutate(target = `2019`/2) %>%
  mutate(`2020` = `2019`) %>%
  mutate(factor = (`2020`-target)/(2030-2020)) %>%
  mutate(`2021` = `2020` - factor) %>%
  mutate(`2022` = `2021` - factor) %>%
  mutate(`2023` = `2022` - factor) %>%
  mutate(`2024` = `2023` - factor) %>%
  mutate(`2025` = `2024` - factor) %>%
  mutate(`2026` = `2025` - factor) %>%
  mutate(`2027` = `2026` - factor) %>%
  mutate(`2028` = `2027` - factor) %>%
  mutate(`2029` = `2028` - factor) %>%
  mutate(`2030` = `2029` - factor) %>%
  pivot_longer(-c(lvl_1:vaccine), names_to = 'year', values_to = 'zerodose_mt') %>%
  filter(year != 'factor', year != 'target') %>%
  mutate(year = as.numeric(year)) %>%
  select(lvl_1:year, unvaccinated_new = zerodose_mt) %>%
  left_join(., zd) %>%
  filter(year %in% c(2020:2030)) %>%
  mutate(diff = unvaccinated - unvaccinated_new,
         measure = "unvaccinated")  %>%
  select(lvl_1:year, measure, unvaccinated_new, unvaccinated, diff)

target_lines <- target %>%
  select(lvl_1:unvaccinated_new) %>%
  rename(unvaccinated = unvaccinated_new) %>%
  mutate(chart = 'line',
         measure = 'IA2030')

ia2030_2019 <- zd2 %>%
  filter(year == 2019) %>%
  mutate(chart = 'line',
         measure = 'IA2030')

zd2 <- rbind(zd2, target_lines, ia2030_2019) %>%
  filter(!(year == 2019 & chart == 'line'))

txt_dta <- zd2 %>%
  filter(year == rev_yr) %>%
  select(Region, measure, unvaccinated) %>%
  pivot_wider(., names_from = measure, values_from = unvaccinated) %>%
  mutate(diff = Estimate - IA2030,
         diff_txt = case_when(diff > 0 ~ "higher than",
                              diff < 0 ~ "lower than",
                              diff == 0 ~ "same as"),
         pcnt_diff = round((Estimate / IA2030 - 1) * 100,0)) %>%
  label_vals_millions(Estimate, "Estimate_lbl") %>%
  label_vals_millions(IA2030, "IA2030_lbl") 
```

<!-- ===================== HEATMAP & LINE PLOT FOR EACH VACCINE ===================== -->

```{r setup-heatmap, include=FALSE}
if (regn == "Africa") {
    by <- "region"
    
  heatmap_df <- wuenic_dta %>%
    filter(lvl_1 == "region",
           lvl_2 %in% c("region_au", "region_au_africa"),
           vaccine %in% c("dtp1", "dtp3", "mcv1"),
           year >= 2000) %>%
    select(lvl_1:vaccine, coverage)
  
} else if (regn == "Global") {
    by <- "region"
    
  heatmap_df <- wuenic_dta %>%
    filter(lvl_1 == "region",
           lvl_2 %in% c("region_unicef_ops", "region_unicef_global_old"),
           vaccine %in% c("dtp1", "dtp3", "mcv1"),
           year >= 2000) %>%
    select(lvl_1:vaccine, coverage)
  
} else {
  by <- "country"
  
  heatmap_df <- df %>%
  filter(vaccine %in% c("dtp1", "dtp3", "mcv1"),
         year >= 2000) %>%
  select(lvl_1:vaccine, coverage)
}

heatmap_df <- heatmap_df %>%
  mutate(cvg_cat = case_when(coverage < 60 ~ '<60',
                             coverage %in% c(60:69) ~ '60-69',
                             coverage %in% c(70:79) ~ '70-79',
                             coverage %in% c(80:89) ~ '80-89',
                             coverage %in% c(90:94) ~ '90-94',
                             coverage >=95 ~ '>=95',
                             TRUE ~ NA_character_)) %>%
  group_by(country, vaccine) %>%
  mutate(rev_yr_cvg = ifelse(year == rev_yr, coverage, NA_real_)) %>%
  fill(rev_yr_cvg, .direction = "updown") %>%
  ungroup()

n_ctry <- heatmap_df %>% mutate(n = n_distinct(country)) %>% pull(n) %>% unique

position <- heatmap_df %>% 
  filter(year == rev_yr) %>%
  arrange(vaccine, coverage, Region == regn) %>%
  group_by(vaccine) %>%
  mutate(cvg_position = seq(n())) %>%
  ungroup()

if (regn == "Africa") {
  position <- position %>%
    filter(Region == regn) %>%
    select(vaccine, cvg_position)
} else {
  position <- position %>%
    filter(lvl_1 == "region") %>%
    select(vaccine, cvg_position)
}

if (regn == "LACR") {
  txt_size <- 2.7
  yaxis_txt <- 9
} else if (n_ctry > 35) {
  txt_size <- 2
  yaxis_txt <- 7
} else if (regn == "ROSA") {
  txt_size <- 4
  yaxis_txt <- 13
} else {
  txt_size <- 3
  yaxis_txt <- 11
}

# HEATMAP FOR EACH VACCINE
for (v in c('dtp1', 'dtp3', 'mcv1')) {
  
  plt_df <- heatmap_df %>% 
    filter(vaccine == v) %>%
    mutate(country = ifelse(country == regn, paste0("**", country, "**"), country)) %>%
    left_join(position) %>%
    distinct() %>% 
    filter(country != "Nicaragua")  # exclude Nicaragua from this plot since it is not shown individually in the report
  
  gg <- plt_df %>%
    ggplot(aes(x = year, y = reorder(country, -rev_yr_cvg), fill = coverage)) +
    geom_tile(color = "white") +
    geom_text(data = plt_df %>% filter(coverage > 80 | coverage < 95), aes(label = paste0(as.character(coverage))), 
              colour = 'black', vjust = 0.5, size = 3) +
    geom_text(data = plt_df %>% filter(coverage <= 80 | coverage >= 95), aes(label = paste0(as.character(coverage))), 
              colour = 'white', vjust = 0.5, size = 3) +
    geom_rect(aes(xmin = 2000 - 0.5, xmax = rev_yr + 0.5, ymin = unique(cvg_position) - 0.5, ymax = unique(cvg_position) + 0.5),
              fill = "transparent", color = "black", size = 1) +
    scale_x_continuous(breaks = seq(2000, rev_yr, by = 1), position = "top", expand = c(0, 0)) +
    scale_y_discrete(limits = rev, expand = c(0, 0)) +
    scale_fill_gradientn(
      colours = c('#B50800', '#E2231A', '#F26A21', '#FFC20E', '#80BD41', '#00833D'),
      values = scales::rescale(c(0, 59, 69, 79, 89, 94, 100)),
      guide = "colorbar", limits = c(0, 100),
      name = str_glue("Vaccine coverage (%)  "),
      breaks = c(0, 59, 69, 79, 89, 100),
      labels = c("0", "60", "70", "80", "90", "100")
    ) +
    theme_minimal() +
    theme(
      strip.placement = "outside",
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.text.y = ggtext::element_markdown(size = yaxis_txt),
      axis.text.x = element_text(size = 11, angle = 90, vjust = 0.5),
      strip.text.x = element_text(size = 12),
      legend.position = "bottom",
      legend.key.width = unit(1.2, "cm"),
      plot.title = element_text(size = 14, hjust = 0.5, face = "bold")
    ) +
    labs(
      x = '', y = '',
      title = str_glue('{toupper(v)} coverage, by {by}, {regn}, 2000-{rev_yr}'),
      caption = str_glue("{wuenic_src}\nNote: Countries ordered based on descending {toupper(v)} coverage in {rev_yr}.")
    )
  
  name <- paste0("heatmap_", plt_df$vaccine, sep = "")
  assign(name , gg)
}

# LINE PLOT FOR EACH VACCINE
for (v in c('dtp1', 'dtp3', 'mcv1')) {
  local({
      v <- v  # force capture of current loop value
      
      plt_vax_df <- df %>% 
        filter(country == regn, vaccine == v) %>%
        select(year, coverage, unvaccinated)
      
      vax_label <- paste0(toupper(v), " coverage")
      
      if (v == "dtp1") {
        count_label <- "Zero-dose children"
      } else if (v == "dtp3") {
        count_label <- "Un- and under-vaccinated children"
      } else if (v == "mcv1") {
        count_label <- "MCV zero-dose children"
      }
      
      max_cvg <- max(plt_vax_df$coverage, na.rm = TRUE)
      max_zd  <- max(plt_vax_df$unvaccinated, na.rm = TRUE)
      scale_factor <- ifelse(max_zd > 0, max_cvg / max_zd, 1)
      
      vax_values <- setNames("#1CABE2", vax_label)
    fill_values <- setNames("#E2231A", count_label)
  
    gg_line <- ggplot(plt_vax_df) +
      geom_col(aes(x = year, y = unvaccinated * scale_factor, fill = count_label),
               alpha = 0.4, width = 0.6) +
      geom_line(aes(x = year, y = coverage, color = vax_label), size = 0.8) +
      geom_point(aes(x = year, y = coverage, color = vax_label), size = 1.5) +
      geom_hline(yintercept = 90, linetype = "dashed", alpha = 0.7) +
      scale_color_manual(name = NULL, values = vax_values, breaks = vax_label) + 
      scale_fill_manual(name = NULL, values = fill_values, breaks = count_label) +
      guides(color = guide_legend(order = 1), fill = guide_legend(order = 2)) +
      scale_x_continuous(breaks = seq(2000, rev_yr, by = 2)) +
      scale_y_continuous(limits = c(0, 100), name = "Coverage (%)",
                         sec.axis = sec_axis(~ . / scale_factor, name = count_label,
                                             labels = scales::label_comma())) +
      labs(
        title = str_glue("{toupper(v)} coverage and {tolower(count_label)}, {regn}, 2000–{rev_yr}"),
        caption = str_glue("Dotted line shows the IA2030 target of 90% coverage\n{wuenic_src}"),
        x = "Year"
      ) +
      theme_unicef() + 
      theme(
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_line(color = "grey90"),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
        plot.title = element_text(size = 14, hjust = 0.5, face = "bold"),
        legend.position = "bottom"
      )
      
    plot_name <- paste0("line_plt_", v)
    assign(plot_name, gg_line, envir = .GlobalEnv)  # must specify envir when inside local()
  })
}
```

<!-- ===================== PAGE 1: OVERVIEW ===================== -->

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{Overview}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:overview}{}")
cat("\\addcontentsline{toc}{section}{Overview}")
```

\vspace{1em}

-   While noting variations across countries and vaccines, `r if (gt90_reg$gt90_sum[1] == 0) "none" else paste(gt90_reg$gt90_sum[1], "out")` of the `r n_distinct(gt90_reg$vaccine)` childhood vaccines achieved an average coverage of 90% or higher within the `r regn` region in `r rev_yr`.

-   Coverage of the first dose of diphtheria-tetanus-pertussis-containing vaccine (DTP1) `r cvg_chng %>% filter(vaccine == 'dtp1') %>% pull(diff_txt)`, and the third dose of DTP-containing vaccine (DTP3) `r cvg_chng %>% filter(vaccine == 'dtp3') %>% pull(diff_txt)`. Coverage of the first dose of the vaccine against measles (MCV1) `r cvg_chng %>% filter(vaccine == 'mcv1') %>% pull(diff_txt)`.

```{r, results='asis', echo=FALSE}
if (nctry_lt90 == 0) {
  cat(str_glue(
    "    -   At the country level, {gt90_ctry_summary %>% filter(ctry_anyvax_gt90 == 1) %>% pull(n)} out of {unique(gt90_ctry_summary$total)} ({gt90_ctry_summary %>% filter(ctry_anyvax_gt90 == 1) %>% pull(pcnt)}%) countries achieved coverage of 90% or higher in at least one of the {n_distinct(gt90_reg$vaccine)} childhood vaccines administered in the region.",
    "\n\\vspace{{1em}}\n"
  ))
} else {
cat(str_glue(
  "-   At the country level, {gt90_ctry_summary %>% filter(ctry_anyvax_gt90 == 1) %>% pull(n)} out of {unique(gt90_ctry_summary$total)} ({gt90_ctry_summary %>% filter(ctry_anyvax_gt90 == 1) %>% pull(pcnt)}%) countries achieved coverage of 90% or higher in at least one of the {n_distinct(gt90_reg$vaccine)} childhood vaccines administered in the region.",
  "\n",
  "-   There {nctry_lt90_txt} that did not achieve at least 90% coverage in any vaccine ({gt90_ctry %>% filter(ctry_anyvax_gt90 == 0) %>% pull(country) %>% knitr::combine_words()}).",
  "\n"
))
}
```

-   In `r rev_yr`, the number of zero-dose children (`r txt_dta$Estimate_lbl`) was approximately `r txt_dta %>% pull(pcnt_diff) %>% abs()`% `r txt_dta %>% pull(diff_txt)` the annual goal to achieve the Immunization Agenda 2030 target of reducing the number of zero-dose children by half by 2030 in `r regn` region (`r txt_dta$IA2030_lbl`).

```{r setup-heatmap-reg, include=FALSE}
df_to_merge <- df %>% select(-region_name)
heatmap_reg_df <- df_to_merge %>%
  rbind(hpv_df %>% filter(lvl_1 == "region", vaccine_code == "PRHPVC_F") %>% select(-vaccine_code)) %>%
  filter(lvl_1 == "region") %>%
  mutate(year = as.character(year),
         cvg_cat = case_when(coverage < 60 ~ '<60',
                             coverage %in% c(60:69) ~ '60-69',
                             coverage %in% c(70:79) ~ '70-79',
                             coverage %in% c(80:89) ~ '80-89',
                             coverage %in% c(90:94) ~ '90-94',
                             coverage >=95 ~ '>=95',
                             TRUE ~ NA_character_)) %>%
  mutate(vaccine = case_when(vaccine == "hepbb" ~ "HepBB",
                             vaccine == "hepb3" ~ "HepB3",
                             vaccine == "hib3" ~ "Hib3",
                             vaccine == "rotac" ~ "RotaC",
                             vaccine == "menga" ~ "MengA",
                             vaccine == "HPVc Females" ~ "HPVc",
                             TRUE ~ toupper(vaccine))) %>%
  group_by(country, vaccine) %>%
  mutate(mean = mean(coverage)) %>%
  ungroup()

heatmap_reg_df$vaccine <- factor(heatmap_reg_df$vaccine, levels = c("BCG", "HepBB", "DTP1", "DTP3", "Hib3", "HepB3", "PCVC", "RotaC", "IPV1", "IPVC", "MCV1", "RCV1", "MCV2", "YFV", "MengA", "HPVc"))

heatmap_reg <- heatmap_reg_df %>%
  ggplot(aes(x = year, y = vaccine, fill = coverage)) +
  geom_tile(color = "white") +
  theme_minimal() +
  geom_text(data = heatmap_reg_df %>% filter(coverage > 80 | coverage < 95), aes(label = paste0(as.character(coverage))), 
              colour = 'black', vjust = 0.5, size = txt_size) +
    geom_text(data = heatmap_reg_df %>% filter(coverage <= 80 | coverage >= 95), aes(label = paste0(as.character(coverage))), 
              colour = 'white', vjust = 0.5, size = txt_size) +
  labs(x = '', y = '',
       title = str_glue("Vaccine coverage (%), by year, {regn}, {min(heatmap_reg_df$year)}-{rev_yr}"),
       caption = str_glue("{wuenic_src}")) +
  scale_x_discrete(position = "top") +
  scale_y_discrete(limits=rev) +
  theme(strip.placement = "outside",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text.x = element_text(size = 12),
        legend.position="bottom",
        legend.key.width = unit(1.2, "cm"),
        plot.title = element_text(size = 14, hjust = 0.5, face = "bold")) +
  scale_fill_gradientn(
    colours = c('#B50800', '#E2231A', '#F26A21', '#FFC20E', '#80BD41', '#00833D') ,
    values = scales::rescale(c(0,59,69,79,89,94, 100)),
    guide = "colorbar", limits=c(0,100),
    name = str_glue("Vaccine coverage (%)  "),
    breaks = c(0,59,69,79,89,100),
    labels = c("0", "60", "70", "80", "90", "100")
  )
```

```{r overview-line-plot, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center'}
# prep plot data from heatmap_ovw_df (already has cleaned vaccine names and categories)
all_vax_plot <- heatmap_reg_df %>%
  mutate(year = as.numeric(year)) %>%
  filter(!is.na(coverage))

# color palette - one color per vaccine
vaccine_colors <- c(
  "BCG"    = "#1CABE2", "HepBB"  = "#0058AB", "DTP1"   = "#00833D",
  "DTP3"   = "#80BD41", "Hib3"   = "#6A1E74", "HepB3"  = "#961A49",
  "PCVC"   = "#E2231A", "RotaC"  = "#F26A21", "IPV1"   = "#B50800", 
  "IPVC"   = "#FF8C00", "MCV1"   = "#2E7EBB", "RCV1"   = "#9E9E9E", 
  "MCV2"   = "#FFB6C1", "YFV"    = "#5B4A9C", "MengA"  = "#C8A900", 
  "HPVc"   = "#D45F9E"
)

# first data point per vaccine for labels
first_points <- all_vax_plot %>%
  group_by(vaccine) %>%
  filter(year == min(year)) %>%
  ungroup()

num_vaccines <- n_distinct(all_vax_plot$vaccine)

min_coverage <- min(all_vax_plot$coverage, na.rm = TRUE)

all_vax_plot <- ggplot(all_vax_plot, aes(x = year, y = coverage, color = vaccine, group = vaccine)) +
  geom_line(size = 0.8, alpha = 0.7) +
  geom_point(size = 1.5, alpha = 0.7) +
  geom_hline(yintercept = 90, linetype = "dashed") +
  geom_text(data = first_points, aes(label = vaccine, color = vaccine, face = "bold"),
            hjust = 1.2, size = 2.5, show.legend = FALSE) +
  scale_color_manual(values = vaccine_colors) +
  scale_x_continuous(breaks = seq(2000, rev_yr, 1), expand = expansion(add = c(2, 0.5))) +
  scale_y_continuous(limits = c(min_coverage - 5, 100), breaks = seq(0, 100, by = 10), name = "Coverage (%)") +
  labs(
    title = str_glue("Vaccine coverage over time, {regn}, 2000–{rev_yr}"),
    caption = paste0("Dotted line shows the IA2030 target of 90% coverage\n", wuenic_src),
    x = "Year", color = NULL
  ) +
  theme_unicef() +
  theme(
    panel.grid.minor = element_blank(), panel.grid.major.x = element_line(),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
    plot.title = element_text(size = 14, hjust = 0.5, face = "bold"),
    legend.position = "right",
    legend.text = element_text(size = 8),
    legend.key.height = unit(0.6, "cm")
  ) +
  guides(color = guide_legend(nrow = num_vaccines))
```

```{r ggplot-allvax-reg, echo=FALSE, fig.width=10, fig.height=6.2}
print(heatmap_reg)
```

```{r all_vax_plot, echo=FALSE, fig.width=10, fig.height=5.5}
print(all_vax_plot)
```

```{r setup-map-africa, include=FALSE, warning=FALSE}
map_dta <- df %>% 
  filter(year == rev_yr,
         vaccine %in% c('dtp1','dtp3','mcv1'),
         lvl_1 == 'country'
  ) %>%
  mutate(coverage = case_when(Region != regn ~ NA_real_,
                              TRUE ~ coverage),
         cvg_cat = case_when(coverage < 60 ~ '<60',
                             coverage %in% c(60:69) ~ '60-69',
                             coverage %in% c(70:79) ~ '70-79',
                             coverage %in% c(80:89) ~ '80-89',
                             coverage %in% c(90:94) ~ '90-94',
                             coverage >=95 ~ '>=95'))

map_df <- left_join(base_map_df, map_dta, by = "iso3c") %>%
  filter(!is.na(coverage))

for (v in c("dtp1","dtp3","mcv1")) {
  
  plt_df <- map_df %>% filter(vaccine == v)
  
  if (regn == "EAPR") {
    # Reproject to Pacific-centered CRS (central meridian = 160°E)
    pacific_crs <- st_crs("+proj=robin +lon_0=160 +x_0=0 +y_0=0 +datum=WGS84 +units=m")
    
    plt_df <- plt_df %>%
      st_wrap_dateline(options = c("WRAPDATELINE=YES", "DATELINEOFFSET=10")) %>%
      st_transform(pacific_crs)
    
    coord_layer <- coord_sf(crs = pacific_crs, expand = FALSE)
  } else {
    coord_layer  <- coord_sf(expand = FALSE)
  }
  
  # grey out nicaragua
  plt_df <- plt_df %>%
    mutate(cvg_cat = if_else(regn == "LACR" & admin == "Nicaragua", "Other countries", as.character(cvg_cat))) %>%
    mutate(cvg_cat = factor(cvg_cat, levels = c('>=95', '90-94', '80-89', '70-79', '60-69', '<60', 'Other countries')))
  
  gg <- ggplot() +
    geom_sf(data = plt_df, aes(fill = cvg_cat), alpha = 0.75, size = 0.5, color = "black") +
    coord_sf() +
    scale_fill_manual(
      name = str_glue(''),
      breaks = c('>=95', '90-94', '80-89', '70-79', '60-69', '<60', 'Other countries'),
      values = c('<60' = '#B50800', '60-69' = '#E2231A', '70-79' = '#F26A21', '80-89' = '#FFC20E',
                 '90-94' = '#80BD41', '>=95' = '#00833D', 'Other countries' = 'lightgrey')) +
    coord_layer +
    theme_minimal(base_size = 12) +
    theme(axis.title = element_blank(),
          axis.text = element_blank(),
          axis.line = element_blank(),
          axis.ticks = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          plot.title = element_text(size = 14, hjust = 0.5, face = "bold"),
          caption = element_text(size = 8, hjust = 0, vjust = 1)
    ) +
    labs(title = str_glue("        {toupper(v)} coverage (%), {regn}, {rev_yr}"),
         caption = str_glue("Source: WHO/UNICEF estimates of national immunization coverage, {rev_yr} revision
                  Note: This map is stylized and based on an approximate scale.This map does not reflect a position\nby UNICEF or WHO on the legal status of any country or territory or the delimitation of any frontiers.")) 
  
  name <- paste0("map_", v, sep = "")
  assign(name , gg)
}
```

<!-- ===================== DTP1 ===================== -->

```{r setup-dtp1, include=FALSE}
v <- "dtp1"

unvac_v <- unvac %>% filter(vaccine == v) %>%
  mutate(ia2030tgt = round(.[[7]]/2, 0),
         ia2030tgt_pdiff = round((.[[9]]/ia2030tgt - 1) * 100),
         ia2030tgt_pdiff_hilo = case_when(ia2030tgt_pdiff > 0 ~ "higher",
                                          ia2030tgt_pdiff < 0 ~ "lower",
                                          TRUE ~ "same")) %>%
  label_vals(ia2030tgt, "ia2030tgt_lbl")

cvg_v <- cvg %>% filter(vaccine == v) 

zd_ctry <- unvac_v %>%
  filter(country != "Nicaragua") %>%  # exclude Nicaragua from this analysis since it is not shown individually in the report
  arrange(desc(.[[9]])) %>%
  filter(lvl_1 == "country") %>%
  mutate(rank = seq(n())) %>%
  slice_max(order_by = .[[9]], n = 2)

cvg_v <- cvg %>%
  filter(vaccine == v, lvl_1 == "country") %>%
  arrange(.[[9]]) %>%
  mutate(rank = dense_rank(.[[9]])) %>%
  mutate(gt90 = ifelse(.[[9]] >= 90, 1, 0),
         n_gt90 = sum(gt90)) %>%
  mutate(gt90_txt = ifelse(
    n_gt90 == 0, 
    "There were no", 
    ifelse(n_gt90 < 10, 
           str_glue(" - {knitr::combine_words(unique(country[.[[9]] >= 90]))}"), 
           "")
  )) %>%
  mutate(n_gt90_txt = ifelse(
    n_gt90 == 1, 
    "There was 1", 
    ifelse(n_gt90 > 1, 
           str_glue("There were {n_gt90}"), 
           "")
  ))

cvg_v_no_nicaragua <- cvg_v %>% filter(country != "Nicaragua")
```

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{DTP1: Reaching zero-dose children}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:dtp1}{}")
cat("\\addcontentsline{toc}{section}{DTP1: Reaching zero-dose children}")
```

-   The number of children vaccinated with DTP1 `r unvac_v %>% filter(lvl_1 == 'region') %>% pull(pchng_txt)`.

-   In `r rev_yr`, the number of zero-dose children (`r txt_dta$Estimate_lbl`) was approximately `r txt_dta %>% pull(pcnt_diff) %>% abs()`% `r txt_dta %>% pull(diff_txt)` the annual goal to achieve the Immunization Agenda 2030 target in `r regn` region (`r txt_dta$IA2030_lbl`).

-   With `r tgt %>% filter(lvl_1 == 'region') %>% pull(diff_abs_lbl)` `r tgt %>% filter(lvl_1 == 'region') %>% pull(moreless)` surviving infants in the region compared to in `r comp_yr`, `r tgt %>% filter(lvl_1 == 'region') %>% pull(to_vax)`. 

-   In `r rev_yr`, `r n_distinct(zd_ctry$iso3c)` countries, `r unique(zd_ctry$country) %>% knitr::combine_words()`, represented approximately `r zd_ctry %>% filter(rank == 1) %>% pull(unvac_ctry_pcnt)`% and `r zd_ctry %>% filter(rank == 2) %>% pull(unvac_ctry_pcnt)`% of all zero-dose children in `r params$region` (n = `r unvac_v %>% filter(lvl_1 == 'region') %>% pull(12)`), respectively. `r zd_ctry %>% filter(rank == 1) %>% pull(country)` had the highest number of zero-dose children at `r zd_ctry %>% filter(rank == 1) %>% pull(12)`, and `r zd_ctry %>% filter(rank == 2) %>% pull(country)` had the second highest at `r zd_ctry %>% filter(rank == 2) %>% pull(12)`. 

-   The lowest DTP1 coverage was observed in `r cvg_v_no_nicaragua %>% filter(rank == 1) %>% pull(country) %>% knitr::combine_words()` at `r cvg_v_no_nicaragua %>% filter(rank == 1) %>% pull(9) %>% unique()`%, followed by `r cvg_v_no_nicaragua %>% filter(rank == 2) %>% pull(country) %>% knitr::combine_words()` at `r cvg_v_no_nicaragua %>% filter(rank == 2) %>% pull(9) %>% unique()`%. 

```{r, results='asis', echo=FALSE}
if (unique(cvg_v$n_gt90) == 0) {
cat(paste0("-   None of the ", nctry, " countries in ", regn, " achieved DTP1 coverage of 90% or higher in ", rev_yr, "."))
} else {
  cat(paste0("-   ", unique(cvg_v$n_gt90_txt), " (", round(unique(cvg_v$n_gt90)/nctry*100,0), "%) out of ", nctry, " countries in the region that achieved DTP1 coverage of 90% or higher", cvg_v %>% pull(gt90_txt) %>% unique() , ".", "\\vspace{1em}\n\n"))
}
```

\vspace{1em}

```{r, results='asis', echo=FALSE}
if (regn %in% c("Africa", "Global")) {
  reg_zd_txt <- heatmap_df
  cat(paste0("    -   ", "", "\n"))
}
```

```{r heatmap-dtp1, echo=FALSE, fig.width=10, fig.height=8.5}
print(heatmap_dtp1)
```

```{r map_dtp1, echo=FALSE, warning=FALSE, fig.width=10, fig.height=6.5}
print(map_dtp1)
```

```{r line_dtp1, echo=FALSE, warning=FALSE, fig.width=10, fig.height=6.5}
print(line_plt_dtp1)
```

<!-- ===================== DTP3 ===================== -->

```{r setup-dtp3, include=FALSE}
v <- "dtp3"
unvac_v <- unvac %>% filter(vaccine == v) 
cvg_v   <- cvg %>% filter(vaccine == v)
vacc_v  <- vacc %>% filter(vaccine == v)

reg_cvg_rev_yr  <- cvg_v %>% filter(lvl_1 == "region") %>% pull(9)   
reg_cvg_comp_yr <- cvg_v %>% filter(lvl_1 == "region") %>% pull(7) 

unvac_ctry <- unvac_v %>%
  arrange(desc(.[[9]])) %>%
  filter(lvl_1 == "country") %>%
  mutate(rank = seq(n())) %>%
  slice_max(order_by = .[[9]], n = 2)

cvg_v <- cvg_v %>%
  filter(vaccine == v,
         lvl_1 == "country") %>%
  arrange(.[[9]]) %>%
  mutate(rank = dense_rank(.[[9]])) %>%
  mutate(gt90 = ifelse(.[[9]] >= 90, 1, 0),
         n_gt90 = sum(gt90)) 

cvg_v_no_nicaragua <- cvg_v %>% filter(country != "Nicaragua")

tgt_chng <- tgt %>%
  filter(lvl_1 == "region") %>%
  mutate(pop_grwth = round((.[[9]] / .[[7]] - 1) * 100, 0)) %>%
  label_vals_millions(.[[9]], "tgt_rev_yr") %>%
  label_vals_millions(.[[7]], "tgt_comp_yr")

if((reg_cvg_rev_yr < reg_cvg_comp_yr) & (tgt_chng[[9]] > tgt_chng[[7]])) {
  pop_chng_txt <- str_glue("However, due to population growth—approximately a {tgt_chng$pop_grwth}% increase between {comp_yr} ({tgt_chng$tgt_comp_yr}) and {rev_yr} ({tgt_chng$tgt_rev_yr})—DTP3 coverage in {rev_yr} ({reg_cvg_rev_yr}%) was lower than in {comp_yr} ({reg_cvg_comp_yr}%)")
} else {
  pop_chng_txt <- ""
}

dropout <- df %>%
  filter(lvl_1 == "region",
         year %in% c(comp_yr, rev_yr),
         vaccine %in% c("dtp1","dtp3")) %>%
  select(lvl_1, lvl_2, Region, iso3c, country, year, vaccine, vaccinated) %>%
  pivot_wider(., names_from = vaccine, values_from = vaccinated) %>%
  mutate(dropout = round(((dtp1-dtp3)/dtp1) * 100, 0),
         dropout_txt = case_when(dropout < 5 ~ "low DTP drop-out rates imply good ability to provide a complete series of vaccines early in life",
                                 dropout %in% c(5:10) ~ "medium DTP drop-out rates imply moderate ability to provide a complete series of vaccines early in life",
                                 dropout > 5 ~ "high DTP drop-out rates imply poor ability to provide a complete series of vaccines early in life"))

dropout_comp <- dropout %>%
  select(-c(dtp1, dtp3, dropout_txt)) %>%
  pivot_wider(., names_from = year, values_from = dropout) %>%
  mutate(diff = .[[7]] - .[[6]],
         diff_txt  = case_when(diff > 0 ~ str_glue("{diff} percentage points higher than in {comp_yr} ({.[[6]]}%)"),
                               diff < 0 ~ str_glue("{abs(diff)} percentage points lower than in {comp_yr} ({.[[6]]}%)"),
                               TRUE ~ str_glue("the same as in {comp_yr}")))

unvac_cvg <- unvac_v %>%
  arrange(desc(.[[9]])) %>%
  filter(lvl_1 == "country") %>%
  mutate(rank = seq(n())) %>%
  select(lvl_1:vaccine, rank, 9, 12) %>%
  rename_with(~"unvaccinated", .cols = 8) %>%
  rename_with(~"unvaccinated_lbl", .cols = 9) %>%
  left_join(., cvg_v %>% select(iso3c, vaccine, 9), by = c("iso3c", "vaccine")) %>%
  rename_with(~"coverage", .cols = 10) %>% 
  filter(country != "Nicaragua")  # exclude Nicaragua from this analysis since it is not shown individually in the report

```

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{DTP3: A marker of routine immunization service delivery to children}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:dtp3}{}")
cat("\\addcontentsline{toc}{section}{DTP3: A marker of routine immunization service delivery to children}")
```

\vspace{1em}

-   The number of children vaccinated with DTP3 `r vacc_v %>% filter(lvl_1 == 'region') %>% pull(pchng_txt)` from `r vacc_v %>% filter(lvl_1 == 'region') %>% pull(vacc_comp_yr_lbl)` in `r comp_yr` to `r vacc_v %>% filter(lvl_1 == 'region') %>% pull(vacc_rev_yr_lbl)` in `r rev_yr`. `r {pop_chng_txt}`

-   In `r rev_yr`, `r dropout %>% filter(year == rev_yr) %>% pull(dropout)`% of children who received DTP1 did not receive DTP3. DTP drop-out was `r dropout_comp %>% pull(diff_txt)`. The `r dropout %>% filter(year == rev_yr) %>% pull(dropout_txt)`. 

-   In `r rev_yr`, DTP3 coverage ranged from `r min(cvg_v_no_nicaragua[[9]], na.rm = TRUE)`% in `r cvg_v_no_nicaragua %>% filter(.[[9]] == min(.[[9]])) %>% pull(country) %>% knitr::combine_words()` to `r max(cvg_v_no_nicaragua[[9]], na.rm = TRUE)`% in `r cvg_v_no_nicaragua %>% filter(.[[9]] == max(.[[9]])) %>% pull(country) %>% knitr::combine_words()`. 

-   `r unvac_cvg %>% filter(rank == 1) %>% pull(country)` had DTP3 coverage of `r unvac_cvg %>% filter(rank == 1) %>% pull(coverage) %>% unique()`% and the highest absolute number of un- and undervaccinated children in the region (`r unvac_cvg %>% filter(rank == 1) %>% pull(unvaccinated_lbl) %>% unique()`), followed by `r unvac_cvg %>% filter(rank == 2) %>% pull(country)` with `r unvac_cvg %>% filter(rank == 2) %>% pull(coverage)`% coverage and `r unvac_cvg %>% filter(rank == 2) %>% pull(unvaccinated_lbl)` un- and undervaccinated children. 

\vspace{1em}

```{r ggplot-dtp3, echo=FALSE, fig.width=10, fig.height=8.5}
print(heatmap_dtp3)
```

```{r ggplot_dtp3, echo=FALSE, warning=FALSE, fig.width=10, fig.height=4.5}
print(map_dtp3)
```

```{r line_dtp3, echo=FALSE, warning=FALSE, fig.width=10, fig.height=6.5}
print(line_plt_dtp3)
```

<!-- ===================== MCV ===================== -->

```{r setup-mcv, include=FALSE}
v <- c("mcv1", "mcv2")
unvac_v <- unvac %>% filter(vaccine %in% c(v)) 
cvg_v   <- cvg %>% filter(vaccine %in% c(v)) 

cvg_mcv1 <- cvg_v %>%
  filter(vaccine == "mcv1") %>%
  mutate(diff = .[[9]] - .[[7]],
         cvg_chng_txt = case_when(diff %in% c(-1:1) ~ str_glue("remained relatively stable, within 1% of the {comp_yr} level"),
                                  diff < -1 ~ str_glue("was lower than in {comp_yr}"),
                                  diff > 1 ~ str_glue("was higher than in {comp_yr}")))

cvg_mcv1_reg <- cvg_mcv1 %>% filter(lvl_1 == "region")

cvg_mcv2 <- cvg_v %>%
  filter(lvl_1== "region",
         vaccine == "mcv2") %>%
  mutate(diff = .[[9]] - .[[7]],
         cvg_chng_txt = case_when(diff %in% c(-1:1) ~ str_glue("remained relatively stable, within 1% of the {comp_yr} level"),
                                  diff < -1 ~ str_glue("decreased from {.[[7]]}% in {comp_yr} to {.[[9]]}% in {rev_yr}"),
                                  diff > 1 ~ str_glue("increased from {.[[7]]}% in {comp_yr} to {.[[9]]}% in {rev_yr}")))

unvac_mcv <- unvac_v %>%
  filter(lvl_1 == "region") %>%
  select(lvl_1:vaccine, 9) %>%
  pivot_wider(., names_from = vaccine, values_from = last_col()) %>%
  mutate(diff = mcv2 - mcv1) %>%
  label_vals_millions(mcv1, 'mcv1_lbl') %>%
  label_vals_millions(mcv2, 'mcv2_lbl') %>%
  label_vals_millions(diff, 'diff_lbl') 

unvac_cvg_mcv2 <- unvac %>%
  filter(lvl_1 == "region",
         vaccine %in% c("mcv1","mcv2")) %>%
  select(lvl_1:vaccine, 10, 12) %>%
  rename_with(~"rev_yr_unvaccinated_lbl", .cols = 8) %>%
  rename_with(~"comp_yr_unvaccinated_lbl", .cols = 7) %>%
  left_join(., cvg_v %>% select(iso3c, vaccine, 9), by = c("iso3c", "vaccine")) %>%
  rename_with(~"coverage", .cols = 9)

unvac_ctry <- unvac_v %>%
  filter(country != "Nicaragua") %>%  # exclude Nicaragua from this analysis since it is not shown individually in the report
    filter(lvl_1 == "country",
           vaccine == "mcv1") %>%
  arrange(desc(.[[9]])) %>%
  mutate(rank = seq(n())) %>%
  slice_max(order_by = .[[9]], n = 1)

nyrs <- 3

if (unique(df$Region) %in% c("ECAR", "LACR")) {
  thrshold <- 80
} else {
thrshold <- 80
}

cvg_trends <- df %>% 
  filter(vaccine == "mcv1",
         year %in% seq((rev_yr - nyrs + 1), rev_yr),
         lvl_1 == "country") %>%
    select(lvl_1, lvl_2, Region, iso3c, country, year, vaccine, coverage) %>%
  arrange(country) %>%
  mutate(ltx = ifelse(coverage < thrshold, 1, 0)) %>%
  group_by(country) %>%
  mutate(sum = sum(ltx)) %>%
  ungroup() %>%
  select(iso3c, country, sum) %>%
  distinct() %>% 
  filter(country != "Nicaragua")  # exclude Nicaragua from this analysis since it is not shown individually in the report

n_ctry_lower_x <- nrow(cvg_trends %>% filter(sum == nyrs))

dropout <- df %>%
  filter(lvl_1 == "region",
         year %in% c(comp_yr, rev_yr),
         vaccine %in% c("dtp1","mcv1")) %>%
  select(lvl_1, lvl_2, Region, iso3c, country, year, vaccine, vaccinated) %>%
  pivot_wider(., names_from = vaccine, values_from = vaccinated) %>%
  mutate(dropout = round(((dtp1-mcv1)/dtp1) * 100, 0),
         dropout_txt = case_when(dropout < 5 ~ "low DTP-MCV drop-out rates imply good retention in immunization programmes and ability to provide a full course of vaccines in infancy (up to one year)",
                                 dropout %in% c(5:10) ~ "medium DTP-MCV drop-out rates imply moderate retention in immunization programmes and ability to provide a full course of vaccines in infancy (up to one year)",
                                 dropout > 5 ~ "high DTP-MCV drop-out rates imply poor retention in immunization programmes and ability to provide a full course of vaccines in infancy (up to one year)"))

dropout_comp <- dropout %>%
  select(-c(dtp1, mcv1, dropout_txt)) %>%
  pivot_wider(., names_from = year, values_from = dropout) %>%
  mutate(diff = .[[7]] - .[[6]],
         diff_txt  = case_when(diff > 0 ~ str_glue("{diff} percentage points higher than in {comp_yr} ({.[[6]]}%)"),
                               diff < 0 ~ str_glue("{abs(diff)} percentage points lower than in {comp_yr} ({.[[6]]}%)"),
                               TRUE ~ str_glue("the same as in {comp_yr}")))
```

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{MCV: Canary in the coal mine}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:mcv}{}")
cat("\\addcontentsline{toc}{section}{MCV: Canary in the coal mine}")
```

\vspace{1em}

-   MCV1 coverage in `r rev_yr` (`r cvg_mcv1_reg[[9]]`%) `r cvg_mcv1_reg$cvg_chng_txt` (`r cvg_mcv1_reg[[7]]`%). 

-   MCV2 coverage `r cvg_mcv2$cvg_chng_txt`. This still leaves `r unvac_mcv$mcv1_lbl` children without any protection and another `r unvac_mcv$diff_lbl` with only partial protection against the measles virus. 

-   In `r rev_yr`, `r dropout %>% filter(year == rev_yr) %>% pull(dropout)`% of children who received DTP1 did not receive MCV1. DTP-MCV drop-out was `r dropout_comp %>% pull(diff_txt)`. The `r dropout %>% filter(year == rev_yr) %>% pull(dropout_txt)`.

```{r, results='asis', echo=FALSE}
if (n_ctry_lower_x == 1) {
    n_ctry_low_txt <- "one country"
    cvg_trends <- cvg_trends %>% filter(sum == nyrs)
cat(str_glue(paste0("-   Over the last {nyrs} years, {n_ctry_low_txt} ({knitr::combine_words(cvg_trends$country)}) consistently had MCV1 coverage below {thrshold}%", ".\n", "\n")))
} else if (n_ctry_lower_x > 1) {
   n_ctry_low_txt <- str_glue("{n_ctry_lower_x} countries") 
   cvg_trends <- cvg_trends %>% filter(sum == nyrs)
   cat(str_glue(paste0("-   Over the last {nyrs} years, {n_ctry_low_txt} ({knitr::combine_words(cvg_trends$country)}) consistently had MCV1 coverage below {thrshold}%", ".\n", "\n")))
} else {
     cat(str_glue(paste0("-   Over the last {nyrs} years, no countries have had MCV1 coverage below {thrshold}%", ".\n", "\\vspace{{1em}}\n")))
}
```

-   `r unvac_ctry %>% pull(country)` had the highest number of children who did not receive the measles vaccine (`r unvac_ctry[[12]]`), representing approximately `r unvac_ctry %>% pull(unvac_ctry_pcnt)`% of all children unprotected against measles in `r unique(df$Region)` (`r unvac_mcv %>% pull(mcv1_lbl)`). 

\vspace{1em}

```{r ggplot-mcv1, echo=FALSE, fig.width=10, fig.height=8.5}
print(heatmap_mcv1)
```

```{r ggplot_mcv1, echo=FALSE, warning=FALSE, fig.width=10, fig.height=4.5}
print(map_mcv1)
```

```{r line_mcv1, echo=FALSE, warning=FALSE, fig.width=10, fig.height=6.5}
print(line_plt_mcv1)
```

<!-- ===================== HPV ===================== -->

```{r setup-hpv, include=FALSE}
hpv <- hpv_df %>%
  select(lvl_1:vaccine_code, coverage) %>%
  filter(year %in% c(hpv_rev_yr, hpv_rev_yr - 1),
         lvl_1 == "region") %>%
  pivot_wider(names_from = year, values_from = coverage) %>%
  mutate(diff = .[[8]] - .[[7]],
         chng = case_when(diff > 0 ~ "increased", diff < 0 ~ "decreased", TRUE ~ "remained constant"),
         diff_txt = case_when(
           chng %in% c("increased", "decreased") ~ str_glue("{chng} from {.[[7]]}% in {hpv_rev_yr-1} to {.[[8]]}% in {hpv_rev_yr}"),
           TRUE ~ str_glue("{chng} at {.[[8]]}% between {hpv_rev_yr-1} and {hpv_rev_yr}")
         ),
         gt90 = case_when(
           .[[8]] < 80 & chng == "increased" ~ "Despite the important increases, HPVc coverage remains below the 90% target",
           .[[8]] < 80 & chng != "increased" ~ "HPVc coverage remains below the 90% target",
           .[[8]] >= 80 & .[[8]] <= 89 & chng == "increased" ~ "Despite the increase, HPVc coverage has not yet reached the 90% target",
           .[[8]] >= 80 & .[[8]] <= 89 & chng != "increased" ~ "HPVc coverage has not yet reached the 90% target",
           .[[8]] == 90 ~ "Coverage of the last dose of HPV vaccine achieved the 90% target",
           .[[8]] > 90 ~ "Coverage of the last dose of HPV vaccine has exceeded the 90% target",
           TRUE ~ NA_character_
         ))

hpv_unvac <- hpv_df %>%
  select(lvl_1:vaccine_code, unvaccinated) %>%
  filter(year %in% c(hpv_rev_yr),
         lvl_1 == "region") %>%
  pivot_wider(., names_from = vaccine_code, values_from = unvaccinated) %>%
  mutate(diff = PRHPVC_F - PRHPV1_F) %>%
  label_vals_millions(PRHPVC_F, "PRHPVC_F_lbl") %>%
  label_vals_millions(PRHPV1_F, "PRHPV1_F_lbl") %>%
  label_vals_millions(diff, "diff_lbl")

n_ctry_in_reg <- df %>% filter(lvl_1 == "country") %>% pull(iso3c) %>% unique()

# fix: use filter(year <= rev_yr) and take the most recent intro year per country
# so countries that introduced HPV before rev_yr are still counted
hpv_intro_y <- wiise_hpv_intro_yrs %>%
  filter(iso3c %in% c(n_ctry_in_reg)) %>%
  filter(nationwide == "yes" | partially == "yes") %>%
  filter(year <= rev_yr) %>%
  group_by(iso3c) %>%
  slice_max(year, n = 1, with_ties = FALSE) %>%
  ungroup()

n_ctry_intro <- n_distinct(hpv_intro_y$iso3c)
n_ctry_total <- length(n_ctry_in_reg)
pct_ctry_intro <- round(n_ctry_intro / n_ctry_total * 100, 0)

# countries that have NOT introduced HPV
hpv_missing_ctry <- df %>%
  filter(lvl_1 == "country") %>%
  select(iso3c, country) %>%
  distinct() %>%
  anti_join(hpv_intro_y, by = "iso3c") %>%
  pull(country)

hpv_missing_flag <- ifelse(length(hpv_missing_ctry) > 0, 1, 0)

# number of countries with WUENIC HPV data for rev_yr
n_ctry_wuenic_hpv <- hpv_df %>%
  filter(lvl_1 == "country",
         year == hpv_rev_yr,
         vaccine_code == "PRHPVC_F") %>%
  summarise(n = n_distinct(iso3c)) %>%
  pull(n)

# df of countries with WUENIC HPV data for rev_yr
hpv_wuenic_ctry <- hpv_df %>%
  filter(lvl_1 == "country",
         year == hpv_rev_yr) %>% 
  select(iso3c, country) %>%
  distinct()

# join countries with hpv data introduced in rev_yr
# get countries that have introduced HPV but do not have WUENIC data in rev_yr for it
hpv_intro_data <- hpv_intro_y %>%
  left_join(hpv_wuenic_ctry, by = "iso3c") %>%
  mutate(has_wuenic_data = ifelse(iso3c %in% hpv_wuenic_ctry$iso3c, "Has WUENIC data", "No WUENIC data")) %>% 
  left_join(wuenic_dta %>% select(iso3c, country) %>% distinct() %>% rename(country_complete = country), by = c("iso3c" = "iso3c"))


pct_ctry_wuenic_hpv <- round(n_ctry_wuenic_hpv / n_ctry_total * 100, 0)

n_ctry_intro_f <- wiise_hpv_intro_yrs %>%
  filter(iso3c %in% c(n_ctry_in_reg) | iso3c == "nic") %>%
  distinct() %>%
  filter(nationwide == "yes" | partially == "yes") %>%
  group_by(year) %>%
  mutate(n_ctry = n_distinct(iso3c)) %>%
  ungroup() %>%
  select(year, n_ctry) %>%
  distinct() %>%
  mutate(sex = "female")
```

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{HPV}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:hpv}{}")
cat("\\addcontentsline{toc}{section}{HPV}")
```

\vspace{1em}

-   Programme coverage of the first dose of HPV vaccine (HPV1) among girls `r hpv %>% filter(vaccine_code == 'PRHPV1_F') %>% pull(diff_txt)` and coverage of the last dose (HPVc) `r hpv %>% filter(vaccine_code == 'PRHPVC_F') %>% pull(diff_txt)`.

-   `r hpv %>% filter(vaccine_code == 'PRHPVC_F') %>% pull(gt90)`.

-   In `r rev_yr`, there were `r hpv_unvac$PRHPV1_F_lbl` girls that did not receive any HPV vaccination.

```{r, results='asis', echo=FALSE}

if (hpv_missing_flag == 1) {
  cat(paste0(
    "-   ", n_ctry_intro, " out of ", n_ctry_total, " countries (", pct_ctry_intro, "%) ",
    "in ", regn, " had introduced HPV vaccination by ", rev_yr, ".\n\n",
    "    -   WUENIC HPV coverage estimates are available for ",
    ifelse(n_ctry_wuenic_hpv == n_ctry_intro, "all ", ""),
    n_ctry_wuenic_hpv, " of these countries in ", hpv_rev_yr, ".\n\n"
  ))
} else {
  cat(paste0(
    "-   All ", n_ctry_total, " countries in ", regn, " had introduced HPV vaccination by ", rev_yr, ".\n\n",
    "    -   WUENIC HPV coverage estimates are available for ", n_ctry_wuenic_hpv,
    " (", pct_ctry_wuenic_hpv, "%) of these countries in ", hpv_rev_yr, ".\n\n"
  ))
}

# indented sub-bullet: countries that introduced HPV but lack WUENIC data
if (n_ctry_intro > n_ctry_wuenic_hpv) {
  hpv_country_df <- hpv_df %>% filter(country != regn)

  hpv_intro_no_wuenic <- hpv_intro_y %>%
    filter(!iso3c %in% (hpv_country_df %>%
                          filter(lvl_1 == "country", year == hpv_rev_yr) %>%
                          pull(iso3c))) %>%
    left_join(
      df %>% filter(lvl_1 == "country") %>% select(iso3c, country) %>% distinct(),
      by = "iso3c"
    ) %>%
    pull(country)

  if (length(hpv_intro_no_wuenic) > 0 && !all(is.na(hpv_intro_no_wuenic))) {
    hpv_intro_no_wuenic <- hpv_intro_no_wuenic[!is.na(hpv_intro_no_wuenic)]
    cat(paste0(
      "    -   ",
      length(hpv_intro_no_wuenic), " ",
      ifelse(length(hpv_intro_no_wuenic) == 1, "country", "countries"),
      " that introduced HPV vaccination by ", rev_yr,
      ifelse(length(hpv_intro_no_wuenic) == 1, " does", " do"),
      " not have WUENIC coverage estimates for ", hpv_rev_yr, ": ",
      knitr::combine_words(hpv_intro_no_wuenic), ".\n\n"
    ))
  }
}

# un-indented bullet: countries that have not introduced HPV at all
if (hpv_missing_flag == 1) {
  cat(paste0(
    "-   The ", length(hpv_missing_ctry), " ",
    ifelse(length(hpv_missing_ctry) == 1, "country", "countries"),
    " that had not introduced HPV vaccination ",
    ifelse(length(hpv_missing_ctry) == 1, "is", "are"), " ",
    knitr::combine_words(hpv_missing_ctry), ".\n\n"
  ))
}
```

```{r hpv-region-line-plot, echo=FALSE, warning=FALSE, message=FALSE}

  # prep plot data
  plot_hpv1 <- hpv_df %>%
    filter(vaccine_code == "PRHPV1_F", country == regn) %>%
    select(year, coverage)

  plot_hpvc <- hpv_df %>%
    filter(vaccine_code == "PRHPVC_F", country == regn) %>%
    select(year, coverage)

  # unvaccinated females scaled to coverage axis
  plot_zd_hpv <- hpv_df %>%
    filter(vaccine_code == "PRHPV1_F", country == regn) %>%
    select(year, unvaccinated)
  
  first_hpv_year <- min(c(plot_hpv1$year, plot_hpvc$year), na.rm = TRUE)

  hpv_scale_factor <- 100 / max(plot_zd_hpv$unvaccinated, na.rm = TRUE)

  hpv_region_plot <- ggplot() +
    geom_col(data = plot_zd_hpv, aes(x = year, y = unvaccinated * hpv_scale_factor, fill = "Unvaccinated females"), alpha = 0.4, width = 0.6) +
    # HPV1 line
    {if (nrow(plot_hpv1) > 1)
      list(geom_line(data  = plot_hpv1, aes(x = year, y = coverage, color = "HPV1 coverage"), size = 0.8),
           geom_point(data = plot_hpv1, aes(x = year, y = coverage, color = "HPV1 coverage"), size = 1.5))
    else
      list(geom_point(data = plot_hpv1, aes(x = year, y = coverage, color = "HPV1 coverage"), size = 1.5))
    } +
    # HPVc line
    {if (nrow(plot_hpvc) > 1)
      list(geom_line(data  = plot_hpvc, aes(x = year, y = coverage, color = "HPVc coverage"), size = 0.8),
           geom_point(data = plot_hpvc, aes(x = year, y = coverage, color = "HPVc coverage"), size = 1.5))
    else
      list(geom_point(data = plot_hpvc, aes(x = year, y = coverage, color = "HPVc coverage"), size = 1.5))
    } +
    geom_hline(yintercept = 90, linetype = "dashed", alpha = 0.7) +
    scale_color_manual(name = NULL, values = c("HPV1 coverage" = "#1CABE2", "HPVc coverage" = "#6A1E72")) +
    scale_fill_manual(name = NULL, values = c("Unvaccinated females" = "#E2231A")) +
    scale_x_continuous(breaks = seq(min(plot_hpvc$year), max(plot_hpvc$year), by = 1)) +
    scale_y_continuous(limits = c(0, 100), name = "Coverage (%)",
                       sec.axis = sec_axis(~ . / hpv_scale_factor, name = "Unvaccinated females (count)", labels = scales::label_comma())) +
    guides(color = guide_legend(order = 1), fill = guide_legend(order = 2)) +
    labs(title = str_glue("HPV1 and HPVc coverage and unvaccinated \nfemales over time, {regn}, {first_hpv_year}–{hpv_rev_yr}"),
         caption = str_glue("Dotted line shows the IA2030 target of 90% coverage\n{wuenic_src}"),
         x = "Year") +
    theme_unicef() +
    theme(
      panel.grid.minor   = element_blank(),
      panel.grid.major.x = element_line(color = "grey90"),
      axis.text.x        = element_text(angle = 45, hjust = 1, size = 9),
      plot.title         = element_text(size = 12, hjust = 0.5, face = "bold"),
      legend.position    = "bottom"
    )

  print(hpv_region_plot)
```

\vspace{1em}

<!-- ===================== BIG CATCH-UP ===================== -->

```{r setup-bcu-ctry, include=FALSE}
bcu_ctry <- c("afg","ben","bfa","bdi","cmr","caf","tcd","tgo","com","civ","cod","eth","gmb","gin","gnb","ken","kgz",
              "mdg","mli","mrt","moz","mmr","npl","ner","nga","pak","slb","som","ssd","sdn","syr","tjk","tza","yem","zmb")

n_bcu_ctry <- df %>%
  select(iso3c) %>%
  distinct() %>%
  summarise(count = sum(iso3c %in% bcu_ctry, na.rm = TRUE)) %>%
  pull(count)

if (n_bcu_ctry > 0) {
  # 1. Improved Data Processing
  recovery <- df %>%
    filter(lvl_1 == 'country',
           iso3c %in% c(bcu_ctry),
           vaccine == 'dtp3',
           year %in% c(2019:rev_yr)) %>%
    select(iso3c, country, vaccine, year, coverage) %>%
    arrange(iso3c, vaccine, year) %>%
    pivot_wider(names_from = year, values_from = coverage) %>%
    # Use rowwise with pmin to handle potential NAs in specific years
    rowwise() %>% 
    mutate(
      # Calculate differences relative to 2019 baseline
      diff_2019_2020 = `2020` - `2019`,
      diff_2019_2021 = `2021` - `2019`,
      diff_2019_2022 = `2022` - `2019`,
      diff_2019_2023 = `2023` - `2019`,
      diff_2019_2024 = `2024` - `2019`,
      # pandemic_impact is the deepest drop relative to 2019
      pandemic_impact = pmin(diff_2019_2020, diff_2019_2021, diff_2019_2022, 
                             diff_2019_2023, diff_2019_2024, na.rm = TRUE),
      # Categorize the severity of the decline
      pandemic_effect = case_when(
        pandemic_impact <= -5 ~ 'substantial decline',
        pandemic_impact < 0   ~ 'small decline',
        TRUE                  ~ 'no decline'
      ),
      # Categorize recovery status
      recovery_status = case_when(
        diff_2019_2024 >= 0 ~ 'recovered',
        # Recovering defined as regaining >3% points from the low
        diff_2019_2024 >= (pandemic_impact + 3) ~ 'recovering',
        TRUE ~ 'not recovered'
      )
    ) %>%
    ungroup()

  # 2. Refined Summary Logic
  recovery_summary <- recovery %>%
    mutate(declined_ever = ifelse(grepl("decline", pandemic_effect), "decline", "no decline")) %>%
    group_by(declined_ever, recovery_status) %>%
    summarise(n_countries = n(), .groups = "drop") %>%
    complete(declined_ever = c("decline", "no decline"), 
             recovery_status = c("recovered", "recovering", "not recovered"), 
             fill = list(n_countries = 0)) 

  # Extract counts for text generation
  n_no_decline <- recovery_summary %>% 
    filter(recovery_status == "recovered", declined_ever == "no decline") %>% 
    summarise(total = sum(n_countries)) %>% pull(total)

  n_recovered <- recovery_summary %>% 
    filter(recovery_status == "recovered", declined_ever == "decline") %>% 
    summarise(total = sum(n_countries)) %>% pull(total)

  n_recovering <- recovery_summary %>% 
    filter(recovery_status == "recovering") %>% 
    summarise(total = sum(n_countries)) %>% pull(total)

  n_no <- recovery_summary %>% 
    filter(recovery_status == "not recovered") %>% 
    summarise(total = sum(n_countries)) %>% pull(total)

  # 3. Helper function for clean text blocks
  build_text <- function(count, status_label, countries_vec, desc) {
    if (count == 0) return(NULL)
    count_txt <- if (count == 1) "One country" else paste(count, "countries")
    names_txt <- knitr::combine_words(countries_vec)
    str_glue("{count_txt} ({names_txt}) {desc}")
  }

  # Generate individual components
  recovery_txt1 <- build_text(n_recovered, "recovered", 
                     recovery %>% filter(recovery_status == 'recovered', grepl('decline', pandemic_effect)) %>% pull(country),
                     "fully recovered to or exceeded their 2019 DTP3 levels.")
  
  recovery_txt2 <- build_text(n_recovering, "recovering", 
                     recovery %>% filter(recovery_status == 'recovering') %>% pull(country),
                     "are on the path to recovery to 2019 DTP3 levels.")
  
  recovery_txt3 <- build_text(n_no, "not recovered", 
                     recovery %>% filter(recovery_status == 'not recovered') %>% pull(country),
                     "have not yet recovered, with stagnation or decline from 2019 DTP3 levels.")
  
  recovery_txt4 <- build_text(n_no_decline, "no decline", 
                     recovery %>% filter(recovery_status == 'recovered', pandemic_effect == 'no decline') %>% pull(country),
                     "did not experience a decline in DTP3 levels compared to 2019.")

  # 4. Final output: Combines only non-empty strings with double newlines
  recovery_txt_final <- paste(Filter(negate(is.null), list(recovery_txt1, recovery_txt2, recovery_txt3, recovery_txt4)), collapse = "\n\n")
}
```

```{r, results='asis', echo=FALSE}
if (n_bcu_ctry > 0) {
  cat("\\newpage\n")
  cat("\\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]\n")
  cat("\\textcolor{white}{\\Large \\textbf{Priority countries supported to recover from the COVID-19 pandemic}}\n")
  cat("\\end{tcolorbox}\n")
  cat("\\phantomsection\n")
  cat("\\hypertarget{sec:bcu}{}\n")
  cat("\\addcontentsline{toc}{section}{Priority countries supported to recover from the COVID-19 pandemic}\n")
  cat("\\vspace{1em}\n")
}
```

```{r, results='asis', echo=FALSE}
if (n_bcu_ctry > 0) {
  cat(paste0("-   In ", regn, " ", n_bcu_ctry, " out of the 35 countries were globally supported for the Big Catch-up. Of these:", "\n"))
}
```

```{r, results='asis', echo=FALSE}
if (n_bcu_ctry > 0 && !is.null(recovery_txt1) && nzchar(recovery_txt1)) {
  cat(paste0("    -   ", recovery_txt1, "\n"))
}
```

```{r, results='asis', echo=FALSE}
if (n_bcu_ctry > 0 && !is.null(recovery_txt2) && nzchar(recovery_txt2)) {
  cat(paste0("    -   ", recovery_txt2, "\n"))
}
```

```{r, results='asis', echo=FALSE}
if (n_bcu_ctry > 0 && !is.null(recovery_txt3) && nzchar(recovery_txt3)) {
  cat(paste0("    -   ", recovery_txt3, "\n"))
}
```

```{r, results='asis', echo=FALSE}
if (n_bcu_ctry > 0 && !is.null(recovery_txt4) && nzchar(recovery_txt4)) {
    cat(paste0("    -   ", recovery_txt4, "\n"))
}
```

```{r bcu-recovery-plot, echo=FALSE, warning=FALSE, fig.width=9, fig.height=max(4, n_bcu_ctry * 0.4 + 1), out.width="100%"}
if (n_bcu_ctry > 0) {

  # Build plot df with recovery status labels
  bcu_plot_df <- recovery %>%
    select(country, iso3c, `2019`, `2024`, pandemic_effect, recovery_status) %>%
    mutate(
      recovery_label = case_when(
        recovery_status == "recovered"     ~ "Recovered",
        recovery_status == "recovering"    ~ "Recovering",
        recovery_status == "not recovered" ~ "Not recovered"
      ),
      pandemic_label = case_when(
        grepl("no decline", pandemic_effect) ~ "No pandemic decline",
        grepl("small", pandemic_effect)      ~ "Small decline",
        grepl("substantial", pandemic_effect) ~ "Substantial decline"
      ),
      # sort by 2024 coverage within recovery group
      country = fct_reorder(country, `2024`)
    )

  # Color palette aligned with recovery status
  status_colors <- c(
    "Recovered"     = "#00833D",
    "Recovering"    = "#80BD41",
    "Not recovered" = "#E2231A"
  )

  gg_bcu <- ggplot(bcu_plot_df) +
    # segment connecting 2019 to 2024
    geom_segment(aes(x = `2019`, xend = `2024`, y = country, yend = country, color = recovery_label), linewidth = 0.7, alpha = 0.5) +
    # 2019 point (hollow)
    geom_point(aes(x = `2019`, y = country), shape = 21, fill = "white", color = "grey40", size = 2.5) +
    # 2024 point (solid, colored by status)
    geom_point(aes(x = `2024`, y = country, color = recovery_label), size = 3) +
    # IA2030 target line
    geom_vline(xintercept = 90, linetype = "dashed", color = "grey40", alpha = 0.7) +
    scale_color_manual(name = "Recovery status", values = status_colors) +
    scale_x_continuous(limits = c(0, 100), breaks = seq(0, 100, 20),
                       labels = function(x) paste0(x, "%")) +
    facet_grid(
      rows = vars(fct_relevel(recovery_label, "Recovered", "Recovering", "Not recovered")),
      scales = "free_y", space = "free_y"
    ) +
    labs(
      title    = str_glue("DTP3 coverage recovery from COVID-19 pandemic, {regn}"),
      subtitle = "Open circle = 2019 (pre-pandemic baseline); filled circle = 2024",
      x        = "DTP3 coverage (%)",
      y        = NULL,
      caption = paste0("Dotted line shows the IA2030 target of 90% coverage.\n", "Countries shown are Big Catch-Up priority countries in ", regn, ".\n", wuenic_src)) +
    theme_unicef() +
    theme(
      legend.position  = "none",      # facet labels make legend redundant
      strip.text       = element_text(face = "bold", size = 10),
      panel.grid.major.y = element_blank(),
      panel.grid.major.x = element_line(color = "grey90"),
      panel.grid.minor   = element_blank(),
      axis.text.y        = element_text(size = 9),
      plot.subtitle      = element_text(size = 9, color = "grey40"),
      plot.title         = element_text(size = 13, face = "bold", hjust = 0.5)
    )

  print(gg_bcu)
}
```

```{r recovery-methodology, echo=FALSE, results='asis'}
cat(paste0('
\\begin{tcolorbox}[
  colback=gray!8,
  colframe=unicefblue,
  arc=4pt,
  boxrule=1pt,
  title=\\textbf{\\color{white} Recovery Methodology},
  colbacktitle=unicefblue
]

\\textbf{Categorization Logic:} This analysis tracks DTP3 coverage trends from 2019 to ', rev_yr, ' to assess post-pandemic recovery. 
Countries are categorized based on their progress toward 2019 baseline levels:
\\begin{itemize}
    \\item \\textbf{Recovered:} 2024 coverage is equal to or higher than 2019 levels.
    \\item \\textbf{Recovering:} Coverage is still below 2019 levels but has improved by at least 3 percentage points from its lowest point during the pandemic.
    \\item \\textbf{Not Recovered:} Coverage remains below 2019 levels and has shown less than 3 percentage points of improvement from the pandemic low.
\\end{itemize}

\\end{tcolorbox}
\\vspace{0.2in}
'))
```

<!-- ===================== APPENDIX ===================== -->

\newpage

\begin{tcolorbox}[colback=unicefblue, colframe=unicefblue, arc=0mm, sharp corners]
\textcolor{white}{\large \textbf{Appendix}}
\end{tcolorbox}

```{r, echo=FALSE, results='asis'}
cat("\\phantomsection")
cat("\\hypertarget{sec:overview}{}")
cat("\\addcontentsline{toc}{section}{Appendix}")
```

\vspace{1em}

\textcolor{cyan}{\large WHO/UNICEF Estimates of National Immunization Coverage (WUENIC)}

\vspace{1em}

Each year, WHO and UNICEF review national immunization data from Member States, including administrative and official coverage, survey estimates and contextual information, and use rules-based data triangulation approach to assess coverage. Estimates are produced independently, without borrowing data from other countries, and are not based on ad hoc adjustments. In some cases, they rely on a single source, typically nationally reported coverage. When data for a specific country-vaccine-year are unavailable, values from adjacent years are interpolated or extrapolated, and conflicting sources are assessed to determine the most reliable estimate, considering potential biases and expert opinion.

\vspace{0.5em}
Methods:

\begin{itemize}
  \item \begin{sloppypar}
        \textcolor{blue}{\href{https://pubmed.ncbi.nlm.nih.gov/19649368/}{Burton et al. 2009. WHO and UNICEF estimates of national infant immunization coverage: methods and processes}}
        \end{sloppypar}
  \item \begin{sloppypar}
        \textcolor{blue}{\href{https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0047806}{Burton et al. 2012. A formal representation of the WHO and UNICEF estimates of national immunization coverage: a computational logic approach}}
        \end{sloppypar}
  \item \begin{sloppypar}
        \textcolor{blue}{\href{https://openpublichealthjournal.com/VOLUME/6/PAGE/73/ABSTRACT/}{Brown et al. 2013. An introduction to the grade of confidence used to characterize uncertainty around the WHO and UNICEF estimates of national immunization coverage}}
        \end{sloppypar}
\end{itemize}

\vspace{0.5em}

These talking points present the latest WUENIC estimates (published 15 July `r rev_yr+1`).

\vspace{1.5em}

\textcolor{cyan}{\fontsize{14}{16}\selectfont Definitions of immunization terms:}

\vspace{0.5em}

-   Bacillus Calmette-Guerin (BCG): vaccine against tuberculosis
\vspace{0.5em}
-   Hepatitis B birth dose, given within 24 hours after birth (HepBB)
\vspace{0.5em}
-   Diphtheria, tetanus, and pertussis vaccine, first dose (DTP1) and third dose (DTP3)
\vspace{0.5em}
-   Hepatitis B vaccine, third dose (HepB3)
\vspace{0.5em}
-   Haemophilus influenzae type B vaccine, third dose (Hib3)
\vspace{0.5em}
-   Inactivated polio vaccine, first dose (IPV1) and last dose of the primary series (IPVC)
\vspace{0.5em}
-   Measles containing vaccine, first dose (MCV1) and second dose (MCV2)
\vspace{0.5em}
-   Rotavirus vaccine, last dose (RotaC)
\vspace{0.5em}
-   Pneumococcal vaccine, second or third dose according to the schedule (PCVC)
\vspace{0.5em}
-   Yellow Fever vaccine (YFV)
\vspace{0.5em}
-   Meningococcal A vaccine (MengA)
\vspace{0.5em}
-   Human Papillomavirus vaccine, first dose (HPV1) and last dose (HPVc)

\vspace{1.5em}

\textcolor{cyan}{\fontsize{14}{16}\selectfont Additional resources:} \textcolor{blue}{\href{https://data.unicef.org/resources/immunization/}{\underline{https://data.unicef.org/resources/immunization/}}}
